---
title: "Skin Temperature (iButtons) Analysis"
subtitle: "Carryover effects in melatonin suppression"
author: "Letizia Woerrlein"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: show
    code-tools: true
    embed-resources: true
    theme: cosmo
    fig-width: 10
    fig-height: 8
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
execute:
  warning: false
  message: false
  cache: false
editor: visual
---

# Overview

This document processes iButton skin temperature data (ankle and neck sensors) for all participants across all laboratory sessions.

**Expected Output Dataset:** ibuttons with columns: - id: Participant ID (P01-P10, excluding P08 and P11) - timestamp: Date-time in format YYYYMMDDTHHMMSS - session: Session identifier (C01_01 = Group 1, Evening 1) - ankle: Skin temperature at ankle (°C) - neck: Skin temperature at neck (°C) - DPG: Distal-Proximal Gradient (ankle - neck, °C)

**Data Processing:** First single files (2 per person per evening) will be merged into one dataset, which will then be used for the subsequent analyses.

**Missing Data:** For session 01 for participant P01 the sampling rate of the iButton sensor on the neck was set to a sampling interval of 1h instead of 1m, therefore data is missing for P01 session 01. Furthermore due to issues with the iButtons software there is missing data for subject P05 for the last experimental session.

# Load Required Packages

```{r setup}
library(dplyr)
library(lubridate)
library(readxl)
library(tidyverse)
library(openxlsx)
library(DT)
library(kableExtra)
library(zoo)
library(patchwork)
library(ggtext)
library(lme4)        
library(lmerTest)   
library(performance) 
library(car)         
library(ggplot2)     
library(purrr)  
library(influence.ME) 
library(clubSandwich) 
library(nlme)         
library(moments)    
library(performance)
```

# Define Output Path

```{r define-output}
# Output directory and filename
output_dir <- "C:/Users/anton/OneDrive/AfterGlow/data/derivatives"
filename <- "01_ibuttons_cleaned_20250923T142900.xlsx"
filepath <- file.path(output_dir, filename)

# Check if file already exists
file_exists <- file.exists(filepath)
```

# Conditional Processing

```{r check-file-existence}
if (file_exists) {
  cat("✓ Cleaned iButtons dataset already exists at:\n")
  cat("  ", filepath, "\n")
  cat("  Skipping data processing.\n\n")
  cat("To regenerate the dataset, delete the existing file first.\n")
} else {
  cat("✗ Cleaned iButtons dataset not found.\n")
  cat("  Starting data processing...\n\n")
}
```

# Data Processing

## Session Times Definition

```{r session-times}
#| eval: !expr '!file_exists'

# Define session start and end times for each participant
P01_session_times <- tibble::tibble(
  session = as.character(101:111),
  start_time = rep("18:30", 11),
  end_time = rep("23:30", 11)
)

P02_session_times <- tibble::tibble(
  session = as.character(101:111),
  start_time = rep("18:30", 11),
  end_time = rep("23:30", 11)
)

P03_session_times <- tibble::tibble(
  session = as.character(101:111),
  start_time = rep("18:30", 11),
  end_time = rep("23:30", 11)
)

P04_session_times <- tibble::tibble(
  session = as.character(101:110),
  start_time = rep("16:30", 10),
  end_time = rep("21:30", 10)
)

P05_session_times <- tibble::tibble(
  session = as.character(101:110),
  start_time = rep("16:30", 10),
  end_time = rep("21:30", 10)
)

P06_session_times <- tibble::tibble(
  session = as.character(101:110),
  start_time = rep("18:00", 10),
  end_time = rep("23:00", 10)
)

P07_session_times <- tibble::tibble(
  session = as.character(101:110),
  start_time = rep("18:00", 10),
  end_time = rep("23:00", 10)
)

P09_session_times <- tibble::tibble(
  session = as.character(101:110),
  start_time = rep("18:00", 10),
  end_time = rep("23:00", 10)
)

P10_session_times <- tibble::tibble(
  session = as.character(101:113),
  start_time = rep("17:30", 13),
  end_time = rep("22:30", 13)
)
```

## Processing Function

```{r processing-function}
#| eval: !expr '!file_exists'

# Function to prepare iButton data with error handling
prepare_ibutton <- function(file_path, session, participant, position, session_times_df) {
  
  # Check if file exists
  if (!file.exists(file_path)) {
    cat("Warning: File does not exist:", file_path, "- Returning NA data\n")
    
    # Create a minimal dataset with NA values
    na_df <- data.frame(
      timestamp = "01.01.2000 00:00",
      stringsAsFactors = FALSE
    )
    na_df[[position]] <- NA_real_
    return(na_df)
  }
  
  # Extract session-specific start/end times
  times <- session_times_df %>% filter(session == !!session)
  start_time <- times$start_time
  end_time   <- times$end_time
  
  # Read Excel and process
  tryCatch({
    df <- read_excel(file_path, skip = 24) %>%
      mutate(
        Date = trimws(Date),
        Time = trimws(Time),
        datetime_full = parse_date_time(paste(Date, Time), orders = c("ymd HMS", "ymd HM")),
        datetime = round_date(datetime_full, unit = "minute"),
        datetime_formatted = format(datetime, "%d.%m.%Y %H:%M")
      ) %>%
      select(-datetime_full) %>%
      rename(timestamp = datetime_formatted) %>%
      select(timestamp, everything(), -Date, -Time, -datetime) %>%
      rename(!!paste0(position) := Value) %>%
      slice(-n()) %>%   
      mutate(
        timestamp_dt = dmy_hm(timestamp),
        time_only = format(timestamp_dt, "%H:%M")
      ) %>%
      filter(time_only >= start_time & time_only <= end_time) %>%
      select(-timestamp_dt, -time_only)
    
    return(df)
  }, error = function(e) {
    cat("Error reading file:", file_path, "- Error:", e$message, "- Returning NA data\n")
    
    # Create a minimal dataset with NA values
    na_df <- data.frame(
      timestamp = "01.01.2000 00:00",
      stringsAsFactors = FALSE
    )
    na_df[[position]] <- NA_real_
    return(na_df)
  })
}
```

## File Paths

```{r file-paths}
#| eval: !expr '!file_exists'

# Define sessions for each group
sessions_group1 <- as.character(101:111)  # P01, P02, P03 (11 sessions)
sessions_group2 <- as.character(101:110)  # P04, P05 (10 sessions)
sessions_group3 <- as.character(101:110)  # P06, P07, P09 (10 sessions)
sessions_group4 <- as.character(101:113)  # P10 (13 sessions)

# P01 file paths
ankle_paths_P01 <- paste0("C:/Users/anton/OneDrive/AfterGlow/data/raw/P01/",
                          sprintf("%02d", 1:11), "_expsession/ibuttons/01_ibutton-ankle_",
                          c("20250616T233000", "20250617T233000", "20250618T233000", "20250619T233000",
                            "20250620T233000", "20250621T233000", "20250622T233000", "20250623T233000",
                            "20250624T233000", "20250625T233000", "20250626T233000"), ".xlsx")

neck_paths_P01 <- paste0("C:/Users/anton/OneDrive/AfterGlow/data/raw/P01/",
                         sprintf("%02d", 1:11), "_expsession/ibuttons/01_ibutton-neck_",
                         c("20250616T233000", "20250617T233000", "20250618T233000", "20250619T233000",
                           "20250620T233000", "20250621T233000", "20250622T233000", "20250623T233000",
                           "20250624T233000", "20250625T233000", "20250626T233000"), ".xlsx")

# P02 file paths
ankle_paths_P02 <- paste0("C:/Users/anton/OneDrive/AfterGlow/data/raw/P02/",
                          sprintf("%02d", 1:11), "_expsession/ibuttons/01_ibutton-ankle_",
                          c("20250616T233000", "20250617T233000", "20250618T233000", "20250619T233000",
                            "20250620T233000", "20250621T233000", "20250622T233000", "20250623T233000",
                            "20250624T233000", "20250625T233000", "20250626T233000"), ".xlsx")

neck_paths_P02 <- paste0("C:/Users/anton/OneDrive/AfterGlow/data/raw/P02/",
                         sprintf("%02d", 1:11), "_expsession/ibuttons/01_ibutton-neck_",
                         c("20250616T233000", "20250617T233000", "20250618T233000", "20250619T233000",
                           "20250620T233000", "20250621T233000", "20250622T233000", "20250623T233000",
                           "20250624T233000", "20250625T233000", "20250626T233000"), ".xlsx")

# P03 file paths
ankle_paths_P03 <- paste0("C:/Users/anton/OneDrive/AfterGlow/data/raw/P03/",
                          sprintf("%02d", 1:11), "_expsession/ibuttons/01_ibutton-ankle_",
                          c("20250616T233000", "20250617T233000", "20250618T233000", "20250619T233000",
                            "20250620T233000", "20250621T233000", "20250622T233000", "20250623T233000",
                            "20250624T233000", "20250625T233000", "20250626T233000"), ".xlsx")

neck_paths_P03 <- paste0("C:/Users/anton/OneDrive/AfterGlow/data/raw/P03/",
                         sprintf("%02d", 1:11), "_expsession/ibuttons/01_ibutton-neck_",
                         c("20250616T233000", "20250617T233000", "20250618T233000", "20250619T233000",
                           "20250620T233000", "20250621T233000", "20250622T233000", "20250623T233000",
                           "20250624T233000", "20250625T233000", "20250626T233000"), ".xlsx")

# P04 file paths
ankle_paths_P04 <- paste0("C:/Users/anton/OneDrive/AfterGlow/data/raw/P04/",
                          sprintf("%02d", 1:10), "_expsession/ibuttons/01_ibutton-ankle_",
                          c("20250630T213000", "20250701T213000", "20250702T213000", "20250703T213000",
                            "20250704T213000", "20250705T213000", "20250706T213000", "20250707T213000",
                            "20250708T213000", "20250709T213000"), ".xlsx")

neck_paths_P04 <- paste0("C:/Users/anton/OneDrive/AfterGlow/data/raw/P04/",
                         sprintf("%02d", 1:10), "_expsession/ibuttons/01_ibutton-neck_",
                         c("20250630T213000", "20250701T213000", "20250702T213000", "20250703T213000",
                           "20250704T213000", "20250705T213000", "20250706T213000", "20250707T213000",
                           "20250708T213000", "20250709T213000"), ".xlsx")

# P05 file paths
ankle_paths_P05 <- paste0("C:/Users/anton/OneDrive/AfterGlow/data/raw/P05/",
                          sprintf("%02d", 1:10), "_expsession/ibuttons/01_ibutton-ankle_",
                          c("20250630T213000", "20250701T213000", "20250702T213000", "20250703T213000",
                            "20250704T213000", "20250705T213000", "20250706T213000", "20250707T213000",
                            "20250708T213000", "20250709T213000"), ".xlsx")

neck_paths_P05 <- paste0("C:/Users/anton/OneDrive/AfterGlow/data/raw/P05/",
                         sprintf("%02d", 1:10), "_expsession/ibuttons/01_ibutton-neck_",
                         c("20250630T213000", "20250701T213000", "20250702T213000", "20250703T213000",
                           "20250704T213000", "20250705T213000", "20250706T213000", "20250707T213000",
                           "20250708T213000", "20250709T213000"), ".xlsx")

# P06 file paths
ankle_paths_P06 <- paste0("C:/Users/anton/OneDrive/AfterGlow/data/raw/P06/",
                          sprintf("%02d", 1:10), "_expsession/ibuttons/01_ibutton-ankle_",
                          c("20250728T230000", "20250729T230000", "20250730T230000", "20250731T230000",
                            "20250801T230000", "20250802T230000", "20250803T230000", "20250804T230000",
                            "20250805T230000", "20250806T230000"), ".xlsx")

neck_paths_P06 <- paste0("C:/Users/anton/OneDrive/AfterGlow/data/raw/P06/",
                         sprintf("%02d", 1:10), "_expsession/ibuttons/01_ibutton-neck_",
                         c("20250728T230000", "20250729T230000", "20250730T230000", "20250731T230000",
                           "20250801T230000", "20250802T230000", "20250803T230000", "20250804T230000",
                           "20250805T230000", "20250806T230000"), ".xlsx")

# P07 file paths
ankle_paths_P07 <- paste0("C:/Users/anton/OneDrive/AfterGlow/data/raw/P07/",
                          sprintf("%02d", 1:10), "_expsession/ibuttons/01_ibutton-ankle_",
                          c("20250728T230000", "20250729T230000", "20250730T230000", "20250731T230000",
                            "20250801T230000", "20250802T230000", "20250803T230000", "20250804T230000",
                            "20250805T230000", "20250806T230000"), ".xlsx")

neck_paths_P07 <- paste0("C:/Users/anton/OneDrive/AfterGlow/data/raw/P07/",
                         sprintf("%02d", 1:10), "_expsession/ibuttons/01_ibutton-neck_",
                         c("20250728T230000", "20250729T230000", "20250730T230000", "20250731T230000",
                           "20250801T230000", "20250802T230000", "20250803T230000", "20250804T230000",
                           "20250805T230000", "20250806T230000"), ".xlsx")

# P09 file paths
ankle_paths_P09 <- paste0("C:/Users/anton/OneDrive/AfterGlow/data/raw/P09/",
                          sprintf("%02d", 1:10), "_expsession/ibuttons/01_ibutton-ankle_",
                          c("20250728T230000", "20250729T230000", "20250730T230000", "20250731T230000",
                            "20250801T230000", "20250802T230000", "20250803T230000", "20250804T230000",
                            "20250805T230000", "20250806T230000"), ".xlsx")

neck_paths_P09 <- paste0("C:/Users/anton/OneDrive/AfterGlow/data/raw/P09/",
                         sprintf("%02d", 1:10), "_expsession/ibuttons/01_ibutton-neck_",
                         c("20250728T230000", "20250729T230000", "20250730T230000", "20250731T230000",
                           "20250801T230000", "20250802T230000", "20250803T230000", "20250804T230000",
                           "20250805T230000", "20250806T230000"), ".xlsx")

# P10 file paths
ankle_paths_P10 <- paste0("C:/Users/anton/OneDrive/AfterGlow/data/raw/P10/",
                          sprintf("%02d", 1:13), "_expsession/ibuttons/01_ibutton-ankle_",
                          c("20250811T223000", "20250812T223000", "20250813T223000", "20250814T223000",
                            "20250815T223000", "20250816T223000", "20250817T223000", "20250818T223000",
                            "20250819T223000", "20250820T223000", "20250821T223000", "20250822T223000",
                            "20250823T223000"), ".xlsx")

neck_paths_P10 <- paste0("C:/Users/anton/OneDrive/AfterGlow/data/raw/P10/",
                         sprintf("%02d", 1:13), "_expsession/ibuttons/01_ibutton-neck_",
                         c("20250811T223000", "20250812T223000", "20250813T223000", "20250814T223000",
                           "20250815T223000", "20250816T223000", "20250817T223000", "20250818T223000",
                           "20250819T223000", "20250820T223000", "20250821T223000", "20250822T223000",
                           "20250823T223000"), ".xlsx")
```

## Helper Functions

```{r helper-functions}
#| eval: !expr '!file_exists'

# Function to get session times for participant
get_session_times <- function(participant) {
  switch(participant,
         "P01" = P01_session_times,
         "P02" = P02_session_times,
         "P03" = P03_session_times,
         "P04" = P04_session_times,
         "P05" = P05_session_times,
         "P06" = P06_session_times,
         "P07" = P07_session_times,
         "P09" = P09_session_times,
         "P10" = P10_session_times,
         stop("Unknown participant: ", participant)
  )
}

# Function to get sessions for participant (different for each group)
get_sessions <- function(participant) {
  if (participant %in% c("P01", "P02", "P03")) {
    return(sessions_group1)  # 101:111 (11 sessions)
  } else if (participant %in% c("P04", "P05")) {
    return(sessions_group2)  # 101:110 (10 sessions)  
  } else if (participant %in% c("P06", "P07", "P09")) {
    return(sessions_group3)  # 101:110 (10 sessions)
  } else if (participant == "P10") {
    return(sessions_group4)  # 101:113 (13 sessions)
  } else {
    stop("Unknown participant: ", participant)
  }
}

# Function to get file paths for participant and position
get_file_paths <- function(participant, position) {
  path_name <- paste0(tolower(position), "_paths_", participant)
  switch(path_name,
         "ankle_paths_P01" = ankle_paths_P01,
         "neck_paths_P01" = neck_paths_P01,
         "ankle_paths_P02" = ankle_paths_P02,
         "neck_paths_P02" = neck_paths_P02,
         "ankle_paths_P03" = ankle_paths_P03,
         "neck_paths_P03" = neck_paths_P03,
         "ankle_paths_P04" = ankle_paths_P04,
         "neck_paths_P04" = neck_paths_P04,
         "ankle_paths_P05" = ankle_paths_P05,
         "neck_paths_P05" = neck_paths_P05,
         "ankle_paths_P06" = ankle_paths_P06,
         "neck_paths_P06" = neck_paths_P06,
         "ankle_paths_P07" = ankle_paths_P07,
         "neck_paths_P07" = neck_paths_P07,
         "ankle_paths_P09" = ankle_paths_P09,
         "neck_paths_P09" = neck_paths_P09,
         "ankle_paths_P10" = ankle_paths_P10,
         "neck_paths_P10" = neck_paths_P10,
         stop("Unknown path combination: ", path_name)
  )
}

# Function: Convert session number to CXX_YY format based on participant
convert_session_format <- function(session_num, participant) {
  session_int <- as.numeric(session_num) - 100
  
  # Determine group based on participant
  if (participant %in% c("P01", "P02", "P03")) {
    return(paste0("C01_", sprintf("%02d", session_int)))
  } else if (participant %in% c("P04", "P05")) {
    return(paste0("C02_", sprintf("%02d", session_int)))
  } else if (participant %in% c("P06", "P07", "P09")) {
    return(paste0("C03_", sprintf("%02d", session_int)))
  } else if (participant == "P10") {
    return(paste0("C04_", sprintf("%02d", session_int)))
  } else {
    stop("Unknown participant group for: ", participant)
  }
}
```

## Process All Participants

```{r process-participants}
#| eval: !expr '!file_exists'

# Initialize storage
all_participants_data <- list()

# Process all participants
for (participant in c("P01", "P02", "P03", "P04", "P05", "P06", "P07", "P09", "P10")) {
  
  cat("\n=== Processing participant:", participant, "===\n")
  
  # Get session times and sessions for current participant
  session_times <- get_session_times(participant)
  participant_sessions <- get_sessions(participant)
  
  # Initialize list for current participant
  participant_skin <- list()
  
  # Process all sessions for current participant
  for (i in seq_along(participant_sessions)) {
    session <- participant_sessions[i]
    
    # Get file paths
    ankle_paths <- get_file_paths(participant, "ankle")
    neck_paths <- get_file_paths(participant, "neck")
    
    # Prepare ankle
    ankle_df <- prepare_ibutton(
      file_path = ankle_paths[i],
      session = session,
      participant = participant,
      position = "ankle",
      session_times_df = session_times
    )
    
    # Prepare neck
    neck_df <- prepare_ibutton(
      file_path = neck_paths[i],
      session = session,
      participant = participant,
      position = "neck",
      session_times_df = session_times
    )
    
    # Merge ankle + neck (handle case where both might be NA datasets)
    if (nrow(ankle_df) == 1 && nrow(neck_df) == 1 && 
        ankle_df$timestamp[1] == "01.01.2000 00:00" && 
        neck_df$timestamp[1] == "01.01.2000 00:00") {
      # Both files missing - create combined NA dataset
      participant_skin[[session]] <- data.frame(
        timestamp = "01.01.2000 00:00",
        ankle = NA_real_,
        neck = NA_real_,
        stringsAsFactors = FALSE
      )
      cat("  Session", session, "- Both files missing, created NA dataset\n")
    } else {
      # At least one file exists - merge normally
      participant_skin[[session]] <- full_join(ankle_df, neck_df, by = "timestamp")
      cat("  Session", session, "processed:", nrow(participant_skin[[session]]), "rows\n")
    }
  }
  
  # Store participant data
  all_participants_data[[participant]] <- participant_skin
}
```

## Format and Merge Data

```{r format-merge}
#| eval: !expr '!file_exists'

# Initialize final data list
ibuttons_list <- list()

# Loop through all participants and sessions
for (participant in names(all_participants_data)) {
  participant_data <- all_participants_data[[participant]]
  
  cat("\n=== Processing participant:", participant, "===\n")
  
  # Loop through all sessions for current participant
  for (session in names(participant_data)) {
    
    # Current dataset
    current_data <- participant_data[[session]]
    
    # Skip empty datasets
    if (nrow(current_data) == 0) {
      cat("Warning: Participant", participant, "Session", session, "is empty. Skipping...\n")
      next
    }
    
    # Format final dataset
    formatted_data <- current_data %>%
      mutate(
        # Add ID
        id = participant,
        
        # Convert session format based on participant group
        session = convert_session_format(session, participant),
        
        # Convert timestamp to YYYYMMDDTHHMMSS format
        timestamp_dt = dmy_hm(timestamp),
        timestamp = format(timestamp_dt, "%Y%m%dT%H%M%S"),
        
        # Ensure ankle and neck are numeric (convert character NAs to numeric NAs)
        ankle = as.numeric(ankle),
        neck = as.numeric(neck)
      ) %>%
      # Select columns and arrange in desired order
      select(id, timestamp, session, ankle, neck) %>%
      # Remove rows where timestamp conversion failed (but keep NA data rows)
      filter(!is.na(timestamp) | (is.na(ankle) & is.na(neck)))
    
    # Create unique identifier for list
    list_key <- paste(participant, session, sep = "_")
    ibuttons_list[[list_key]] <- formatted_data
    
    cat("Participant", participant, "Session", session, "processed:", nrow(formatted_data), "rows\n")
  }
}

# Merge all sessions and participants into a single dataset
ibuttons <- bind_rows(ibuttons_list)
```

## Calculate DPG

```{r calculate-dpg}
#| eval: !expr '!file_exists'

# Calculate temperature gradient: Distal-Proximal Gradient (DPG)
# DPG = ankle - neck
# If either ankle or neck is NA, gradient will be NA
ibuttons <- ibuttons %>%
  mutate(
    DPG = ankle - neck
  )
```

# Save Dataset

```{r save-dataset}
#| eval: !expr '!file_exists'

# Create output directory if it doesn't exist
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
  cat("Created directory:", output_dir, "\n")
}

# Convert entire dataset to character to preserve NAs as "NA" text
ibuttons_for_excel <- ibuttons %>%
  mutate(
    id = as.character(id),
    timestamp = as.character(timestamp),
    session = as.character(session),
    ankle = ifelse(is.na(ankle), "NA", as.character(ankle)),
    neck = ifelse(is.na(neck), "NA", as.character(neck)),
    DPG = ifelse(is.na(DPG), "NA", as.character(DPG))
  )

# Write to Excel
write.xlsx(ibuttons_for_excel, filepath)
cat("\n✓ Dataset saved successfully to:\n")
cat("  ", filepath, "\n")
```

# Load Cleaned Dataset

```{r load-data}
# Load cleaned iButtons dataset
ibuttons <- read_excel("C:/Users/anton/OneDrive/AfterGlow/data/derivatives/01_ibuttons_cleaned_20250923T142900.xlsx")

# Convert columns back to proper types
ibuttons <- ibuttons %>%
  mutate(
    id = as.character(id),
    timestamp = as.character(timestamp),
    session = as.character(session),
    ankle = as.numeric(ankle),
    neck = as.numeric(neck),
    DPG = as.numeric(DPG)
  )
```

# Clenaed Dataset Overview

```{r interactive-table}
# Create table
datatable(
  ibuttons,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    searching = TRUE,
    ordering = TRUE
  ),
  caption = "iButtons Skin Temperature Dataset"
)
```

# Calculate Rate of Change (ROC) for iButton data

Following a similar (slightly different) approach as van Marken Lichtenbelt et al. (2006) cleaning procedure

Step 1: Rate of Change (ROC) Detection: The rate of change (ROC) was calculated for each consecutive temperature measurement within each participant and session Session-specific interquartile ranges (IQR) were computed based on the 25th and 75th percentiles of ROC values Temperature changes exceeding 1.5 × IQR were flagged as artifacts, indicating sudden, implausible fluctuations

Step 2: Level-Based Outlier Detection: Absolute temperature values were evaluated using session-specific IQRs (25th and 75th percentiles) Values falling outside the range \[Q1 - 1.5×IQR, Q3 + 1.5×IQR\] were flagged as outliers

Step 3: Artifact Removal: If either Step 1 (ROC-based) or Step 2 (level-based) flagged a measurement as an artifact, the corresponding ankle and neck temperature values were set to NA Distal-proximal gradient (DPG) values were recalculated after artifact removal, with DPG set to NA if either ankle or neck temperature was missing

```{r calculate-roc}
# Calculate ROC separately for each participant, session, and measurement site
ibuttons <- ibuttons %>%
  group_by(id, session) %>%
  arrange(id, session, timestamp) %>%
  mutate(
    # ROC for ankle (change per minute)
    roc_ankle = ankle - lag(ankle),
    
    # ROC for neck (change per minute)
    roc_neck = neck - lag(neck)
  ) %>%
  ungroup()
```

# ROC by Participant and Session

```{r roc-by-session}
# Summary by participant and session
roc_summary <- ibuttons %>%
  group_by(id, session) %>%
  summarise(
    ankle_mean_roc = mean(abs(roc_ankle), na.rm = TRUE),
    ankle_max_roc = max(abs(roc_ankle), na.rm = TRUE),
    neck_mean_roc = mean(abs(roc_neck), na.rm = TRUE),
    neck_max_roc = max(abs(roc_neck), na.rm = TRUE),
    n_obs = n(),
    .groups = "drop"
  )

print(roc_summary)
```

# Step 1: ROC-based Outlier Detection (adapted for laboratory conditions)

Approach: Using 1.5×IQR instead of 1×IQR (van Marken Lichtenbelt et al. 2006 used 1×IQR for field studies)

```{r step1-roc-outliers}
# Calculate thresholds per participant and session (1.5×IQR for lab conditions)
roc_thresholds <- ibuttons %>%
  group_by(id, session) %>%
  summarise(
    ankle_lower = quantile(roc_ankle, 0.25, na.rm = TRUE) - 1.5 * IQR(roc_ankle, na.rm = TRUE),
    ankle_upper = quantile(roc_ankle, 0.75, na.rm = TRUE) + 1.5 * IQR(roc_ankle, na.rm = TRUE),
    neck_lower = quantile(roc_neck, 0.25, na.rm = TRUE) - 1.5 * IQR(roc_neck, na.rm = TRUE),
    neck_upper = quantile(roc_neck, 0.75, na.rm = TRUE) + 1.5 * IQR(roc_neck, na.rm = TRUE),
    .groups = "drop"
  )

# Flag outliers
ibuttons <- ibuttons %>%
  left_join(roc_thresholds, by = c("id", "session")) %>%
  mutate(
    outlier_roc_ankle = roc_ankle < ankle_lower | roc_ankle > ankle_upper,
    outlier_roc_neck = roc_neck < neck_lower | roc_neck > neck_upper
  ) %>%
  select(-ends_with("_lower"), -ends_with("_upper"))  # GEÄNDERT

cat("ROC outliers - Ankle:", sum(ibuttons$outlier_roc_ankle, na.rm = TRUE), "/", sum(!is.na(ibuttons$outlier_roc_ankle)), "\n")
cat("ROC outliers - Neck:", sum(ibuttons$outlier_roc_neck, na.rm = TRUE), "/", sum(!is.na(ibuttons$outlier_roc_neck)), "\n")
```

# Step 2: Level-based Outlier Detection

```{r step2-level-outliers}
# Calculate thresholds for absolute temperature values (1.5×IQR for lab conditions)
level_thresholds <- ibuttons %>%
  group_by(id, session) %>%
  summarise(
    ankle_lower = quantile(ankle, 0.25, na.rm = TRUE) - 1.5 * IQR(ankle, na.rm = TRUE),
    ankle_upper = quantile(ankle, 0.75, na.rm = TRUE) + 1.5 * IQR(ankle, na.rm = TRUE),
    neck_lower = quantile(neck, 0.25, na.rm = TRUE) - 1.5 * IQR(neck, na.rm = TRUE),
    neck_upper = quantile(neck, 0.75, na.rm = TRUE) + 1.5 * IQR(neck, na.rm = TRUE),
    .groups = "drop"
  )

# Flag level-based outliers
ibuttons <- ibuttons %>%
  left_join(level_thresholds, by = c("id", "session")) %>%
  mutate(
    outlier_level_ankle = ankle < ankle_lower | ankle > ankle_upper,
    outlier_level_neck = neck < neck_lower | neck > neck_upper
  ) %>%
  select(-ends_with("_lower"), -ends_with("_upper"))  # GEÄNDERT

cat("Level outliers - Ankle:", sum(ibuttons$outlier_level_ankle, na.rm = TRUE), "/", sum(!is.na(ibuttons$outlier_level_ankle)), "\n")
cat("Level outliers - Neck:", sum(ibuttons$outlier_level_neck, na.rm = TRUE), "/", sum(!is.na(ibuttons$outlier_level_neck)), "\n")
```

# Combine outlier flags

```{r combine-outliers}
ibuttons <- ibuttons %>%
  mutate(
    outlier_ankle = outlier_roc_ankle | outlier_level_ankle,
    outlier_neck = outlier_roc_neck | outlier_level_neck
  )

# Final summary
cat("=== Final Outlier Summary ===\n")
cat("Ankle:", sum(ibuttons$outlier_ankle, na.rm = TRUE), 
    "/", sum(!is.na(ibuttons$outlier_ankle)), 
    "(", round(100 * sum(ibuttons$outlier_ankle, na.rm = TRUE) / sum(!is.na(ibuttons$outlier_ankle)), 1), "%)\n")
cat("Neck:", sum(ibuttons$outlier_neck, na.rm = TRUE), 
    "/", sum(!is.na(ibuttons$outlier_neck)), 
    "(", round(100 * sum(ibuttons$outlier_neck, na.rm = TRUE) / sum(!is.na(ibuttons$outlier_neck)), 1), "%)\n")
```

# Calculate combined outlier rate

```{r calculate-combined-outliers}
ibuttons_with_any_outlier <- ibuttons %>%
  mutate(any_outlier = outlier_ankle | outlier_neck)

n_any_outlier <- sum(ibuttons_with_any_outlier$any_outlier, na.rm = TRUE)
n_total <- sum(!is.na(ibuttons_with_any_outlier$any_outlier))

cat("Rows with ANY outlier (ankle OR neck):", n_any_outlier, "/", n_total, 
    "(", round(100 * n_any_outlier / n_total, 1), "%)\n")
```

# Outlier distribution by session

```{r analyze-outliers-by-session}
outlier_by_session <- ibuttons %>%
  mutate(any_outlier = outlier_ankle | outlier_neck) %>%
  group_by(id, session) %>%
  summarise(
    n_total = n(),
    n_outliers = sum(any_outlier, na.rm = TRUE),
    pct_outliers = round(100 * n_outliers / n_total, 1),
    .groups = "drop"
  ) %>%
  arrange(desc(pct_outliers))

print(head(outlier_by_session, 20))
```

Interpretation: Especially P06 shows high numbers of outliers, therefore a sensitivity analysis will be conducted excluding P06 to test robustness of findings.

# Remove outliers by setting to NA

```{r remove-outliers-set-na}
ibuttons <- ibuttons %>%
  mutate(
    # Set ankle to NA if outlier
    ankle = if_else(outlier_ankle, NA_real_, ankle),
    
    # Set neck to NA if outlier
    neck = if_else(outlier_neck, NA_real_, neck),
    
    # Recalculate DPG with cleaned data
    DPG = ankle - neck
  )
```

# Interactive Table

```{r interactive-table-cleaned}
# Select relevant columns for display
ibuttons_display <- ibuttons %>%
  select(id, timestamp, session, ankle, neck, DPG)

datatable(
  ibuttons_display,
  options = list(
    pageLength = 25,
    scrollX = TRUE,
    searching = TRUE,
    ordering = TRUE
  ),
  caption = "Cleaned iButtons Dataset (outliers set to NA)",
  filter = "top"
)
```

# Calculate 30-minute block means for DPG

```{r calculate-dpg-30min-blocks}
ibuttons <- ibuttons %>%
  mutate(
    time = paste0(substr(timestamp, 10, 11), ":", substr(timestamp, 12, 13)),
    group = case_when(
      id %in% c("P01", "P02", "P03") ~ 1,
      id %in% c("P04", "P05") ~ 2,
      id %in% c("P06", "P07", "P09") ~ 3,
      id %in% c("P10") ~ 4
    ),
    # Assign time blocks based on group
    time_block = case_when(
      # Group 1: 18:30-23:30
      group == 1 & time >= "18:30" & time < "19:00" ~ "18:30-19:00",
      group == 1 & time >= "19:00" & time < "19:30" ~ "19:00-19:30",
      group == 1 & time >= "19:30" & time < "20:00" ~ "19:30-20:00",
      group == 1 & time >= "20:00" & time < "20:30" ~ "20:00-20:30",
      group == 1 & time >= "20:30" & time < "21:00" ~ "20:30-21:00",
      group == 1 & time >= "21:00" & time < "21:30" ~ "21:00-21:30",
      group == 1 & time >= "21:30" & time < "22:00" ~ "21:30-22:00",
      group == 1 & time >= "22:00" & time < "22:30" ~ "22:00-22:30",
      group == 1 & time >= "22:30" & time < "23:00" ~ "22:30-23:00",
      group == 1 & time >= "23:00" & time <= "23:30" ~ "23:00-23:30",
      
      # Group 2: 16:30-21:30
      group == 2 & time >= "16:30" & time < "17:00" ~ "16:30-17:00",
      group == 2 & time >= "17:00" & time < "17:30" ~ "17:00-17:30",
      group == 2 & time >= "17:30" & time < "18:00" ~ "17:30-18:00",
      group == 2 & time >= "18:00" & time < "18:30" ~ "18:00-18:30",
      group == 2 & time >= "18:30" & time < "19:00" ~ "18:30-19:00",
      group == 2 & time >= "19:00" & time < "19:30" ~ "19:00-19:30",
      group == 2 & time >= "19:30" & time < "20:00" ~ "19:30-20:00",
      group == 2 & time >= "20:00" & time < "20:30" ~ "20:00-20:30",
      group == 2 & time >= "20:30" & time < "21:00" ~ "20:30-21:00",
      group == 2 & time >= "21:00" & time <= "21:30" ~ "21:00-21:30",
      
      # Group 3: 18:00-23:00
      group == 3 & time >= "18:00" & time < "18:30" ~ "18:00-18:30",
      group == 3 & time >= "18:30" & time < "19:00" ~ "18:30-19:00",
      group == 3 & time >= "19:00" & time < "19:30" ~ "19:00-19:30",
      group == 3 & time >= "19:30" & time < "20:00" ~ "19:30-20:00",
      group == 3 & time >= "20:00" & time < "20:30" ~ "20:00-20:30",
      group == 3 & time >= "20:30" & time < "21:00" ~ "20:30-21:00",
      group == 3 & time >= "21:00" & time < "21:30" ~ "21:00-21:30",
      group == 3 & time >= "21:30" & time < "22:00" ~ "21:30-22:00",
      group == 3 & time >= "22:00" & time < "22:30" ~ "22:00-22:30",
      group == 3 & time >= "22:30" & time <= "23:00" ~ "22:30-23:00",
      
      # Group 4: 17:30-22:30
      group == 4 & time >= "17:30" & time < "18:00" ~ "17:30-18:00",
      group == 4 & time >= "18:00" & time < "18:30" ~ "18:00-18:30",
      group == 4 & time >= "18:30" & time < "19:00" ~ "18:30-19:00",
      group == 4 & time >= "19:00" & time < "19:30" ~ "19:00-19:30",
      group == 4 & time >= "19:30" & time < "20:00" ~ "19:30-20:00",
      group == 4 & time >= "20:00" & time < "20:30" ~ "20:00-20:30",
      group == 4 & time >= "20:30" & time < "21:00" ~ "20:30-21:00",
      group == 4 & time >= "21:00" & time < "21:30" ~ "21:00-21:30",
      group == 4 & time >= "21:30" & time < "22:00" ~ "21:30-22:00",
      group == 4 & time >= "22:00" & time <= "22:30" ~ "22:00-22:30",
      
      TRUE ~ NA_character_
    )
  ) %>%
  group_by(id, session, time_block) %>%
  mutate(DPG_30min = mean(DPG, na.rm = TRUE)) %>%
  ungroup()
```

# Create interactive table with 30-minute block means for DPG

```{r dpg-30min-table}
ibuttons %>%
  group_by(id, session, time_block) %>%
  slice(1) %>%
  ungroup() %>%
  select(id, session, time_block, DPG_30min) %>%
  filter(!is.na(time_block)) %>%
  arrange(id, session, time_block) %>%
  datatable(
    filter = 'top',
    options = list(
      pageLength = 50,
      order = list(list(0, 'asc'), list(1, 'asc'), list(2, 'asc')),
      scrollX = TRUE
    ),
    rownames = FALSE,
    caption = "Distal-Proximal Gradient (DPG) - 30-Minute Block Means per Participant and Session"
  ) %>%
  formatRound('DPG_30min', 2)
```

# DPG Visualization - 30-minute block means per participant and session

```{r dpg-visualization-setup}
#| label: dpg-visualization-setup

# Define participant information for light exposure patterns
participant_info <- list(
  P01 = list(hbt = 0, light_pattern = c("D", "D", "D", "B", "B", "B", "D", "D", "B", "D", "B"), group = 1),
  P02 = list(hbt = 0, light_pattern = c("D", "D", "D", "B", "B", "B", "D", "D", "B", "D", "B"), group = 1),
  P03 = list(hbt = 0, light_pattern = c("D", "D", "D", "B", "B", "B", "D", "D", "B", "D", "B"), group = 1),
  P04 = list(hbt = 22, light_pattern = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"), group = 2),
  P05 = list(hbt = 22, light_pattern = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"), group = 2),
  P06 = list(hbt = 23.5, light_pattern = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"), group = 3),
  P07 = list(hbt = 23.5, light_pattern = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"), group = 3),
  P09 = list(hbt = 23.5, light_pattern = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"), group = 3),
  P10 = list(hbt = 23, light_pattern = c("D", "D", "D", "B", NA, "D", "B", "B", "B", "D", "B", "D", "D"), group = 4)
)

# Function to create DPG plot for one participant
create_dpg_plot <- function(participant_id, info) {
  
  # Prepare data
  dpg_plot <- ibuttons %>%
    filter(id == participant_id) %>%
    filter(!(participant_id == "P10" & session == "C04_05")) %>%
    mutate(
      session_num = as.numeric(gsub("C0[1-4]_", "", session))
    ) %>%
    mutate(
      light_condition = info$light_pattern[session_num]
    ) %>%
    group_by(session_num, time_block) %>%
    slice(1) %>%
    ungroup() %>%
    mutate(
      start_hour = as.numeric(substr(time_block, 1, 2)),
      start_min = as.numeric(substr(time_block, 4, 5)),
      time_decimal_mid = start_hour + start_min/60 + 0.25,
      hours_before_hbt = case_when(
        info$hbt == 0 ~ 24 - time_decimal_mid,
        info$hbt == 22 ~ ifelse(time_decimal_mid >= 16.5, 22 - time_decimal_mid, NA_real_),
        info$hbt == 23 ~ ifelse(time_decimal_mid >= 17.5, 23 - time_decimal_mid, NA_real_),
        info$hbt == 23.5 ~ ifelse(time_decimal_mid >= 18, 23.5 - time_decimal_mid, NA_real_),
        TRUE ~ NA_real_
      )
    ) %>%
    filter(hours_before_hbt >= 0.5 & hours_before_hbt <= 5.5) %>%
    arrange(session_num, hours_before_hbt)
  
  if(nrow(dpg_plot) == 0) {
    warning(paste("No data available for", participant_id))
    return(NULL)
  }
  
  # Determine y-axis limits
  y_min <- floor(min(dpg_plot$DPG_30min, na.rm = TRUE) * 10) / 10 - 0.1
  y_max <- ceiling(max(dpg_plot$DPG_30min, na.rm = TRUE) * 10) / 10 + 0.1
  
  # X-axis labels with HBT times (odd blocks only)
  x_labels <- c(
    "B1\n(5.25h)", "B2", "B3\n(4.25h)", "B4", "B5\n(3.25h)",
    "B6", "B7\n(2.25h)", "B8", "B9\n(1.25h)", "B10"
  )
  
  # Create plot
  p <- ggplot(dpg_plot, aes(x = hours_before_hbt, y = DPG_30min)) +
    
    # Only yellow background for B evenings, no background for D evenings
    geom_rect(data = subset(dpg_plot, light_condition == "B") %>%
                group_by(session_num) %>% slice(1) %>% ungroup(),
              aes(xmin = 2.5, xmax = 1, ymin = -Inf, ymax = Inf),
              fill = "#FFF9E6", alpha = 0.8, inherit.aes = FALSE) +
    
    geom_line(color = "black", linewidth = 0.8) +
    geom_point(color = "black", size = 2.5) +
    
    scale_x_reverse(
      breaks = c(5.25, 4.75, 4.25, 3.75, 3.25, 2.75, 2.25, 1.75, 1.25, 0.75),
      labels = x_labels,
      limits = c(5.5, 0.5),
      expand = c(0, 0)
    ) +
    scale_y_continuous(
      limits = c(y_min, y_max),
      expand = expansion(mult = c(0.05, 0.05))
    ) +
    
    facet_wrap(~ session_num, ncol = 4,
               labeller = labeller(session_num = function(x) {
                 sapply(x, function(session) {
                   session_num <- as.numeric(session)
                   light_label <- if(session_num <= length(info$light_pattern) &&
                                    !is.na(info$light_pattern[session_num])) {
                     info$light_pattern[session_num]
                   } else {
                     "?"
                   }
                   paste0("Evening ", session, " (", light_label, ")")
                 })
               })) +
    
    labs(
      title = paste0("Distal-proximal skin temperature gradient per evening (", participant_id, ")"),
      x = "Block number (hours before HBT)",
      y = "DPG [°C]"
    ) +
    
    theme_classic(base_size = 11) +
    theme(
      aspect.ratio = 1,
      plot.title = element_text(size = 15, hjust = 0, face = "plain", margin = margin(b = 10)),
      axis.text = element_text(color = "black", size = 10),
      axis.text.x = element_text(size = 7),
      axis.title = element_text(size = 11),
      axis.line = element_line(linewidth = 0.5),
      axis.ticks = element_line(linewidth = 0.5),
      panel.grid = element_blank(),
      strip.background = element_rect(fill = "white", color = "grey80", linewidth = 0.5),
      strip.text = element_text(size = 10, color = "black"),
      panel.background = element_rect(fill = "white", color = "grey80", linewidth = 0.5),
      plot.margin = margin(10, 10, 10, 10),
      legend.position = "none"
    )
  
  return(p)
}
```

# DPG Plots for All Participants

```{r dpg-plots-all-participants, fig.width=12, fig.height=9}
#| label: fig-dpg-all-participants
#| fig-cap: "Distal-Proximal Gradient (DPG) per evening - all participants"
#| fig-width: 12
#| fig-height: 9

for (pid in names(participant_info)) {
  plot <- create_dpg_plot(pid, participant_info[[pid]])
  if (!is.null(plot)) {
    print(plot)
  }
}
```

# Calculate DPG_entire and DPG_intervention from cleaned minute-level DPG values

DPG_entire = mean of DPG values for entire 5h spent in the controlled climate chamber DPG_intervention = mean of DPG values for light intervention time window

```{r dpg-calculate-windows-from-raw}
ibuttons <- ibuttons %>%
  mutate(
    time = paste0(substr(timestamp, 10, 11), ":", substr(timestamp, 12, 13)),
    
    # Calculate hours before HBT per group
    start_hour = as.numeric(substr(timestamp, 10, 11)),
    start_min = as.numeric(substr(timestamp, 12, 13)),
    time_decimal = start_hour + start_min/60,
    
    hours_before_hbt = case_when(
      # Group 1: HBT = 00:00 (midnight)
      group == 1 ~ 24 - time_decimal,
      # Group 2: HBT = 22:00
      group == 2 ~ ifelse(time_decimal >= 16.5, 22 - time_decimal, NA_real_),
      # Group 3: HBT = 23:30
      group == 3 ~ ifelse(time_decimal >= 18, 23.5 - time_decimal, NA_real_),
      # Group 4: HBT = 23:00
      group == 4 ~ ifelse(time_decimal >= 17.5, 23 - time_decimal, NA_real_),
      TRUE ~ NA_real_
    ),
    
    # Define windows based on hours before HBT
    in_entire_window = (hours_before_hbt >= 0.5 & hours_before_hbt <= 5.5),
    in_intervention_window = (hours_before_hbt >= 1.0 & hours_before_hbt <= 2.5)
  ) %>%
  # Calculate means per session from minute-level DPG values 
  group_by(id, session) %>%
  mutate(
    DPG_entire = mean(DPG[in_entire_window], na.rm = TRUE),
    DPG_intervention = mean(DPG[in_intervention_window], na.rm = TRUE)
  ) %>%
  ungroup() %>%
  # Remove helper variables
  select(-start_hour, -start_min, -time_decimal, -hours_before_hbt, 
         -in_entire_window, -in_intervention_window)
```

```{r dpg-windows-table-from-raw}
# Create interactive table
ibuttons %>%
  group_by(id, session) %>%
  slice(1) %>%
  ungroup() %>%
  select(id, session, group, DPG_entire, DPG_intervention) %>%
  arrange(id, session) %>%
  datatable(
    filter = 'top',
    options = list(
      pageLength = 25,
      order = list(list(0, 'asc'), list(1, 'asc')),
      scrollX = TRUE
    ),
    rownames = FALSE,
    caption = "Distal-Proximal Gradient (DPG) - Entire 5h Window and Intervention Window (1.5h) Means per Participant and Session (calculated from minute-level cleaned data)"
  ) %>%
  formatRound(c('DPG_entire', 'DPG_intervention'), 2)
```

# DPG Descriptive Statistics 

```{r}
# DPG Descriptive Statistics 
# Define light patterns
light_patterns <- list(
  P01 = strsplit("DDDBBBDDBDB", "")[[1]],
  P02 = strsplit("DDDBBBDDBDB", "")[[1]],
  P03 = strsplit("DDDBBBDDBDB", "")[[1]],
  P04 = strsplit("DBBBDDDBDB", "")[[1]],
  P05 = strsplit("DBBBDDDBDB", "")[[1]],
  P06 = strsplit("DBBBDDDBDB", "")[[1]],
  P09 = strsplit("DBBBDDDBDB", "")[[1]],
  P10 = strsplit("DDDBXDBBBDBDD", "")[[1]]
)

# Add light condition to data
dpg_data <- ibuttons %>%
  filter(id != "P07") %>%
  mutate(
    session_num = as.numeric(sub("C\\d+_(\\d+)", "\\1", session)),
    light_condition = sapply(1:n(), function(i) light_patterns[[id[i]]][session_num[i]])
  )

# Aggregate to session level first
dpg_sessions <- dpg_data %>%
  filter(light_condition != "X") %>%  # Exclude P10 session 5
  group_by(id, session, light_condition) %>%
  slice(1) %>%  # Take one row per session
  ungroup()

# Calculate mean ± SD by light condition
dpg_summary <- dpg_sessions %>%
  group_by(light_condition) %>%
  summarise(
    DPG_entire = sprintf("%.2f ± %.2f", 
                         mean(DPG_entire, na.rm = TRUE), 
                         sd(DPG_entire, na.rm = TRUE)),
    DPG_intervention = sprintf("%.2f ± %.2f", 
                               mean(DPG_intervention, na.rm = TRUE), 
                               sd(DPG_intervention, na.rm = TRUE)),
    n_sessions = n()
  )

print(dpg_summary)
```

# DPG Doublet Analysis - 5h Exposure Window

```{r dpg-doublet-setup}
# Define light patterns for groups to make sure they are available
light_g1 <- unlist(strsplit("DDDBBBDDBDB", ""))
light_g2 <- unlist(strsplit("DBBBDDDBDB", ""))
light_g4_orig <- unlist(strsplit("DDDBXDBBBDBDD", ""))

# Create doublet dataframes
df_doublets_g1 <- data.frame(
  session_start = 1:(length(light_g1) - 1),
  session_end = 2:length(light_g1),
  doublet = paste0(light_g1[-length(light_g1)], light_g1[-1])
)

df_doublets_g2 <- data.frame(
  session_start = 1:(length(light_g2) - 1),
  session_end = 2:length(light_g2),
  doublet = paste0(light_g2[-length(light_g2)], light_g2[-1])
)

df_doublets_g4_orig <- data.frame(
  session_start = 1:(length(light_g4_orig) - 1),
  session_end = 2:length(light_g4_orig),
  doublet = paste0(light_g4_orig[-length(light_g4_orig)], light_g4_orig[-1])
)

# Filter out doublets containing X (excluded session)
df_doublets_g4 <- df_doublets_g4_orig %>% 
  filter(!str_detect(doublet, "X"))
```

```{r dpg-doublet-calculate-5h}
# Calculate doublet DPG for all participants - entire 5h window
participants <- c("P01", "P02", "P03", "P04", "P05", "P06", "P07", "P09", "P10")
doublets <- c("DD", "DB", "BD", "BB")

df_plot_dpg <- data.frame(
  doublet = character(),
  participant = character(),
  mean_dpg = numeric(),
  stringsAsFactors = FALSE
)

for(p in participants) {
  # Select appropriate doublet mapping
  if(p %in% c("P01", "P02", "P03")) {
    df_doublets <- df_doublets_g1
  } else if(p %in% c("P04", "P05", "P06", "P07", "P09")) {
    df_doublets <- df_doublets_g2
  } else {
    df_doublets <- df_doublets_g4
  }
  
  # Get DPG data (one row per session with DPG_entire)
  dpg_data <- ibuttons %>%
    filter(id == p) %>%
    # Exclude P10 session 5
    filter(!(p == "P10" & session == "C04_05")) %>%
    group_by(session) %>%
    slice(1) %>%
    ungroup() %>%
    mutate(exp_session = as.numeric(gsub("C0[1-4]_", "", session))) %>%
    select(exp_session, DPG_entire)
  
  # Join with doublets
  dpg_doublets <- df_doublets %>%
    rename(exp_session = session_end) %>%
    left_join(dpg_data, by = "exp_session")
  
  # Calculate mean per doublet
  mean_dpg_doublets <- dpg_doublets %>%
    group_by(doublet) %>%
    summarise(mean_dpg = mean(DPG_entire, na.rm = TRUE), .groups = "drop")
  
  # Add to plot data
  for(d in doublets) {
    if(d %in% mean_dpg_doublets$doublet) {
      df_plot_dpg <- rbind(df_plot_dpg, data.frame(
        doublet = d,
        participant = p,
        mean_dpg = mean_dpg_doublets$mean_dpg[mean_dpg_doublets$doublet == d]
      ))
    }
  }
}

df_plot_dpg$doublet <- factor(df_plot_dpg$doublet, levels = c("DD", "BD", "DB", "BB"))
```

```{r dpg-doublet-plot-5h}
#| label: fig-doublet-box-dpg-5h
#| fig-cap: "Mean DPG per doublet type (5-h exposure, P07 excluded)"
#| fig-width: 10
#| fig-height: 8

# Labels with bold second letter
doublet_labels <- c(
  "DD" = expression("D"*bold("D")),
  "BD" = expression("B"*bold("D")),
  "DB" = expression("D"*bold("B")),
  "BB" = expression("B"*bold("B"))
)

# Colors based on second letter
doublet_colors <- c(
  "DD" = "#4A90E2", "BD" = "#4A90E2",
  "DB" = "#F5A623", "BB" = "#F5A623"
)

# Exclude P07
df_plot_dpg_noP07 <- df_plot_dpg %>% filter(participant != "P07")

ggplot(df_plot_dpg_noP07, aes(x = doublet, y = mean_dpg)) +
  geom_boxplot(aes(fill = doublet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15, seed = 42),
             size = 2.5, alpha = 0.7, shape = 16) +
  scale_fill_manual(values = doublet_colors) +
  scale_x_discrete(labels = doublet_labels) +
  labs(
    title = "Distal-proximal gradient under light exposure doublets",
    subtitle = "Mean second-day DPG per doublet type (5-h exposure; P07 excluded)",
    x = "Doublet",
    y = "Mean Distal-Proximal Gradient (°C)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )
```

```{r dpg-doublet-table-5h}
# Create detailed overview table
doublet_overview_dpg <- df_plot_dpg_noP07 %>%
  arrange(doublet, participant) %>%
  select(participant, doublet, mean_dpg)

DT::datatable(
  doublet_overview_dpg,
  options = list(
    pageLength = 20,
    scrollX = TRUE,
    order = list(list(1, 'asc'), list(0, 'asc'))
  ),
  caption = "Individual doublet observations for DPG (5-h exposure, P07 excluded)",
  rownames = FALSE
) %>%
  DT::formatRound(columns = 'mean_dpg', digits = 2) %>%
  DT::formatStyle(
    'doublet',
    backgroundColor = DT::styleEqual(
      c('DD', 'BD', 'DB', 'BB'),
      c('#e3f2fd', '#e3f2fd', '#fff3e0', '#fff3e0')
    )
  )
```

# DPG Doublet Analysis - 1.5h Intervention Window

```{r dpg-doublet-calculate-intervention}
# Calculate doublet DPG for all participants - intervention window (1.5h)
participants_int <- c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10")
doublets <- c("DD", "DB", "BD", "BB")

df_plot_dpg_int <- data.frame(
  doublet = character(),
  participant = character(),
  mean_dpg_int = numeric(),
  stringsAsFactors = FALSE
)

for(p in participants_int) {
  # Select appropriate doublet mapping
  if(p %in% c("P01", "P02", "P03")) {
    df_doublets <- df_doublets_g1
  } else if(p %in% c("P04", "P05", "P06", "P09")) {
    df_doublets <- df_doublets_g2
  } else {
    df_doublets <- df_doublets_g4
  }
  
  # Get DPG intervention data (one row per session with DPG_intervention)
  dpg_data_int <- ibuttons %>%
    filter(id == p) %>%
    # Exclude P10 session 5
    filter(!(p == "P10" & session == "C04_05")) %>%
    group_by(session) %>%
    slice(1) %>%
    ungroup() %>%
    mutate(exp_session = as.numeric(gsub("C0[1-4]_", "", session))) %>%
    select(exp_session, DPG_intervention)
  
  # Join with doublets
  dpg_doublets_int <- df_doublets %>%
    rename(exp_session = session_end) %>%
    left_join(dpg_data_int, by = "exp_session")
  
  # Calculate mean per doublet
  mean_dpg_doublets_int <- dpg_doublets_int %>%
    group_by(doublet) %>%
    summarise(mean_dpg_int = mean(DPG_intervention, na.rm = TRUE), .groups = "drop")
  
  # Add to plot data
  for(d in doublets) {
    if(d %in% mean_dpg_doublets_int$doublet) {
      df_plot_dpg_int <- rbind(df_plot_dpg_int, data.frame(
        doublet = d,
        participant = p,
        mean_dpg_int = mean_dpg_doublets_int$mean_dpg_int[mean_dpg_doublets_int$doublet == d]
      ))
    }
  }
}

df_plot_dpg_int$doublet <- factor(df_plot_dpg_int$doublet, levels = c("DD", "BD", "DB", "BB"))
```

```{r dpg-doublet-plot-intervention}
#| label: fig-doublet-box-dpg-1.5h
#| fig-cap: "Mean DPG per doublet type (1.5-h intervention, P07 excluded)"
#| fig-width: 10
#| fig-height: 8

# Labels with bold second letter
doublet_labels <- c(
  "DD" = expression("D"*bold("D")),
  "BD" = expression("B"*bold("D")),
  "DB" = expression("D"*bold("B")),
  "BB" = expression("B"*bold("B"))
)

# Colors based on second letter
doublet_colors <- c(
  "DD" = "#4A90E2", "BD" = "#4A90E2",
  "DB" = "#F5A623", "BB" = "#F5A623"
)

# Exclude P07
df_plot_dpg_int_noP07 <- df_plot_dpg_int %>% filter(participant != "P07")

ggplot(df_plot_dpg_int_noP07, aes(x = doublet, y = mean_dpg_int)) +
  geom_boxplot(aes(fill = doublet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15, seed = 42),
             size = 2.5, alpha = 0.7, shape = 16) +
  scale_fill_manual(values = doublet_colors) +
  scale_x_discrete(labels = doublet_labels) +
  labs(
    title = "Distal-proximal gradient under light exposure doublets",
    subtitle = "Mean second-day DPG per doublet type (1.5-h intervention window; P07 excluded)",
    x = "Doublet",
    y = "Mean Distal-Proximal Gradient (°C)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )
```

```{r dpg-doublet-table-intervention}
# Create detailed overview table
doublet_overview_dpg_int <- df_plot_dpg_int_noP07 %>%
  arrange(doublet, participant) %>%
  select(participant, doublet, mean_dpg_int)

DT::datatable(
  doublet_overview_dpg_int,
  options = list(
    pageLength = 20,
    scrollX = TRUE,
    order = list(list(1, 'asc'), list(0, 'asc'))
  ),
  caption = "Individual doublet observations for DPG (1.5-h intervention window, P07 excluded)",
  rownames = FALSE
) %>%
  DT::formatRound(columns = 'mean_dpg_int', digits = 2) %>%
  DT::formatStyle(
    'doublet',
    backgroundColor = DT::styleEqual(
      c('DD', 'BD', 'DB', 'BB'),
      c('#e3f2fd', '#e3f2fd', '#fff3e0', '#fff3e0')
    )
  )
```

# DPG Doublet Position Analysis - 5h Exposure Window with Standard Error Bars

```{r dpg-doublet-position-setup}
# Create doublet position mappings
doublet_positions_g1 <- df_doublets_g1 %>%
  group_by(doublet) %>%
  summarise(
    position1_sessions = list(session_start),
    position2_sessions = list(session_end),
    .groups = "drop"
  )

doublet_positions_g2 <- df_doublets_g2 %>%
  group_by(doublet) %>%
  summarise(
    position1_sessions = list(session_start),
    position2_sessions = list(session_end),
    .groups = "drop"
  )

doublet_positions_g4 <- df_doublets_g4 %>%
  group_by(doublet) %>%
  summarise(
    position1_sessions = list(session_start),
    position2_sessions = list(session_end),
    .groups = "drop"
  )

# Function to calculate doublet positions for DPG (entire window)
calc_doublet_positions_dpg <- function(participant_id, dpg_data, doublet_map) {
  results <- doublet_map %>%
    rowwise() %>%
    mutate(
      pos1_dpg = mean(dpg_data$DPG_entire[dpg_data$exp_session %in% position1_sessions], 
                      na.rm = TRUE),
      pos2_dpg = mean(dpg_data$DPG_entire[dpg_data$exp_session %in% position2_sessions], 
                      na.rm = TRUE),
      participant = participant_id
    ) %>%
    ungroup() %>%
    select(participant, doublet, pos1_dpg, pos2_dpg)
  
  return(results)
}
```

```{r dpg-doublet-position-calculate-5h}
# Calculate position data for all participants (excluding P07) - 5h window
participants_noP07 <- c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10")
doublets <- c("DD", "DB", "BD", "BB")

mean_doublet_pos_dpg_noP07 <- data.frame(
  doublet = character(),
  position = character(),
  mean_dpg = numeric(),
  sd_dpg = numeric(),
  se_dpg = numeric(),
  n_participants = integer(),
  stringsAsFactors = FALSE
)

for(d in doublets) {
  for(pos in c("pos1", "pos2")) {
    values <- c()
    for(p in participants_noP07) {
      # Select mapping
      if(p %in% c("P01", "P02", "P03")) {
        doublet_map <- doublet_positions_g1
      } else if(p %in% c("P04", "P05", "P06", "P09")) {
        doublet_map <- doublet_positions_g2
      } else {
        doublet_map <- doublet_positions_g4
      }
      
      # Get DPG data
      dpg_data <- ibuttons %>%
        filter(id == p) %>%
        filter(!(p == "P10" & session == "C04_05")) %>%
        group_by(session) %>%
        slice(1) %>%
        ungroup() %>%
        mutate(exp_session = as.numeric(gsub("C0[1-4]_", "", session))) %>%
        select(exp_session, DPG_entire)
      
      # Calculate positions
      pos_dpg <- calc_doublet_positions_dpg(p, dpg_data, doublet_map)
      
      # Extract value
      if(d %in% pos_dpg$doublet) {
        if(pos == "pos1") {
          values <- c(values, pos_dpg$pos1_dpg[pos_dpg$doublet == d])
        } else {
          values <- c(values, pos_dpg$pos2_dpg[pos_dpg$doublet == d])
        }
      }
    }
    
    if(length(values) > 0) {
      n <- length(values)
      mean_doublet_pos_dpg_noP07 <- rbind(mean_doublet_pos_dpg_noP07, data.frame(
        doublet = d,
        position = pos,
        mean_dpg = mean(values, na.rm = TRUE),
        sd_dpg = sd(values, na.rm = TRUE),
        se_dpg = sd(values, na.rm = TRUE) / sqrt(n),  # Standard Error
        n_participants = n
      ))
    }
  }
}

# Prepare for plotting
df_plot_positions_dpg_noP07 <- mean_doublet_pos_dpg_noP07 %>%
  mutate(
    position_label = ifelse(position == "pos1", "Position 1", "Position 2"),
    position_num = ifelse(position == "pos1", 1, 2),
    light_type = ifelse(position == "pos1", 
                        substr(doublet, 1, 1),
                        substr(doublet, 2, 2)),
    light_label = ifelse(light_type == "D", "dim", "bright")
  ) %>%
  mutate(doublet = factor(doublet, levels = c("DD", "BD", "DB", "BB")))
```

```{r dpg-doublet-position-with-error-panel}
# Define color scheme
light_colors <- c("dim" = "#4A90E2", "bright" = "#F5A623")

# Main plot without error bars
p_main <- ggplot(df_plot_positions_dpg_noP07, 
                 aes(x = position_num, y = mean_dpg, group = doublet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_label), size = 5, shape = 16) +
  scale_color_manual(values = light_colors,
                     labels = c("dim" = "Dim (10 lx)", "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2), 
                     labels = c("Position 1", "Position 2"),
                     limits = c(0.8, 2.2)) +
  labs(
    title = "Distal-proximal gradient across light exposure doublets",
    subtitle = "Mean DPG across all participants (5-h exposure; P07 excluded)",
    x = "Doublet position",
    y = "Mean Distal-Proximal Gradient (°C)",
    color = "Light condition"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "right"
  )

# Error Bar Panel (right)
p_error <- ggplot(df_plot_positions_dpg_noP07, 
                  aes(x = doublet, y = mean_dpg, color = light_label)) +
  geom_errorbar(aes(ymin = mean_dpg - se_dpg, ymax = mean_dpg + se_dpg),
                width = 0.3, linewidth = 0.8) +
  geom_point(size = 3, shape = 16) +
  facet_wrap(~ position_label, ncol = 1, scales = "free_y") +
  scale_color_manual(values = light_colors) +
  labs(
    title = "SE per Doublet",
    y = NULL,
    x = "Doublet"
  ) +
  theme_classic(base_size = 11) +
  theme(
    aspect.ratio = 2,
    plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
    axis.text.x = element_text(size = 9, angle = 0),
    axis.text.y = element_text(size = 9),
    strip.text = element_text(size = 10, face = "bold"),
    strip.background = element_rect(fill = "grey90"),
    legend.position = "none",
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3)
  )

# Combine both plots
p_main + p_error + plot_layout(widths = c(3, 1))
```

# DPG Doublet Position Analysis - 1.5h Intervention Window with Standard Error Panel

```{r dpg-doublet-position-intervention-function}
# Function to calculate doublet positions for DPG (intervention window)
calc_doublet_positions_dpg_int <- function(participant_id, dpg_data_int, doublet_map) {
  results <- doublet_map %>%
    rowwise() %>%
    mutate(
      pos1_dpg_int = mean(dpg_data_int$DPG_intervention[dpg_data_int$exp_session %in% position1_sessions], 
                          na.rm = TRUE),
      pos2_dpg_int = mean(dpg_data_int$DPG_intervention[dpg_data_int$exp_session %in% position2_sessions], 
                          na.rm = TRUE),
      participant = participant_id
    ) %>%
    ungroup() %>%
    select(participant, doublet, pos1_dpg_int, pos2_dpg_int)
  
  return(results)
}
```

```{r dpg-doublet-position-calculate-intervention}
# Calculate position data for all participants (excluding P07) - intervention window
participants_noP07 <- c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10")
doublets <- c("DD", "DB", "BD", "BB")

mean_doublet_pos_dpg_int_noP07 <- data.frame(
  doublet = character(),
  position = character(),
  mean_dpg_int = numeric(),
  sd_dpg_int = numeric(),
  se_dpg_int = numeric(),
  n_participants = integer(),
  stringsAsFactors = FALSE
)

for(d in doublets) {
  for(pos in c("pos1", "pos2")) {
    values <- c()
    for(p in participants_noP07) {
      # Select mapping
      if(p %in% c("P01", "P02", "P03")) {
        doublet_map <- doublet_positions_g1
      } else if(p %in% c("P04", "P05", "P06", "P09")) {
        doublet_map <- doublet_positions_g2
      } else {
        doublet_map <- doublet_positions_g4
      }
      
      # Get DPG intervention data
      dpg_data_int <- ibuttons %>%
        filter(id == p) %>%
        filter(!(p == "P10" & session == "C04_05")) %>%
        group_by(session) %>%
        slice(1) %>%
        ungroup() %>%
        mutate(exp_session = as.numeric(gsub("C0[1-4]_", "", session))) %>%
        select(exp_session, DPG_intervention)
      
      # Calculate positions
      pos_dpg_int <- calc_doublet_positions_dpg_int(p, dpg_data_int, doublet_map)
      
      # Extract value
      if(d %in% pos_dpg_int$doublet) {
        if(pos == "pos1") {
          values <- c(values, pos_dpg_int$pos1_dpg_int[pos_dpg_int$doublet == d])
        } else {
          values <- c(values, pos_dpg_int$pos2_dpg_int[pos_dpg_int$doublet == d])
        }
      }
    }
    
    if(length(values) > 0) {
      n <- length(values)
      mean_doublet_pos_dpg_int_noP07 <- rbind(mean_doublet_pos_dpg_int_noP07, data.frame(
        doublet = d,
        position = pos,
        mean_dpg_int = mean(values, na.rm = TRUE),
        sd_dpg_int = sd(values, na.rm = TRUE),
        se_dpg_int = sd(values, na.rm = TRUE) / sqrt(n),
        n_participants = n
      ))
    }
  }
}

# Prepare for plotting
df_plot_positions_dpg_int_noP07 <- mean_doublet_pos_dpg_int_noP07 %>%
  mutate(
    position_label = ifelse(position == "pos1", "Position 1", "Position 2"),
    position_num = ifelse(position == "pos1", 1, 2),
    light_type = ifelse(position == "pos1", 
                        substr(doublet, 1, 1),
                        substr(doublet, 2, 2)),
    light_label = ifelse(light_type == "D", "dim", "bright")
  ) %>%
  mutate(doublet = factor(doublet, levels = c("DD", "BD", "DB", "BB")))
```

```{r dpg-doublet-position-intervention-with-error-panel}
# Define colors
light_colors <- c("dim" = "#4A90E2", "bright" = "#F5A623")

# Main plot without error bars
p_main_int <- ggplot(df_plot_positions_dpg_int_noP07, 
                     aes(x = position_num, y = mean_dpg_int, group = doublet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_label), size = 5, shape = 16) +
  scale_color_manual(values = light_colors,
                     labels = c("dim" = "Dim (10 lx)", "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2), 
                     labels = c("Position 1", "Position 2"),
                     limits = c(0.8, 2.2)) +
  labs(
    title = "Distal-proximal gradient across light exposure doublets",
    subtitle = "Mean DPG across all participants (1.5-h intervention window; P07 excluded)",
    x = "Doublet position",
    y = "Mean Distal-Proximal Gradient (°C)",
    color = "Light condition"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "right"
  )

# Error bar panel (right)
p_error_int <- ggplot(df_plot_positions_dpg_int_noP07, 
                      aes(x = doublet, y = mean_dpg_int, color = light_label)) +
  geom_errorbar(aes(ymin = mean_dpg_int - se_dpg_int, ymax = mean_dpg_int + se_dpg_int),
                width = 0.3, linewidth = 0.8) +
  geom_point(size = 3, shape = 16) +
  facet_wrap(~ position_label, ncol = 1, scales = "free_y") +
  scale_color_manual(values = light_colors) +
  labs(
    title = "SE per Doublet",
    y = NULL,
    x = "Doublet"
  ) +
  theme_classic(base_size = 11) +
  theme(
    aspect.ratio = 2,
    plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
    axis.text.x = element_text(size = 9, angle = 0),
    axis.text.y = element_text(size = 9),
    strip.text = element_text(size = 10, face = "bold"),
    strip.background = element_rect(fill = "grey90"),
    legend.position = "none",
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3)
  )

# Combine both plots
p_main_int + p_error_int + plot_layout(widths = c(3, 1))
```

# DPG Triplet Analysis - 5h Exposure Window

```{r dpg-triplet-setup}
# Define light patterns for triplets to make sure they are avialable 
light_g1 <- unlist(strsplit("DDDBBBDDBDB", ""))
light_g2 <- unlist(strsplit("DBBBDDDBDB", ""))
light_g4_orig <- unlist(strsplit("DDDBXDBBBDBDD", ""))

# Create triplet dataframes
df_triplets_g1 <- data.frame(
  session_start = 1:(length(light_g1) - 2),
  session_mid = 2:(length(light_g1) - 1),
  session_end = 3:length(light_g1),
  triplet = paste0(light_g1[1:(length(light_g1) - 2)], 
                   light_g1[2:(length(light_g1) - 1)], 
                   light_g1[3:length(light_g1)])
)

df_triplets_g2 <- data.frame(
  session_start = 1:(length(light_g2) - 2),
  session_mid = 2:(length(light_g2) - 1),
  session_end = 3:length(light_g2),
  triplet = paste0(light_g2[1:(length(light_g2) - 2)], 
                   light_g2[2:(length(light_g2) - 1)], 
                   light_g2[3:length(light_g2)])
)

df_triplets_g4_orig <- data.frame(
  session_start = 1:(length(light_g4_orig) - 2),
  session_mid = 2:(length(light_g4_orig) - 1),
  session_end = 3:length(light_g4_orig),
  triplet = paste0(light_g4_orig[1:(length(light_g4_orig) - 2)], 
                   light_g4_orig[2:(length(light_g4_orig) - 1)], 
                   light_g4_orig[3:length(light_g4_orig)])
)

# Filter out triplets containing X (excluded session 5 for P10)
df_triplets_g4 <- df_triplets_g4_orig %>% 
  filter(!str_detect(triplet, "X"))
```

```{r dpg-triplet-calculate-5h}
# Calculate triplet DPG for all participants - entire 5h window
participants <- c("P01", "P02", "P03", "P04", "P05", "P06", "P07", "P09", "P10")
triplets <- c("DDD", "DDB", "DBD", "DBB", "BDD", "BDB", "BBD", "BBB")

df_plot_dpg_triplet <- data.frame(
  triplet = character(),
  participant = character(),
  mean_dpg = numeric(),
  stringsAsFactors = FALSE
)

for(p in participants) {
  # Select appropriate triplet mapping
  if(p %in% c("P01", "P02", "P03")) {
    df_triplets <- df_triplets_g1
  } else if(p %in% c("P04", "P05", "P06", "P07", "P09")) {
    df_triplets <- df_triplets_g2
  } else {
    df_triplets <- df_triplets_g4
  }
  
  # Get DPG data
  dpg_data <- ibuttons %>%
    filter(id == p) %>%
    filter(!(p == "P10" & session == "C04_05")) %>%
    group_by(session) %>%
    slice(1) %>%
    ungroup() %>%
    mutate(exp_session = as.numeric(gsub("C0[1-4]_", "", session))) %>%
    select(exp_session, DPG_entire)
  
  # Join with triplets
  dpg_triplets <- df_triplets %>%
    rename(exp_session = session_end) %>%
    left_join(dpg_data, by = "exp_session")
  
  # Calculate mean per triplet
  mean_dpg_triplets <- dpg_triplets %>%
    group_by(triplet) %>%
    summarise(mean_dpg = mean(DPG_entire, na.rm = TRUE), .groups = "drop")
  
  # Add to plot data
  for(t in triplets) {
    if(t %in% mean_dpg_triplets$triplet) {
      df_plot_dpg_triplet <- rbind(df_plot_dpg_triplet, data.frame(
        triplet = t,
        participant = p,
        mean_dpg = mean_dpg_triplets$mean_dpg[mean_dpg_triplets$triplet == t]
      ))
    }
  }
}

df_plot_dpg_triplet$triplet <- factor(df_plot_dpg_triplet$triplet, 
                                      levels = c("DDD", "BDD", "DBD", "BBD", 
                                                "DDB", "BDB", "DBB", "BBB"))
```

```{r dpg-triplet-plot-5h}
#| label: fig-triplet-box-dpg-5h
#| fig-cap: "Mean DPG per triplet type (5-h exposure, P07 excluded)"
#| fig-width: 10
#| fig-height: 8

# Labels with bold third letter
triplet_labels <- c(
  "DDD" = expression("DD"*bold("D")),
  "BDD" = expression("BD"*bold("D")),
  "DBD" = expression("DB"*bold("D")),
  "BBD" = expression("BB"*bold("D")),
  "DDB" = expression("DD"*bold("B")),
  "BDB" = expression("BD"*bold("B")),
  "DBB" = expression("DB"*bold("B")),
  "BBB" = expression("BB"*bold("B"))
)

# Colors based on third letter
triplet_colors <- c(
  "DDD" = "#4A90E2", "BDD" = "#4A90E2", "DBD" = "#4A90E2", "BBD" = "#4A90E2",
  "DDB" = "#F5A623", "BDB" = "#F5A623", "DBB" = "#F5A623", "BBB" = "#F5A623"
)

# Exclude P07
df_plot_dpg_triplet_noP07 <- df_plot_dpg_triplet %>% filter(participant != "P07")

ggplot(df_plot_dpg_triplet_noP07, aes(x = triplet, y = mean_dpg)) +
  geom_boxplot(aes(fill = triplet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15, seed = 42),
             size = 2.5, alpha = 0.7, shape = 16) +
  scale_fill_manual(values = triplet_colors) +
  scale_x_discrete(labels = triplet_labels) +
  labs(
    title = "Distal-proximal gradient under light exposure triplets",
    subtitle = "Mean third-day DPG per triplet type (5-h exposure; P07 excluded)",
    x = "Triplet",
    y = "Mean Distal-Proximal Gradient (°C)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )
```

```{r dpg-triplet-table-5h}
# Create detailed overview table
triplet_overview_dpg <- df_plot_dpg_triplet_noP07 %>%
  arrange(triplet, participant) %>%
  select(participant, triplet, mean_dpg)

DT::datatable(
  triplet_overview_dpg,
  options = list(
    pageLength = 20,
    scrollX = TRUE,
    order = list(list(1, 'asc'), list(0, 'asc'))
  ),
  caption = "Individual triplet observations for DPG (5-h exposure, P07 excluded)",
  rownames = FALSE
) %>%
  DT::formatRound(columns = 'mean_dpg', digits = 2) %>%
  DT::formatStyle(
    'triplet',
    backgroundColor = DT::styleEqual(
      c('DDD', 'BDD', 'DBD', 'BBD', 'DDB', 'BDB', 'DBB', 'BBB'),
      c('#e3f2fd', '#e3f2fd', '#e3f2fd', '#e3f2fd', 
        '#fff3e0', '#fff3e0', '#fff3e0', '#fff3e0')
    )
  )
```

# DPG Triplet Analysis - 1.5h Intervention Window

```{r dpg-triplet-calculate-intervention}
# Calculate triplet DPG for all participants - intervention window (1.5h)
participants_int <- c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10")
triplets <- c("DDD", "DDB", "DBD", "DBB", "BDD", "BDB", "BBD", "BBB")

df_plot_dpg_triplet_int <- data.frame(
  triplet = character(),
  participant = character(),
  mean_dpg_int = numeric(),
  stringsAsFactors = FALSE
)

for(p in participants_int) {
  # Select appropriate triplet mapping
  if(p %in% c("P01", "P02", "P03")) {
    df_triplets <- df_triplets_g1
  } else if(p %in% c("P04", "P05", "P06", "P09")) {
    df_triplets <- df_triplets_g2
  } else {
    df_triplets <- df_triplets_g4
  }
  
  # Get DPG intervention data
  dpg_data_int <- ibuttons %>%
    filter(id == p) %>%
    filter(!(p == "P10" & session == "C04_05")) %>%
    group_by(session) %>%
    slice(1) %>%
    ungroup() %>%
    mutate(exp_session = as.numeric(gsub("C0[1-4]_", "", session))) %>%
    select(exp_session, DPG_intervention)
  
  # Join with triplets
  dpg_triplets_int <- df_triplets %>%
    rename(exp_session = session_end) %>%
    left_join(dpg_data_int, by = "exp_session")
  
  # Calculate mean per triplet
  mean_dpg_triplets_int <- dpg_triplets_int %>%
    group_by(triplet) %>%
    summarise(mean_dpg_int = mean(DPG_intervention, na.rm = TRUE), .groups = "drop")
  
  # Add to plot data
  for(t in triplets) {
    if(t %in% mean_dpg_triplets_int$triplet) {
      df_plot_dpg_triplet_int <- rbind(df_plot_dpg_triplet_int, data.frame(
        triplet = t,
        participant = p,
        mean_dpg_int = mean_dpg_triplets_int$mean_dpg_int[mean_dpg_triplets_int$triplet == t]
      ))
    }
  }
}

df_plot_dpg_triplet_int$triplet <- factor(df_plot_dpg_triplet_int$triplet, 
                                          levels = c("DDD", "BDD", "DBD", "BBD", 
                                                    "DDB", "BDB", "DBB", "BBB"))
```

```{r dpg-triplet-plot-intervention}
#| label: fig-triplet-box-dpg-1.5h
#| fig-cap: "Mean DPG per triplet type (1.5-h intervention, P07 excluded)"
#| fig-width: 10
#| fig-height: 8

# Labels with bold third letter
triplet_labels <- c(
  "DDD" = expression("DD"*bold("D")),
  "BDD" = expression("BD"*bold("D")),
  "DBD" = expression("DB"*bold("D")),
  "BBD" = expression("BB"*bold("D")),
  "DDB" = expression("DD"*bold("B")),
  "BDB" = expression("BD"*bold("B")),
  "DBB" = expression("DB"*bold("B")),
  "BBB" = expression("BB"*bold("B"))
)

# Colors based on third letter
triplet_colors <- c(
  "DDD" = "#4A90E2", "BDD" = "#4A90E2", "DBD" = "#4A90E2", "BBD" = "#4A90E2",
  "DDB" = "#F5A623", "BDB" = "#F5A623", "DBB" = "#F5A623", "BBB" = "#F5A623"
)

# Exclude P07
df_plot_dpg_triplet_int_noP07 <- df_plot_dpg_triplet_int %>% filter(participant != "P07")

ggplot(df_plot_dpg_triplet_int_noP07, aes(x = triplet, y = mean_dpg_int)) +
  geom_boxplot(aes(fill = triplet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15, seed = 42),
             size = 2.5, alpha = 0.7, shape = 16) +
  scale_fill_manual(values = triplet_colors) +
  scale_x_discrete(labels = triplet_labels) +
  labs(
    title = "Distal-proximal gradient under light exposure triplets",
    subtitle = "Mean third-day DPG per triplet type (1.5-h intervention window; P07 excluded)",
    x = "Triplet",
    y = "Mean Distal-Proximal Gradient (°C)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )
```

```{r dpg-triplet-table-intervention}
# Create detailed overview table
triplet_overview_dpg_int <- df_plot_dpg_triplet_int_noP07 %>%
  arrange(triplet, participant) %>%
  select(participant, triplet, mean_dpg_int)

DT::datatable(
  triplet_overview_dpg_int,
  options = list(
    pageLength = 20,
    scrollX = TRUE,
    order = list(list(1, 'asc'), list(0, 'asc'))
  ),
  caption = "Individual triplet observations for DPG (1.5-h intervention window, P07 excluded)",
  rownames = FALSE
) %>%
  DT::formatRound(columns = 'mean_dpg_int', digits = 2) %>%
  DT::formatStyle(
    'triplet',
    backgroundColor = DT::styleEqual(
      c('DDD', 'BDD', 'DBD', 'BBD', 'DDB', 'BDB', 'DBB', 'BBB'),
      c('#e3f2fd', '#e3f2fd', '#e3f2fd', '#e3f2fd', 
        '#fff3e0', '#fff3e0', '#fff3e0', '#fff3e0')
    )
  )
```

# DPG Triplet Position Analysis - 5h Exposure Window

```{r dpg-triplet-position-setup}
# Create triplet position mappings
triplet_positions_g1 <- df_triplets_g1 %>%
  group_by(triplet) %>%
  summarise(
    position1_sessions = list(session_start),
    position2_sessions = list(session_mid),
    position3_sessions = list(session_end),
    .groups = "drop"
  )

triplet_positions_g2 <- df_triplets_g2 %>%
  group_by(triplet) %>%
  summarise(
    position1_sessions = list(session_start),
    position2_sessions = list(session_mid),
    position3_sessions = list(session_end),
    .groups = "drop"
  )

triplet_positions_g4 <- df_triplets_g4 %>%
  group_by(triplet) %>%
  summarise(
    position1_sessions = list(session_start),
    position2_sessions = list(session_mid),
    position3_sessions = list(session_end),
    .groups = "drop"
  )

# Function to calculate triplet positions for DPG (entire window)
calc_triplet_positions_dpg <- function(participant_id, dpg_data, triplet_map) {
  results <- triplet_map %>%
    rowwise() %>%
    mutate(
      pos1_dpg = mean(dpg_data$DPG_entire[dpg_data$exp_session %in% position1_sessions], 
                      na.rm = TRUE),
      pos2_dpg = mean(dpg_data$DPG_entire[dpg_data$exp_session %in% position2_sessions], 
                      na.rm = TRUE),
      pos3_dpg = mean(dpg_data$DPG_entire[dpg_data$exp_session %in% position3_sessions], 
                      na.rm = TRUE),
      participant = participant_id
    ) %>%
    ungroup() %>%
    select(participant, triplet, pos1_dpg, pos2_dpg, pos3_dpg)
  
  return(results)
}

# Function for intervention window
calc_triplet_positions_dpg_int <- function(participant_id, dpg_data_int, triplet_map) {
  results <- triplet_map %>% 
    rowwise() %>%
    mutate(
      pos1_dpg_int = mean(dpg_data_int$DPG_intervention[dpg_data_int$exp_session %in% position1_sessions], 
                          na.rm = TRUE),
      pos2_dpg_int = mean(dpg_data_int$DPG_intervention[dpg_data_int$exp_session %in% position2_sessions], 
                          na.rm = TRUE),
      pos3_dpg_int = mean(dpg_data_int$DPG_intervention[dpg_data_int$exp_session %in% position3_sessions], 
                          na.rm = TRUE),
      participant = participant_id
    ) %>%
    ungroup() %>%
    select(participant, triplet, pos1_dpg_int, pos2_dpg_int, pos3_dpg_int)
  
  return(results)
}
```

```{r dpg-triplet-position-calculate-5h}
# Calculate position data for all participants (excluding P07) - 5h window
participants_noP07 <- c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10")
triplets <- c("DDD", "DDB", "DBD", "DBB", "BDD", "BDB", "BBD", "BBB")

mean_triplet_pos_dpg_noP07 <- data.frame(
  triplet = character(),
  position = character(),
  mean_dpg = numeric(),
  sd_dpg = numeric(),
  se_dpg = numeric(),
  n_participants = integer(),
  stringsAsFactors = FALSE
)

for(t in triplets) {
  for(pos in c("pos1", "pos2", "pos3")) {
    values <- c()
    for(p in participants_noP07) {
      # Select mapping
      if(p %in% c("P01", "P02", "P03")) {
        triplet_map <- triplet_positions_g1
      } else if(p %in% c("P04", "P05", "P06", "P09")) {
        triplet_map <- triplet_positions_g2
      } else {
        triplet_map <- triplet_positions_g4
      }
      
      # Get DPG data
      dpg_data <- ibuttons %>%
        filter(id == p) %>%
        filter(!(p == "P10" & session == "C04_05")) %>%
        group_by(session) %>%
        slice(1) %>%
        ungroup() %>%
        mutate(exp_session = as.numeric(gsub("C0[1-4]_", "", session))) %>%
        select(exp_session, DPG_entire)
      
      # Calculate positions
      pos_dpg <- calc_triplet_positions_dpg(p, dpg_data, triplet_map)
      
      # Extract value
      if(t %in% pos_dpg$triplet) {
        if(pos == "pos1") {
          values <- c(values, pos_dpg$pos1_dpg[pos_dpg$triplet == t])
        } else if(pos == "pos2") {
          values <- c(values, pos_dpg$pos2_dpg[pos_dpg$triplet == t])
        } else {
          values <- c(values, pos_dpg$pos3_dpg[pos_dpg$triplet == t])
        }
      }
    }
    
    if(length(values) > 0) {
      n <- length(values)
      mean_triplet_pos_dpg_noP07 <- rbind(mean_triplet_pos_dpg_noP07, data.frame(
        triplet = t,
        position = pos,
        mean_dpg = mean(values, na.rm = TRUE),
        sd_dpg = sd(values, na.rm = TRUE),
        se_dpg = sd(values, na.rm = TRUE) / sqrt(n),
        n_participants = n
      ))
    }
  }
}

# Prepare for plotting
df_plot_positions_triplet_dpg_noP07 <- mean_triplet_pos_dpg_noP07 %>%
  mutate(
    position_label = case_when(
      position == "pos1" ~ "Position 1",
      position == "pos2" ~ "Position 2",
      position == "pos3" ~ "Position 3"
    ),
    position_num = case_when(
      position == "pos1" ~ 1,
      position == "pos2" ~ 2,
      position == "pos3" ~ 3
    ),
    light_type = case_when(
      position == "pos1" ~ substr(triplet, 1, 1),
      position == "pos2" ~ substr(triplet, 2, 2),
      position == "pos3" ~ substr(triplet, 3, 3)
    ),
    light_label = ifelse(light_type == "D", "dim", "bright")
  ) %>%
  mutate(triplet = factor(triplet, levels = c("DDD", "BDD", "DBD", "BBD", 
                                              "DDB", "BDB", "DBB", "BBB")))
```

```{r dpg-triplet-position-plot-5h-with-error-panel}
# Define colors
light_colors <- c("dim" = "#4A90E2", "bright" = "#F5A623")

# Main plot without error bars
p_main_triplet <- ggplot(df_plot_positions_triplet_dpg_noP07, 
                         aes(x = position_num, y = mean_dpg, group = triplet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_label), size = 5, shape = 16) +
  scale_color_manual(values = light_colors,
                     labels = c("dim" = "Dim (10 lx)", "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2, 3), 
                     labels = c("Position 1", "Position 2", "Position 3"),
                     limits = c(0.8, 3.2)) +
  labs(
    title = "Distal-proximal gradient across light exposure triplets",
    subtitle = "Mean DPG across all participants (5-h exposure; P07 excluded)",
    x = "Triplet position",
    y = "Mean Distal-Proximal Gradient (°C)",
    color = "Light condition"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "right"
  )

# Error Bar Panel 
p_error_triplet <- ggplot(df_plot_positions_triplet_dpg_noP07, 
                          aes(x = triplet, y = mean_dpg, color = light_label)) +
  geom_errorbar(aes(ymin = mean_dpg - se_dpg, ymax = mean_dpg + se_dpg),
                width = 0.3, linewidth = 0.8) +
  geom_point(size = 3, shape = 16) +
  facet_wrap(~ position_label, ncol = 1, scales = "free_y") +
  scale_color_manual(values = light_colors) +
  labs(
    title = "SE per Triplet",
    y = NULL,
    x = "Triplet"
  ) +
  theme_classic(base_size = 11) +
  theme(
    aspect.ratio = 1.5, 
    plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
    axis.text.x = element_text(size = 8, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 9),
    strip.text = element_text(size = 10, face = "bold"),
    strip.background = element_rect(fill = "grey90"),
    legend.position = "none",
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3)
  )

# Combine both plots
p_main_triplet + p_error_triplet + plot_layout(widths = c(3, 1))
```

```{r dpg-triplet-position-calculate-intervention}
# Calculate position data for intervention window (excluding P07) - 1.5h window
participants_int <- c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10")

mean_triplet_pos_dpg_int_noP07 <- data.frame(
  triplet = character(),
  position = character(),
  mean_dpg_int = numeric(),
  sd_dpg_int = numeric(),
  se_dpg_int = numeric(),
  n_participants = integer(),
  stringsAsFactors = FALSE
)

for(t in triplets) {
  for(pos in c("pos1", "pos2", "pos3")) {
    values <- c()
    for(p in participants_int) {
      # Select mapping
      if(p %in% c("P01", "P02", "P03")) {
        triplet_map <- triplet_positions_g1
      } else if(p %in% c("P04", "P05", "P06", "P09")) {
        triplet_map <- triplet_positions_g2
      } else {
        triplet_map <- triplet_positions_g4
      }
      
      # Get intervention DPG data
      dpg_data_int <- ibuttons %>%
        filter(id == p) %>%
        filter(!(p == "P10" & session == "C04_05")) %>%
        group_by(session) %>%
        slice(1) %>%
        ungroup() %>%
        mutate(exp_session = as.numeric(gsub("C0[1-4]_", "", session))) %>%
        select(exp_session, DPG_intervention)
      
      # Calculate positions
      pos_dpg_int <- calc_triplet_positions_dpg_int(p, dpg_data_int, triplet_map)
      
      # Extract value
      if(t %in% pos_dpg_int$triplet) {
        if(pos == "pos1") {
          values <- c(values, pos_dpg_int$pos1_dpg_int[pos_dpg_int$triplet == t])
        } else if(pos == "pos2") {
          values <- c(values, pos_dpg_int$pos2_dpg_int[pos_dpg_int$triplet == t])
        } else {
          values <- c(values, pos_dpg_int$pos3_dpg_int[pos_dpg_int$triplet == t])
        }
      }
    }
    
    if(length(values) > 0) {
      n <- length(values)
      mean_triplet_pos_dpg_int_noP07 <- rbind(mean_triplet_pos_dpg_int_noP07, data.frame(
        triplet = t,
        position = pos,
        mean_dpg_int = mean(values, na.rm = TRUE),
        sd_dpg_int = sd(values, na.rm = TRUE),
        se_dpg_int = sd(values, na.rm = TRUE) / sqrt(n),
        n_participants = n
      ))
    }
  }
}

# Prepare for plotting
df_plot_positions_triplet_dpg_int_noP07 <- mean_triplet_pos_dpg_int_noP07 %>%
  mutate(
    position_label = case_when(
      position == "pos1" ~ "Position 1",
      position == "pos2" ~ "Position 2",
      position == "pos3" ~ "Position 3"
    ),
    position_num = case_when(
      position == "pos1" ~ 1,
      position == "pos2" ~ 2,
      position == "pos3" ~ 3
    ),
    light_type = case_when(
      position == "pos1" ~ substr(triplet, 1, 1),
      position == "pos2" ~ substr(triplet, 2, 2),
      position == "pos3" ~ substr(triplet, 3, 3)
    ),
    light_label = ifelse(light_type == "D", "dim", "bright")
  ) %>%
  mutate(triplet = factor(triplet, levels = c("DDD", "BDD", "DBD", "BBD", 
                                              "DDB", "BDB", "DBB", "BBB")))
```

```{r dpg-triplet-position-plot-intervention-with-error-panel}
# Main plot without error bars
p_main_triplet_int <- ggplot(df_plot_positions_triplet_dpg_int_noP07, 
                             aes(x = position_num, y = mean_dpg_int, group = triplet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_label), size = 5, shape = 16) +
  scale_color_manual(values = light_colors,
                     labels = c("dim" = "Dim (10 lx)", "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2, 3), 
                     labels = c("Position 1", "Position 2", "Position 3"),
                     limits = c(0.8, 3.2)) +
  labs(
    title = "Distal-proximal gradient across light exposure triplets",
    subtitle = "Mean DPG across all participants (1.5-h intervention window; P07 excluded)",
    x = "Triplet position",
    y = "Mean Distal-Proximal Gradient (°C)",
    color = "Light condition"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "right"
  )

# Error bar panel (right)
p_error_triplet_int <- ggplot(df_plot_positions_triplet_dpg_int_noP07, 
                              aes(x = triplet, y = mean_dpg_int, color = light_label)) +
  geom_errorbar(aes(ymin = mean_dpg_int - se_dpg_int, ymax = mean_dpg_int + se_dpg_int),
                width = 0.3, linewidth = 0.8) +
  geom_point(size = 3, shape = 16) +
  facet_wrap(~ position_label, ncol = 1, scales = "free_y") +
  scale_color_manual(values = light_colors) +
  labs(
    title = "SE per Triplet",
    y = NULL,
    x = "Triplet"
  ) +
  theme_classic(base_size = 11) +
  theme(
    aspect.ratio = 1.5,  
    plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
    axis.text.x = element_text(size = 8, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 9),
    strip.text = element_text(size = 10, face = "bold"),
    strip.background = element_rect(fill = "grey90"),
    legend.position = "none",
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3)
  )

# Combine both plots
p_main_triplet_int + p_error_triplet_int + plot_layout(widths = c(3, 1))
```

# DPG Doublet Descriptive Statistics

```{r dpg-doublet-stats}
# Calculate statistics for doublet 5-h exposure
doublet_stats_dpg_5h <- df_plot_dpg_noP07 %>%
  group_by(doublet) %>%
  summarise(
    n_5h = n(),
    median_5h = median(mean_dpg, na.rm = TRUE),
    IQR_5h = IQR(mean_dpg, na.rm = TRUE),
    .groups = "drop"
  )

# Calculate statistics for doublet 1.5-h exposure
doublet_stats_dpg_1.5h <- df_plot_dpg_int_noP07 %>%
  group_by(doublet) %>%
  summarise(
    n_1.5h = n(),
    median_1.5h = median(mean_dpg_int, na.rm = TRUE),
    IQR_1.5h = IQR(mean_dpg_int, na.rm = TRUE),
    .groups = "drop"
  )

# Combine both exposure windows
doublet_stats_dpg_combined <- doublet_stats_dpg_5h %>%
  full_join(doublet_stats_dpg_1.5h, by = "doublet") %>%
  arrange(doublet)

# Display as formatted table
knitr::kable(
  doublet_stats_dpg_combined,
  digits = 2,
  col.names = c("Doublet", "N", "Median", "IQR", "N", "Median", "IQR"),
  caption = "Descriptive statistics for doublet second-day DPG across exposure windows (P07 excluded)",
  align = c("l", rep("r", 6))
) %>%
  kableExtra::add_header_above(c(" " = 1, "5-h Exposure" = 3, "1.5-h Exposure" = 3))
```

```{r dpg-triplet-stats}
# Calculate statistics for triplet 5-h exposure
triplet_stats_dpg_5h <- df_plot_dpg_triplet_noP07 %>%
  group_by(triplet) %>%
  summarise(
    n_5h = n(),
    median_5h = median(mean_dpg, na.rm = TRUE),
    IQR_5h = IQR(mean_dpg, na.rm = TRUE),
    .groups = "drop"
  )

# Calculate statistics for triplet 1.5-h exposure
triplet_stats_dpg_1.5h <- df_plot_dpg_triplet_int_noP07 %>%
  group_by(triplet) %>%
  summarise(
    n_1.5h = n(),
    median_1.5h = median(mean_dpg_int, na.rm = TRUE),
    IQR_1.5h = IQR(mean_dpg_int, na.rm = TRUE),
    .groups = "drop"
  )

# Combine both exposure windows
triplet_stats_dpg_combined <- triplet_stats_dpg_5h %>%
  full_join(triplet_stats_dpg_1.5h, by = "triplet") %>%
  arrange(triplet)

# Display as formatted table
knitr::kable(
  triplet_stats_dpg_combined,
  digits = 2,
  col.names = c("Triplet", "N", "Median", "IQR", "N", "Median", "IQR"),
  caption = "Descriptive statistics for triplet third-day DPG across exposure windows (P07 excluded)",
  align = c("l", rep("r", 6))
) %>%
  kableExtra::add_header_above(c(" " = 1, "5-h Exposure" = 3, "1.5-h Exposure" = 3))
```

# DPG Triplet Time-Course Analysis

## Data Preparation

```{r dpg-triplet-timecourse-setup}
# Get session 3 data for each triplet 
participants_noP07 <- c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10")
triplets_all <- c("DDD", "DDB", "DBD", "DBB", "BDD", "BDB", "BBD", "BBB")

# Function to get session 3 for each triplet type
get_triplet_session3 <- function(p, triplet_type) {
  if(p %in% c("P01", "P02", "P03")) {
    df_triplets <- df_triplets_g1
  } else if(p %in% c("P04", "P05", "P06", "P09")) {
    df_triplets <- df_triplets_g2
  } else {
    df_triplets <- data.frame(
      session_start = 1:(length(light_g4_orig)-2),
      session_mid = 2:(length(light_g4_orig)-1),
      session_end = 3:length(light_g4_orig),
      triplet = sapply(1:(length(light_g4_orig)-2), function(i) 
        paste0(light_g4_orig[i:(i+2)], collapse = ""))
    ) %>%
      filter(!str_detect(triplet, "X")) %>%
      filter(session_start != 5 & session_mid != 5 & session_end != 5)
  }
  
  sessions_3 <- df_triplets %>%
    filter(triplet == triplet_type) %>%
    pull(session_end)
  
  return(sessions_3)
}

# Function to get all participant-session combinations for a triplet
get_triplet_sessions_overview <- function(triplet_type) {
  overview <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      for(s in sessions_3) {
        overview <- rbind(overview, data.frame(
          participant_id = p,
          session = s
        ))
      }
    }
  }
  
  return(overview %>% arrange(session, participant_id))
}

# Function to get group ID from participant ID
get_group_id <- function(p) {
  if(p %in% c("P01", "P02", "P03")) return(1)
  if(p %in% c("P04", "P05")) return(2)
  if(p %in% c("P06", "P07", "P09")) return(3)
  if(p == "P10") return(4)
  return(NA)
}

# Function to create DPG plot for one triplet type
plot_triplet_timecourse_dpg <- function(triplet_type) {
  df_triplet_dpg <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      group_id <- get_group_id(p)
      session_ids <- paste0("C0", group_id, "_", sprintf("%02d", sessions_3))
      
      p_data <- ibuttons %>%
        filter(id == p, session %in% session_ids) %>%
        select(id, session, time_block, DPG_30min) %>%
        filter(!is.na(time_block)) %>%
        group_by(session, time_block) %>%
        slice(1) %>%
        ungroup()
      
      df_triplet_dpg <- rbind(df_triplet_dpg, p_data)
    }
  }
  
  if(nrow(df_triplet_dpg) == 0) {
    warning(paste("No data found for triplet", triplet_type))
    return(NULL)
  }
  
  df_triplet_dpg <- df_triplet_dpg %>%
    group_by(id, session) %>%
    arrange(time_block) %>%
    mutate(block_num = row_number()) %>%
    ungroup()
  
  triplet_summary <- df_triplet_dpg %>%
    group_by(block_num) %>%
    summarise(
      mean_dpg = mean(DPG_30min, na.rm = TRUE),
      sd_dpg = sd(DPG_30min, na.rm = TRUE),
      n = n(),
      .groups = "drop"
    )
  
  y_min <- floor(min(triplet_summary$mean_dpg - triplet_summary$sd_dpg, 
                     na.rm = TRUE) * 10) / 10 - 0.1
  y_max <- ceiling(max(triplet_summary$mean_dpg + triplet_summary$sd_dpg, 
                       na.rm = TRUE) * 10) / 10 + 0.1
  
  third_letter <- substr(triplet_type, 3, 3)
  intervention_color <- ifelse(third_letter == "D", "#F5F5F5", "#FFF9E6")
  intervention_alpha <- ifelse(third_letter == "D", 0.4, 0.8)
  intervention_label <- ifelse(third_letter == "D", "dim\n(10 lx)", "bright\n(1000 lx)")
  
  triplet_first_two <- substr(triplet_type, 1, 2)
  triplet_third <- substr(triplet_type, 3, 3)
  
  p <- ggplot(triplet_summary, aes(x = block_num, y = mean_dpg)) +
    annotate("rect", xmin = 0.5, xmax = 6.5, ymin = -Inf, ymax = Inf, 
             fill = "#F5F5F5", alpha = 0.4) +
    annotate("rect", xmin = 6.5, xmax = 9.5, ymin = -Inf, ymax = Inf, 
             fill = intervention_color, alpha = intervention_alpha) +
    annotate("rect", xmin = 9.5, xmax = 10.5, ymin = -Inf, ymax = Inf, 
             fill = "#F5F5F5", alpha = 0.4) +
    annotate("text", x = 3.5, y = y_max - 0.05, label = "dim\n(10 lx)", 
             hjust = 0.5, vjust = 1, size = 3.5, color = "grey50", fontface = "italic") +
    annotate("text", x = 8, y = y_max - 0.05, label = intervention_label, 
             hjust = 0.5, vjust = 1, size = 3.5, 
             color = ifelse(third_letter == "D", "grey50", "#B8860B"), 
             fontface = "italic") +
    annotate("text", x = 10, y = y_max - 0.05, label = "dim\n(10 lx)", 
             hjust = 0.5, vjust = 1, size = 3.5, color = "grey50", fontface = "italic") +
    geom_ribbon(aes(ymin = mean_dpg - sd_dpg,
                    ymax = mean_dpg + sd_dpg),
                alpha = 0.15, fill = "grey40") +
    geom_line(linewidth = 0.8, color = "black") +
    geom_point(size = 3.5, shape = 16, color = "black") +
    scale_x_continuous(
      breaks = 1:10,
      labels = c(
        "B1\n(5.25h)", "B2", "B3\n(4.25h)", "B4", "B5\n(3.25h)", 
        "B6", "B7\n(2.25h)", "B8", "B9\n(1.25h)", "B10"
      ),
      limits = c(0.5, 10.5),
      expand = c(0, 0)
    ) +
    scale_y_continuous(limits = c(y_min, y_max), expand = c(0, 0)) +
    labs(
      title = bquote("Distal-proximal gradient time course - Triplet " * 
                     .(triplet_first_two) * bold(.(triplet_third))),
      subtitle = paste0("Grand average third-day (n=", 
                       length(unique(df_triplet_dpg$id)), 
                       " participants; P07 excluded)"),
      x = "30-minute block (hours before HBT)",
      y = "Mean distal-proximal gradient (°C)"
    ) +
    theme_classic(base_size = 14) +
    theme(
      aspect.ratio = 1,
      plot.title = element_text(size = 16, hjust = 0),
      plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
      axis.text = element_text(color = "black", size = 13),
      axis.text.x = element_text(size = 9, margin = margin(t = 5)),
      axis.title = element_text(size = 14),
      axis.title.x = element_text(margin = margin(t = 12)),
      axis.title.y = element_text(margin = margin(r = 8)),
      panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
      plot.margin = margin(15, 10, 10, 10)
    )
  
  return(p)
}
```

## DDD Triplet

```{r dpg-triplet-DDD-sessions}
knitr::kable(get_triplet_sessions_overview("DDD"),
             caption = "Sessions used for Triplet DDD (DPG)")
```

```{r dpg-triplet-DDD-plot}
plot_triplet_timecourse_dpg("DDD")
```

## DDB Triplet

```{r dpg-triplet-DDB-sessions}
knitr::kable(get_triplet_sessions_overview("DDB"),
             caption = "Sessions used for Triplet DDB (DPG)")
```

```{r dpg-triplet-DDB-plot}
plot_triplet_timecourse_dpg("DDB")
```

## DBD Triplet

```{r dpg-triplet-DBD-sessions}
knitr::kable(get_triplet_sessions_overview("DBD"),
             caption = "Sessions used for Triplet DBD (DPG)")
```

```{r dpg-triplet-DBD-plot}
plot_triplet_timecourse_dpg("DBD")
```

## DBB Triplet

```{r dpg-triplet-DBB-sessions}
knitr::kable(get_triplet_sessions_overview("DBB"),
             caption = "Sessions used for Triplet DBB (DPG)")
```

```{r dpg-triplet-DBB-plot}
plot_triplet_timecourse_dpg("DBB")
```

## BDD Triplet

```{r dpg-triplet-BDD-sessions}
knitr::kable(get_triplet_sessions_overview("BDD"),
             caption = "Sessions used for Triplet BDD (DPG)")
```

```{r dpg-triplet-BDD-plot}
plot_triplet_timecourse_dpg("BDD")
```

## BDB Triplet

```{r dpg-triplet-BDB-sessions}
knitr::kable(get_triplet_sessions_overview("BDB"),
             caption = "Sessions used for Triplet BDB (DPG)")
```

```{r dpg-triplet-BDB-plot}
plot_triplet_timecourse_dpg("BDB")
```

## BBD Triplet

```{r dpg-triplet-BBD-sessions}
knitr::kable(get_triplet_sessions_overview("BBD"),
             caption = "Sessions used for Triplet BBD (DPG)")
```

```{r dpg-triplet-BBD-plot}
plot_triplet_timecourse_dpg("BBD")
```

## BBB Triplet

```{r dpg-triplet-BBB-sessions}
knitr::kable(get_triplet_sessions_overview("BBB"),
             caption = "Sessions used for Triplet BBB (DPG)")
```

```{r dpg-triplet-BBB-plot}
plot_triplet_timecourse_dpg("BBB")
```

# DPG Triplet Time-Course Summary Table

```{r dpg-triplet-timecourse-summary-table}
# Combine all triplet data for the unified table
triplets_all_list <- c("DDD", "DDB", "DBD", "DBB", "BDD", "BDB", "BBD", "BBB")

df_all_triplets_dpg_unified <- data.frame()

for(triplet_type in triplets_all_list) {
  df_triplet_dpg <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      group_id <- get_group_id(p)
      session_ids <- paste0("C0", group_id, "_", sprintf("%02d", sessions_3))
      
      p_data <- ibuttons %>%
        filter(id == p, session %in% session_ids) %>%
        select(id, session, time_block, DPG_30min) %>%
        filter(!is.na(time_block)) %>%
        group_by(session, time_block) %>%
        slice(1) %>%
        ungroup() %>%
        mutate(triplet = triplet_type)
      
      df_triplet_dpg <- rbind(df_triplet_dpg, p_data)
    }
  }
  
  # Assign block numbers
  df_triplet_dpg <- df_triplet_dpg %>%
    group_by(id, session) %>%
    arrange(time_block) %>%
    mutate(block_num = row_number()) %>%
    ungroup()
  
  df_all_triplets_dpg_unified <- rbind(df_all_triplets_dpg_unified, df_triplet_dpg)
}

# Calculate summary statistics per triplet and block
triplet_summary_dpg_unified <- df_all_triplets_dpg_unified %>%
  group_by(triplet, block_num) %>%
  summarise(
    mean_dpg = mean(DPG_30min, na.rm = TRUE),
    sd_dpg = sd(DPG_30min, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(triplet = factor(triplet, levels = triplets_all_list))

# Create table with mean ± SD for all triplets
triplet_timecourse_dpg_table <- triplet_summary_dpg_unified %>%
  mutate(
    mean_sd = sprintf("%.2f ± %.2f", mean_dpg, sd_dpg)
  ) %>%
  select(triplet, block_num, mean_sd) %>%
  pivot_wider(
    names_from = block_num,
    values_from = mean_sd,
    names_prefix = "B"
  )

knitr::kable(
  triplet_timecourse_dpg_table,
  caption = "Triplet time-course: Mean ± SD distal-proximal gradient (°C) per 30-minute block",
  align = c("l", rep("r", 10))
) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                           full_width = FALSE) %>%
  kableExtra::row_spec(0, bold = TRUE)
```

# DPG Grand Average Time Course

```{r dpg-grand-average-timecourse}
# Map light conditions to each session for DPG (making sure the light patterns are available)
light_patterns_dpg <- list(
  P01 = c("D", "D", "D", "B", "B", "B", "D", "D", "B", "D", "B"),
  P02 = c("D", "D", "D", "B", "B", "B", "D", "D", "B", "D", "B"),
  P03 = c("D", "D", "D", "B", "B", "B", "D", "D", "B", "D", "B"),
  P04 = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"),
  P05 = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"),
  P06 = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"),
  P07 = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"),
  P09 = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B")
)

# Add P10 pattern (12 sessions without session 5)
light_patterns_dpg[["P10"]] <- c("D", "D", "D", "B", "D", "B", "B", "B", "D", "B", "D", "D")

df_light_dpg <- bind_rows(
  # All participants except P10
  bind_rows(lapply(names(light_patterns_dpg), function(p) {
    pattern <- light_patterns_dpg[[p]]
    data.frame(
      id = p,
      exp_session = seq_along(pattern),
      light = pattern,
      stringsAsFactors = FALSE
    )
  })),
  # P10 
  data.frame(
    id = "P10",
    exp_session = 1:13,
    light = ifelse(1:13 %in% c(1,2,3,6,10,12,13), "D", "B"),
    stringsAsFactors = FALSE
  )
)

# Add group information and create session IDs
df_light_dpg <- df_light_dpg %>%
  mutate(
    group = case_when(
      id %in% c("P01", "P02", "P03") ~ 1,
      id %in% c("P04", "P05") ~ 2,
      id %in% c("P06", "P07", "P09") ~ 3,
      id == "P10" ~ 4
    ),
    session = paste0("C0", group, "_", sprintf("%02d", exp_session))
  )

# Join with DPG data - get one row per session and time_block
df_dpg_light <- ibuttons %>%
  filter(id %in% c("P01", "P02", "P03", "P04", "P05", "P06", "P07", "P09", "P10")) %>%
  filter(!is.na(time_block)) %>%
  group_by(id, session, time_block) %>%
  slice(1) %>%
  ungroup() %>%
  inner_join(df_light_dpg, by = c("id", "session"))

# Assign block numbers (1-10)
df_dpg_light <- df_dpg_light %>%
  group_by(id, session) %>%
  arrange(time_block) %>%
  mutate(block_num = row_number()) %>%
  ungroup()

# Calculate per-participant summary
df_summary_dpg_per_participant <- df_dpg_light %>%
  group_by(id, block_num, light) %>%
  summarise(
    mean_dpg = mean(DPG_30min, na.rm = TRUE),
    sd_dpg = sd(DPG_30min, na.rm = TRUE),
    n = n(),
    se_dpg = sd_dpg / sqrt(n),
    .groups = "drop"
  ) %>%
  mutate(light = factor(light, levels = c("D", "B"), labels = c("Dim", "Bright")))

# Grand average excluding P07
df_summary_dpg_noP07 <- df_summary_dpg_per_participant %>%
  filter(id != "P07")

grand_summary_dpg_manual <- data.frame()

for(b in unique(df_summary_dpg_noP07$block_num)) {
  for(l in levels(df_summary_dpg_noP07$light)) {
    vals <- df_summary_dpg_noP07 %>%
      filter(block_num == b, light == l) %>%
      pull(mean_dpg)
    
    n_val <- length(vals)
    mean_val <- mean(vals, na.rm = TRUE)
    sd_val <- sd(vals, na.rm = TRUE)
    se_val <- sd_val / sqrt(n_val)
    
    grand_summary_dpg_manual <- rbind(grand_summary_dpg_manual, data.frame(
      block_num = b,
      light = l,
      mean_dpg = mean_val,
      sd_dpg = sd_val,
      n = n_val,
      se_dpg = se_val
    ))
  }
}

grand_summary_dpg_manual$light <- factor(grand_summary_dpg_manual$light, 
                                         levels = c("Bright", "Dim"))
```

```{r dpg-grand-average-plot}
# Ensure block_num is numeric
grand_summary_dpg_manual <- grand_summary_dpg_manual %>%
  mutate(block_num = as.numeric(block_num))

# Define labels
hbt_values <- c("5.25", "4.75", "4.25", "3.75", "3.25", "2.75", "2.25", "1.75", "1.25", "0.75")

# Create x-axis labels
x_labels <- sapply(1:10, function(i) {
  if (i %% 2 == 1) {  
    paste0("B", i, "\n(", hbt_values[i], "h)")
  } else {  
    paste0("B", i)
  }
})

# Y-axis limits
y_min <- floor(min(grand_summary_dpg_manual$mean_dpg - grand_summary_dpg_manual$se_dpg, 
                   na.rm = TRUE) * 10) / 10 - 0.1
y_max <- ceiling(max(grand_summary_dpg_manual$mean_dpg + grand_summary_dpg_manual$se_dpg, 
                     na.rm = TRUE) * 10) / 10 + 0.1

# Create plot
ggplot(grand_summary_dpg_manual, 
       aes(x = block_num, y = mean_dpg, 
           color = light, fill = light, group = light)) +
  # Background shading
  annotate("rect", xmin = 0.5, xmax = 6.5, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  annotate("rect", xmin = 6.5, xmax = 9.5, ymin = -Inf, ymax = Inf, 
           fill = "#FFF9E6", alpha = 0.8) +
  annotate("rect", xmin = 9.5, xmax = 10.5, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  annotate("text", x = 3.75, y = y_max - 0.05, label = "dim\n(10 lx)",  
           hjust = 0.5, vjust = 1, size = 3.5, color = "grey50", fontface = "italic") +
  annotate("text", x = 8, y = y_max - 0.05, label = "intervention\n(10 or 1000 lx)",  
           hjust = 0.5, vjust = 1, size = 3.5, color = "#B8860B", fontface = "italic") +
  annotate("text", x = 10, y = y_max - 0.05, label = "dim\n(10 lx)", 
           hjust = 0.5, vjust = 1, size = 3.5, color = "grey50", fontface = "italic") + 
  # Ribbon for SE
  geom_ribbon(aes(ymin = mean_dpg - se_dpg,
                  ymax = mean_dpg + se_dpg),
              alpha = 0.1, color = NA, show.legend = FALSE) +
  # Line and points
  geom_line(linewidth = 0.8) +
  geom_point(size = 3.5, shape = 16) +
  # Colors
  scale_color_manual(
    values = c("Bright" = "#F5A623", "Dim" = "#4A90E2"),
    labels = c("Bright" = "Bright (1000 lx)", "Dim" = "Dim (10 lx)")
  ) +
  scale_fill_manual(
    values = c("Bright" = "#F5A623", "Dim" = "#4A90E2"),
    labels = c("Bright" = "Bright (1000 lx)", "Dim" = "Dim (10 lx)")
  ) +
  # Axes 
  scale_x_continuous(
    breaks = 1:10,
    labels = x_labels,
    limits = c(0.5, 10.5),
    expand = c(0, 0)
  ) +
  scale_y_continuous(limits = c(y_min, y_max), expand = c(0, 0)) +
  # Labels
  labs(
    title = "Distal-proximal gradient time course",
    subtitle = "Grand average across participants (P07 excluded, n=8)",
    x = "30-minute block (hours before HBT)",
    y = "Mean Distal-Proximal Gradient (°C)",
    color = "Light condition",
    fill = "Light condition"
  ) +
  # Theme
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.text.x = element_text(size = 9, margin = margin(t = 5)),
    axis.title = element_text(size = 14),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 8)),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "right",
    plot.margin = margin(15, 10, 10, 10)
  )
```

# DPG Panel View - Triplet and Grand Average Time Courses

```{r dpg-panel-view-setup}
#| label: fig-dpg-triplet-panel
#| fig-cap: "Distal-proximal gradient time course - Grand average and triplet patterns (P07 excluded)"
#| fig-width: 14
#| fig-height: 8

# 1. Prepare data for all triplets
triplets_all_panel <- c("DDD", "BDD", "DBD", "BBD", "DDB", "BDB", "DBB", "BBB")

df_all_triplets_dpg_for_panel <- data.frame()

for(triplet_type in triplets_all_panel) {
  df_triplet_dpg <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      group_id <- get_group_id(p)
      session_ids <- paste0("C0", group_id, "_", sprintf("%02d", sessions_3))
      
      p_data <- ibuttons %>%
        filter(id == p, session %in% session_ids) %>%
        select(id, session, time_block, DPG_30min) %>%
        filter(!is.na(time_block)) %>%
        group_by(session, time_block) %>%
        slice(1) %>%
        ungroup() %>%
        mutate(triplet = triplet_type)
      
      df_triplet_dpg <- rbind(df_triplet_dpg, p_data)
    }
  }
  
  df_triplet_dpg <- df_triplet_dpg %>%
    group_by(id, session) %>%
    arrange(time_block) %>%
    mutate(block_num = row_number()) %>%
    ungroup()
  
  df_all_triplets_dpg_for_panel <- rbind(df_all_triplets_dpg_for_panel, df_triplet_dpg)
}

# 2. Calculate summary per triplet
triplet_summary_dpg_for_panel <- df_all_triplets_dpg_for_panel %>%
  group_by(triplet, block_num) %>%
  summarise(
    mean_dpg = mean(DPG_30min, na.rm = TRUE),
    sd_dpg = sd(DPG_30min, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    triplet = factor(triplet, levels = triplets_all_panel),
    triplet_type = ifelse(substr(triplet, 3, 3) == "D", "Dim", "Bright")
  )

# 3. Unified Y-axis limits
y_min_dpg_unified <- floor(min(triplet_summary_dpg_for_panel$mean_dpg - triplet_summary_dpg_for_panel$sd_dpg, 
                               na.rm = TRUE) * 10) / 10 - 0.1
y_max_dpg_unified <- ceiling(max(triplet_summary_dpg_for_panel$mean_dpg + triplet_summary_dpg_for_panel$sd_dpg, 
                                 na.rm = TRUE) * 10) / 10 + 0.1

# 4. Function to create triplet plot
create_triplet_plot_dpg <- function(triplet_name, show_x = FALSE, show_y = TRUE) {
  data_triplet <- triplet_summary_dpg_for_panel %>%
    filter(triplet == triplet_name)
  
  triplet_third <- substr(triplet_name, 3, 3)
  intervention_color <- ifelse(triplet_third == "D", "#F5F5F5", "#FFF9E6")
  intervention_alpha <- ifelse(triplet_third == "D", 0.4, 0.8)
  
  triplet_first_two <- substr(triplet_name, 1, 2)
  triplet_label <- paste0(triplet_first_two, "**", triplet_third, "**")
  
  x_labels_with_hbt <- c(
    "B1\n(5.25h)", "B2", "B3\n(4.25h)", "B4", "B5\n(3.25h)", 
    "B6", "B7\n(2.25h)", "B8", "B9\n(1.25h)", "B10"
  )
  
  ggplot(data_triplet, aes(x = block_num, y = mean_dpg)) +
    annotate("rect", xmin = 0.5, xmax = 6.5, ymin = -Inf, ymax = Inf, 
             fill = "#F5F5F5", alpha = 0.4) +
    annotate("rect", xmin = 6.5, xmax = 9.5, ymin = -Inf, ymax = Inf, 
             fill = intervention_color, alpha = intervention_alpha) +
    annotate("rect", xmin = 9.5, xmax = 10.5, ymin = -Inf, ymax = Inf, 
             fill = "#F5F5F5", alpha = 0.4) +
    geom_ribbon(aes(ymin = mean_dpg - sd_dpg, 
                    ymax = mean_dpg + sd_dpg),
                alpha = 0.15, fill = "grey40") +
    geom_line(linewidth = 0.8, color = "black") +
    geom_point(size = 2.5, shape = 16, color = "black") +
    scale_x_continuous(
      breaks = 1:10,
      labels = if(show_x) x_labels_with_hbt else 1:10,
      limits = c(0.5, 10.5),
      expand = c(0, 0)
    ) +
    scale_y_continuous(limits = c(y_min_dpg_unified, y_max_dpg_unified), expand = c(0, 0)) +
    labs(
      title = triplet_label,
      x = NULL,
      y = if(show_y) "Mean DPG ± SD (°C)" else NULL
    ) +
    theme_classic(base_size = 11) +
    theme(
      plot.title = element_markdown(size = 11, hjust = 0.5, margin = margin(b = 5)),
      axis.text = element_text(color = "black", size = 9),
      axis.text.x = element_text(size = if(show_x) 7 else 8, margin = margin(t = 3)),
      axis.title = element_text(size = 10),
      axis.title.y = element_text(margin = margin(r = 5)),
      panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
      plot.margin = margin(5, 5, if(show_x) 8 else 5, 5)
    )
}

# 5. Create D-ending triplet plots (no x-axis labels)
p_DDD_dpg <- create_triplet_plot_dpg("DDD", show_x = FALSE, show_y = TRUE)
p_BDD_dpg <- create_triplet_plot_dpg("BDD", show_x = FALSE, show_y = FALSE)
p_DBD_dpg <- create_triplet_plot_dpg("DBD", show_x = FALSE, show_y = FALSE)
p_BBD_dpg <- create_triplet_plot_dpg("BBD", show_x = FALSE, show_y = FALSE)

# 6. Create B-ending triplet plots (with x-axis labels)
p_DDB_dpg <- create_triplet_plot_dpg("DDB", show_x = TRUE, show_y = TRUE)
p_BDB_dpg <- create_triplet_plot_dpg("BDB", show_x = TRUE, show_y = FALSE)
p_DBB_dpg <- create_triplet_plot_dpg("DBB", show_x = TRUE, show_y = FALSE)
p_BBB_dpg <- create_triplet_plot_dpg("BBB", show_x = TRUE, show_y = FALSE)

# 7. Grand Average plot
p_grand_dpg <- ggplot(grand_summary_dpg_manual,
       aes(x = block_num, y = mean_dpg,
           color = light, fill = light, group = light)) +
  annotate("rect", xmin = 0.5, xmax = 6.5, ymin = -Inf, ymax = Inf,
           fill = "#F5F5F5", alpha = 0.4) +
  annotate("rect", xmin = 9.5, xmax = 10.5, ymin = -Inf, ymax = Inf,
           fill = "#F5F5F5", alpha = 0.4) +
  annotate("rect", xmin = 6.5, xmax = 9.5, ymin = -Inf, ymax = Inf,
           fill = "#FFF9E6", alpha = 0.8) +
  geom_ribbon(aes(ymin = mean_dpg - se_dpg,
                  ymax = mean_dpg + se_dpg),
              alpha = 0.1, color = NA, show.legend = FALSE) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.5, shape = 16) +
  scale_color_manual(
    values = c("Bright" = "#F5A623", "Dim" = "#4A90E2"),
    labels = c("Bright" = "Bright (1000lx)", "Dim" = "Dim (10lx)")
  ) +
  scale_fill_manual(
    values = c("Bright" = "#F5A623", "Dim" = "#4A90E2"),
    labels = c("Bright" = "Bright (1000lx)", "Dim" = "Dim (10lx)")
  ) +
  scale_x_continuous(
    breaks = 1:10,
    labels = 1:10,
    limits = c(0.5, 10.5),
    expand = c(0, 0)
  ) +
  scale_y_continuous(limits = c(y_min_dpg_unified, y_max_dpg_unified), expand = c(0, 0)) +
  labs(
    title = "Grand Average",
    x = NULL,
    y = NULL,
    color = "Light condition"
  ) +
  guides(
    color = guide_legend(override.aes = list(shape = 16, size = 2.5)),
    fill = "none"
  ) +
  theme_classic(base_size = 11) +
  theme(
    plot.title = element_text(size = 11, hjust = 0.5, margin = margin(b = 5)),
    axis.text = element_text(color = "black", size = 9),
    axis.text.x = element_text(size = 8, margin = margin(t = 3)),
    axis.title = element_text(size = 10),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "bottom",
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 7),
    plot.margin = margin(5, 5, 5, 5)
  )

# 8. Empty spacer
p_spacer_dpg <- ggplot() + theme_void()

# 9. Combine in layout
(p_DDD_dpg | p_BDD_dpg | p_DBD_dpg | p_BBD_dpg | p_grand_dpg) /
(p_DDB_dpg | p_BDB_dpg | p_DBB_dpg | p_BBB_dpg | p_spacer_dpg) +
  plot_annotation(
    title = "Distal-proximal gradient time course - Grand average and triplet patterns",
    subtitle = "Grand average and third-day triplet patterns (P07 excluded)",
    caption = "30-minute block (hours before HBT)",
    theme = theme(
      plot.title = element_text(size = 15, hjust = 0, face = "plain"),
      plot.subtitle = element_text(size = 12, hjust = 0, margin = margin(b = 8)),
      plot.caption = element_text(size = 11, hjust = 0.5, margin = margin(t = 10))
    )
  )
```

# Data Preparation for DPG Statistical Analysis

## Define Study Parameters for DPG

```{r dpg-lmm-setup}
# Define participant list and light patterns
participant_list_dpg <- c("P01", "P02", "P03", "P04", "P05", "P06", "P07", "P09", "P10")

# Light patterns for DPG (making sure light patterns are available)
light_patterns_dpg <- list(
  P01 = c("D", "D", "D", "B", "B", "B", "D", "D", "B", "D", "B"),
  P02 = c("D", "D", "D", "B", "B", "B", "D", "D", "B", "D", "B"),
  P03 = c("D", "D", "D", "B", "B", "B", "D", "D", "B", "D", "B"),
  P04 = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"),
  P05 = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"),
  P06 = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"),
  P07 = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"),
  P09 = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B")
)

# Combine DPG data with light information
ibuttons_lmm <- map_dfr(participant_list_dpg, ~{
  df <- ibuttons %>% 
    filter(id == .x) %>%
    # Extract session number from session variable (e.g., "C01_01" -> 1, "C01_02" -> 2)
    mutate(exp_session = as.integer(gsub("C0[1-4]_", "", session))) %>%
    # Keep only DPG_intervention values (one per session)
    filter(!is.na(DPG_intervention)) %>%
    distinct(id, exp_session, DPG_intervention, .keep_all = TRUE) %>%
    select(id, exp_session, DPG_intervention) %>%
    rename(participant_id = id)
  
  # Special light pattern assignment for P10 (light automation malfunction session 5)
  if(.x == "P10") {
    # Manual assignment for each actual session number
    light_df <- tibble(
      exp_session = c(1, 2, 3, 4, 6, 7, 8, 9, 10, 11, 12, 13),
      Light_Lag0 = c("D", "D", "D", "B", "D", "B", "B", "B", "D", "B", "D", "D")
    )
  } else {
    light_df <- tibble(
      exp_session = seq_along(light_patterns_dpg[[.x]]),
      Light_Lag0 = light_patterns_dpg[[.x]]
    )
  }
  
  df <- df %>% 
    left_join(light_df, by = "exp_session")
  
  return(df)
})

# Build lagged light variables and filter valid triplets
ibuttons_lmm <- ibuttons_lmm %>%
  group_by(participant_id) %>%
  arrange(exp_session) %>%
  mutate(
    Light_Lag1 = lag(Light_Lag0, 1),
    Light_Lag2 = lag(Light_Lag0, 2)
  ) %>%
  ungroup() %>%
  filter(!is.na(Light_Lag0) & !is.na(Light_Lag1) & !is.na(Light_Lag2)) %>%
  filter(!(participant_id == "P10" & exp_session %in% c(6, 7))) %>%
  mutate(
    Light_Lag0_num = ifelse(Light_Lag0 == "D", 0, 1),
    Light_Lag1_num = ifelse(Light_Lag1 == "D", 0, 1),
    Light_Lag2_num = ifelse(Light_Lag2 == "D", 0, 1)
  )

# Add participant demographics
participant_demographics_dpg <- tibble(
  participant_id = c("P01", "P02", "P03", "P04", "P05", "P06", "P07", "P09", "P10"),
  sex = c("female", "male", "female", "female", "male", "female", "female", "female", "male"),
  age = c(24, 21, 24, 23, 31, 23, 24, 28, 32),
  bmi = c(25.82645, 18.83239, 28.01022, 24.67550, 20.28651, 22.23099, 23.23346, 24.50895, 27.14304)
)

ibuttons_lmm <- ibuttons_lmm %>%
  left_join(participant_demographics_dpg, by = "participant_id")

# Apply exclusion criteria
ibuttons_lmm <- ibuttons_lmm %>%
  filter(participant_id != "P07")  # Exclude P07

# Keep only necessary variables
ibuttons_lmm <- ibuttons_lmm %>%
  dplyr::select(
    participant_id, exp_session, DPG_intervention,
    Light_Lag0, Light_Lag1, Light_Lag2,
    Light_Lag0_num, Light_Lag1_num, Light_Lag2_num,
    sex, age, bmi
  )
```

## Dataset Overview

```{r dpg-lmm-structure}
# Display structure
str(ibuttons_lmm)
```

```{r dpg-lmm-head}
# Display first rows
head(ibuttons_lmm, 10)
```

```{r dpg-lmm-summary}
# Summary
cat("Total observations:", nrow(ibuttons_lmm), "\n")
cat("Participants:", length(unique(ibuttons_lmm$participant_id)), "\n")
cat("Sessions range:", min(ibuttons_lmm$exp_session), "-", max(ibuttons_lmm$exp_session), "\n")
```

## Dataset ibuttons_lmm

```{r dpg-lmm-datatable}
DT::datatable(
  ibuttons_lmm,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    order = list(list(0, 'asc'), list(1, 'asc'))
  ),
  caption = "Complete DPG dataset for statistical analysis (P07 excluded)",
  rownames = FALSE
) %>%
  DT::formatRound(columns = 'DPG_intervention', digits = 3) %>%
  DT::formatRound(columns = 'bmi', digits = 2) %>%
  DT::formatStyle(
    'Light_Lag0',
    backgroundColor = DT::styleEqual(c('D', 'B'), c('#e3f2fd', '#fff3e0'))
  )
```

# Descriptive Statistics for DPG

## Summary Statistics

```{r dpg-lmm-summary-stats}
summary(ibuttons_lmm$DPG_intervention)
```

```{r dpg-lmm-summary-table}
data.frame(
  N = nrow(ibuttons_lmm),
  N_participants = length(unique(ibuttons_lmm$participant_id)),
  Mean = round(mean(ibuttons_lmm$DPG_intervention), 2),
  SD = round(sd(ibuttons_lmm$DPG_intervention), 2),
  Min = round(min(ibuttons_lmm$DPG_intervention), 2),
  Max = round(max(ibuttons_lmm$DPG_intervention), 2)
)
```

## Distribution Visualization

```{r dpg-lmm-distribution}
ggplot(ibuttons_lmm, aes(x = DPG_intervention)) +
  geom_histogram(aes(y = after_stat(density)), bins = 15, 
                 fill = "#4A90E2", alpha = 0.7, color = "black") +
  geom_density(color = "red", linewidth = 1) +
  geom_vline(xintercept = mean(ibuttons_lmm$DPG_intervention), 
             linetype = "dashed", color = "red") +
  labs(title = "Distribution of Distal-Proximal Gradient during Intervention", 
       x = "Distal-Proximal Gradient (°C)", 
       y = "Density") +
  theme_classic()
```

Interpretation: Despite the light left-skewness of the raw data, a LMM with Gaussian identity link function will be applied. However, within the diagnostics checks the normality assumption for residuals will be evaluated and different sensitivity analyses applied if necessary.

## DPG Trajectories Across Experimental Sessions

Individual participant trajectories with LOESS smooth (full dataset, n=67 observations)

```{r dpg-lmm-trajectories}
# Prepare data - count observations per session
session_counts_dpg <- ibuttons_lmm %>%
  group_by(exp_session) %>%
  summarise(n = n(), .groups = "drop")

# Create plot
ggplot(ibuttons_lmm, 
       aes(x = exp_session, y = DPG_intervention, 
           color = participant_id, group = participant_id)) +
  
  # Individual trajectories
  geom_line(alpha = 0.6, linewidth = 0.8) +
  geom_point(size = 2.5, alpha = 0.7) +
  
  # LOESS smooth (overall trend)
  geom_smooth(aes(group = 1), method = "loess", span = 0.75,
              color = "black", linetype = "dashed", linewidth = 1.2,
              se = TRUE, fill = "grey50", alpha = 0.3) +
  
  # Add sample size annotations for reduced sessions
  geom_text(data = session_counts_dpg %>% filter(n < 8),
            aes(x = exp_session, y = max(ibuttons_lmm$DPG_intervention) * 0.998, 
                label = paste0("n=", n)),
            color = "black", size = 3, inherit.aes = FALSE) +
  
  # Vertical line for data cutoff (session 10)
  geom_vline(xintercept = 10.5, linetype = "dotted", color = "grey50", linewidth = 0.8) +
  
  # Scales
  scale_x_continuous(
    breaks = 3:13,
    limits = c(2.5, 13.5),
    expand = c(0.01, 0)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.05))
  ) +
  
  # Colors
  scale_color_manual(
    values = c("P01" = "#E41A1C", "P02" = "#377EB8", "P03" = "#4DAF4A",
               "P04" = "#984EA3", "P05" = "#FF7F00", "P06" = "#FFFF33",
               "P09" = "#A65628", "P10" = "#F781BF"),
    name = "Participant"
  ) +
  
  # Labels
  labs(
    title = "Distal-proximal gradient across experimental sessions",
    subtitle = "Individual participant trajectories with LOESS smooth (full dataset, n=67, P07 excluded)",
    x = "Experimental Session",
    y = "Distal-Proximal Gradient (°C)"
  ) +
  
  # Theme
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, face = "plain"),
    plot.subtitle = element_text(size = 12, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 12),
    axis.title = element_text(size = 13),
    legend.position = "right",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    plot.margin = margin(15, 15, 10, 10)
  )
```

## Light Condition Balance

```{r dpg-lmm-light-balance}
ibuttons_lmm %>%
  summarise(
    Lag0_Dim = sum(Light_Lag0_num == 0),
    Lag0_Bright = sum(Light_Lag0_num == 1),
    Lag1_Dim = sum(Light_Lag1_num == 0),
    Lag1_Bright = sum(Light_Lag1_num == 1),
    Lag2_Dim = sum(Light_Lag2_num == 0),
    Lag2_Bright = sum(Light_Lag2_num == 1)
  )
```

## Observations per Session

```{r dpg-lmm-sessions}
table(ibuttons_lmm$exp_session)
```

## Observations per Participant

```{r dpg-lmm-participants}
table(ibuttons_lmm$participant_id)
```

# Statistical Analysis for Distal-Proximal Gradient

## Hierarchical Model Building Strategy

### Model m0: Baseline (Random Intercept Only)

This baseline model quantifies between-participant variability without any predictors.

```{r dpg-m0}
m0_dpg <- lmer(
  DPG_intervention ~ 1 + (1 | participant_id),
  data = ibuttons_lmm,
  REML = FALSE
)

summary(m0_dpg)

# Wald Confidence Intervals
ci_m0 <- confint(m0_dpg, method = "Wald")
print(ci_m0)
```

### Model m1a: Add Light Lag 0 (Current Day)

This model adds the current day light exposure variable as a fixed effect with Light_Lag0_num.

```{r dpg-m1a}
m1a_dpg <- lmer(
  DPG_intervention ~ 
    Light_Lag0_num +
    (1 | participant_id),
  data = ibuttons_lmm,
  REML = FALSE
)

summary(m1a_dpg)

# Wald Confidence Intervals
ci_m1a <- confint(m1a_dpg, method = "Wald")
print(ci_m1a)
```

### Model m1b: Add Light Lag 1 Additionally (One Day Before)

This model adds the light exposure one day before as a fixed effect using Light_Lag1_num.

```{r dpg-m1b}
m1b_dpg <- lmer(
  DPG_intervention ~ 
    Light_Lag0_num + Light_Lag1_num +
    (1 | participant_id),
  data = ibuttons_lmm,
  REML = FALSE
)

summary(m1b_dpg)

# Wald Confidence Intervals
ci_m1b <- confint(m1b_dpg, method = "Wald")
print(ci_m1b)
```

### Model m1c: Add Light Lag 2 Additionally (Two Days Before)

This model adds the light exposure two days before as a fixed effect with Light_Lag2_num.

```{r dpg-m1c}
m1c_dpg <- lmer(
  DPG_intervention ~ 
    Light_Lag0_num + Light_Lag1_num + Light_Lag2_num +
    (1 | participant_id),
  data = ibuttons_lmm,
  REML = FALSE
)

summary(m1c_dpg)

# Wald Confidence Intervals
ci_m1c <- confint(m1c_dpg, method = "Wald")
print(ci_m1c)
```

### Likelihood Ratio Tests: Lag Effects

```{r likelihood-ratio-tests-lag-effects}
# LRT 1: Does adding Lag 0 improve fit over baseline?
lrt_m0_vs_m1a_dpg <- anova(m0_dpg, m1a_dpg)

# LRT 2: Does adding Lag 1 improve fit over Lag 0 only?
lrt_m1a_vs_m1b_dpg <- anova(m1a_dpg, m1b_dpg)

# LRT 3: Does adding Lag 2 improve fit over Lag 0 + Lag 1?
lrt_m1b_vs_m1c_dpg <- anova(m1b_dpg, m1c_dpg)

# Display results
lrt_m0_vs_m1a_dpg
lrt_m1a_vs_m1b_dpg
lrt_m1b_vs_m1c_dpg
```

Interpretation Likelihood Ratio Tests Lag Effects:

-   LRT 1: Adding Lag 0 did not improve fit over the baseline model only (m0 vs. m1a: χ²=1.0387, p=0.3081)

-   LRT 2: Adding Lag 1 did not significantly improve fit over model m1a_dpg (m1a vs. m1b: χ²=0.2919, p=0.589)

-   LRT 3: Adding Lag 2 did not significantly improve fit over model m1b_dpg (m1b vs. m1c: χ²=0.0754, p=0.7836)

### Model m2: Add Linear Session Effect

This model adds a linear session term to account for potential temporal trends across the study.

```{r model-m2}
m2_dpg <- lmer(
  DPG_intervention ~ 
    exp_session +
    (1 | participant_id),
  data = ibuttons_lmm,
  REML = FALSE
)

summary(m2_dpg)

# Wald Confidence Intervals
ci_m2 <- confint(m2_dpg, method = "Wald")
print(ci_m2)
```

### Model m3: Add Quadratic Session Effekct

This model tests for non-linear temporal trends (e.g., U-shaped patterns) across sessions.

```{r model-m3}
m3_dpg <- lmer(
  DPG_intervention ~ 
    exp_session + I(exp_session^2) +
    (1 | participant_id),
  data = ibuttons_lmm,
  REML = FALSE
)

summary(m3_dpg)

# Wald Confidence Intervals
ci_m3 <- confint(m3_dpg, method = "Wald")
print(ci_m3)
```

### Likelihood Ratio Tests: Session Effects

```{r likelihood-ratio-tests-session-effects}
# LRT 1: Does adding exp_session (linear) improve fit over baseline?
lrt_m0_vs_m2_dpg <- anova(m0_dpg, m2_dpg)

# LRT 2: Does adding quadratic term improve fit over linear only?
lrt_m2_vs_m3_dpg <- anova(m2_dpg, m3_dpg)

lrt_m0_vs_m2_dpg
lrt_m2_vs_m3_dpg
```

Interpretation Likelihood Ratio Tests Session Effects: - Adding a linear session term did not improve model fit over the baseline only model (m0 vs. m2: χ²=2.3853, p=0.1225)

-   Adding a quadratic session term did not improve model fit over model m2 (m2 vs. m3: χ²=0.8772, p=0.349)

## Model Selection Summary

Based on LRT test results, model m0_dpg with the random intercept only was chosen as the final model. All other models did not find significant light or session effects of DPG values in this group.

-   Model m1a_dpg: Lag 0 (β=0.2325, SE=0.2271, p=0.31); AIC=199.0, BIC=207.7

-   Model m1b_dpg: Lag 0 (β=0.2274, SE=0.2268, p=0.320); Lag 1 (β=-0.1222, SE=0.2260, p=0.591); AIC=200.7, BIC=211.6

-   Model m1c_dpg: Lag 0 (β=0.22274, SE=0.22726, p=0.331); Lag 1 (β=-0.11959, SE=0.22602, p=0.599); Lag 2 (β=-0.06224, SE=0.22651, p=0.784); AIC=202.6, BIC=215.7

-   Model m2_dpg: Linear Session Term (β=-0.06925, SE=0.04425, p=0.123); AIC=197.6, BIC=206.3

-   Model m3_dpg: Quadratic Session Term (β=0.01634, SE=0.01739, p=0.351012); AIC=198.8, BIC=209.6

## Diagnostics Summary

### Check 1: Normality Assumption for Residuals

```{r normality-residuals}
# Q-Q Plot Residuals
qqnorm(resid(m0_dpg), main = "Q-Q Plot: Residuals")
qqline(resid(m0_dpg), col = "red")

# Shapiro-Wilk Test 
shapiro.test(resid(m0_dpg))
```

Interpretation Check 1 Normality Assumption for Residuals: Normality assumption for residuals seems to be adequately met. Both visual inspection of the QQ-plot shows good alignment with the theoretical normal distribution with minor deviations in the lower and upper tails, likely attributable to the relatively small sample size. Also the Shapiro-Wilk Test did not find evidence against the normality of residuals (W=0.98497, p=0.6175).

### Check 2: Distribution of Residuals

```{r distribution-residuals}
# Histogram with Density Curve
ggplot(data.frame(residuals = resid(m0_dpg)), aes(x = residuals)) +
  geom_histogram(aes(y = after_stat(density)), bins = 20, 
                 fill = "#4A90E2", alpha = 0.7, color = "black") +
  geom_density(color = "red", linewidth = 1) +
  labs(title = "Distribution of Residuals (m0_dpg)",
       x = "Residuals", y = "Density") +
  theme_classic()
```

Interpretation Check 2 Distribution of Residuals: Consistent with the Q-Q plot findings, the histogram shows approximately normally distributed residuals with no extreme outliers.

### Check 3: Normality of Random Effects

```{r normality-random-effects}
# Q-Q Plot of Random Effects (Intercepts)
qqnorm(ranef(m0_dpg)$participant_id[,1], 
       main = "Q-Q Plot: Random Intercepts")
qqline(ranef(m0_dpg)$participant_id[,1], col = "red")
```

Interpretation Check 3 Normality of Random Effects: Acknowledging the limited interpretability due to the small number of random intercepts (n = 8), the assumption of normally distributed random effects appears to be reasonably met.

### Check 4: Variance Components and ICC

```{r variance-compoenents-icc}
# Intraclass Correlation Coefficient
icc(m0_dpg)

# Model Performance
model_performance(m0_dpg)
```

Interpretation Check 4 ICC: Approximately 54% of the total variance in DPG values is attributeable to between-participant differences, while the remaining 46% reflect the within-participant differences.

### Check 5: Heteroscedasticity

```{r heteroscedasticity}
# Check for heteroscedasticity across participants
df_resid <- data.frame(
  participant_id = ibuttons_lmm$participant_id,
  residuals = resid(m0_dpg)
)

ggplot(df_resid, aes(x = participant_id, y = residuals)) +
  geom_boxplot(fill = "#4A90E2", alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals by Participant (m0_dpg)",
       x = "Participant", y = "Residuals") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Interpretation Check 5 Heteroscedasticity: Residual variances appeared reasonably consistent across participants, supporting model assumptions.

### Check 6: Temporal Patterns

```{r temporal-patterns}
# Check for temporal patterns
df_resid_time <- data.frame(
  exp_session = ibuttons_lmm$exp_session,
  residuals = resid(m0_dpg)
)

ggplot(df_resid_time, aes(x = exp_session, y = residuals)) +
  geom_point(alpha = 0.6, size = 2.5) +
  geom_smooth(method = "loess", se = TRUE, color = "red") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Residuals across Sessions (m0_dpg)",
       x = "Experimental Session", y = "Residuals") +
  theme_classic()
```

Interpretation Check 6 Temporal Patterns: From Session 3 to around Session 8 a systematic downward trend is visible with a plateau at the end. Whereas session 11-13 are not well suited for interpretation due to the sparse data (n=3 for session 11, 12 and n=1 for session 13). The LOESS smooth confirms this pattern, suggesting a potential session effect not captured by the baseline model. This temporal pattern indicates that DPG may systematically decrease across experimental sessions, which was not detected as statistically significant in likelihood ratio tests (m0 vs m2: p=0.12), possibly due to limited statistical power (n=8 participants).

# Sensitivity Analysis

## Sensitivity Analysis 1: Exclude P06 (Problematic Outliers)

Rationale: P06 showed extensive outliers during data cleaning (30-50%), potentially biasing model estimates. This sensitivity analysis tests model robustness by excluding P06 (n=7 participants remaining).

```{r sensitivity-exclude-p06-data}
# Create subset without P06
ibuttons_lmm_sens <- ibuttons_lmm %>%
  filter(participant_id != "P06")

# Summary
cat("Original dataset: n =", nrow(ibuttons_lmm), "observations,", 
    length(unique(ibuttons_lmm$participant_id)), "participants\n")
cat("Sensitivity dataset: n =", nrow(ibuttons_lmm_sens), "observations,", 
    length(unique(ibuttons_lmm_sens$participant_id)), "participants\n")
```

```{r sensitivity-exclude-p06-summary}
# Descriptive statistics comparison
data.frame(
  Dataset = c("Original (n=8)", "Sensitivity (n=7, P06 excluded)"),
  N_obs = c(nrow(ibuttons_lmm), nrow(ibuttons_lmm_sens)),
  Mean = c(mean(ibuttons_lmm$DPG_intervention), 
           mean(ibuttons_lmm_sens$DPG_intervention)),
  SD = c(sd(ibuttons_lmm$DPG_intervention), 
         sd(ibuttons_lmm_sens$DPG_intervention)),
  Min = c(min(ibuttons_lmm$DPG_intervention), 
          min(ibuttons_lmm_sens$DPG_intervention)),
  Max = c(max(ibuttons_lmm$DPG_intervention), 
          max(ibuttons_lmm_sens$DPG_intervention))
) %>%
  kable(digits = 2, 
        caption = "Descriptive statistics: Original vs. Sensitivity Analysis (P06 excluded)")
```

```{r sensitivity-exclude-p06-model}
# Fit baseline model without P06
m0_dpg_sens <- lmer(
  DPG_intervention ~ 1 + (1 | participant_id),
  data = ibuttons_lmm_sens,
  REML = FALSE
)

summary(m0_dpg_sens)

# Wald Confidence Intervals
ci_m0_sens <- confint(m0_dpg_sens, method = "Wald")
print(ci_m0_sens)

# Add Lag 0 (without P06)
m1a_dpg_sens <- lmer(
  DPG_intervention ~ 
    Light_Lag0_num +
    (1 | participant_id),
  data = ibuttons_lmm_sens,
  REML = FALSE
)

summary(m1a_dpg_sens)

# Wald Confidence Intervals
ci_m1a_sens <- confint(m1a_dpg_sens, method = "Wald")
print(ci_m1a_sens)

# Add Lag 1 (without P06)
m1b_dpg_sens <- lmer(
  DPG_intervention ~ 
    Light_Lag0_num + Light_Lag1_num + 
    (1 | participant_id),
  data = ibuttons_lmm_sens,
  REML = FALSE
)

summary(m1b_dpg_sens)

# Wald Confidence Intervals
ci_m1b_sens <- confint(m1b_dpg_sens, method = "Wald")
print(ci_m1b_sens)

# Add Lag 2 (without P06)
m1c_dpg_sens <- lmer(
  DPG_intervention ~ 
    Light_Lag0_num + Light_Lag1_num + Light_Lag2_num +
    (1 | participant_id),
  data = ibuttons_lmm_sens,
  REML = FALSE
)

summary(m1c_dpg_sens)

# Wald Confidence Intervals
ci_m1c_sens <- confint(m1c_dpg_sens, method = "Wald")
print(ci_m1c_sens)

# LRT 1: Does adding Lag 0 improve fit over baseline (without P06)?
lrt_m0_vs_m1a_dpg_sens <- anova(m0_dpg_sens, m1a_dpg_sens)

# LRT 2: Does adding Lag 1 improve fit over Lag 0 only (without P06)?
lrt_m1a_vs_m1b_dpg_sens <- anova(m1a_dpg_sens, m1b_dpg_sens)

# LRT 3: Does adding Lag 2 improve fit over Lag 0 + Lag 1 (without P06)?
lrt_m1b_vs_m1c_dpg_sens <- anova(m1b_dpg_sens, m1c_dpg_sens)

# Display results
lrt_m0_vs_m1a_dpg_sens
lrt_m1a_vs_m1b_dpg_sens
lrt_m1b_vs_m1c_dpg_sens

# Add Linear Session Term to Model m0_dpg_sens (without P06)
m2_dpg_sens <- lmer(
  DPG_intervention ~ 
    exp_session +
    (1 | participant_id),
  data = ibuttons_lmm_sens,
  REML = FALSE
)

summary(m2_dpg_sens)

# Wald Confidence Intervals
ci_m2_sens <- confint(m2_dpg_sens, method = "Wald")
print(ci_m2_sens)

# Add Quadratic Session Term to Model m0_dpg_sens (without P06)
m3_dpg_sens <- lmer(
  DPG_intervention ~ 
    exp_session + I(exp_session^2) +
    (1 | participant_id),
  data = ibuttons_lmm_sens,
  REML = FALSE
)

summary(m3_dpg_sens)

# Wald Confidence Intervals
ci_m3_sens <- confint(m3_dpg_sens, method = "Wald")
print(ci_m3_sens)

# LRT 1: Does adding exp_session (linear) improve fit over baseline (without P06)?
lrt_m0_vs_m2_dpg_sens <- anova(m0_dpg_sens, m2_dpg_sens)

# LRT 2: Does adding quadratic term improve fit over linear only (without P06)?
lrt_m2_vs_m3_dpg_sens <- anova(m2_dpg_sens, m3_dpg_sens)

lrt_m0_vs_m2_dpg_sens
lrt_m2_vs_m3_dpg_sens
```

Interpretation Sensitivity 1 (Excluding P06 out of the analysis due to a lot of outliers identified): Conclusions drawn from the analysis and hierarchiechal model building using the entire dataset were not changed comapred to using the dataset, which excludes P06 from the analysis.

# Exploratory Analyses

## Exploratory Analysis 1: Random Slopes for Light_Lag0

Rationale: The primary models (m1a-m1c) assumed that light effects are identical across all participants (fixed slopes). This exploratory analysis tests whether participants vary in their response to current-day light exposure by fitting random slopes for Light_Lag0.

```{r dpg-random-slope-lag0}
# Fit model with random slopes for Light_Lag0
m1c_dpg_rs_lag0 <- lmer(
  DPG_intervention ~ Light_Lag0_num + Light_Lag1_num + Light_Lag2_num +
    (1 + Light_Lag0_num | participant_id),
  data = ibuttons_lmm, 
  REML = FALSE, 
  control = lmerControl(optimizer = "bobyqa")
)

summary(m1c_dpg_rs_lag0)

# Wald Confidence Intervals
ci_m1c_dpg_rs_lag0 <- confint(m1c_dpg_rs_lag0, method = "Wald")
print(ci_m1c_dpg_rs_lag0)
```

Interpretation Exploratory Analysis 1 Random Slopes for Light_Lag0: The model with random slopes for Light_Lag0 resulted in a singular fit (correlation between random intercept and slope = 1.00), indicating that the data do not support individual variation in current-day light exposure effects. The random slope variance was minimal (SD = 0.18) and fully confounded with the random intercept structure.

## Exploratory Analysis 2: Age and Sex Moderation

Rationale: To test whether the skin temperature gradient is influenced by biological sex or age.

```{r age-sex-moderation}
# Fit model with age and sex 
# Center age
ibuttons_lmm <- ibuttons_lmm %>%
  mutate(age_c = age - mean(age, na.rm = TRUE))

# Moderation model 
dpg_moderation <- lmer(
  DPG_intervention ~ 
    sex + age_c +
    (1 | participant_id),
  data = ibuttons_lmm,
  REML = FALSE
)

summary(dpg_moderation)

# Wald Confidence Intervals
ci_dpg_moderation <- confint(dpg_moderation, method = "Wald")
print(ci_dpg_moderation)
```

Interpretation Exploratory Analysis 2 Age and Sex Moderation: No statistically significant main effect was found of biological sex (male with female as reference category: β=0.35296, SE=0.84021, p=0.686) or age (centered: β=-0.02889, SE=0.10711, p=0.794). As no statistically signficant effects of the different light exposure variables or the experimental session number were found, no interactions between demographic factors and such fixed effects could be investigated.

## Exploratory Analysis 3: Light Lag Interactions

Rationale: Despite finding no significant main effects of light exposure on the current day or one or two days prior, all two-way interactions between the different Light Lags will be tested.

```{r lag-interactions}
dpg_interactions <- lmer(
  DPG_intervention ~ 
    Light_Lag0_num * Light_Lag1_num + 
    Light_Lag0_num * Light_Lag2_num +
    Light_Lag1_num * Light_Lag2_num +
    (1 | participant_id),
  data = ibuttons_lmm,
  REML = FALSE
)

summary(dpg_interactions)

# Wald Confidence Intervals
ci_dpg_interactions <- confint(dpg_interactions, method = "Wald")
print(ci_dpg_interactions)
```

Interpretation Exploratory 3 Light Lag Interactions: When including the interaction terms between the different Light Lags, the following results emerged:

-   Light Lag 2: significant effect of Lag 2 for the DPG with bright light two days prior suggesting, that bright light exposure two days prior is associated with a more negative DPG of around -0.8°C (β=-0.797319, SE=0.379887, p=0.0403) when controlling for interactions.

-   Light Lag 0 x Light Lag 2 Interaction: A significant positive interaction term between Lag 0 and Lag 2 was found (β=1.118357, SE=0.429289, p=0.0117). This interaction indicates that the effect of Lag2 depends on current-day light exposure: when both Lag0 and Lag2 are bright, the combined effect is less negative than would be predicted by adding the individual effects (non-additive relationship).

All other effects and interactions were not statistically significant (p\>0.3). Despite the exploratory nature of this Light Lag Interactions model, a rather synergistic effect of current evenings light exposure with the light exposure 2 days prior on current evenings DPG values can be found, whereby current bright light appears to attenuate the effect of bright light from two days ago.

## Exploratory Analysis 4: Persistence Analysis

Rationale: The persistence analysis for DPG addresses the question: Does the number of consecutive days of the same light condition (run length) modulate DPG responses? This tests whether:

-   Dim nights: Multiple consecutive dim evenings produce cumulative effects on DPG
-   Bright nights: Multiple consecutive bright evenings produce cumulative effects on DPG

Key distinction from lag analysis: - Lag analysis: Tests whether Day\[t-1\], Day\[t-2\] each add independent effects - Persistence analysis: Tests whether being on the 1st vs. 2nd vs. 3rd+ consecutive day of the same condition matters

## Dataset Preparation

```{r dpg-persistence-prep}
participant_list_dpg <- c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10")

# Light patterns for DPG 
light_patterns_dpg <- list(
  P01 = c("D", "D", "D", "B", "B", "B", "D", "D", "B", "D", "B"),
  P02 = c("D", "D", "D", "B", "B", "B", "D", "D", "B", "D", "B"),
  P03 = c("D", "D", "D", "B", "B", "B", "D", "D", "B", "D", "B"),
  P04 = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"),
  P05 = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"),
  P06 = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"),
  P07 = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"),
  P09 = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"),
  P10 = c("D", "D", "D", "B", "D", "B", "B", "B", "D", "B", "D", "D")
)

dpg_persistence <- map_dfr(participant_list_dpg, ~{
  df <- ibuttons %>% 
    filter(id == .x) %>%
    # Extract session number from session variable (e.g., "C01_01" -> 1)
    mutate(exp_session = as.integer(gsub("C0[1-4]_", "", session))) %>%
    filter(!is.na(DPG_intervention)) %>%
    # Aggregate by session: take mean DPG_intervention per session
    group_by(id, exp_session) %>%
    summarise(DPG_intervention = mean(DPG_intervention, na.rm = TRUE), .groups = "drop") %>%
    rename(participant_id = id)
  
  # Get light pattern
  pattern_vec <- light_patterns_dpg[[.x]]
  
  if (.x == "P10") {
    # For P10: create indices for sessions excluding position 5
    # P10 has 13 positions in total, but position 5 is excluded
    # Sessions are: 1,2,3,4,6,7,8,9,10,11,12,13
    valid_sessions <- c(1, 2, 3, 4, 6, 7, 8, 9, 10, 11, 12, 13)
    light_df <- tibble(
      exp_session = valid_sessions, 
      Light_Lag0 = pattern_vec
    )
  } else {
    light_df <- tibble(
      exp_session = seq_along(pattern_vec), 
      Light_Lag0 = pattern_vec
    )
  }
  
  df %>% left_join(light_df, by = "exp_session")
})

# Add demographics 
dpg_persistence <- dpg_persistence %>%
  left_join(participant_demographics_dpg, by = "participant_id") %>%
  filter(!is.na(Light_Lag0)) %>%
  filter(!(participant_id == "P10" & exp_session == 5)) %>%
  filter(participant_id != "P07") %>%  # Exclude P07
  arrange(participant_id, exp_session) %>%
  group_by(participant_id) %>%
  mutate(
    condition_change = Light_Lag0 != lag(Light_Lag0, default = first(Light_Lag0)),
    run_id = cumsum(condition_change),
    run_length = ave(run_id, run_id, FUN = seq_along),
    run_length_dim = ifelse(Light_Lag0 == "D", run_length, NA),
    run_length_bright = ifelse(Light_Lag0 == "B", run_length, NA),
    Light_Lag0_num = ifelse(Light_Lag0 == "D", 0, 1)
  ) %>%
  ungroup() %>%
  mutate(
    run_category_dim = factor(case_when(
      run_length_dim == 1 ~ "First",
      run_length_dim == 2 ~ "Second",
      run_length_dim >= 3 ~ "Third"
    ), levels = c("First", "Second", "Third")),
    run_category_bright = factor(case_when(
      run_length_bright == 1 ~ "First",
      run_length_bright == 2 ~ "Second",
      run_length_bright >= 3 ~ "Third"
    ), levels = c("First", "Second", "Third"))
  ) %>%
  select(participant_id, exp_session, DPG_intervention, Light_Lag0, Light_Lag0_num,
         run_length, run_length_dim, run_length_bright, 
         run_category_dim, run_category_bright, sex, age, bmi)

# Verify dataset
cat("Total observations:", nrow(dpg_persistence), "\n")
cat("Participants:", paste(unique(dpg_persistence$participant_id), collapse = ", "), "\n")
cat("P10 sessions:", paste(sort(dpg_persistence$exp_session[dpg_persistence$participant_id == "P10"]), collapse = ", "), "\n")
```

## Persistence Dataset

```{r dpg-persistence-table}
DT::datatable(
  dpg_persistence,
  options = list(
    pageLength = 20, 
    scrollX = TRUE, 
    order = list(list(0, 'asc'), list(1, 'asc'))
  ),
  caption = "DPG Persistence dataset with run-length encoding (P07 excluded, all sessions)",
  rownames = FALSE
) %>%
  DT::formatRound(columns = 'DPG_intervention', digits = 3) %>%
  DT::formatRound(columns = 'bmi', digits = 2) %>%
  DT::formatStyle(
    'Light_Lag0', 
    backgroundColor = DT::styleEqual(c('D', 'B'), c('#e3f2fd', '#fff3e0'))
  )
```

## Descriptive Visualizations

```{r dpg-persistence-plots, fig.width=12, fig.height=10}
color_dim <- "#4A90E2"
color_bright <- "#F5A623"

# Prepare plotting datasets
plot_data_dim_s1 <- dpg_persistence %>% 
  filter(!is.na(run_length_dim)) %>% 
  mutate(run_length_dim = factor(run_length_dim))

plot_data_bright_s1 <- dpg_persistence %>% 
  filter(!is.na(run_length_bright)) %>% 
  mutate(run_length_bright = factor(run_length_bright))

plot_data_dim_no_s1 <- dpg_persistence %>% 
  filter(!is.na(run_length_dim), exp_session > 1) %>% 
  mutate(run_length_dim = factor(run_length_dim))

plot_data_bright_no_s1 <- dpg_persistence %>% 
  filter(!is.na(run_length_bright), exp_session > 1) %>% 
  mutate(run_length_bright = factor(run_length_bright))

# Create plots
p_dim_s1 <- ggplot(plot_data_dim_s1, aes(x = run_length_dim, y = DPG_intervention)) +
  geom_boxplot(fill = color_dim, outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16) +
  labs(
    title = "Dim Nights (with Session 1)", 
    subtitle = paste0("n = ", nrow(plot_data_dim_s1)),
    x = "Consecutive Nights", 
    y = "DPG Intervention (°C)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13), 
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3)
  )

p_bright_s1 <- ggplot(plot_data_bright_s1, aes(x = run_length_bright, y = DPG_intervention)) +
  geom_boxplot(fill = color_bright, outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16) +
  labs(
    title = "Bright Nights (with Session 1)", 
    subtitle = paste0("n = ", nrow(plot_data_bright_s1)),
    x = "Consecutive Nights", 
    y = "DPG Intervention (°C)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13), 
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3)
  )

p_dim_no_s1 <- ggplot(plot_data_dim_no_s1, aes(x = run_length_dim, y = DPG_intervention)) +
  geom_boxplot(fill = color_dim, outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16) +
  labs(
    title = "Dim Nights (without Session 1)", 
    subtitle = paste0("n = ", nrow(plot_data_dim_no_s1)),
    x = "Consecutive Nights", 
    y = "DPG Intervention (°C)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13), 
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3)
  )

p_bright_no_s1 <- ggplot(plot_data_bright_no_s1, aes(x = run_length_bright, y = DPG_intervention)) +
  geom_boxplot(fill = color_bright, outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16) +
  labs(
    title = "Bright Nights (without Session 1)", 
    subtitle = paste0("n = ", nrow(plot_data_bright_no_s1)),
    x = "Consecutive Nights", 
    y = "DPG Intervention (°C)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13), 
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3)
  )

# Combine plots
(p_dim_s1 + p_bright_s1) / (p_dim_no_s1 + p_bright_no_s1) +
  plot_layout(heights = c(1, 1))
```

## Model 1: Linear Persistence Models, Dim Evenings

```{r dpg-persistence-dim-linear}
# Dim Evenings (all sessions), linear
dpg_persistence_dim <- dpg_persistence %>% filter(!is.na(run_length_dim))

model_persistence_dim_numeric <- lmer(
  DPG_intervention ~ run_length_dim + (1 | participant_id),
  data = dpg_persistence_dim,
  REML = FALSE
)

summary(model_persistence_dim_numeric)

# Wald Confidence Intervals
ci_model_persistence_dim_numeric <- confint(model_persistence_dim_numeric, method = "Wald")
print(ci_model_persistence_dim_numeric)

# Dim Evenings WITHOUT Session 1, linear
dpg_persistence_dim_no_s1 <- dpg_persistence %>% filter(!is.na(run_length_dim), exp_session > 1)

model_persistence_dim_no_s1 <- lmer(
  DPG_intervention ~ run_length_dim + (1 | participant_id),
  data = dpg_persistence_dim_no_s1,
  REML = FALSE
)

summary(model_persistence_dim_no_s1)

# Wald Confidence Intervals
ci_model_persistence_dim_no_s1 <- confint(model_persistence_dim_no_s1, method = "Wald")
print(ci_model_persistence_dim_no_s1)
```

Interpretation Exploratory 4 Persistence, Linear, Dim: To test whether consecutive dim light evenings produce cumulative effects on the distal-proximal skin temperature gradient during the light intervention window, the linear relationship between run length (number of consecutive dim evenings) and DPG was examined. Two models were fitted, whilst the first model includes all observations, the second model excludes Session 1, as there is no controlled light exposure on the previous day, which might influence results.

Including Session 1 for dim evening analysis did not find a linear effect of consecutive dim evenings for average skin temperature gradient (β=0.1282, SE=0.1947, p=0.515).

Excluding Session 1 for dim evening analysis did not find a linear effect of consecutive dim evenings for average skin temperature gradient (0.2968, SE=0.1985, p=0.146). Despite the visual impression that the DPG gradient is less negative, particularly on the third day of dim light, this visual trend could not be confirmed in the statistical analysis.

The positive (non-significant) coefficients suggest that with increasing number of consecutive dim nights, the DPG graident becomes less negative. However, findings did not reach statistical significance.

# Model 2: Linear Persistence Models, Bright Evenings

```{r dpg-persistence-bright-linear}
# Bright Evenings (all sessions), linear
dpg_persistence_bright <- dpg_persistence %>% filter(!is.na(run_length_bright))

model_persistence_bright_numeric <- lmer(
  DPG_intervention ~ run_length_bright + (1 | participant_id),
  data = dpg_persistence_bright,
  REML = FALSE
)

summary(model_persistence_bright_numeric)

# Wald Confidence Intervals
ci_model_persistence_bright_numeric <- confint(model_persistence_bright_numeric, method = "Wald")
print(ci_model_persistence_bright_numeric)

# Bright Evenings WITHOUT Session 1, linear
dpg_persistence_bright_no_s1 <- dpg_persistence %>% filter(!is.na(run_length_bright), exp_session > 1)

model_persistence_bright_no_s1 <- lmer(
  DPG_intervention ~ run_length_bright + (1 | participant_id),
  data = dpg_persistence_bright_no_s1,
  REML = FALSE
)

summary(model_persistence_bright_no_s1)

# Wald Confidence Intervals
ci_model_persistence_bright_no_s1 <- confint(model_persistence_bright_no_s1, method = "Wald")
print(ci_model_persistence_bright_no_s1)
```

Interpretation Exploratory 4 Persistence, Linear, Bright: To test whether consecutive bright light evenings produce cumulative effects on average skin temperature gradient during the light intervention window, the linear relationship between run length (number of consecutive dim evenings) and DPG was examined. As the first evening light exposure was dim for all participants, the analysis including and excluding session 1 are the same.

No statistically significant linear persistence effect was found for consecutive bright evenings (β=-0.0163, SE=0.1663, p=0.923).

# Model 3: Categorical Persistence Models, Dim Evenings

```{r dpg-persistence-dim-categorical}
# Dim Evenings (all sessions), categorical
dpg_persistence_dim <- dpg_persistence %>% filter(!is.na(run_category_dim))

model_persistence_dim_categorical <- lmer(
  DPG_intervention ~ run_category_dim + (1 | participant_id),
  data = dpg_persistence_dim,
  REML = FALSE
)

summary(model_persistence_dim_categorical)

# Wald Confidence Intervals
ci_model_persistence_dim_categorical <- confint(model_persistence_dim_categorical, method = "Wald")
print(ci_model_persistence_dim_categorical)

# Dim Evenings WITHOUT Session 1, categorical 
dpg_persistence_dim_no_s1 <- dpg_persistence %>% filter(!is.na(run_category_dim), exp_session > 1)

model_persistence_dim_categorical_no_s1 <- lmer(
  DPG_intervention ~ run_category_dim + (1 | participant_id),
  data = dpg_persistence_dim_no_s1,
  REML = FALSE
)

summary(model_persistence_dim_categorical_no_s1)

# Wald Confidence Intervals
ci_model_persistence_dim_categorical_no_s1 <- confint(model_persistence_dim_categorical_no_s1, method = "Wald")
print(ci_model_persistence_dim_categorical_no_s1)
```

Interpretation Exploratory 4 Persistence, Categorical, Dim: In order to examine potential non-linear patterns of cumulative dim light exposure on mean distal-proximal skin temperature gradient, a model was fitted using the categories second (2nd day cumulative dim light exposure) and third (3rd day cumulative dim light exposure) compared to the first dim night (reference category) to predict mean DPG values during the light intervention window. As in the linear persistence analysis, two models (first with, second without first experimental session) were fitted.

Including Session 1 for dim evening analysis did not find non-linear effects of consecutive dim evenings for DPG values: Second (β=-0.3801, SE=0.3246, p=0.249), Third (β=0.5141, SE=0.3954, p=0.202).

Excluding Session 1 for dim evening analysis did not find statistically significant non-linear effects of consecutive dim evenings for DPG values, however the effect of the third bright light evening almost reached statistical significance level: Second (β=-0.1461, SE=0.32, p=0.6587), Third (β=0.7745, SE=0.3948, p=0.0594).

# Model 4: Categorical Persistence Models, Bright Evenings

```{r dpg-persistence-bright-categorical}
# Bright Evenings (all sessions), categorical
dpg_persistence_bright <- dpg_persistence %>% filter(!is.na(run_category_bright))

model_persistence_bright_categorical <- lmer(
  DPG_intervention ~ run_category_bright + (1 | participant_id),
  data = dpg_persistence_bright,
  REML = FALSE
)

summary(model_persistence_bright_categorical)

# Wald Confidence Intervals
ci_model_persistence_bright_categorical <- confint(model_persistence_bright_categorical, method = "Wald")
print(ci_model_persistence_bright_categorical)

# Bright Evenings WITHOUT Session 1, categorical 
pg_persistence_bright_no_s1 <- dpg_persistence %>% filter(!is.na(run_category_bright), exp_session > 1)

model_persistence_bright_categorical_no_s1 <- lmer(
  DPG_intervention ~ run_category_bright + (1 | participant_id),
  data = dpg_persistence_bright_no_s1,
  REML = FALSE
)

summary(model_persistence_bright_categorical_no_s1)

# Wald Confidence Intervals
ci_model_persistence_bright_categorical_no_s1 <- confint(model_persistence_bright_categorical_no_s1, method = "Wald")
print(ci_model_persistence_bright_categorical_no_s1)
```

Interpretation: In order to examine potential non-linear patterns of cumulative bright light exposure on mean skin temperature gradient during light intervention period, a model was fitted using the categories second (2nd day cumulative bright light exposure) and third (3rd day cumulative bright light exposure) compared to the first bright night (reference category) to predict DPG values As the first evening light exposure was dim for all participants, the analysis including and excluding session 1 are the same.

No statistically significant effects of the second consecutive bright evening was found for the second (β=-0.31031, SE=0.33753, p=0.365), and the third consecutive bright evening did not yield statistical significance (β=0.04873, SE=0.33753, p=0.886).

# Exploratory Analysis: Climate Chamber Temperature

An exploratory model including the temperature data from the climate chamber during the intervention window will be used to identify a potential impact of the climate temperature on distal-proximal gradient values. Despite efforts to keep the chamber at a constant temperature throughout and across sessions, some deviations happened, hence this analysis examines the robustness of findings to the inclusion of the climate chamber data.

## Load Climate Data and Use UNIX Column (Correct Timestamps)

```{r climate-load}
# Load climate data
climate_lab <- read_excel("C:/Users/anton/OneDrive/AfterGlow/data/raw/group/climate/01_climate_20250831T100000.xlsx")

str(climate_lab)
```

```{r climate-convert-unix}
# Convert UNIX timestamp to MESZ (UTC+2 for Munich summer time)
climate_lab$unix_converted <- as_datetime(climate_lab$UNIX, tz = "Europe/Berlin")

head(climate_lab)
```

## Group 1 Data Extraction & Processing

```{r climate-group1}
# Filter data
group1_dates <- seq(as.Date("2025-06-16"), by = "day", length.out = 11)
climate_group1_all <- climate_lab %>%
  filter(
    as.Date(unix_converted) %in% group1_dates,
    hour(unix_converted) == 21 & minute(unix_converted) >= 30 |
    hour(unix_converted) == 22 |
    hour(unix_converted) == 23 & minute(unix_converted) == 0
  ) %>%
  mutate(
    unix_converted = floor_date(unix_converted, unit = "minute"),
    temp_felt = (tg + tdb) / 2
  )

# Check missing data
gaps <- climate_group1_all %>%
  arrange(unix_converted) %>%
  group_by(as.Date(unix_converted)) %>%
  mutate(time_diff = as.numeric(difftime(unix_converted, lag(unix_converted), units = "mins"))) %>%
  filter(time_diff > 1) %>%
  ungroup()

cat("Total obs:", nrow(climate_group1_all), "| Gaps > 1 min:", nrow(gaps), "\n\n")
table(as.Date(climate_group1_all$unix_converted))

# Daily average
temp_felt_daily <- climate_group1_all %>%
  group_by(date_only = as.Date(unix_converted)) %>%
  summarise(temp_felt_mean = mean(temp_felt, na.rm = TRUE), n_obs = n())
print(temp_felt_daily)
```

## Group 2 Data Extraction & Processing

```{r climate-group2}
# Filter data
group2_dates <- seq(as.Date("2025-06-30"), by = "day", length.out = 10)
climate_group2_all <- climate_lab %>%
  filter(
    as.Date(unix_converted) %in% group2_dates,
    hour(unix_converted) == 19 & minute(unix_converted) >= 30 |
    hour(unix_converted) == 20 |
    hour(unix_converted) == 21 & minute(unix_converted) == 0
  ) %>%
  mutate(
    unix_converted = floor_date(unix_converted, unit = "minute"),
    temp_felt = (tg + tdb) / 2
  )

# Check missing data
gaps <- climate_group2_all %>%
  arrange(unix_converted) %>%
  group_by(as.Date(unix_converted)) %>%
  mutate(time_diff = as.numeric(difftime(unix_converted, lag(unix_converted), units = "mins"))) %>%
  filter(time_diff > 1) %>%
  ungroup()

cat("Total obs:", nrow(climate_group2_all), "| Gaps > 1 min:", nrow(gaps), "\n\n")
table(as.Date(climate_group2_all$unix_converted))

# Daily average
temp_felt_daily_group2 <- climate_group2_all %>%
  group_by(date_only = as.Date(unix_converted)) %>%
  summarise(temp_felt_mean = mean(temp_felt, na.rm = TRUE), n_obs = n())
print(temp_felt_daily_group2)
```

## Group 3 Data Extraction & Processing

```{r climate-group3}
# Filter data
group3_dates <- seq(as.Date("2025-07-28"), by = "day", length.out = 10)
climate_group3_all <- climate_lab %>%
  filter(
    as.Date(unix_converted) %in% group3_dates,
    hour(unix_converted) == 21 |
    hour(unix_converted) == 22 & minute(unix_converted) <= 30
  ) %>%
  mutate(
    unix_converted = floor_date(unix_converted, unit = "minute"),
    temp_felt = (tg + tdb) / 2
  )

# Check missing data
gaps <- climate_group3_all %>%
  arrange(unix_converted) %>%
  group_by(as.Date(unix_converted)) %>%
  mutate(time_diff = as.numeric(difftime(unix_converted, lag(unix_converted), units = "mins"))) %>%
  filter(time_diff > 1) %>%
  ungroup()

cat("Total obs:", nrow(climate_group3_all), "| Gaps > 1 min:", nrow(gaps), "\n\n")
table(as.Date(climate_group3_all$unix_converted))

# Daily average
temp_felt_daily_group3 <- climate_group3_all %>%
  group_by(date_only = as.Date(unix_converted)) %>%
  summarise(temp_felt_mean = mean(temp_felt, na.rm = TRUE), n_obs = n())
print(temp_felt_daily_group3)
```

## Group 4 Data Extraction & Processing

```{r climate-group4}
# Filter data
group4_dates <- seq(as.Date("2025-08-11"), by = "day", length.out = 13)
climate_group4_all <- climate_lab %>%
  filter(
    as.Date(unix_converted) %in% group4_dates,
    hour(unix_converted) == 20 & minute(unix_converted) >= 30 |
    hour(unix_converted) == 21 |
    hour(unix_converted) == 22 & minute(unix_converted) == 0
  ) %>%
  mutate(
    unix_converted = floor_date(unix_converted, unit = "minute"),
    temp_felt = (tg + tdb) / 2
  )

# Check missing data
gaps <- climate_group4_all %>%
  arrange(unix_converted) %>%
  group_by(as.Date(unix_converted)) %>%
  mutate(time_diff = as.numeric(difftime(unix_converted, lag(unix_converted), units = "mins"))) %>%
  filter(time_diff > 1) %>%
  ungroup()

cat("Total obs:", nrow(climate_group4_all), "| Gaps > 1 min:", nrow(gaps), "\n\n")
table(as.Date(climate_group4_all$unix_converted))

# Daily average
temp_felt_daily_group4 <- climate_group4_all %>%
  group_by(date_only = as.Date(unix_converted)) %>%
  summarise(temp_felt_mean = mean(temp_felt, na.rm = TRUE), n_obs = n())
print(temp_felt_daily_group4)
```

## Add Climate Temperature to ibuttons_lmm Dataset

```{r climate-join-ibuttons}
# Remove temperature columns if they exist
ibuttons_lmm <- ibuttons_lmm %>%
  select(-contains("temp_felt_mean"), -contains("date_only"))

# Create temperature lookup
temp_lookup_dpg <- bind_rows(
  temp_felt_daily %>% mutate(group = 1, participants = list(c("P01", "P02", "P03"))),
  temp_felt_daily_group2 %>% mutate(group = 2, participants = list(c("P04", "P05"))),
  temp_felt_daily_group3 %>% mutate(group = 3, participants = list(c("P06", "P07", "P09"))),
  temp_felt_daily_group4 %>% mutate(group = 4, participants = list(c("P10")))
) %>%
  unnest(participants) %>%
  group_by(participants, group) %>%
  mutate(exp_session = row_number()) %>%
  ungroup() %>%
  select(participant_id = participants, exp_session, temp_felt_mean, date_only)

# Join temperature data
ibuttons_lmm <- ibuttons_lmm %>%
  left_join(temp_lookup_dpg, by = c("participant_id", "exp_session"))

# Check result
ibuttons_lmm %>% 
  filter(participant_id == "P02", exp_session == 3) %>%
  select(participant_id, exp_session, DPG_intervention, temp_felt_mean, date_only)
```

```{r climate-datatable-ibuttons}
DT::datatable(
  ibuttons_lmm,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    order = list(list(0, 'asc'), list(1, 'asc'))
  ),
  caption = "Complete DPG dataset with temperature data",
  rownames = FALSE
) %>%
  DT::formatRound(columns = 'DPG_intervention', digits = 3) %>%
  DT::formatRound(columns = 'temp_felt_mean', digits = 2) %>%
  DT::formatRound(columns = 'bmi', digits = 2) %>%
  DT::formatStyle(
    'Light_Lag0',
    backgroundColor = DT::styleEqual(c('D', 'B'), c('#e3f2fd', '#fff3e0'))
  )
```

## Fit LMM with Temperature as Fixed Effect

```{r climate-models-dpg}
# Model 1 m_temp: Temperature as predictor
m_temp_dpg <- lmer(
  DPG_intervention ~ temp_felt_mean + (1 | participant_id),
  data = ibuttons_lmm,
  REML = FALSE
)

summary(m_temp_dpg)

# Wald Confidence Intervals
ci_m_temp_dpg <- confint(m_temp_dpg, method = "Wald")
print(ci_m_temp_dpg)

# Model 2: Session + Temperature
m_session_temp_dpg <- lmer(
  DPG_intervention ~ exp_session + temp_felt_mean + (1 | participant_id),
  data = ibuttons_lmm,
  REML = FALSE
)

summary(m_session_temp_dpg)

# Wald Confidence Intervals
ci_m_session_temp_dpg <- confint(m_session_temp_dpg, method = "Wald")
print(ci_m_session_temp_dpg)

# Model 3: Lag 0 + Temperature 
m_lag0_temp_dpg <- lmer(
  DPG_intervention ~ Light_Lag0_num + temp_felt_mean + (1 | participant_id),
  data = ibuttons_lmm,
  REML = FALSE
)

summary(m_lag0_temp_dpg)

# Wald Confidence Intervals
ci_m_lag0_temp_dpg <- confint(m_lag0_temp_dpg, method = "Wald")
print(ci_m_lag0_temp_dpg)

# Model 4: Lag 1 + Temperature 
m_lag1_temp_dpg <- lmer(
  DPG_intervention ~ Light_Lag1_num + temp_felt_mean + (1 | participant_id),
  data = ibuttons_lmm,
  REML = FALSE
)

summary(m_lag1_temp_dpg)

# Wald Confidence Intervals
ci_m_lag1_temp_dpg <- confint(m_lag1_temp_dpg, method = "Wald")
print(ci_m_lag1_temp_dpg)

# Model 5: Lag 2 + Temperature 
m_lag2_temp_dpg <- lmer(
  DPG_intervention ~ Light_Lag2_num + temp_felt_mean + (1 | participant_id),
  data = ibuttons_lmm,
  REML = FALSE
)

summary(m_lag2_temp_dpg)

# Wald Confidence Intervals
ci_m_lag2_temp_dpg <- confint(m_lag2_temp_dpg, method = "Wald")
print(ci_m_lag2_temp_dpg)
```

Interpretation Exploratory 5 Climate Data: To examine potential effects of laboratory sensed temperature on DPG values, sensitivity analyses including climate chamber temperature as a covariate were conducted:

-   Model m_temp_dpg: A significant effect of the sensed temperature on DPG values was observed (β=0.6542, SE=0.2436, p=0.00933). Thereby an increase in sensed temperature by 1°C was associated with an increase in around 0.65°C in the distal-proximal skin temperature gradient (i.e., DPG became less negative, indicating greater distal vasodilation).

-   Model m_session_temp_dpg: A significant effect for both sensed temperature (β=0.72502, SE=0.23627, p=0.00322) as well as experimental session number (β=-0.08596, SE=0.04093, p=0.04010). Thus an increase in 1°C of sensed temperature was associated with approximately an increase in 0.73°C of DPG, whereas an increase in experimental session number was associated with a decrease in DPG by around -0.086°C (more negative gradient).

-   Model m_lag0_temp_dpg: When controlling for sensed temperature (β=0.70046, SE=0.28172, p=0.0156), no effect of current evening light exposure could be found on mean DPG values (β=-0.07833, SE=0.24723, p=0.7525).

-   Model m_lag1_temp_dpg: Similarly when controlling for sensed temperature (β=0.64546, SE=0.24525, p=0.01075), no statistically significant effect of pior evenings light exposure could be found on DPG values (β=-0.06276, SE=0.21452, p=0.77093).

-   Model m_lag2_temp_dpg: Similarly when controlling for sensed temperature (β=0.65086, SE=0.24471, p=0.00998), no statistically significant effect of the light exposure two days before could be found for mean DPG values (β=-0.03041, SE=0.21405, p=0.88755 ).

In summary, despite consistent significant effects of an increase of sensed temperature being associated with an increase in DPG values, no significant effects of light exposure history (Lag 0, Lag 1, Lag 2) could be found wen controlling for ambient temperature.

# Statistical Analysis Summary Table

## Model Comparison: All Tested Models for DPG

## Helper Functions

```{r dpg-summary-helpers}
# Helper functions to extract coefficients from LMM
get_coef_lmm <- function(model, coef_name) {
  coefs <- summary(model)$coefficients
  if (coef_name %in% rownames(coefs)) {
    return(coefs[coef_name, "Estimate"])
  } else {
    return(NA)
  }
}

get_p_lmm <- function(model, coef_name) {
  coefs <- summary(model)$coefficients
  if (coef_name %in% rownames(coefs)) {
    return(coefs[coef_name, "Pr(>|t|)"])
  } else {
    return(NA)
  }
}
```

## Model Comparison: All Tested Models for DPG

```{r dpg-summary-table}
# Create extended comparison table for DPG
model_comparison_dpg <- data.frame(
  Model = c(
    "m0 (Baseline)",
    "m1a (Lag 0)",
    "m1b (Lag 0+1)",
    "m1c (Lag 0+1+2)",
    "m2 (+ Session)",
    "m3 (+ Session²)",
    "m0_sens (P06 excluded)",
    "m1a_sens (P06 excluded)",
    "m1b_sens (P06 excluded)",
    "m1c_sens (P06 excluded)",
    "m2_sens (P06 excluded)",
    "m3_sens (P06 excluded)",
    "Random Slopes Lag0",
    "Light Interactions",
    "Persistence Dim (linear)",
    "Persistence Bright (linear)",
    "Persistence Dim (categorical)",
    "Persistence Bright (categorical)",
    "Age/Sex Moderation Light",
    "Temperature Only",
    "Session + Temperature",
    "Lag0 + Temperature",
    "Lag1 + Temperature",
    "Lag2 + Temperature"
  ),
  
  Description = c(
    "Intercept only",
    "Light Lag 0 only",
    "Lag 0 + Lag 1",
    "Lag 0 + Lag 1 + Lag 2",
    "Session (linear)",
    "Session + Session²",
    "Baseline (n=7, P06+P07 excluded)",
    "Lag 0 only (n=7, P06+P07 excluded)",
    "Lag 0 + Lag 1 (n=7, P06+P07 excluded)",
    "Lag 0 + Lag 1 + Lag 2 (n=7, P06+P07 excluded)",
    "Session (linear, n=7, P06+P07 excluded)",
    "Session + Session² (n=7, P06+P07 excluded)",
    "m1c + random slope Lag0",
    "Lag 0 × Lag 1 × Lag 2 interactions",
    "Run length dim (continuous)",
    "Run length bright (continuous)",
    "Run category dim (First/Second/Third)",
    "Run category bright (First/Second/Third)",
    "m1c + age + sex + light interactions",
    "Temperature as predictor",
    "Session + Temperature",
    "Lag 0 + Temperature",
    "Lag 1 + Temperature",
    "Lag 2 + Temperature"
  ),
  
  npar = c(
    attr(logLik(m0_dpg), "df"),
    attr(logLik(m1a_dpg), "df"),
    attr(logLik(m1b_dpg), "df"),
    attr(logLik(m1c_dpg), "df"),
    attr(logLik(m2_dpg), "df"),
    attr(logLik(m3_dpg), "df"),
    attr(logLik(m0_dpg_sens), "df"),
    attr(logLik(m1a_dpg_sens), "df"),
    attr(logLik(m1b_dpg_sens), "df"),
    attr(logLik(m1c_dpg_sens), "df"),
    attr(logLik(m2_dpg_sens), "df"),
    attr(logLik(m3_dpg_sens), "df"),
    attr(logLik(m1c_dpg_rs_lag0), "df"),
    attr(logLik(dpg_interactions), "df"),
    attr(logLik(model_persistence_dim_numeric), "df"),
    attr(logLik(model_persistence_bright_numeric), "df"),
    attr(logLik(model_persistence_dim_categorical), "df"),
    attr(logLik(model_persistence_bright_categorical), "df"),
    NA,  # Age/Sex moderation if fitted
    attr(logLik(m_temp_dpg), "df"),
    attr(logLik(m_session_temp_dpg), "df"),
    attr(logLik(m_lag0_temp_dpg), "df"),
    attr(logLik(m_lag1_temp_dpg), "df"),
    attr(logLik(m_lag2_temp_dpg), "df")
  ),
  
  N_obs = c(
    rep(nrow(ibuttons_lmm), 6),  # Primary models (n=65)
    rep(nrow(ibuttons_lmm_sens), 6),  # Sensitivity models (n=57)
    nrow(ibuttons_lmm),  # Random slopes
    nrow(ibuttons_lmm),  # Interactions
    nrow(dpg_persistence_dim),  # Persistence dim
    nrow(dpg_persistence_bright),  # Persistence bright
    nrow(dpg_persistence_dim),  # Persistence dim categorical
    nrow(dpg_persistence_bright),  # Persistence bright categorical
    nrow(ibuttons_lmm),  # Age/Sex moderation
    rep(nrow(ibuttons_lmm), 5)  # Temperature models
  ),
  
  AIC = c(
    round(AIC(m0_dpg), 1),
    round(AIC(m1a_dpg), 1),
    round(AIC(m1b_dpg), 1),
    round(AIC(m1c_dpg), 1),
    round(AIC(m2_dpg), 1),
    round(AIC(m3_dpg), 1),
    round(AIC(m0_dpg_sens), 1),
    round(AIC(m1a_dpg_sens), 1),
    round(AIC(m1b_dpg_sens), 1),
    round(AIC(m1c_dpg_sens), 1),
    round(AIC(m2_dpg_sens), 1),
    round(AIC(m3_dpg_sens), 1),
    round(AIC(m1c_dpg_rs_lag0), 1),
    round(AIC(dpg_interactions), 1),
    round(AIC(model_persistence_dim_numeric), 1),
    round(AIC(model_persistence_bright_numeric), 1),
    round(AIC(model_persistence_dim_categorical), 1),
    round(AIC(model_persistence_bright_categorical), 1),
    NA,  # Age/Sex moderation
    round(AIC(m_temp_dpg), 1),
    round(AIC(m_session_temp_dpg), 1),
    round(AIC(m_lag0_temp_dpg), 1),
    round(AIC(m_lag1_temp_dpg), 1),
    round(AIC(m_lag2_temp_dpg), 1)
  ),
  
  BIC = c(
    round(BIC(m0_dpg), 1),
    round(BIC(m1a_dpg), 1),
    round(BIC(m1b_dpg), 1),
    round(BIC(m1c_dpg), 1),
    round(BIC(m2_dpg), 1),
    round(BIC(m3_dpg), 1),
    round(BIC(m0_dpg_sens), 1),
    round(BIC(m1a_dpg_sens), 1),
    round(BIC(m1b_dpg_sens), 1),
    round(BIC(m1c_dpg_sens), 1),
    round(BIC(m2_dpg_sens), 1),
    round(BIC(m3_dpg_sens), 1),
    round(BIC(m1c_dpg_rs_lag0), 1),
    round(BIC(dpg_interactions), 1),
    round(BIC(model_persistence_dim_numeric), 1),
    round(BIC(model_persistence_bright_numeric), 1),
    round(BIC(model_persistence_dim_categorical), 1),
    round(BIC(model_persistence_bright_categorical), 1),
    NA,  # Age/Sex moderation
    round(BIC(m_temp_dpg), 1),
    round(BIC(m_session_temp_dpg), 1),
    round(BIC(m_lag0_temp_dpg), 1),
    round(BIC(m_lag1_temp_dpg), 1),
    round(BIC(m_lag2_temp_dpg), 1)
  ),
  
  Beta_Main = c(
    NA,
    round(get_coef_lmm(m1a_dpg, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m1b_dpg, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m1c_dpg, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m2_dpg, "exp_session"), 3),
    round(get_coef_lmm(m3_dpg, "exp_session"), 3),
    NA,
    round(get_coef_lmm(m1a_dpg_sens, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m1b_dpg_sens, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m1c_dpg_sens, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m2_dpg_sens, "exp_session"), 3),
    round(get_coef_lmm(m3_dpg_sens, "exp_session"), 3),
    round(get_coef_lmm(m1c_dpg_rs_lag0, "Light_Lag0_num"), 3),
    round(get_coef_lmm(dpg_interactions, "Light_Lag2_num"), 3),
    round(get_coef_lmm(model_persistence_dim_numeric, "run_length_dim"), 3),
    round(get_coef_lmm(model_persistence_bright_numeric, "run_length_bright"), 3),
    round(get_coef_lmm(model_persistence_dim_categorical, "run_category_dimSecond"), 3),
    round(get_coef_lmm(model_persistence_bright_categorical, "run_category_brightSecond"), 3),
    NA,  # Age/Sex moderation
    round(get_coef_lmm(m_temp_dpg, "temp_felt_mean"), 3),
    round(get_coef_lmm(m_session_temp_dpg, "exp_session"), 3),
    round(get_coef_lmm(m_lag0_temp_dpg, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m_lag1_temp_dpg, "Light_Lag1_num"), 3),
    round(get_coef_lmm(m_lag2_temp_dpg, "Light_Lag2_num"), 3)
  ),
  
  p_Main = c(
    NA,
    format.pval(get_p_lmm(m1a_dpg, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1b_dpg, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1c_dpg, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m2_dpg, "exp_session"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m3_dpg, "exp_session"), digits = 3, eps = 0.001),
    NA,
    format.pval(get_p_lmm(m1a_dpg_sens, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1b_dpg_sens, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1c_dpg_sens, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m2_dpg_sens, "exp_session"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m3_dpg_sens, "exp_session"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1c_dpg_rs_lag0, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(dpg_interactions, "Light_Lag2_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(model_persistence_dim_numeric, "run_length_dim"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(model_persistence_bright_numeric, "run_length_bright"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(model_persistence_dim_categorical, "run_category_dimSecond"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(model_persistence_bright_categorical, "run_category_brightSecond"), digits = 3, eps = 0.001),
    NA,  # Age/Sex moderation
    format.pval(get_p_lmm(m_temp_dpg, "temp_felt_mean"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m_session_temp_dpg, "exp_session"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m_lag0_temp_dpg, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m_lag1_temp_dpg, "Light_Lag1_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m_lag2_temp_dpg, "Light_Lag2_num"), digits = 3, eps = 0.001)
  ),
  
  Status = c(
    "FINAL MODEL",
    "Step 1 (non-sig)",
    "Step 2 (non-sig)",
    "Step 3 (non-sig)",
    "Not selected (trend p=0.12)",
    "Not selected (non-sig)",
    "Sensitivity (n=7)",
    "Sensitivity (n=7, non-sig)",
    "Sensitivity (n=7, non-sig)",
    "Sensitivity (n=7, non-sig)",
    "Sensitivity (n=7, non-sig)",
    "Sensitivity (n=7, non-sig)",
    "Exploratory (singular fit)",
    "Exploratory (Lag2 sig, Lag0×Lag2 sig)",
    "Exploratory (non-sig)",
    "Exploratory (non-sig)",
    "Exploratory (non-sig)",
    "Exploratory (non-sig)",
    "Exploratory (non-sig)",
    "Exploratory (temp sig)",
    "Exploratory (session sig, temp sig)",
    "Exploratory (temp sig)",
    "Exploratory (temp sig)",
    "Exploratory (temp sig)"
  )
)

kable(model_comparison_dpg, 
      caption = "Model Comparison: All Tested Models for Distal-Proximal Gradient (DPG)",
      align = c('l', 'l', 'c', 'c', 'c', 'c', 'c', 'c', 'l'),
      col.names = c("Model", "Description", "npar", "N", "AIC", "BIC", 
                    "β Main", "p Main", "Status")) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  row_spec(1, bold = TRUE, background = "#D6EAF8") %>%
  pack_rows("Primary Model Selection (n=65)", 1, 6) %>%
  pack_rows("Sensitivity Analyses (P06 excluded, n=57)", 7, 12) %>%
  pack_rows("Exploratory Analyses", 13, 24)
```

## Final Model Summary: m0_dpg

```{r dpg-final-model-summary}
# Extract components from m0_dpg
m0_dpg_coef <- summary(m0_dpg)$coefficients

# Variance components
vc_final_dpg <- as.data.frame(VarCorr(m0_dpg))

cat("**Final Model: m0_dpg (Baseline - Random Intercept Only)**\n\n")

cat("**Fixed Effects:**\n")
cat("Intercept:", round(m0_dpg_coef["(Intercept)", "Estimate"], 3), "°C\n")
cat("  SE:", round(m0_dpg_coef["(Intercept)", "Std. Error"], 3), "\n")
cat("  p:", format.pval(m0_dpg_coef["(Intercept)", "Pr(>|t|)"], digits = 3, eps = 0.001), "\n\n")

cat("**Variance Components:**\n")
cat("Between-participant (Random Intercept):", round(vc_final_dpg$vcov[1], 3), "\n")
cat("Within-participant (Residual):", round(vc_final_dpg$vcov[2], 3), "\n\n")

# Calculate ICC
icc_final_dpg <- vc_final_dpg$vcov[1] / (vc_final_dpg$vcov[1] + vc_final_dpg$vcov[2])
cat("ICC (Intraclass Correlation):", round(icc_final_dpg, 3), 
    paste0("(", round(icc_final_dpg * 100, 1), "% of variance between participants)"), "\n\n")

cat("**Model Fit:**\n")
cat("AIC:", round(AIC(m0_dpg), 1), "| BIC:", round(BIC(m0_dpg), 1), "\n")
cat("N obs:", nrow(ibuttons_lmm), "| N participants:", 
    length(unique(ibuttons_lmm$participant_id)), "\n")
cat("Family: Gaussian (identity link)\n\n")
```

# Combined figure of boxplots 
```{r}
#| label: fig-combined-dpg-boxplots
#| fig-cap: "DPG under light exposure doublets and triplets (P07 excluded)"
#| fig-width: 20
#| fig-height: 16

# Labels with bold second letter
doublet_labels <- c(
  "DD" = expression("D"*bold("D")),
  "BD" = expression("B"*bold("D")),
  "DB" = expression("D"*bold("B")),
  "BB" = expression("B"*bold("B"))
)

# Colors based on second letter
doublet_colors <- c(
  "DD" = "#4A90E2", "BD" = "#4A90E2",
  "DB" = "#F5A623", "BB" = "#F5A623"
)

# Labels with bold third letter
triplet_labels <- c(
  "DDD" = expression("DD"*bold("D")),
  "BDD" = expression("BD"*bold("D")),
  "DBD" = expression("DB"*bold("D")),
  "BBD" = expression("BB"*bold("D")),
  "DDB" = expression("DD"*bold("B")),
  "BDB" = expression("BD"*bold("B")),
  "DBB" = expression("DB"*bold("B")),
  "BBB" = expression("BB"*bold("B"))
)

# Colors based on third letter
triplet_colors <- c(
  "DDD" = "#4A90E2", "BDD" = "#4A90E2", "DBD" = "#4A90E2", "BBD" = "#4A90E2",
  "DDB" = "#F5A623", "BDB" = "#F5A623", "DBB" = "#F5A623", "BBB" = "#F5A623"
)

# Compute shared y-axis range across all four datasets
all_values <- c(
  df_plot_dpg_noP07$mean_dpg,
  df_plot_dpg_int_noP07$mean_dpg_int,
  df_plot_dpg_triplet_noP07$mean_dpg,
  df_plot_dpg_triplet_int_noP07$mean_dpg_int
)
y_min <- floor(min(all_values, na.rm = TRUE))
y_max <- ceiling(max(all_values, na.rm = TRUE))
y_breaks <- pretty(c(y_min, y_max), n = 6)

# --- Plot 1: Doublet 5h ---
p1 <- ggplot(df_plot_dpg_noP07, aes(x = doublet, y = mean_dpg)) +
  geom_boxplot(aes(fill = doublet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15, seed = 42),
             size = 2.5, alpha = 0.7, shape = 16) +
  scale_fill_manual(values = doublet_colors) +
  scale_x_discrete(labels = doublet_labels) +
  scale_y_continuous(limits = c(y_min, y_max), breaks = y_breaks) +
  labs(
    title = "Doublets \u2013 5-h exposure",
    x = "Doublet",
    y = "Mean Distal-Proximal Gradient (\u00b0C)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )

# --- Plot 2: Doublet 1.5h ---
p2 <- ggplot(df_plot_dpg_int_noP07, aes(x = doublet, y = mean_dpg_int)) +
  geom_boxplot(aes(fill = doublet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15, seed = 42),
             size = 2.5, alpha = 0.7, shape = 16) +
  scale_fill_manual(values = doublet_colors) +
  scale_x_discrete(labels = doublet_labels) +
  scale_y_continuous(limits = c(y_min, y_max), breaks = y_breaks) +
  labs(
    title = "Doublets \u2013 1.5-h intervention",
    x = "Doublet",
    y = "Mean Distal-Proximal Gradient (\u00b0C)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )

# --- Plot 3: Triplet 5h ---
p3 <- ggplot(df_plot_dpg_triplet_noP07, aes(x = triplet, y = mean_dpg)) +
  geom_boxplot(aes(fill = triplet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15, seed = 42),
             size = 2.5, alpha = 0.7, shape = 16) +
  scale_fill_manual(values = triplet_colors) +
  scale_x_discrete(labels = triplet_labels) +
  scale_y_continuous(limits = c(y_min, y_max), breaks = y_breaks) +
  labs(
    title = "Triplets \u2013 5-h exposure",
    x = "Triplet",
    y = "Mean Distal-Proximal Gradient (\u00b0C)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )

# --- Plot 4: Triplet 1.5h ---
p4 <- ggplot(df_plot_dpg_triplet_int_noP07, aes(x = triplet, y = mean_dpg_int)) +
  geom_boxplot(aes(fill = triplet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15, seed = 42),
             size = 2.5, alpha = 0.7, shape = 16) +
  scale_fill_manual(values = triplet_colors) +
  scale_x_discrete(labels = triplet_labels) +
  scale_y_continuous(limits = c(y_min, y_max), breaks = y_breaks) +
  labs(
    title = "Triplets \u2013 1.5-h intervention",
    x = "Triplet",
    y = "Mean Distal-Proximal Gradient (\u00b0C)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )

# --- Combine ---
(p1 | p2) / (p3 | p4) +
  plot_annotation(
    title = "Distal-proximal skin temperature gradient under light exposure doublets and triplets",
    subtitle = "Mean DPG of the second / third day per sequence type (5-h exposure & 1.5-h intervention; P07 excluded)",
    theme = theme(
      plot.title = element_text(size = 24, hjust = 0),
      plot.subtitle = element_text(size = 18, hjust = 0, margin = margin(b = 10))
    )
  )
```
