---
title: "Carryover Effects in Melatonin Suppression"
subtitle: "Karolinska Sleepiness Scale (KSS) Analysis"
author: "Letizia Woerrlein"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: show
    code-tools: true
    df-print: paged
    embed-resources: true
    theme: cosmo
    fig-width: 10
    fig-height: 8
execute:
  warning: false
  message: false
  cache: false
---

# Study Overview {.unnumbered}

## Design

**Study Type**: Within-subject controlled trial

**Total Participants**: 9 (P01-P07, P09-P11)

**Exclusions**:

- P11: Complete exclusion (study dropout due to infection)
- P10 Session 5: Excluded (light automation malfunction)
- P07: Excluded from statistical analyses (negative melatonin suppression)

## Study Phases

### Phase 1: Pre-Ambulatory Period

- **Duration**: 7 days
- **Setting**: At home
- **Measurements**: Sleep logs, ActLumus, ActiGraph, CGM

### Phase 2: In-Laboratory Period

- **Duration**: 10-13 days 
- **Daily protocol**: Arrival 5.5h before habitual bedtime (HBT); Departure 0.5h before HBT
- **Light conditions**:
  - D (Dim): 10 lx continuously
  - B (Bright): 10 lx → 1000 lx (1.5h) → 10 lx
- **KSS measurements**: Every 30 minutes throughout chamber period (11 measurements per session, questionnaires 1-11)

### Phase 3: Post-Ambulatory Period

- **Duration**: 7 days
- **Setting**: At home (identical to Phase 1)

## Participant Groups

### Group 1 (HBT 00:00)

- **Participants**: P01, P02, P03
- **Light sequence**: DDDBBBDDBDB (11 sessions)

### Group 2 (HBT 22:00)

- **Participants**: P04, P05
- **Light sequence**: DBBBDDDBDB (10 sessions)

### Group 3 (HBT 23:30)

- **Participants**: P06, P07, P09
- **Light sequence**: DBBBDDDBDB (10 sessions)

### Group 4 (HBT 23:00)

- **Participant**: P10
- **Light sequence**: DDDBXDBBBDBDD (12 valid sessions, session 5 excluded due to light automation malfunction)

# Setup

## Load Required Packages

```{r}
#| label: setup-packages

library(tidyverse)
library(readxl)
library(lme4)
library(lmerTest)
library(nlme)
library(car)
library(ggplot2)
library(gridExtra)
library(grid)
library(cowplot)
library(scales)
library(knitr)
library(kableExtra)
library(broom.mixed)
library(purrr)
library(DHARMa)
library(DT)
library(patchwork)
library(ggtext)
library(devtools)
```

## Define Analysis Parameters

```{r}
#| label: setup-parameters

# Participant list (excluding P07, P11)
participant_list <- c("P01", "P02", "P03", "P04", "P05", "P07", "P06", "P09", "P10")

# Light exposure patterns by participant
light_patterns <- list(
  "P01" = c("D", "D", "D", "B", "B", "B", "D", "D", "B", "D", "B"),
  "P02" = c("D", "D", "D", "B", "B", "B", "D", "D", "B", "D", "B"),
  "P03" = c("D", "D", "D", "B", "B", "B", "D", "D", "B", "D", "B"),
  "P04" = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"),
  "P05" = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"),
  "P06" = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"),
  "P07" = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"),
  "P09" = c("D", "B", "B", "B", "D", "D", "D", "B", "D", "B"),
  "P10" = c("D", "D", "D", "B", "X", "D", "B", "B", "B", "D", "B", "D", "D")
)

# Color scheme for visualizations
light_colors <- c("dim" = "#4A90E2", "bright" = "#F5A623")

doublet_colors <- c(
  "DD" = "#4A90E2", "BD" = "#4A90E2",
  "DB" = "#F5A623", "BB" = "#F5A623"
)

triplet_colors <- c(
  "DDD" = "#4A90E2", "BDD" = "#4A90E2", "DBD" = "#4A90E2", "BBD" = "#4A90E2",
  "DDB" = "#F5A623", "BDB" = "#F5A623", "DBB" = "#F5A623", "BBB" = "#F5A623"
)

# Labels for doublets and triplets (using expression for bold formatting)
doublet_labels <- c(
  "DD" = expression("D"*bold("D")),
  "BD" = expression("B"*bold("D")),
  "DB" = expression("D"*bold("B")),
  "BB" = expression("B"*bold("B"))
)

triplet_labels <- c(
  "DDD" = expression("DD"*bold("D")),
  "BDD" = expression("BD"*bold("D")),
  "DBD" = expression("DB"*bold("D")),
  "BBD" = expression("BB"*bold("D")),
  "DDB" = expression("DD"*bold("B")),
  "BDB" = expression("BD"*bold("B")),
  "DBB" = expression("DB"*bold("B")),
  "BBB" = expression("BB"*bold("B"))
)
```

# Data Processing 

## Manual Data Cleaning: In-Session Surveys

This analysis uses pre-processed in-session survey data. The raw data was downloaded from REDCap and underwent manual cleaning before analysis.

### Input and Output Files

- **Input**: Raw in-session survey data from *data/raw/group/group1-4/*
- **Output**: *01_insessionsurvey_cleaned_YYYYMMDDTHHMMSS.xlsx* in *data/derivatives/*

### Manual Processing Steps

#### 1. Column Removal

Unnecessary / doublet columns were removed from the raw data:
- `A` (record_id)
- `B` (redcap_survey_identifier)
- `F` (general_info_complete)
- `G` (session_survey_timestamp)
- `H` (start_time)
- `I` (participant_id2)
- `J` (exp_session2)
- `AI` (language_used_from)
- `AJ` (session_survey_complete)

#### 2. Column Renaming

Columns were renamed for clarity and consistency:

| Original Name | New Name |
|--------------|----------|
| general_info_timestamp | timestamp |
| participant_id | ID |
| exp_session | session |
| kss_response | kss |
| light_rating | light_comfort |
| panas_upset | mood_upset |
| panas_host | mood_hostile |
| panas_alert | mood_alert |
| panas_ashmd | mood_ashamed |
| panas_insprd | mood_inspired |
| panas_nervs | mood_nervous |
| panas_determd | mood_determined |
| panas_attent | mood_attentive |
| panas_afraid | mood_afraid |
| panas_active | mood_active |
| session_therm_1 | thermal_comfort_now |
| session_thermal_2 | thermal_comfort_assess |
| session_thermal_3 | thermal_comfort_preference |

#### 3. Data Standardization

The following standardization procedures were applied:

- Removal of incomplete entries
- Standardization of participant ID format (P01, P02, etc.)
- Replacement of missing values with NA
- Conversion of session numbers (1 → 101: group number + experimental evening)
- Addition of questionnaire number column (ques_numb)
- Addition of light condition column (dim/bright based on experimental protocol)

#### 4. Data Integration

Merging of all group data (groups 1-4)

### Final Dataframe Structure

The cleaned dataset (In_session_Surveys) contains the following columns:

- **A**: timestamp
- **B**: ID (participant identifier)
- **C**: session
- **D**: ques_numb (questionnaire number)
- **E**: kss (Karolinska Sleepiness Scale score)
- **F**: light_comfort
- **G-O**: Symptom ratings (discomfort, headache, blur_vis, double_vis, eye_strain, eye_burn, eye_water, light_sens, nausea)
- **P-X**: PANAS mood ratings (upset, hostile, alert, ashamed, inspired, nervous, determined, attentive, afraid, active)
- **Y**: mood_afraid
- **Z**: thermal_comfort_now
- **AA**: thermal_comfort_assess
- **AB**: thermal_comfort_preference

----

# Data Import

## Load and Standardize Data

```{r}
#| label: data-import

# Load cleaned in-session survey data
insess_surv <- read_excel(
  "C:/Users/anton/OneDrive/AfterGlow/data/derivatives/01_insessionsurvey_cleaned_20250901T174900.xlsx"
)

# Standardize participant IDs, convert data types, and calculate temporal variables
insess_surv <- insess_surv %>%
  mutate(
    # Ensure consistent participant ID formatting
    id = case_when(
      id %in% c("p01", "P01") ~ "P01",
      id %in% c("p02", "P02") ~ "P02", 
      id %in% c("p03", "P03") ~ "P03", 
      id %in% c("p04", "P04") ~ "P04",
      id %in% c("p05", "P05") ~ "P05",
      id %in% c("p06", "P06") ~ "P06",
      id %in% c("p07", "P07") ~ "P07",
      id %in% c("p09", "P09") ~ "P09",
      id %in% c("p10", "P10") ~ "P10",
      TRUE ~ as.character(id)
    ),
    
    # Convert to appropriate data types
    session = as.character(session),
    ques_numb = as.numeric(ques_numb),
    light = tolower(trimws(as.character(light))),
    kss = as.numeric(kss),
    
    # Calculate temporal position relative to habitual bedtime
    hours_before_hbt = case_when(
      ques_numb == 1 ~ 5.5, ques_numb == 2 ~ 5.0, ques_numb == 3 ~ 4.5,
      ques_numb == 4 ~ 4.0, ques_numb == 5 ~ 3.5, ques_numb == 6 ~ 3.0,
      ques_numb == 7 ~ 2.5, ques_numb == 8 ~ 2.0, ques_numb == 9 ~ 1.5,
      ques_numb == 10 ~ 1.0, ques_numb == 11 ~ 0.5
    )
  ) %>%
  
  # Exclude invalid observations
  filter(
    !is.na(id), !is.na(session), !is.na(ques_numb), !is.na(light), 
    !is.na(kss), !is.na(hours_before_hbt), 
    ques_numb <= 11, light %in% c("dim", "bright")
  )

```

# Calculate KSS Metrics

KSS scores were averaged within each evening session to create a single sleepiness index per participant per evening. Questionnaires 7-9 (HBT-2.5h to HBT-1.5h) were selected to isolate the 1.5-hour intervention window where light conditions differed (10 lx vs 1000 lx), excluding pre-intervention baseline (Q1-Q6) and post-intervention recovery periods (Q10-Q11) where both conditions returned to 10 lx.** A sensitivity analysis (at the very end of this analysis) using Q8-Q10 tests robustness to this choice.
```{r}
#| label: kss-metrics

# Mean KSS across all questionnaires (full 5-hour session)
kss_per_evening <- insess_surv %>%
  mutate(
    session_raw = as.integer(str_extract(session, "\\d+$")),
    exp_session = case_when(
      id %in% c("P01", "P02", "P03") ~ session_raw - 100,
      id %in% c("P04", "P05") ~ session_raw - 200,
      id %in% c("P06", "P07", "P09") ~ session_raw - 300,
      id == "P10" ~ session_raw - 400,
      TRUE ~ NA_real_
    )
  ) %>%
  group_by(id, exp_session, session) %>%
  summarise(
    mean_kss = mean(kss, na.rm = TRUE),
    n_measurements = n(),
    .groups = "drop"
  ) %>%
  rename(participant_id = id)

# Mean KSS during intervention window only (Q7-Q9, 1.5 hours)
kss_per_evening_q789 <- insess_surv %>%
  filter(ques_numb %in% c(7, 8, 9)) %>%
  mutate(
    session_raw = as.integer(str_extract(session, "\\d+$")),
    exp_session = case_when(
      id %in% c("P01", "P02", "P03") ~ session_raw - 100,
      id %in% c("P04", "P05") ~ session_raw - 200,
      id %in% c("P06", "P07", "P09") ~ session_raw - 300,
      id == "P10" ~ session_raw - 400,
      TRUE ~ NA_real_
    )
  ) %>%
  group_by(id, exp_session, session) %>%
  summarise(
    mean_kss = mean(kss, na.rm = TRUE),
    n_measurements = n(),
    .groups = "drop"
  ) %>%
  rename(participant_id = id)

# Display descriptive statistics
kss_per_evening %>%
  group_by(participant_id) %>%
  summarise(
    n_sessions = n(),
    M = round(mean(mean_kss), 2),
    SD = round(sd(mean_kss), 2),
    .groups = "drop"
  ) %>%
  print()

kss_per_evening_q789 %>%
  group_by(participant_id) %>%
  summarise(
    n_sessions = n(),
    M = round(mean(mean_kss), 2),
    SD = round(sd(mean_kss), 2),
    .groups = "drop"
  ) %>%
  print()
```
```{r}
#| label: view-kss-data

# Add light condition by extracting character at position exp_session
kss_display <- kss_per_evening %>%
  select(participant_id, exp_session, mean_kss_full = mean_kss) %>%
  left_join(
    kss_per_evening_q789 %>% 
      select(participant_id, exp_session, mean_kss_intervention = mean_kss),
    by = c("participant_id", "exp_session")
  ) %>%
  mutate(
    light_condition = map2_chr(participant_id, exp_session, function(pid, session) {
      pattern_string <- paste(light_patterns[[pid]], collapse = "")
      str_sub(pattern_string, session, session)
    }),
    light_condition = ifelse(light_condition == "X", "excluded", light_condition)
  ) %>%
  arrange(participant_id, exp_session) %>%
  select(
    Participant = participant_id,
    Session = exp_session,
    `Light Condition` = light_condition,
    `KSS (5h window, Q1-11)` = mean_kss_full,
    `KSS (1.5h intervention, Q7-9)` = mean_kss_intervention
  )

# Display table
kss_display
```

# Descriptive Statistics 
```{r}
# Function to extract light condition
get_light_condition <- function(participant_id, exp_session) {
  pattern <- light_patterns[[participant_id]]
  if (exp_session <= length(pattern)) {
    condition <- pattern[exp_session]
    if (condition == "X") return(NA_character_)
    return(condition)
  }
  return(NA_character_)
}

# A) KSS across the 5h window (all questionnaires)
kss_by_light_5h <- kss_per_evening %>%
  filter(participant_id %in% participant_list) %>%
  rowwise() %>%
  mutate(light_condition = get_light_condition(participant_id, exp_session)) %>%
  ungroup() %>%
  filter(!is.na(light_condition)) %>%
  group_by(participant_id, light_condition) %>%
  summarise(
    n_sessions = n(),
    M = round(mean(mean_kss, na.rm = TRUE), 2),
    SD = round(sd(mean_kss, na.rm = TRUE), 2),
    .groups = "drop"
  ) %>%
  arrange(participant_id, light_condition)

print("KSS Scores - 5 Hour Window (Q1-11):")
print(kss_by_light_5h)  

# B) KSS during intervention period (Q7-9)
kss_by_light_intervention <- kss_per_evening_q789 %>%
  filter(participant_id %in% participant_list) %>%
  rowwise() %>%
  mutate(light_condition = get_light_condition(participant_id, exp_session)) %>%
  ungroup() %>%
  filter(!is.na(light_condition)) %>%
  group_by(participant_id, light_condition) %>%
  summarise(
    n_sessions = n(),
    M = round(mean(mean_kss, na.rm = TRUE), 2),
    SD = round(sd(mean_kss, na.rm = TRUE), 2),
    .groups = "drop"
  ) %>%
  arrange(participant_id, light_condition)

print("KSS Scores - Intervention Period (Q7-9):")
print(kss_by_light_intervention)  

# Summary across all participants 
print("\nOverall means across all participants (n=8, excluding P07):")

print("5h Window:")
kss_by_light_5h %>%
  filter(participant_id != "P07") %>%  
  group_by(light_condition) %>%
  summarise(
    n_total = sum(n_sessions),
    M_entire = round(mean(M), 2),
    SD_entire = round(sd(M), 2),
    .groups = "drop"
  ) %>%
  print()

print("Intervention Period:")
kss_by_light_intervention %>%
  filter(participant_id != "P07") %>% 
  group_by(light_condition) %>%
  summarise(
    n_total = sum(n_sessions),
    M_intervention = round(mean(M), 2),
    SD_intervention = round(sd(M), 2),
    .groups = "drop"
  ) %>%
  print()
```

# Visualization

## Grand Average Time Course

### Calculate Grand Average

Following the melatonin analysis approach, grand averages are calculated by first computing per-participant means for each questionnaire and light condition, then averaging across participants to account for between-participant variability in the standard error calculation.
```{r}
#| label: grand-average-calculation

# Step 1: Add exp_session and light pattern to raw data
insess_surv_with_pattern <- insess_surv %>%
  mutate(
    session_raw = as.integer(str_extract(session, "\\d+$")),
    exp_session = case_when(
      id %in% c("P01", "P02", "P03") ~ session_raw - 100,
      id %in% c("P04", "P05") ~ session_raw - 200,
      id %in% c("P06", "P07", "P09") ~ session_raw - 300,
      id == "P10" ~ session_raw - 400,
      TRUE ~ NA_real_
    ),
    light_pattern = map2_chr(id, exp_session, function(pid, sess) {
      if (is.na(sess) || sess < 1) return(NA_character_)
      pattern <- light_patterns[[pid]]
      if (sess > length(pattern)) return(NA_character_)
      pattern[sess]
    })
  ) %>%
  filter(!is.na(light_pattern), light_pattern != "X")

# Step 2: Calculate per-participant means by questionnaire and light
kss_summary_per_participant <- insess_surv_with_pattern %>%
  group_by(id, ques_numb, light_pattern) %>%
  summarise(
    mean_kss = mean(kss, na.rm = TRUE),
    sd_kss = sd(kss, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  rename(participant_id = id) %>%
  mutate(light_pattern = factor(light_pattern, levels = c("D", "B"), 
                                labels = c("DIM", "BRIGHT")))

# Step 3: Exclude P07
kss_summary_noP07 <- kss_summary_per_participant %>%
  filter(participant_id != "P07")

# Step 4: Grand average across participants (manual calculation for clarity)
grand_summary_all <- data.frame()

for(q in unique(kss_summary_noP07$ques_numb)) {
  for(l in levels(kss_summary_noP07$light_pattern)) {
    # Extract all participant means for this questionnaire and light
    vals <- kss_summary_noP07 %>%
      filter(ques_numb == q, light_pattern == l) %>%
      pull(mean_kss)
    
    n_participants <- length(vals)
    mean_val <- mean(vals, na.rm = TRUE)
    sd_val <- sd(vals, na.rm = TRUE)
    se_val <- sd_val / sqrt(n_participants)
    
    # Append to grand summary
    grand_summary_all <- rbind(
      grand_summary_all,
      data.frame(
        ques_numb = q,
        light_type = l,
        mean_kss = mean_val,
        sd_kss = sd_val,
        n_participants = n_participants,
        se_kss = se_val
      )
    )
  }
}

# Ensure proper factor ordering
grand_summary_all$light_type <- factor(grand_summary_all$light_type, 
                                       levels = c("BRIGHT", "DIM"))

# Display summary
cat("Grand average calculated across", length(unique(kss_summary_noP07$participant_id)), 
    "participants (P07 excluded)\n\n")
print(grand_summary_all)
```

### Plot Grand Average
```{r}
#| label: fig-grand-average
#| fig-cap: "Grand average KSS time course across all participants (n=8, P07 excluded). Orange line shows bright evenings (1000 lx during intervention), blue line shows dim evenings (10 lx). Shaded regions indicate ±1 SE."
#| fig-width: 10
#| fig-height: 8

# Prepare axis labels
hbt_values <- c("5.5h", "5h", "4.5h", "4h", "3.5h", "3h", 
                "2.5h", "2h", "1.5h", "1h", "0.5h")

# Calculate y-axis limits
y_vals <- grand_summary_all$mean_kss[is.finite(grand_summary_all$mean_kss)]
se_vals <- grand_summary_all$se_kss[is.finite(grand_summary_all$se_kss)]

if (length(y_vals) > 0 && length(se_vals) > 0) {
  y_range <- range(c(y_vals - se_vals, y_vals + se_vals), na.rm = TRUE)
  y_min <- floor(y_range[1] * 2) / 2
  y_max <- ceiling(y_range[2] * 2) / 2 + 0.3
} else {
  y_min <- 1
  y_max <- 9
}

if (!is.finite(y_min)) y_min <- 1
if (!is.finite(y_max)) y_max <- 9

# Create plot
ggplot(grand_summary_all, 
       aes(x = ques_numb, y = mean_kss, 
           color = light_type, fill = light_type, group = light_type)) +
  
  # Background shading: Dim phases
  annotate("rect", xmin = 1, xmax = 7, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  annotate("rect", xmin = 10, xmax = 11, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  
  # Background shading: Intervention phase
  annotate("rect", xmin = 7, xmax = 10, ymin = -Inf, ymax = Inf, 
           fill = "#FFF9E6", alpha = 0.8) +
  
  # Phase labels
  annotate("text", x = 3.75, y = y_max - 0.05, 
           label = "dim\n(10 lx)", hjust = 0.5, vjust = 1, 
           size = 3.5, color = "grey50", fontface = "italic") +
  annotate("text", x = 8.5, y = y_max - 0.05, 
           label = "intervention\n(10 or 1000 lx)", hjust = 0.5, vjust = 1, 
           size = 3.5, color = "#B8860B", fontface = "italic") +
  annotate("text", x = 10.5, y = y_max - 0.05, 
           label = "dim\n(10 lx)", hjust = 0.5, vjust = 1, 
           size = 3.5, color = "grey50", fontface = "italic") +
  
  # Confidence ribbon
  geom_ribbon(aes(ymin = mean_kss - se_kss, ymax = mean_kss + se_kss), 
              alpha = 0.1, color = NA, show.legend = FALSE) +
  
  # Line and points
  geom_line(linewidth = 0.8) +
  geom_point(size = 3.5, shape = 16) +
  
  # Color scale
  scale_color_manual(
    values = c("BRIGHT" = "#F5A623", "DIM" = "#4A90E2"),
    labels = c("BRIGHT" = "Bright (1000 lx)", "DIM" = "Dim (10 lx)")
  ) +
  scale_fill_manual(
    values = c("BRIGHT" = "#F5A623", "DIM" = "#4A90E2"),
    labels = c("BRIGHT" = "Bright (1000 lx)", "DIM" = "Dim (10 lx)")
  ) +
  
  # Axes
  scale_x_continuous(
    breaks = 1:11,
    labels = paste0("Q", 1:11, "\n(", hbt_values, ")"),
    limits = c(0.5, 11.5),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    limits = c(y_min, y_max),
    breaks = seq(y_min, y_max, 0.5),
    expand = c(0, 0)
  ) +
  
  # Labels
  labs(
    title = "Karolinska sleepiness scale (KSS) time course",
    subtitle = "Grand average across all participants (P07 excluded, n=8)",
    x = "KSS question number (hours before HBT)",
    y = expression(atop("Mean KSS ± SE", 
                        scriptstyle("(1 = Extremely alert; 9 = Very sleepy)"))),
    color = "Light condition",
    fill = "Light condition"
  ) +
  
  # Theme
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0, face = "plain"),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "right",
    plot.margin = margin(15, 10, 10, 10)
  )
```
# Individual KSS trajectories acoross all experimental sessions

```{r}
#| label: fig-kss-individual-profiles
#| fig-cap: "Karolinska Sleepiness Scale (KSS) per evening - all participants"
#| fig-width: 12
#| fig-height: 9

# Function to create KSS profile plot for one participant
create_kss_profile <- function(participant_id, light_pattern) {
  
  # Prepare data for this participant
  p_data <- insess_surv_with_pattern %>%
    filter(id == !!participant_id) %>%
    filter(light_pattern != "X") %>%
    mutate(
      light_condition = light_pattern,
      exp_session = as.integer(exp_session)
    ) %>%
    arrange(exp_session, ques_numb)
  
  # Get mean KSS per evening (for labeling)
  p_mean <- kss_per_evening_q789 %>%
    filter(participant_id == !!participant_id) %>%
    mutate(
      exp_session = as.integer(exp_session),
      light_condition = map2_chr(participant_id, exp_session, function(pid, sess) {
        pattern <- light_patterns[[pid]]
        if (sess > length(pattern)) return(NA_character_)
        pattern[sess]
      }),
      mean_label = paste0("M=", round(mean_kss, 1)),
      
      label_x = case_when(
        (participant_id == "P01" & exp_session == 5) ~ 11,
        (participant_id == "P05" & exp_session == 2) ~ 11,
        (participant_id == "P09" & exp_session == 9) ~ 11,
        (participant_id == "P10" & exp_session == 10) ~ 11,
        (participant_id == "P10" & exp_session == 12) ~ 11,
        (participant_id == "P04" & exp_session == 9) ~ 11,
        (participant_id == "P10" & exp_session == 1) ~ 11,
        TRUE ~ 1
      ),
      
      label_y = case_when(
        (participant_id == "P04" & exp_session == 9) ~ -Inf,
        (participant_id == "P10" & exp_session == 1) ~ -Inf,
        TRUE ~ Inf
      ),
      
      label_hjust = case_when(
        (participant_id == "P01" & exp_session == 5) ~ 1,
        (participant_id == "P04" & exp_session == 9) ~ 1,
        (participant_id == "P05" & exp_session == 2) ~ 1,
        (participant_id == "P09" & exp_session == 9) ~ 1,
        (participant_id == "P10" & exp_session == 1) ~ 1,
        (participant_id == "P10" & exp_session == 10) ~ 1,
        (participant_id == "P10" & exp_session == 12) ~ 1,
        TRUE ~ 0
      ),
      
      label_vjust = case_when(
        (participant_id == "P04" & exp_session == 9) ~ -0.3,
        (participant_id == "P10" & exp_session == 1) ~ -0.3,
        TRUE ~ 1.3
      )
    ) %>%
    filter(!is.na(light_condition), light_condition != "X")
  
  # Determine y-axis limit
  max_y <- max(p_data$kss, na.rm = TRUE)
  
  if (!is.finite(max_y) || is.na(max_y)) {
    max_y <- 9
  }
  
  y_limit <- min(9, ceiling(max_y))
  
  if (!is.finite(y_limit) || is.na(y_limit) || y_limit < 1) {
    y_limit <- 9
  }
  
  # Create plot
  p <- ggplot(p_data, aes(x = ques_numb, y = kss)) +
    
    # Only yellow background for B evenings, no background for D evenings
    geom_rect(data = p_mean %>% filter(light_condition == "B"),
              aes(xmin = 7, xmax = 10, ymin = -Inf, ymax = Inf),
              fill = "#FFF9E6", alpha = 0.8, inherit.aes = FALSE) +
    
    # Area plot
    geom_area(fill = "#4A90E2", alpha = 0.3) +
    
    # Line and points
    geom_line(color = "black", linewidth = 0.8) +
    geom_point(color = "black", size = 2.5, shape = 16) +
    
    # Mean KSS label
    geom_text(data = p_mean,
              aes(x = label_x, y = label_y, label = mean_label,
                  hjust = label_hjust, vjust = label_vjust),
              size = 3.5,
              inherit.aes = FALSE) +
    
    # Scales
    scale_x_continuous(
      breaks = c(1, 3, 5, 7, 9, 11),
      labels = c("Q1\n(5.5h)", "Q3\n(4.5h)", "Q5\n(3.5h)",
                 "Q7\n(2.5h)", "Q9\n(1.5h)", "Q11\n(0.5h)"),
      limits = c(0.5, 11.5),
      expand = c(0, 0)
    ) +
    scale_y_continuous(
      limits = c(1, y_limit),
      breaks = seq(1, y_limit, by = 1),
      expand = expansion(mult = c(0.05, 0.05))
    ) +
    
    # Facet by evening
    facet_wrap(~ exp_session, ncol = 4,
               labeller = labeller(exp_session = function(x) {
                 idx <- as.numeric(x)
                 light <- light_pattern[idx]
                 paste0("Evening ", x, " (", light, ")")
               })) +
    
    # Labels
    labs(
      title = paste0("Karolinska Sleepiness Scale (KSS) per evening (", participant_id, ")"),
      x = "Questionnaire number (hours before HBT)",
      y = "KSS score\n(1 = Extremely alert; 9 = Very sleepy)"
    ) +
    
    # Theme
    theme_classic(base_size = 11) +
    theme(
      aspect.ratio = 1,
      plot.title = element_text(size = 15, hjust = 0, face = "plain", margin = margin(b = 10)),
      axis.text = element_text(color = "black", size = 10),
      axis.text.x = element_text(size = 7),
      axis.title = element_text(size = 11),
      axis.title.y = element_text(margin = margin(r = 8)),
      axis.line = element_line(linewidth = 0.5),
      axis.ticks = element_line(linewidth = 0.5),
      panel.grid = element_blank(),
      strip.background = element_rect(fill = "white", color = "grey80", linewidth = 0.5),
      strip.text = element_text(size = 10, color = "black"),
      panel.background = element_rect(fill = "white", color = "grey80", linewidth = 0.5),
      plot.margin = margin(10, 10, 10, 10),
      legend.position = "none"
    )
  
  return(p)
}

# Group 1 (HBT 00:00)
p01_kss <- create_kss_profile("P01", light_patterns[["P01"]])
p02_kss <- create_kss_profile("P02", light_patterns[["P02"]])
p03_kss <- create_kss_profile("P03", light_patterns[["P03"]])

# Group 2 (HBT 22:00)
p04_kss <- create_kss_profile("P04", light_patterns[["P04"]])
p05_kss <- create_kss_profile("P05", light_patterns[["P05"]])

# Group 3 (HBT 23:30)
p06_kss <- create_kss_profile("P06", light_patterns[["P06"]])
p07_kss <- create_kss_profile("P07", light_patterns[["P07"]])
p09_kss <- create_kss_profile("P09", light_patterns[["P09"]])

# Group 4 (HBT 23:00)
p10_kss <- create_kss_profile("P10", light_patterns[["P10"]])

print(p01_kss)
print(p02_kss)
print(p03_kss)
print(p04_kss)
print(p05_kss)
print(p06_kss)
print(p07_kss)
print(p09_kss)
print(p10_kss)
```

# Sequence Analysis

## Define Sequential Patterns

```{r}
#| label: define-sequences
#| code-fold: true

# Group 1 (P01-P03): DDDBBBDDBDB
doublet_positions_g1 <- tibble(
  doublet = c("DD", "DB", "BD", "BB"),
  position1_sessions = list(c(1, 2, 7), c(3, 8, 10), c(6, 9), c(4, 5)),
  position2_sessions = list(c(2, 3, 8), c(4, 9, 11), c(7, 10), c(5, 6))
)

triplet_positions_g1 <- tibble(
  triplet = c("DDD", "DDB", "DBD", "DBB", "BDD", "BDB", "BBD", "BBB"),
  position1_sessions = list(c(1), c(2, 7), c(8), c(3), c(6), c(9), c(5), c(4)),
  position2_sessions = list(c(2), c(3, 8), c(9), c(4), c(7), c(10), c(6), c(5)),
  position3_sessions = list(c(3), c(4, 9), c(10), c(5), c(8), c(11), c(7), c(6))
)

# Groups 2 & 3 (P04-P06, P09): DBBBDDDBDB
doublet_positions_g2_g3 <- tibble(
  doublet = c("DD", "DB", "BD", "BB"),
  position1_sessions = list(c(5, 6), c(1, 7, 9), c(4, 8), c(2, 3)),
  position2_sessions = list(c(6, 7), c(2, 8, 10), c(5, 9), c(3, 4))
)

triplet_positions_g2_g3 <- tibble(
  triplet = c("DDD", "DDB", "DBD", "DBB", "BDD", "BDB", "BBD", "BBB"),
  position1_sessions = list(c(5), c(6), c(7), c(1), c(4), c(8), c(3), c(2)),
  position2_sessions = list(c(6), c(7), c(8), c(2), c(5), c(9), c(4), c(3)),
  position3_sessions = list(c(7), c(8), c(9), c(3), c(6), c(10), c(5), c(4))
)

# Group 4 (P10): DDDBDBBBDBDD (session 5 excluded)
doublet_positions_g4 <- tibble(
  doublet = c("DD", "DB", "BD", "BB"),
  position1_sessions = list(c(1, 2, 12), c(3, 6, 10), c(9, 11), c(7, 8)),
  position2_sessions = list(c(2, 3, 13), c(4, 7, 11), c(10, 12), c(8, 9))
)

triplet_positions_g4 <- tibble(
  triplet = c("DDD", "DDB", "DBD", "DBB", "BDD", "BDB", "BBD", "BBB"),
  position1_sessions = list(c(1), c(2), c(10), c(6), c(11), c(9), c(8), c(7)),
  position2_sessions = list(c(2), c(3), c(11), c(7), c(12), c(10), c(9), c(8)),
  position3_sessions = list(c(3), c(4), c(12), c(8), c(13), c(11), c(10), c(9))
)
```

## Calculate Sequence Metrics

Mean KSS scores were calculated for each doublet and triplet position across all occurrences of that sequence for each participant.

```{r}
#| label: sequence-calculation
#| code-fold: true

# Function to calculate doublet positions
calc_doublet_positions_kss <- function(participant_id, kss_data, doublet_map) {
  doublet_map %>%
    rowwise() %>%
    mutate(
      pos1_kss = mean(kss_data$mean_kss[kss_data$exp_session %in% position1_sessions], 
                      na.rm = TRUE),
      pos2_kss = mean(kss_data$mean_kss[kss_data$exp_session %in% position2_sessions], 
                      na.rm = TRUE),
      participant = participant_id
    ) %>%
    ungroup() %>%
    select(participant, doublet, pos1_kss, pos2_kss)
}

# Function to calculate triplet positions
calc_triplet_positions_kss <- function(participant_id, kss_data, triplet_map) {
  triplet_map %>%
    rowwise() %>%
    mutate(
      pos1_kss = if(length(position1_sessions) > 0) {
        mean(kss_data$mean_kss[kss_data$exp_session %in% position1_sessions], na.rm = TRUE)
      } else NA_real_,
      pos2_kss = if(length(position2_sessions) > 0) {
        mean(kss_data$mean_kss[kss_data$exp_session %in% position2_sessions], na.rm = TRUE)
      } else NA_real_,
      pos3_kss = if(length(position3_sessions) > 0) {
        mean(kss_data$mean_kss[kss_data$exp_session %in% position3_sessions], na.rm = TRUE)
      } else NA_real_,
      participant = participant_id
    ) %>%
    ungroup() %>%
    select(participant, triplet, pos1_kss, pos2_kss, pos3_kss)
}

# Extract KSS data per participant (intervention period Q7-Q9 only)
kss_list <- list(
  P01 = kss_per_evening_q789 %>% filter(participant_id == "P01"),
  P02 = kss_per_evening_q789 %>% filter(participant_id == "P02"),
  P03 = kss_per_evening_q789 %>% filter(participant_id == "P03"),
  P04 = kss_per_evening_q789 %>% filter(participant_id == "P04"),
  P05 = kss_per_evening_q789 %>% filter(participant_id == "P05"),
  P06 = kss_per_evening_q789 %>% filter(participant_id == "P06"),
  P09 = kss_per_evening_q789 %>% filter(participant_id == "P09"),
  P10 = kss_per_evening_q789 %>% filter(participant_id == "P10")
)

# Calculate sequences for each group
doublets_g1 <- bind_rows(
  calc_doublet_positions_kss("P01", kss_list$P01, doublet_positions_g1),
  calc_doublet_positions_kss("P02", kss_list$P02, doublet_positions_g1),
  calc_doublet_positions_kss("P03", kss_list$P03, doublet_positions_g1)
)

doublets_g2_g3 <- bind_rows(
  calc_doublet_positions_kss("P04", kss_list$P04, doublet_positions_g2_g3),
  calc_doublet_positions_kss("P05", kss_list$P05, doublet_positions_g2_g3),
  calc_doublet_positions_kss("P06", kss_list$P06, doublet_positions_g2_g3),
  calc_doublet_positions_kss("P09", kss_list$P09, doublet_positions_g2_g3)
)

doublets_g4 <- calc_doublet_positions_kss("P10", kss_list$P10, doublet_positions_g4)

triplets_g1 <- bind_rows(
  calc_triplet_positions_kss("P01", kss_list$P01, triplet_positions_g1),
  calc_triplet_positions_kss("P02", kss_list$P02, triplet_positions_g1),
  calc_triplet_positions_kss("P03", kss_list$P03, triplet_positions_g1)
)

triplets_g2_g3 <- bind_rows(
  calc_triplet_positions_kss("P04", kss_list$P04, triplet_positions_g2_g3),
  calc_triplet_positions_kss("P05", kss_list$P05, triplet_positions_g2_g3),
  calc_triplet_positions_kss("P06", kss_list$P06, triplet_positions_g2_g3),
  calc_triplet_positions_kss("P09", kss_list$P09, triplet_positions_g2_g3)
)

triplets_g4 <- calc_triplet_positions_kss("P10", kss_list$P10, triplet_positions_g4)

# Combine all participants
pos_kss_all_q789 <- bind_rows(doublets_g1, doublets_g2_g3, doublets_g4)
triplet_kss_all_q789 <- bind_rows(triplets_g1, triplets_g2_g3, triplets_g4)
```

```{r}
#| label: sequence-calculation-5h
#| code-fold: true

# Calculate doublets and triplets for FULL 5-hour window (using kss_per_evening)

# Extract KSS data per participant (full 5h window)
kss_list_5h <- list(
  P01 = kss_per_evening %>% filter(participant_id == "P01"),
  P02 = kss_per_evening %>% filter(participant_id == "P02"),
  P03 = kss_per_evening %>% filter(participant_id == "P03"),
  P04 = kss_per_evening %>% filter(participant_id == "P04"),
  P05 = kss_per_evening %>% filter(participant_id == "P05"),
  P06 = kss_per_evening %>% filter(participant_id == "P06"),
  P09 = kss_per_evening %>% filter(participant_id == "P09"),
  P10 = kss_per_evening %>% filter(participant_id == "P10")
)

# Calculate doublets for each group (5h window)
doublets_g1_5h <- bind_rows(
  calc_doublet_positions_kss("P01", kss_list_5h$P01, doublet_positions_g1),
  calc_doublet_positions_kss("P02", kss_list_5h$P02, doublet_positions_g1),
  calc_doublet_positions_kss("P03", kss_list_5h$P03, doublet_positions_g1)
)

doublets_g2_g3_5h <- bind_rows(
  calc_doublet_positions_kss("P04", kss_list_5h$P04, doublet_positions_g2_g3),
  calc_doublet_positions_kss("P05", kss_list_5h$P05, doublet_positions_g2_g3),
  calc_doublet_positions_kss("P06", kss_list_5h$P06, doublet_positions_g2_g3),
  calc_doublet_positions_kss("P09", kss_list_5h$P09, doublet_positions_g2_g3)
)

doublets_g4_5h <- calc_doublet_positions_kss("P10", kss_list_5h$P10, doublet_positions_g4)

# Calculate triplets for each group (5h window)
triplets_g1_5h <- bind_rows(
  calc_triplet_positions_kss("P01", kss_list_5h$P01, triplet_positions_g1),
  calc_triplet_positions_kss("P02", kss_list_5h$P02, triplet_positions_g1),
  calc_triplet_positions_kss("P03", kss_list_5h$P03, triplet_positions_g1)
)

triplets_g2_g3_5h <- bind_rows(
  calc_triplet_positions_kss("P04", kss_list_5h$P04, triplet_positions_g2_g3),
  calc_triplet_positions_kss("P05", kss_list_5h$P05, triplet_positions_g2_g3),
  calc_triplet_positions_kss("P06", kss_list_5h$P06, triplet_positions_g2_g3),
  calc_triplet_positions_kss("P09", kss_list_5h$P09, triplet_positions_g2_g3)
)

triplets_g4_5h <- calc_triplet_positions_kss("P10", kss_list_5h$P10, triplet_positions_g4)

# Combine all participants (5h window)
pos_kss_all <- bind_rows(doublets_g1_5h, doublets_g2_g3_5h, doublets_g4_5h)
triplet_kss_all <- bind_rows(triplets_g1_5h, triplets_g2_g3_5h, triplets_g4_5h)
```

## Sequence Analysis Plots

### Doublet Analysis

#### Full Session (5-hour exposure)

**KSS levels under light exposure doublets**
```{r}
#| label: fig-doublet-box-5h
#| fig-cap: "Mean KSS per doublet type (5-h exposure, P07 excluded)"
#| fig-width: 10
#| fig-height: 8

df_doublet_box_5h <- pos_kss_all %>%
  filter(participant != "P07") %>%
  select(participant, doublet, mean_kss = pos2_kss) %>%
  mutate(doublet = factor(doublet, levels = c("DD", "BD", "DB", "BB")))

y_min <- floor(min(df_doublet_box_5h$mean_kss[is.finite(df_doublet_box_5h$mean_kss)]) * 2) / 2
y_max <- ceiling(max(df_doublet_box_5h$mean_kss[is.finite(df_doublet_box_5h$mean_kss)]) * 2) / 2

ggplot(df_doublet_box_5h, aes(x = doublet, y = mean_kss)) +
  geom_boxplot(aes(fill = doublet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15, seed = 42),
             size = 2.5, alpha = 0.7, shape = 16) +
  scale_fill_manual(values = doublet_colors) +
  scale_x_discrete(labels = doublet_labels) +
  scale_y_continuous(limits = c(y_min, y_max),
                     breaks = seq(1, 9, 0.5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  labs(
    title = "Karolinska Sleepiness Scale under light exposure doublets",
    subtitle = "Mean second-day KSS scores per doublet type (5-h exposure; P07 excluded)",
    x = "Doublet",
    y = expression(atop("Mean KSS",
                        scriptstyle("(1 = Extremely alert; 9 = Very sleepy)")))
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )
```

##### Data Overview: Doublet Observations
```{r}
#| label: doublet-overview-5h

# Create detailed overview table
doublet_overview_5h <- pos_kss_all %>%
  filter(participant != "P07") %>%
  select(participant, doublet, pos1_kss, pos2_kss) %>%
  arrange(participant, doublet)

doublet_overview_5h
```

#### Intervention Window (1.5-hour exposure)

**KSS levels under light exposure doublets**
```{r}
#| label: fig-doublet-box-intervention
#| fig-cap: "Mean KSS per doublet type (1.5-h intervention window, P07 excluded)"
#| fig-width: 10
#| fig-height: 8
df_doublet_box_q789 <- pos_kss_all_q789 %>%
  filter(participant != "P07") %>%
  select(participant, doublet, mean_kss = pos2_kss) %>%
  mutate(doublet = factor(doublet, levels = c("DD", "BD", "DB", "BB")))

y_min <- floor(min(df_doublet_box_q789$mean_kss[is.finite(df_doublet_box_q789$mean_kss)]) * 2) / 2
y_max <- ceiling(max(df_doublet_box_q789$mean_kss[is.finite(df_doublet_box_q789$mean_kss)]) * 2) / 2

ggplot(df_doublet_box_q789, aes(x = doublet, y = mean_kss)) +
  geom_boxplot(aes(fill = doublet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15, seed = 42),
             size = 2.5, alpha = 0.7, shape = 16) +
  scale_fill_manual(values = doublet_colors) +
  scale_x_discrete(labels = doublet_labels) +
  scale_y_continuous(limits = c(y_min, y_max),
                     breaks = seq(1, 9, 0.5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  labs(
    title = "Karolinska Sleepiness Scale under light exposure doublets",
    subtitle = "Mean second-day KSS scores per doublet type (1.5-h intervention; P07 excluded)",
    x = "Doublet",
    y = expression(atop("Mean KSS",
                        scriptstyle("(1 = Extremely alert; 9 = Very sleepy)")))
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )
```

##### Data Overview: Doublet Observations
```{r}
#| label: doublet-overview-intervention

# Create detailed overview table
doublet_overview_q789 <- pos_kss_all_q789 %>%
  filter(participant != "P07") %>%
  select(participant, doublet, pos1_kss, pos2_kss) %>%
  arrange(participant, doublet)

doublet_overview_q789
```

### Doublet Position Trajectory

#### Full Session (5-hour exposure)

**Mean KSS trajectory across doublet positions**
```{r}
#| label: fig-doublet-position-5h
#| fig-cap: "Mean KSS trajectory across doublet positions (5-h exposure, P07 excluded)"
#| fig-width: 8
#| fig-height: 8

# Prepare position data using full 5h window
mean_doublet_pos_5h <- pos_kss_all %>%
  filter(participant != "P07") %>%
  group_by(doublet) %>%
  summarise(
    pos1_mean = mean(pos1_kss, na.rm = TRUE),
    pos2_mean = mean(pos2_kss, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(pos1_mean, pos2_mean), 
               names_to = "position", values_to = "mean_kss") %>%
  mutate(
    position_num = ifelse(position == "pos1_mean", 1, 2),
    light_type = ifelse(position == "pos1_mean", 
                       substr(doublet, 1, 1), substr(doublet, 2, 2)),
    light_label = ifelse(light_type == "D", "dim", "bright"),
    doublet = factor(doublet, levels = c("DD", "BD", "DB", "BB"))
  )

# Calculate y-axis limits
y_min <- floor(min(mean_doublet_pos_5h$mean_kss[is.finite(mean_doublet_pos_5h$mean_kss)]) * 2) / 2
y_max <- ceiling(max(mean_doublet_pos_5h$mean_kss[is.finite(mean_doublet_pos_5h$mean_kss)]) * 2) / 2

# Create plot
ggplot(mean_doublet_pos_5h, aes(x = position_num, y = mean_kss, group = doublet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_label), size = 3.5, shape = 16) +
  scale_color_manual(values = light_colors,
                     labels = c("dim" = "Dim (10 lx)", 
                                "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2), 
                     labels = c("Position 1", "Position 2"), 
                     limits = c(0.8, 2.2)) +
  scale_y_continuous(limits = c(y_min, y_max),
                     breaks = seq(1, 9, 0.5),
                     expand = expansion(mult = c(0.05, 0.05))) +
  labs(
    title = "Karolinska Sleepiness Scale across light exposure doublets",
    subtitle = "Mean KSS scores across all participants (5-h exposure; P07 excluded)",
    x = "Doublet position", 
    y = expression(atop("Mean KSS", 
                        scriptstyle("(1 = Extremely alert; 9 = Very sleepy)"))),
    color = "Light condition"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,

    # Title and subtitle size adjustments
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),

    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),

    legend.position = "right",
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 12),

    plot.margin = margin(15, 10, 10, 10)
  )
```

#### Intervention Window (1.5-hour exposure)

**Mean KSS trajectory across doublet positions**
```{r}
#| label: fig-doublet-position-intervention
#| fig-cap: "Mean KSS trajectory across doublet positions (1.5-h intervention window, P07 excluded)"
#| fig-width: 8
#| fig-height: 8

# Prepare position data using intervention window
mean_doublet_pos_q789 <- pos_kss_all_q789 %>%
  filter(participant != "P07") %>%
  group_by(doublet) %>%
  summarise(
    pos1_mean = mean(pos1_kss, na.rm = TRUE),
    pos2_mean = mean(pos2_kss, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(pos1_mean, pos2_mean), 
               names_to = "position", values_to = "mean_kss") %>%
  mutate(
    position_num = ifelse(position == "pos1_mean", 1, 2),
    light_type = ifelse(position == "pos1_mean", 
                       substr(doublet, 1, 1), substr(doublet, 2, 2)),
    light_label = ifelse(light_type == "D", "dim", "bright"),
    doublet = factor(doublet, levels = c("DD", "BD", "DB", "BB"))
  )

# Calculate y-axis limits
y_min <- floor(min(mean_doublet_pos_q789$mean_kss[is.finite(mean_doublet_pos_q789$mean_kss)]) * 2) / 2
y_max <- ceiling(max(mean_doublet_pos_q789$mean_kss[is.finite(mean_doublet_pos_q789$mean_kss)]) * 2) / 2

# Create plot
ggplot(mean_doublet_pos_q789, aes(x = position_num, y = mean_kss, group = doublet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_label), size = 3.5, shape = 16) +
  scale_color_manual(values = light_colors,
                     labels = c("dim" = "Dim (10 lx)", 
                                "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2), 
                     labels = c("Position 1", "Position 2"), 
                     limits = c(0.8, 2.2)) +
  scale_y_continuous(limits = c(y_min, y_max),
                     breaks = seq(1, 9, 0.5),
                     expand = expansion(mult = c(0.05, 0.05))) +
  labs(
    title = "Karolinska Sleepiness Scale across light exposure doublets",
    subtitle = "Mean KSS scores across all participants (1.5-h exposure; P07 excluded)",
    x = "Doublet position", 
    y = expression(atop("Mean KSS", 
                        scriptstyle("(1 = Extremely alert; 9 = Very sleepy)"))),
    color = "Light condition"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,

    # Title & subtitle sizes added
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),

    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),

    legend.position = "right",
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 12),

    plot.margin = margin(15, 10, 10, 10)
  )
```

### Doublet Intervention Window Summary Statistics (Boxplots)
```{r}
# Calculate statistics for 5-h exposure (full session)
doublet_stats_5h <- pos_kss_all %>%
  filter(participant != "P07") %>%
  select(doublet, pos2_kss) %>%
  group_by(doublet) %>%
  summarise(
    n_5h = n(),
    median_5h = median(pos2_kss, na.rm = TRUE),
    IQR_5h = IQR(pos2_kss, na.rm = TRUE),
    .groups = "drop"
  )

# Calculate statistics for 1.5-h exposure (intervention window)
doublet_stats_1.5h <- pos_kss_all_q789 %>%
  filter(participant != "P07") %>%
  select(doublet, pos2_kss) %>%
  group_by(doublet) %>%
  summarise(
    n_1.5h = n(),
    median_1.5h = median(pos2_kss, na.rm = TRUE),
    IQR_1.5h = IQR(pos2_kss, na.rm = TRUE),
    .groups = "drop"
  )

# Combine both exposure windows
doublet_stats_combined <- doublet_stats_5h %>%
  full_join(doublet_stats_1.5h, by = "doublet") %>%
  mutate(doublet = factor(doublet, levels = c("DD", "BD", "DB", "BB"))) %>%
  arrange(doublet)

# Display as formatted table
knitr::kable(
  doublet_stats_combined,
  digits = 2,
  col.names = c("Doublet", "N", "Median", "IQR", "N", "Median", "IQR"),
  caption = "Descriptive statistics for doublet second-day KSS scores across exposure windows (P07 excluded)",
  align = c("l", rep("r", 6))
) %>%
  kableExtra::add_header_above(c(" " = 1, "5-h Exposure" = 3, "1.5-h Exposure" = 3))
```

### Triplet Analysis

#### Full Session (5-hour exposure)

**KSS levels under light exposure triplets**
```{r}
#| label: fig-triplet-box-5h
#| fig-cap: "Mean KSS per triplet type (5-h exposure, P07 excluded)"
#| fig-width: 10
#| fig-height: 8
df_triplet_box_5h <- triplet_kss_all %>%
  filter(participant != "P07", !is.na(pos3_kss)) %>%
  select(participant, triplet, mean_kss = pos3_kss) %>%
  mutate(triplet = factor(triplet, levels = c("DDD", "BDD", "DBD", "BBD",
                                              "DDB", "BDB", "DBB", "BBB")))

y_min <- floor(min(df_triplet_box_5h$mean_kss[is.finite(df_triplet_box_5h$mean_kss)]) * 2) / 2
y_max <- ceiling(max(df_triplet_box_5h$mean_kss[is.finite(df_triplet_box_5h$mean_kss)]) * 2) / 2

ggplot(df_triplet_box_5h, aes(x = triplet, y = mean_kss)) +
  geom_boxplot(aes(fill = triplet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15, seed = 42),
             size = 2.5, alpha = 0.7, shape = 16) +
  scale_fill_manual(values = triplet_colors) +
  scale_x_discrete(labels = triplet_labels) +
  scale_y_continuous(limits = c(y_min, y_max),
                     breaks = seq(1, 9, 0.5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  labs(
    title = "Karolinska Sleepiness Scale under light exposure triplets",
    subtitle = "Mean third-day KSS scores per triplet type (5-h exposure; P07 excluded)",
    x = "Triplet",
    y = expression(atop("Mean KSS",
                        scriptstyle("(1 = Extremely alert; 9 = Very sleepy)")))
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )
```

##### Data Overview: Triplet Observations
```{r}
#| label: triplet-overview-5h

# Create detailed overview table
triplet_overview_5h <- triplet_kss_all %>%
  filter(participant != "P07", !is.na(pos3_kss)) %>%
  select(participant, triplet, pos1_kss, pos2_kss, pos3_kss) %>%
  arrange(participant, triplet)

triplet_overview_5h
```

#### Intervention Window (1.5-hour exposure)

**KSS levels under light exposure triplets**
```{r}
#| label: fig-triplet-box-intervention
#| fig-cap: "Mean KSS per triplet type (1.5-h intervention window, P07 excluded)"
#| fig-width: 10
#| fig-height: 8
df_triplet_box_q789 <- triplet_kss_all_q789 %>%
  filter(participant != "P07", !is.na(pos3_kss)) %>%
  select(participant, triplet, mean_kss = pos3_kss) %>%
  mutate(triplet = factor(triplet, levels = c("DDD", "BDD", "DBD", "BBD",
                                              "DDB", "BDB", "DBB", "BBB")))

y_min <- floor(min(df_triplet_box_q789$mean_kss[is.finite(df_triplet_box_q789$mean_kss)]) * 2) / 2
y_max <- ceiling(max(df_triplet_box_q789$mean_kss[is.finite(df_triplet_box_q789$mean_kss)]) * 2) / 2

ggplot(df_triplet_box_q789, aes(x = triplet, y = mean_kss)) +
  geom_boxplot(aes(fill = triplet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15, seed = 42),
             size = 2.5, alpha = 0.7, shape = 16) +
  scale_fill_manual(values = triplet_colors) +
  scale_x_discrete(labels = triplet_labels) +
  scale_y_continuous(limits = c(y_min, y_max),
                     breaks = seq(1, 9, 0.5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  labs(
    title = "Karolinska Sleepiness Scale under light exposure triplets",
    subtitle = "Mean third-day KSS scores per triplet type (1.5-h intervention; P07 excluded)",
    x = "Triplet",
    y = expression(atop("Mean KSS",
                        scriptstyle("(1 = Extremely alert; 9 = Very sleepy)")))
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )
```

##### Data Overview: Triplet Observations
```{r}
#| label: triplet-overview-intervention

# Create detailed overview table
triplet_overview_q789 <- triplet_kss_all_q789 %>%
  filter(participant != "P07", !is.na(pos3_kss)) %>%
  select(participant, triplet, pos1_kss, pos2_kss, pos3_kss) %>%
  arrange(participant, triplet)

triplet_overview_q789
```

### Triplet Position Trajectory

#### Full Session (5-hour exposure)

**Mean KSS trajectory across triplet positions**
```{r}
#| label: fig-triplet-position-5h
#| fig-cap: "Mean KSS trajectory across triplet positions (5-h exposure, P07 excluded)"
#| fig-width: 8
#| fig-height: 8

# Prepare position data using full 5h window
mean_triplet_pos_5h <- triplet_kss_all %>%
  filter(participant != "P07") %>%
  group_by(triplet) %>%
  summarise(
    pos1_mean = mean(pos1_kss, na.rm = TRUE),
    pos2_mean = mean(pos2_kss, na.rm = TRUE),
    pos3_mean = mean(pos3_kss, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(pos1_mean, pos2_mean, pos3_mean), 
               names_to = "position", values_to = "mean_kss") %>%
  mutate(
    position_num = case_when(
      position == "pos1_mean" ~ 1,
      position == "pos2_mean" ~ 2,
      position == "pos3_mean" ~ 3
    ),
    light_type = substr(triplet, position_num, position_num),
    light_label = ifelse(light_type == "D", "dim", "bright"),
    triplet = factor(triplet, levels = c("DDD", "BDD", "DBD", "BBD", 
                                        "DDB", "BDB", "DBB", "BBB"))
  ) %>%
  filter(!is.na(mean_kss))

# Calculate y-axis limits
y_min <- floor(min(mean_triplet_pos_5h$mean_kss[is.finite(mean_triplet_pos_5h$mean_kss)]) * 2) / 2
y_max <- ceiling(max(mean_triplet_pos_5h$mean_kss[is.finite(mean_triplet_pos_5h$mean_kss)]) * 2) / 2

# Create plot
ggplot(mean_triplet_pos_5h, aes(x = position_num, y = mean_kss, group = triplet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_label), size = 3.5, shape = 16) +
  scale_color_manual(values = light_colors,
                     labels = c("dim" = "Dim (10 lx)", 
                                "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2, 3), 
                     labels = c("Position 1", "Position 2", "Position 3"), 
                     limits = c(0.8, 3.2)) +
  scale_y_continuous(limits = c(y_min, y_max),
                     breaks = seq(1, 9, 0.5),
                     expand = expansion(mult = c(0.05, 0.05))) +
  labs(
    title = "Karolinska Sleepiness Scale across light exposure triplets",
    subtitle = "Mean KSS scores across all participants (5-h exposure; P07 excluded)",
    x = "Triplet position", 
    y = expression(atop("Mean KSS", 
                        scriptstyle("(1 = Extremely alert; 9 = Very sleepy)"))),
    color = "Light condition"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,

    # Title & subtitle sizes aligned with all previous plots
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),

    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),

    legend.position = "right",
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 12),

    plot.margin = margin(15, 10, 10, 10)
  )
```

#### Intervention Window (1.5-hour exposure)

**Mean KSS trajectory across triplet positions**
```{r}
#| label: fig-triplet-position-intervention
#| fig-cap: "Mean KSS trajectory across triplet positions (1.5-h intervention window, P07 excluded)"
#| fig-width: 8
#| fig-height: 8

# Prepare position data using intervention window
mean_triplet_pos_q789 <- triplet_kss_all_q789 %>%
  filter(participant != "P07") %>%
  group_by(triplet) %>%
  summarise(
    pos1_mean = mean(pos1_kss, na.rm = TRUE),
    pos2_mean = mean(pos2_kss, na.rm = TRUE),
    pos3_mean = mean(pos3_kss, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(pos1_mean, pos2_mean, pos3_mean), 
               names_to = "position", values_to = "mean_kss") %>%
  mutate(
    position_num = case_when(
      position == "pos1_mean" ~ 1,
      position == "pos2_mean" ~ 2,
      position == "pos3_mean" ~ 3
    ),
    light_type = substr(triplet, position_num, position_num),
    light_label = ifelse(light_type == "D", "dim", "bright"),
    triplet = factor(triplet, levels = c("DDD", "BDD", "DBD", "BBD", 
                                        "DDB", "BDB", "DBB", "BBB"))
  ) %>%
  filter(!is.na(mean_kss))

# Calculate y-axis limits
y_min <- floor(min(mean_triplet_pos_q789$mean_kss[is.finite(mean_triplet_pos_q789$mean_kss)]) * 2) / 2
y_max <- ceiling(max(mean_triplet_pos_q789$mean_kss[is.finite(mean_triplet_pos_q789$mean_kss)]) * 2) / 2

# Create plot
ggplot(mean_triplet_pos_q789, aes(x = position_num, y = mean_kss, group = triplet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_label), size = 3.5, shape = 16) +
  scale_color_manual(values = light_colors,
                     labels = c("dim" = "Dim (10 lx)", 
                               "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2, 3), 
                     labels = c("Position 1", "Position 2", "Position 3"), 
                     limits = c(0.8, 3.2)) +
  scale_y_continuous(limits = c(y_min, y_max),
                     breaks = seq(1, 9, 0.5),
                     expand = expansion(mult = c(0.05, 0.05))) +
  labs(title = "Karolinska Sleepiness Scale across light exposure triplets",
    subtitle = "Mean KSS scores across all participants (1.5-h exposure; P07 excluded)",
    x = "Triplet position", 
       y = expression(atop("Mean KSS", 
                           scriptstyle("(1 = Extremely alert; 9 = Very sleepy)"))),
       color = "Light condition") +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "right",
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 12),
    plot.margin = margin(15, 10, 10, 10)
  )
```

### Triplet Intervention Window Summary Statistics (Boxplots)
```{r}
# Calculate statistics for 5-h exposure (full session)
triplet_stats_5h <- triplet_kss_all %>%
  filter(participant != "P07", !is.na(pos3_kss)) %>%
  select(triplet, pos3_kss) %>%
  group_by(triplet) %>%
  summarise(
    n_5h = n(),
    median_5h = median(pos3_kss, na.rm = TRUE),
    IQR_5h = IQR(pos3_kss, na.rm = TRUE),
    .groups = "drop"
  )

# Calculate statistics for 1.5-h exposure (intervention window)
triplet_stats_1.5h <- triplet_kss_all_q789 %>%
  filter(participant != "P07", !is.na(pos3_kss)) %>%
  select(triplet, pos3_kss) %>%
  group_by(triplet) %>%
  summarise(
    n_1.5h = n(),
    median_1.5h = median(pos3_kss, na.rm = TRUE),
    IQR_1.5h = IQR(pos3_kss, na.rm = TRUE),
    .groups = "drop"
  )

# Combine both exposure windows
triplet_stats_combined <- triplet_stats_5h %>%
  full_join(triplet_stats_1.5h, by = "triplet") %>%
  mutate(triplet = factor(triplet, levels = c("DDD", "BDD", "DBD", "BBD",
                                              "DDB", "BDB", "DBB", "BBB"))) %>%
  arrange(triplet)

# Display as formatted table
knitr::kable(
  triplet_stats_combined,
  digits = 2,
  col.names = c("Triplet", "N", "Median", "IQR", "N", "Median", "IQR"),
  caption = "Descriptive statistics for triplet third-day KSS scores across exposure windows (P07 excluded)",
  align = c("l", rep("r", 6))
) %>%
  kableExtra::add_header_above(c(" " = 1, "5-h Exposure" = 3, "1.5-h Exposure" = 3))
```

## Triplet Time-Course Analysis

### Data Preparation
```{r}
# Participants and triplets (already defined in setup)
participants_noP07 <- c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10")
triplets_all <- c("DDD", "DDB", "DBD", "DBB", "BDD", "BDB", "BBD", "BBB")

# Reuse triplet definition functions 
# Define light patterns for groups 
light_g1 <- unlist(strsplit("DDDBBBDDBDB", ""))
light_g2 <- unlist(strsplit("DBBBDDDBDB", ""))
light_g4_orig <- unlist(strsplit("DDDBXDBBBDBDD", ""))

# Create triplet dataframes for each group
df_triplets_g1 <- data.frame(
  session_1 = 1:(length(light_g1)-2),
  session_2 = 2:(length(light_g1)-1),
  session_3 = 3:length(light_g1),
  triplet = sapply(1:(length(light_g1)-2), function(i) paste0(light_g1[i:(i+2)], collapse = ""))
)

df_triplets_g2 <- data.frame(
  session_1 = 1:(length(light_g2)-2),
  session_2 = 2:(length(light_g2)-1),
  session_3 = 3:length(light_g2),
  triplet = sapply(1:(length(light_g2)-2), function(i) paste0(light_g2[i:(i+2)], collapse = ""))
)

# Function to get session 3 for each triplet type
get_triplet_session3_kss <- function(p, triplet_type) {
  if(p %in% c("P01", "P02", "P03")) {
    df_triplets <- df_triplets_g1
  } else if(p %in% c("P04", "P05", "P06", "P09")) {
    df_triplets <- df_triplets_g2
  } else {
    df_triplets <- data.frame(
      session_1 = 1:(length(light_g4_orig)-2),
      session_2 = 2:(length(light_g4_orig)-1),
      session_3 = 3:length(light_g4_orig),
      triplet = sapply(1:(length(light_g4_orig)-2), function(i) 
        paste0(light_g4_orig[i:(i+2)], collapse = ""))
    ) %>%
      filter(!str_detect(triplet, "X")) %>%
      filter(session_1 != 5 & session_2 != 5 & session_3 != 5)
  }
  
  sessions_3 <- df_triplets %>%
    filter(triplet == triplet_type) %>%
    pull(session_3)
  
  return(sessions_3)
}

# Function to get all participant-session combinations for a triplet
get_triplet_sessions_overview_kss <- function(triplet_type) {
  overview <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3_kss(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      for(s in sessions_3) {
        overview <- rbind(overview, data.frame(
          participant_id = p,
          session = s
        ))
      }
    }
  }
  
  return(overview %>% arrange(session, participant_id))
}

# Function to create plot for one triplet type
plot_triplet_timecourse_kss <- function(triplet_type) {
  df_triplet_kss <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3_kss(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      # Get KSS data for these sessions
      p_data <- insess_surv %>%
        mutate(
          session_raw = as.integer(str_extract(session, "\\d+$")),
          exp_session = case_when(
            id %in% c("P01", "P02", "P03") ~ session_raw - 100,
            id %in% c("P04", "P05") ~ session_raw - 200,
            id %in% c("P06", "P07", "P09") ~ session_raw - 300,
            id == "P10" ~ session_raw - 400,
            TRUE ~ NA_real_
          )
        ) %>%
        filter(id == p, exp_session %in% sessions_3) %>%
        select(id, ques_numb, kss)
      
      df_triplet_kss <- rbind(df_triplet_kss, p_data)
    }
  }
  
  triplet_summary <- df_triplet_kss %>%
    group_by(ques_numb) %>%
    summarise(
      mean_kss = mean(kss, na.rm = TRUE),
      sd_kss = sd(kss, na.rm = TRUE),
      n = n(),
      .groups = "drop"
    ) %>%
    mutate(ques_numb = as.numeric(ques_numb))
  
  y_min <- floor(min(triplet_summary$mean_kss - triplet_summary$sd_kss, 
                     na.rm = TRUE) * 2) / 2 - 0.1
  y_max <- ceiling(max(triplet_summary$mean_kss + triplet_summary$sd_kss, 
                       na.rm = TRUE) * 2) / 2 + 0.3
  
  hbt_values <- c("5.5h", "5h", "4.5h", "4h", "3.5h", "3h", 
                  "2.5h", "2h", "1.5h", "1h", "0.5h")
  
  # Determine third letter and intervention color/label
  third_letter <- substr(triplet_type, 3, 3)
  intervention_color <- ifelse(third_letter == "D", "#F5F5F5", "#FFF9E6")
  intervention_alpha <- ifelse(third_letter == "D", 0.4, 0.8)
  intervention_label <- ifelse(third_letter == "D", "dim (10 lx)", "bright (1000 lx)")
  
  # Create title with bold third letter
  triplet_first_two <- substr(triplet_type, 1, 2)
  triplet_third <- substr(triplet_type, 3, 3)
  
  p <- ggplot(triplet_summary, aes(x = ques_numb, y = mean_kss)) +
    annotate("rect", xmin = 1, xmax = 7, ymin = -Inf, ymax = Inf, 
             fill = "#F5F5F5", alpha = 0.4) +
    annotate("rect", xmin = 7, xmax = 10, ymin = -Inf, ymax = Inf, 
             fill = intervention_color, alpha = intervention_alpha) +
    annotate("rect", xmin = 10, xmax = 11, ymin = -Inf, ymax = Inf, 
             fill = "#F5F5F5", alpha = 0.4) +
    annotate("text", x = 3.75, y = y_max - 0.05, label = "dim (10 lx)", 
             hjust = 0.5, vjust = 1, size = 3.5, color = "grey50", fontface = "italic") +
    annotate("text", x = 8.5, y = y_max - 0.05, label = intervention_label, 
             hjust = 0.5, vjust = 1, size = 3.5, 
             color = ifelse(third_letter == "D", "grey50", "#B8860B"), 
             fontface = "italic") +
    annotate("text", x = 10.5, y = y_max - 0.05, label = "dim (10 lx)", 
             hjust = 0.5, vjust = 1, size = 3.5, color = "grey50", fontface = "italic") +
    geom_ribbon(aes(ymin = mean_kss - sd_kss,
                    ymax = mean_kss + sd_kss),
                alpha = 0.15, fill = "grey40") +
    geom_line(linewidth = 0.8, color = "black") +
    geom_point(size = 3.5, shape = 16, color = "black") +
    scale_x_continuous(
      breaks = 1:11,
      labels = paste0("Q", 1:11, "\n(", hbt_values, ")"),
      limits = c(0.5, 11.5),
      expand = c(0, 0)
    ) +
    scale_y_continuous(limits = c(y_min, y_max), expand = c(0, 0)) +
    labs(
      title = bquote("KSS time course - Triplet " * 
                     .(triplet_first_two) * bold(.(triplet_third))),
      subtitle = paste0("Grand average third-day (n=", 
                       length(unique(df_triplet_kss$id)), 
                       " participants; P07 excluded)"),
      x = "KSS question number (hours before HBT)",
      y = expression(atop("Mean KSS ± SD", 
                          scriptstyle("(1 = Extremely alert; 9 = Very sleepy)")))
    ) +
    theme_classic(base_size = 14) +
    theme(
      aspect.ratio = 1,
      plot.title = element_text(size = 16, hjust = 0),
      plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
      axis.text = element_text(color = "black", size = 13),
      axis.title = element_text(size = 14),
      axis.title.x = element_text(margin = margin(t = 8)),
      axis.title.y = element_text(margin = margin(r = 8)),
      panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
      plot.margin = margin(15, 10, 10, 10)
    )
  
  return(p)
}
```

### DDD Triplet
```{r}
knitr::kable(get_triplet_sessions_overview_kss("DDD"),
             caption = "Sessions used for Triplet DDD")
```
```{r}
plot_triplet_timecourse_kss("DDD")
```

### DDB Triplet
```{r}
knitr::kable(get_triplet_sessions_overview_kss("DDB"),
             caption = "Sessions used for Triplet DDB")
```
```{r}
plot_triplet_timecourse_kss("DDB")
```

### DBD Triplet
```{r}
knitr::kable(get_triplet_sessions_overview_kss("DBD"),
             caption = "Sessions used for Triplet DBD")
```
```{r}
plot_triplet_timecourse_kss("DBD")
```

### DBB Triplet
```{r}
knitr::kable(get_triplet_sessions_overview_kss("DBB"),
             caption = "Sessions used for Triplet DBB")
```
```{r}
plot_triplet_timecourse_kss("DBB")
```

### BDD Triplet
```{r}
knitr::kable(get_triplet_sessions_overview_kss("BDD"),
             caption = "Sessions used for Triplet BDD")
```
```{r}
plot_triplet_timecourse_kss("BDD")
```

### BDB Triplet
```{r}
knitr::kable(get_triplet_sessions_overview_kss("BDB"),
             caption = "Sessions used for Triplet BDB")
```
```{r}
plot_triplet_timecourse_kss("BDB")
```

### BBD Triplet
```{r}
knitr::kable(get_triplet_sessions_overview_kss("BBD"),
             caption = "Sessions used for Triplet BBD")
```
```{r}
plot_triplet_timecourse_kss("BBD")
```

### BBB Triplet
```{r}
knitr::kable(get_triplet_sessions_overview_kss("BBB"),
             caption = "Sessions used for Triplet BBB")
```
```{r}
plot_triplet_timecourse_kss("BBB")
```

### Panel Plot: Dim-Ending Triplets (DDD, BDD, DBD, BBD)
```{r}
# Prepare data for all D-ending triplets
triplets_D_panel <- c("DDD", "BDD", "DBD", "BBD")

df_all_D_triplets_kss <- data.frame()

for(triplet_type in triplets_D_panel) {
  df_triplet_kss <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3_kss(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      p_data <- insess_surv %>%
        mutate(
          session_raw = as.integer(str_extract(session, "\\d+$")),
          exp_session = case_when(
            id %in% c("P01", "P02", "P03") ~ session_raw - 100,
            id %in% c("P04", "P05") ~ session_raw - 200,
            id %in% c("P06", "P07", "P09") ~ session_raw - 300,
            id == "P10" ~ session_raw - 400,
            TRUE ~ NA_real_
          )
        ) %>%
        filter(id == p, exp_session %in% sessions_3) %>%
        select(id, ques_numb, kss) %>%
        mutate(triplet = triplet_type)
      
      df_triplet_kss <- rbind(df_triplet_kss, p_data)
    }
  }
  
  df_all_D_triplets_kss <- rbind(df_all_D_triplets_kss, df_triplet_kss)
}

# Calculate summary per triplet
triplet_summary_D_kss <- df_all_D_triplets_kss %>%
  group_by(triplet, ques_numb) %>%
  summarise(
    mean_kss = mean(kss, na.rm = TRUE),
    sd_kss = sd(kss, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    ques_numb = as.numeric(ques_numb),
    triplet = factor(triplet, levels = triplets_D_panel)
  )

# Y-axis limits for D-ending triplets
y_min_panel_D <- floor(min(triplet_summary_D_kss$mean_kss - triplet_summary_D_kss$sd_kss, 
                           na.rm = TRUE) * 2) / 2 - 0.1
y_max_panel_D <- ceiling(max(triplet_summary_D_kss$mean_kss + triplet_summary_D_kss$sd_kss, 
                             na.rm = TRUE) * 2) / 2 + 0.3

hbt_values <- c("5.5h", "5h", "4.5h", "4h", "3.5h", "3h", 
                "2.5h", "2h", "1.5h", "1h", "0.5h")

# Create panel plot
ggplot(triplet_summary_D_kss, aes(x = ques_numb, y = mean_kss)) +
  annotate("rect", xmin = 1, xmax = 7, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  annotate("rect", xmin = 7, xmax = 10, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  annotate("rect", xmin = 10, xmax = 11, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  geom_ribbon(aes(ymin = mean_kss - sd_kss,
                  ymax = mean_kss + sd_kss),
              alpha = 0.15, fill = "grey40") +
  geom_line(linewidth = 0.8, color = "black") +
  geom_point(size = 2.5, shape = 16, color = "black") +
  scale_x_continuous(
    breaks = 1:11,
    labels = paste0("Q", 1:11, "\n(", hbt_values, ")"),
    limits = c(0.5, 11.5),
    expand = c(0, 0)
  ) +
  scale_y_continuous(limits = c(y_min_panel_D, y_max_panel_D), expand = c(0, 0)) +
  facet_wrap(~ triplet, ncol = 2, 
             labeller = labeller(triplet = function(x) {
               sapply(x, function(t) {
                 first_two <- substr(t, 1, 2)
                 third <- substr(t, 3, 3)
                 paste0(first_two, third)
               })
             })) +
  labs(
    title = "KSS time course - Dim-ending triplets",
    subtitle = "Grand average third-day (P07 excluded)",
    x = "KSS question number (hours before HBT)",
    y = expression(atop("Mean KSS", 
                        scriptstyle("(1 = Extremely alert; 9 = Very sleepy)")))
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 11),
    axis.title = element_text(size = 13),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    strip.background = element_rect(fill = "white", color = "grey80", linewidth = 0.5),
    strip.text = element_text(size = 12, color = "black", face = "bold"),
    plot.margin = margin(15, 10, 10, 10)
  )
```

### Panel Plot: Bright-Ending Triplets (DDB, BDB, DBB, BBB)
```{r}
# Prepare data for all B-ending triplets
triplets_B_panel <- c("DDB", "BDB", "DBB", "BBB")

df_all_B_triplets_kss <- data.frame()

for(triplet_type in triplets_B_panel) {
  df_triplet_kss <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3_kss(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      p_data <- insess_surv %>%
        mutate(
          session_raw = as.integer(str_extract(session, "\\d+$")),
          exp_session = case_when(
            id %in% c("P01", "P02", "P03") ~ session_raw - 100,
            id %in% c("P04", "P05") ~ session_raw - 200,
            id %in% c("P06", "P07", "P09") ~ session_raw - 300,
            id == "P10" ~ session_raw - 400,
            TRUE ~ NA_real_
          )
        ) %>%
        filter(id == p, exp_session %in% sessions_3) %>%
        select(id, ques_numb, kss) %>%
        mutate(triplet = triplet_type)
      
      df_triplet_kss <- rbind(df_triplet_kss, p_data)
    }
  }
  
  df_all_B_triplets_kss <- rbind(df_all_B_triplets_kss, df_triplet_kss)
}

# Calculate summary per triplet
triplet_summary_B_kss <- df_all_B_triplets_kss %>%
  group_by(triplet, ques_numb) %>%
  summarise(
    mean_kss = mean(kss, na.rm = TRUE),
    sd_kss = sd(kss, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    ques_numb = as.numeric(ques_numb),
    triplet = factor(triplet, levels = triplets_B_panel)
  )

# Y-axis limits for B-ending triplets
y_min_panel_B <- floor(min(triplet_summary_B_kss$mean_kss - triplet_summary_B_kss$sd_kss, 
                           na.rm = TRUE) * 2) / 2 - 0.1
y_max_panel_B <- ceiling(max(triplet_summary_B_kss$mean_kss + triplet_summary_B_kss$sd_kss, 
                             na.rm = TRUE) * 2) / 2 + 0.3

hbt_values <- c("5.5h", "5h", "4.5h", "4h", "3.5h", "3h", 
                "2.5h", "2h", "1.5h", "1h", "0.5h")

# Create panel plot
ggplot(triplet_summary_B_kss, aes(x = ques_numb, y = mean_kss)) +
  annotate("rect", xmin = 1, xmax = 7, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  annotate("rect", xmin = 7, xmax = 10, ymin = -Inf, ymax = Inf, 
           fill = "#FFF9E6", alpha = 0.8) +
  annotate("rect", xmin = 10, xmax = 11, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  geom_ribbon(aes(ymin = mean_kss - sd_kss,
                  ymax = mean_kss + sd_kss),
              alpha = 0.15, fill = "grey40") +
  geom_line(linewidth = 0.8, color = "black") +
  geom_point(size = 2.5, shape = 16, color = "black") +
  scale_x_continuous(
    breaks = 1:11,
    labels = paste0("Q", 1:11, "\n(", hbt_values, ")"),
    limits = c(0.5, 11.5),
    expand = c(0, 0)
  ) +
  scale_y_continuous(limits = c(y_min_panel_B, y_max_panel_B), expand = c(0, 0)) +
  facet_wrap(~ triplet, ncol = 2, 
             labeller = labeller(triplet = function(x) {
               sapply(x, function(t) {
                 first_two <- substr(t, 1, 2)
                 third <- substr(t, 3, 3)
                 paste0(first_two, third)
               })
             })) +
  labs(
    title = "KSS time course - Bright-ending triplets",
    subtitle = "Grand average third-day (P07 excluded)",
    x = "KSS question number (hours before HBT)",
    y = expression(atop("Mean KSS", 
                        scriptstyle("(1 = Extremely alert; 9 = Very sleepy)")))
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 11),
    axis.title = element_text(size = 13),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    strip.background = element_rect(fill = "white", color = "grey80", linewidth = 0.5),
    strip.text = element_text(size = 12, color = "black", face = "bold"),
    plot.margin = margin(15, 10, 10, 10)
  )
```

### Grand Average with D-ending and B-ending rows
```{r fig.width=14, fig.height=8}
#| label: fig-kss-triplet-panel
#| fig-cap: "KSS time course - Grand average and triplet patterns (P07 excluded)"
#| fig-width: 14
#| fig-height: 8

# Prepare data for all triplets
triplets_all_panel <- c("DDD", "BDD", "DBD", "BBD", "DDB", "BDB", "DBB", "BBB")

df_all_triplets_unified_kss <- data.frame()

for(triplet_type in triplets_all_panel) {
  df_triplet_kss <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3_kss(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      p_data <- insess_surv %>%
        mutate(
          session_raw = as.integer(str_extract(session, "\\d+$")),
          exp_session = case_when(
            id %in% c("P01", "P02", "P03") ~ session_raw - 100,
            id %in% c("P04", "P05") ~ session_raw - 200,
            id %in% c("P06", "P07", "P09") ~ session_raw - 300,
            id == "P10" ~ session_raw - 400,
            TRUE ~ NA_real_
          )
        ) %>%
        filter(id == p, exp_session %in% sessions_3) %>%
        select(id, ques_numb, kss) %>%
        mutate(triplet = triplet_type)
      
      df_triplet_kss <- rbind(df_triplet_kss, p_data)
    }
  }
  
  df_all_triplets_unified_kss <- rbind(df_all_triplets_unified_kss, df_triplet_kss)
}

# Calculate summary per triplet
triplet_summary_unified_kss <- df_all_triplets_unified_kss %>%
  group_by(triplet, ques_numb) %>%
  summarise(
    mean_kss = mean(kss, na.rm = TRUE),
    sd_kss = sd(kss, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    ques_numb = as.numeric(ques_numb),
    triplet = factor(triplet, levels = triplets_all_panel)
  )

# Unified Y-axis limits
y_min_unified <- floor(min(triplet_summary_unified_kss$mean_kss - triplet_summary_unified_kss$sd_kss, 
                           na.rm = TRUE) * 2) / 2 - 0.1
y_max_unified <- ceiling(max(triplet_summary_unified_kss$mean_kss + triplet_summary_unified_kss$sd_kss, 
                             na.rm = TRUE) * 2) / 2 + 0.3

# Function to create triplet plot
create_triplet_plot <- function(triplet_name, show_x = FALSE, show_y = TRUE) {
  data_triplet <- triplet_summary_unified_kss %>%
    filter(triplet == triplet_name)
  
  triplet_third <- substr(triplet_name, 3, 3)
  intervention_color <- ifelse(triplet_third == "D", "#F5F5F5", "#FFF9E6")
  intervention_alpha <- ifelse(triplet_third == "D", 0.4, 0.8)
  
  triplet_first_two <- substr(triplet_name, 1, 2)
  triplet_label <- paste0(triplet_first_two, "**", triplet_third, "**")
  
  ggplot(data_triplet, aes(x = ques_numb, y = mean_kss)) +
    annotate("rect", xmin = 1, xmax = 7, ymin = -Inf, ymax = Inf, 
             fill = "#F5F5F5", alpha = 0.4) +
    annotate("rect", xmin = 7, xmax = 10, ymin = -Inf, ymax = Inf, 
             fill = intervention_color, alpha = intervention_alpha) +
    annotate("rect", xmin = 10, xmax = 11, ymin = -Inf, ymax = Inf, 
             fill = "#F5F5F5", alpha = 0.4) +
    geom_ribbon(aes(ymin = mean_kss - sd_kss, ymax = mean_kss + sd_kss),
                alpha = 0.15, fill = "grey40") +
    geom_line(linewidth = 0.8, color = "black") +
    geom_point(size = 2.5, shape = 16, color = "black") +
    scale_x_continuous(
      breaks = 1:11,
      labels = if(show_x) c("Q1\n(5.5h)", "Q2", "Q3\n(4.5h)", "Q4", "Q5\n(3.5h)", 
                             "Q6", "Q7\n(2.5h)", "Q8", "Q9\n(1.5h)", "Q10", "Q11\n(0.5h)") else 1:11,
      limits = c(0.5, 11.5),
      expand = c(0, 0)
    ) +
    scale_y_continuous(limits = c(y_min_unified, y_max_unified), expand = c(0, 0)) +
    labs(
      title = triplet_label,
      x = NULL,
      y = if(show_y) expression(atop("Mean KSS ± SD", 
                        scriptstyle("(1 = Extremely alert; 9 = Very sleepy)"))) else NULL
    ) +
    theme_classic(base_size = 11) +
    theme(
      plot.title = element_markdown(size = 11, hjust = 0.5, margin = margin(b = 5)),
      axis.text = element_text(color = "black", size = 9),
      axis.text.x = element_text(size = 8),
      axis.title = element_text(size = 10),
      axis.title.y = element_text(margin = margin(r = 5)),
      panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
      plot.margin = margin(5, 5, 5, 5)
    )
}

# Create D-ending triplet plots (no x-axis labels)
p_DDD <- create_triplet_plot("DDD", show_x = FALSE, show_y = TRUE)
p_BDD <- create_triplet_plot("BDD", show_x = FALSE, show_y = FALSE)
p_DBD <- create_triplet_plot("DBD", show_x = FALSE, show_y = FALSE)
p_BBD <- create_triplet_plot("BBD", show_x = FALSE, show_y = FALSE)

# Create B-ending triplet plots (with x-axis labels)
p_DDB <- create_triplet_plot("DDB", show_x = TRUE, show_y = TRUE)
p_BDB <- create_triplet_plot("BDB", show_x = TRUE, show_y = FALSE)
p_DBB <- create_triplet_plot("DBB", show_x = TRUE, show_y = FALSE)
p_BBB <- create_triplet_plot("BBB", show_x = TRUE, show_y = FALSE)

# Grand Average plot
p_grand <- ggplot(grand_summary_all, 
       aes(x = ques_numb, y = mean_kss, 
           color = light_type, fill = light_type, group = light_type)) +
  annotate("rect", xmin = 1, xmax = 7, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  annotate("rect", xmin = 10, xmax = 11, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  annotate("rect", xmin = 7, xmax = 10, ymin = -Inf, ymax = Inf, 
           fill = "#FFF9E6", alpha = 0.8) +
  geom_ribbon(aes(ymin = mean_kss - se_kss, ymax = mean_kss + se_kss), 
              alpha = 0.1, color = NA, show.legend = FALSE) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.5, shape = 16) +
  scale_color_manual(
    values = c("BRIGHT" = "#F5A623", "DIM" = "#4A90E2"),
    labels = c("BRIGHT" = "Bright (1000lx)", "DIM" = "Dim (10lx)")
  ) +
  scale_fill_manual(
    values = c("BRIGHT" = "#F5A623", "DIM" = "#4A90E2"),
    labels = c("BRIGHT" = "Bright (1000lx)", "DIM" = "Dim (10lx)")
  ) +
  scale_x_continuous(
    breaks = 1:11,
    labels = 1:11,
    limits = c(0.5, 11.5),
    expand = c(0, 0)
  ) +
  scale_y_continuous(limits = c(y_min_unified, y_max_unified), expand = c(0, 0)) +
  labs(
    title = "Grand Average",
    x = NULL,
    y = NULL,
    color = "Light condition"
  ) +
  guides(
    color = guide_legend(override.aes = list(shape = 16, size = 2.5)),
    fill = "none"
  ) +
  theme_classic(base_size = 11) +
  theme(
    plot.title = element_text(size = 11, hjust = 0.5, margin = margin(b = 5)),
    axis.text = element_text(color = "black", size = 9),
    axis.text.x = element_text(size = 9),
    axis.title = element_text(size = 10),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
     legend.position = "bottom",
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 7),
    plot.margin = margin(5, 5, 5, 5)
  )

# Empty spacer
p_spacer <- ggplot() + theme_void()

# Combine in layout
(p_DDD | p_BDD | p_DBD | p_BBD | p_grand) /
(p_DDB | p_BDB | p_DBB | p_BBB | p_spacer) +
  plot_annotation(
    title = "KSS time course - Grand average and triplet patterns",
    subtitle = "Grand average and third-day triplet patterns (P07 excluded)",
    caption = "KSS questionnaire number (hours before HBT)",
    theme = theme(
      plot.title = element_text(size = 15, hjust = 0, face = "plain"),
      plot.subtitle = element_text(size = 12, hjust = 0, margin = margin(b = 8)),
      plot.caption = element_text(size = 11, hjust = 0.5, margin = margin(t = 8))
    )
  )
```


# Data Preparaton for Statistical Analysis

## Prepare Data for Linear Mixed Models

```{r}
#| label: lmm-data-prep

# Add participant demographics (WITHOUT P07)
participant_demographics <- tibble(
  participant_id = c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10"),
  sex = c("female", "male", "female", "female", "male", 
          "female", "female", "male"),
  age = c(24, 21, 24, 23, 31, 23, 28, 32)
)

# Prepare LMM dataset with lagged light exposure
kss_lmm <- map_dfr(participant_list, function(pid) {
  # Get KSS data
  df <- kss_per_evening_q789 %>%
    filter(participant_id == pid) %>%
    select(participant_id, exp_session, mean_kss)
  
  # Add light conditions
  light_df <- tibble(
    exp_session = seq_along(light_patterns[[pid]]),
    Light_Lag0 = light_patterns[[pid]]
  )
  
  df %>% left_join(light_df, by = "exp_session")
})

# Create lagged variables and numeric coding
kss_lmm <- kss_lmm %>%
  group_by(participant_id) %>%
  arrange(exp_session) %>%
  mutate(
    # Lag 1 = previous evening
    Light_Lag1 = lag(Light_Lag0, 1),
    # Lag 2 = two evenings ago
    Light_Lag2 = lag(Light_Lag0, 2),
    # Numeric coding: Dim=0, Bright=1
    Light_Lag0_num = ifelse(Light_Lag0 == "D", 0, 1),
    Light_Lag1_num = ifelse(Light_Lag1 == "D", 0, 1),
    Light_Lag2_num = ifelse(Light_Lag2 == "D", 0, 1)
  ) %>%
  ungroup() %>%
  # Remove rows with missing lags or excluded sessions
  filter(!is.na(Light_Lag0) & !is.na(Light_Lag1) & !is.na(Light_Lag2)) %>%
  filter(!(participant_id == "P10" & exp_session %in% c(6, 7))) %>%
  filter(participant_id != "P07") %>%
  # Add demographics
  left_join(participant_demographics, by = "participant_id") %>%
  # Center age for interpretability
  mutate(age_c = age - mean(age))

# Display prepared data
cat("LMM dataset prepared (P07 excluded from statistical analysis):\n")
cat("Total observations:", nrow(kss_lmm), "\n")
cat("Participants:", paste(sort(unique(kss_lmm$participant_id)), collapse = ", "), "\n\n")

kss_lmm %>% 
  group_by(participant_id) %>% 
  summarise(n_sessions = n(), .groups = "drop") %>%
  print()

# Display full dataset kss_lmm
DT::datatable(kss_lmm, 
              options = list(pageLength = 10, scrollX = TRUE),
              caption = "Complete LMM dataset with lagged light exposure variables (P07 excluded)")
```

## KSS Trajectories Across Experimental Sessions

Individual participant trajectories with LOESS smooth (full dataset, n=67 observations)
```{r}
#| label: fig-kss-trajectories
#| fig-cap: "KSS across experimental sessions - Individual trajectories"
#| fig-width: 10
#| fig-height: 8

# Prepare data - count observations per session
session_counts <- kss_per_evening_q789 %>%
  filter(participant_id != "P07") %>%  # Exclude P07
  group_by(exp_session) %>%
  summarise(n = n(), .groups = "drop")

# Create plot
ggplot(kss_per_evening_q789 %>% filter(participant_id != "P07"), 
       aes(x = exp_session, y = mean_kss, color = participant_id, group = participant_id)) +
  
  # Individual trajectories
  geom_line(alpha = 0.6, linewidth = 0.8) +
  geom_point(size = 2.5, alpha = 0.7) +
  
  # LOESS smooth (overall trend)
  geom_smooth(aes(group = 1), method = "loess", span = 0.75,
              color = "black", linetype = "dashed", linewidth = 1.2,
              se = TRUE, fill = "grey50", alpha = 0.3) +
  
  # Add sample size annotations
  geom_text(data = session_counts %>% filter(n < 8),
            aes(x = exp_session, y = 9, label = paste0("n=", n)),
            color = "black", size = 3, inherit.aes = FALSE) +
  
  # Vertical line for data cutoff 
  geom_vline(xintercept = 11.5, linetype = "dotted", color = "grey50", linewidth = 0.8) +
  
  # Scales
  scale_x_continuous(
    breaks = 1:13,
    limits = c(2.5, 13.5),
    expand = c(0.01, 0)
  ) +
  scale_y_continuous(
    limits = c(0, 9.5),
    breaks = seq(0, 9, 1),
    expand = c(0, 0)
  ) +
  
  # Colors
  scale_color_manual(
    values = c("P01" = "#E41A1C", "P02" = "#377EB8", "P03" = "#4DAF4A",
               "P04" = "#984EA3", "P05" = "#FF7F00", "P06" = "#FFFF33",
               "P09" = "#A65628", "P10" = "#F781BF"),
    name = "Participant"
  ) +
  
  # Labels
  labs(
    title = "KSS across experimental sessions",
    subtitle = "Individual participant trajectories with LOESS smooth (full dataset, n=67)",
    x = "Experimental Session",
    y = expression(atop("Mean KSS (intervention window)",
                        scriptstyle("(1 = Extremely alert; 9 = Very sleepy)")))
  ) +
  
  # Theme
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, face = "plain"),
    plot.subtitle = element_text(size = 12, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 12),
    axis.title = element_text(size = 13),
    legend.position = "right",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    plot.margin = margin(15, 15, 10, 10)
  )
```

## Distribution of KSS Scores

### Observed Distribution
```{r}
#| label: fig-kss-distribution
#| fig-cap: "Distribution of mean KSS scores"
#| fig-width: 10
#| fig-height: 6

# Prepare data
kss_dist_data <- kss_per_evening_q789 %>%
  filter(participant_id != "P07")

# Calculate statistics
mean_kss <- mean(kss_dist_data$mean_kss, na.rm = TRUE)
median_kss <- median(kss_dist_data$mean_kss, na.rm = TRUE)

# Create density plot
ggplot(kss_dist_data, aes(x = mean_kss)) +
  
  # Histogram
  geom_histogram(aes(y = after_stat(density)), 
                 bins = 20, fill = "#4A90E2", alpha = 0.6, 
                 color = "black", linewidth = 0.3) +
  
  # Density curve
  geom_density(color = "#E41A1C", linewidth = 1.5, adjust = 1.2) +
  
  # Mean line
  geom_vline(xintercept = mean_kss, color = "black", 
             linetype = "dashed", linewidth = 1) +
  annotate("text", x = mean_kss + 0.3, y = 0.35, 
           label = paste0("Mean = ", round(mean_kss, 2)),
           color = "black", size = 4, hjust = 0) +
  
  # Median line
  geom_vline(xintercept = median_kss, color = "darkgreen", 
             linetype = "dotted", linewidth = 1) +
  annotate("text", x = median_kss + 0.3, y = 0.32, 
           label = paste0("Median = ", round(median_kss, 2)),
           color = "darkgreen", size = 4, hjust = 0) +
  
  # Scales
  scale_x_continuous(
    limits = c(1, 9),
    breaks = 1:9,
    expand = c(0.01, 0)
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  
  # Labels
  labs(
    title = "Distribution of Mean KSS Scores",
    subtitle = paste0("n = ", nrow(kss_dist_data), " observations from ", 
                     length(unique(kss_dist_data$participant_id)), " participants"),
    x = "Mean KSS (1 = Extremely alert; 9 = Very sleepy)",
    y = "Density"
  ) +
  
  # Theme
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, face = "plain"),
    plot.subtitle = element_text(size = 12, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 12),
    axis.title = element_text(size = 13),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    plot.margin = margin(15, 15, 10, 10)
  )

# Print descriptive statistics
cat("\nDescriptive Statistics:\n")
cat("Mean:", round(mean_kss, 2), "\n")
cat("Median:", round(median_kss, 2), "\n")
cat("SD:", round(sd(kss_dist_data$mean_kss, na.rm = TRUE), 2), "\n")
cat("Range:", round(min(kss_dist_data$mean_kss, na.rm = TRUE), 2), "-", 
    round(max(kss_dist_data$mean_kss, na.rm = TRUE), 2), "\n")
cat("Skewness:", round(moments::skewness(kss_dist_data$mean_kss), 2), "\n")
cat("Kurtosis:", round(moments::kurtosis(kss_dist_data$mean_kss), 2), "\n")
```

### Model Specification Rationale

**Interpretation:** The distribution of mean KSS scores shows bimodal characteristics. A Linear Mixed Model (LMM) with Gaussian distribution was selected for the following reasons:

**1. Aggregation Creates Quasi-Continuous Data:**
Although individual KSS responses are ordinal (9-point scale), averaging across three measurements (Q7-Q9) within the intervention window produces quasi-continuous values. The Central Limit Theorem ensures that such averaged scores approximate a normal distribution, even when based on ordinal data. 

**2. Ordinal vs Beta Distribution:**
A Generalized Linear Mixed Model (GLMM) with beta distribution was considered but rejected because:
   - Beta regression requires continuous data on the (0,1) interval with equal intervals
   - KSS is ordinal with potentially unequal intervals (e.g., distance between scores 1-2 ≠ 8-9)
   - Transforming KSS scores to (0,1) via [(KSS-1)/8] would impose artificial continuity and equal spacing assumptions that misrepresent the scale's measurement properties

**3. Robustness of Gaussian LMM:**
Linear Mixed Models are robust to moderate deviations from normality, particularly when:
   - Sample size is adequate (n = 67 observations from 8 participants)
   - Interest focuses on mean differences between conditions rather than distributional shape
   - Residual diagnostics confirm approximate normality (Shapiro-Wilk p = 0.060)
   - The observed bimodality reflects between-participant heterogeneity (captured by random intercepts) rather than model misspecification

----

# Statistical Analysis

## Hierarchical Model Building

Models were built hierarchically, adding complexity only when justified by likelihood ratio tests (α = 0.05).

----

## Step 1: Baseline Model (Intercept-only)

### Model Specification
```{r}
#| label: baseline-model

m0 <- lmer(mean_kss ~ 1 + (1 | participant_id), data = kss_lmm, REML = FALSE)
summary(m0)

# Wald Confidence Intervals
ci_m0 <- confint(m0, method = "Wald")
print(ci_m0)
```

### Fit Statistics and Variance Components
```{r}
#| label: baseline-statistics

# Model fit
cat("AIC:", round(AIC(m0), 2), "| BIC:", round(BIC(m0), 2), 
    "| LogLik:", round(as.numeric(logLik(m0)), 2), "\n\n")

# Variance components
vc <- as.data.frame(VarCorr(m0))
icc <- vc$vcov[1] / sum(vc$vcov)

cat("Between-participant variance (τ²):", round(vc$vcov[1], 3), "\n")
cat("Within-participant variance (σ²): ", round(vc$vcov[2], 3), "\n")
cat("ICC:", round(icc, 3), "(", round(icc * 100, 1), "% between participants)\n")
```

### Participant-Specific Means
```{r}
#| label: baseline-plot
#| fig-width: 8
#| fig-height: 5

ranef_m0 <- ranef(m0)$participant_id
ranef_m0$participant_id <- rownames(ranef_m0)
names(ranef_m0)[1] <- "random_intercept"
grand_mean <- fixef(m0)["(Intercept)"]
ranef_m0$predicted_mean <- grand_mean + ranef_m0$random_intercept
ranef_m0 <- ranef_m0[order(ranef_m0$predicted_mean), ]
ranef_m0$participant_id <- factor(ranef_m0$participant_id, levels = ranef_m0$participant_id)

ggplot(ranef_m0, aes(x = participant_id, y = predicted_mean)) +
  geom_hline(yintercept = grand_mean, linetype = "dashed", color = "red", linewidth = 1) +
  geom_point(size = 4, color = "#4A90E2") +
  geom_segment(aes(xend = participant_id, y = grand_mean, yend = predicted_mean),
               color = "#4A90E2", linewidth = 1) +
  labs(x = "Participant", y = "Predicted Mean KSS",
       title = "Participant-Specific Mean KSS (Baseline Model)") +
  scale_y_continuous(limits = c(1, 9), breaks = 1:9) +
  theme_classic(base_size = 13) +
  theme(panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3))
```

### Residual Diagnostics
```{r}
#| label: baseline-diagnostics
#| fig-width: 10
#| fig-height: 5

par(mfrow = c(1, 2))
hist(resid(m0), breaks = 20, main = "Residuals", xlab = "Residuals", 
     col = "lightblue", border = "black")
qqnorm(resid(m0), main = "Q-Q Plot", pch = 16, col = rgb(0, 0, 0, 0.5))
qqline(resid(m0), col = "red", lwd = 2)
par(mfrow = c(1, 1))
```

**Interpretation:** The baseline model establishes that `r round(icc * 100, 1)`% of variance in KSS is attributable to participant differences (ICC = `r round(icc, 3)`). The next step tests whether light exposure variables improve model fit.

## Step 2: Add Light Exposure Lags (Sequential Testing)

### Model m1a: Add Light Lag 0 Only
```{r}
#| label: step2-m1a-model

m1a <- lmer(
  mean_kss ~ Light_Lag0_num + (1 | participant_id),
  data = kss_lmm,
  REML = FALSE
)

summary(m1a)

# Wald Confidence Intervals
ci_m1a <- confint(m1a, method = "Wald")
print(ci_m1a)
```

### Model m1b: Add Light Lag 1 Additionally (One Day Before)
```{r}
#| label: step2-m1b-model

m1b <- lmer(
  mean_kss ~ Light_Lag0_num + Light_Lag1_num + (1 | participant_id),
  data = kss_lmm,
  REML = FALSE
)

summary(m1b)

# Wald Confidence Intervals
ci_m1b <- confint(m1b, method = "Wald")
print(ci_m1b)
```

### Model m1c: Add Light Lag 2 Additionally (Two Days Before)
```{r}
#| label: step2-m1c-model

m1c <- lmer(
  mean_kss ~ Light_Lag0_num + Light_Lag1_num + Light_Lag2_num + 
    (1 | participant_id),
  data = kss_lmm,
  REML = FALSE
)

summary(m1c)

# Wald Confidence Intervals
ci_m1c <- confint(m1c, method = "Wald")
print(ci_m1c)
```

### Likelihood Ratio Tests: Lag Effects
```{r}
#| label: step2-lrt-tests

# LRT 1: Does adding Lag 0 improve fit over baseline?
lrt_m0_vs_m1a <- anova(m0, m1a)
print(lrt_m0_vs_m1a)

# LRT 2: Does adding Lag 1 improve fit over Lag 0 only?
lrt_m1a_vs_m1b <- anova(m1a, m1b)
print(lrt_m1a_vs_m1b)

# LRT 3: Does adding Lag 2 improve fit over Lag 0 + Lag 1?
lrt_m1b_vs_m1c <- anova(m1b, m1c)
print(lrt_m1b_vs_m1c)
```

### Model Comparison Table
```{r}
#| label: step2-comparison-table

data.frame(
  Model = c("m0 (Baseline)", 
            "m1a (+ Lag 0)", 
            "m1b (+ Lag 0+1)", 
            "m1c (+ Lag 0+1+2)"),
  npar = c(attr(logLik(m0), "df"), 
           attr(logLik(m1a), "df"),
           attr(logLik(m1b), "df"),
           attr(logLik(m1c), "df")),
  AIC = c(AIC(m0), AIC(m1a), AIC(m1b), AIC(m1c)),
  BIC = c(BIC(m0), BIC(m1a), BIC(m1b), BIC(m1c)),
  LogLik = c(as.numeric(logLik(m0)), 
             as.numeric(logLik(m1a)),
             as.numeric(logLik(m1b)),
             as.numeric(logLik(m1c)))
) %>%
  kable(digits = 2, caption = "Sequential Model Comparison: Testing Light Lag Effects") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

### Interpretation

Based on the LRT results:

- **LRT 1 (m0 vs m1a):** Adding Lag 0 significantly improved fit over baseline (χ²(1) = 11.23, p < 0.001, ΔAIC = -9.2) 
- **LRT 2 (m1a vs m1b):** Adding Lag 1 did not improve fit over Lag 0 only (χ²(1) = 0.32, p = 0.572, ΔAIC = +1.7) 
- **LRT 3 (m1b vs m1c):** Adding Lag 2 did not improve fit over Lag 0 + Lag 1 (χ²(1) = 0.60, p = 0.438, ΔAIC = +1.4) 

**Decision:** Proceed with **m1a** (Light_Lag0_num only). Only current evening light exposure significantly predicts subjective sleepiness; prior nights show no statistically significant carryover effects.

----

## Step 3: Test Session Effects

### Linear Session Term
```{r}
#| label: step3-linear

m2 <- lmer(
  mean_kss ~ Light_Lag0_num + 
    exp_session + (1 | participant_id),
  data = kss_lmm,
  REML = FALSE
)

summary(m2)

# Wald Confidence Intervals
ci_m2 <- confint(m2, method = "Wald")
print(ci_m2)
```

### Quadratic Session Term
```{r}
#| label: step3-quadratic

m3 <- lmer(
  mean_kss ~ Light_Lag0_num + 
    exp_session + I(exp_session^2) + (1 | participant_id),
  data = kss_lmm,
  REML = FALSE
)

summary(m3)

# Wald Confidence Intervals
ci_m3 <- confint(m3, method = "Wald")
print(ci_m3)
```

### Likelihood Ratio Tests: Session Effects
```{r}
#| label: step3-lrt

# LRT 1: Does adding exp_session (linear) improve fit over Lag 0 only?
lrt_m1a_vs_m2 <- anova(m1a, m2)
print(lrt_m1a_vs_m2)

# LRT 2: Does adding quadratic term improve fit over linear only?
lrt_m2_vs_m3 <- anova(m2, m3)
print(lrt_m2_vs_m3)
```

### Interpretation

Based on the LRT results, model fit was not significantly improved by adding linear exp_session (χ²(1) = 0.23, p = 0.634, ΔAIC = +1.8), nor by adding the quadratic term (χ²(1) = 0.05, p = 0.823, ΔAIC = +1.9). Therefore, model **m1a** (Lag 0 only) was retained as the final model, indicating that subjective sleepiness shows no systematic temporal trend in this experiment (habituation or sensitization) across experimental sessions.

**Final Primary Model:** m1a (Light_Lag0_num + random intercept)

----

## Primary Model: Diagnostics

Assumption checks for the selected primary model (m1a).

### Check 1: Multicollinearity (VIF)
```{r}
#| label: primary-vif
#| fig-width: 8
#| fig-height: 6

# Count number of fixed effects (excluding intercept)
n_predictors <- length(fixef(m1a)) - 1  # -1 for intercept

if (n_predictors >= 2) {
  vif_m1a <- vif(m1a)
  
  # Create barplot
  barplot(vif_m1a, 
          main = "Variance Inflation Factors (Primary Model)",
          ylab = "VIF",
          col = ifelse(vif_m1 > 5, "red", 
                      ifelse(vif_m1 > 2, "orange", "lightblue")),
          las = 2,
          ylim = c(0, max(vif_m1) * 1.2))
  abline(h = 5, col = "red", lty = 2, lwd = 2)
  abline(h = 10, col = "darkred", lty = 2, lwd = 2)
  legend("topright", 
         legend = c("VIF < 2 (excellent)", "VIF 2-5 (acceptable)", "VIF > 5 (concern)"),
         fill = c("lightblue", "orange", "red"),
         cex = 0.8)
  
  print(round(vif_m1a, 2))
  
  if (all(vif_m1 < 5)) {
    cat("\n✓ All VIF < 5: No multicollinearity issues.\n")
  } else {
    cat("\n⚠ Warning: Some VIF ≥ 5, indicating potential multicollinearity.\n")
  }
} else {
  cat("Final model (m1) contains only 1 predictor (Light_Lag0_num).\n")
  cat("VIF not applicable - multicollinearity assessment requires at least 2 predictors.\n")
  cat("\n No multicollinearity possible with single predictor.\n")
}
```

### Check 2: Residual Plots
```{r}
#| label: primary-residuals
#| fig-width: 10
#| fig-height: 8

par(mfrow = c(2, 2))

# Panel 1: Residuals vs Fitted
plot(fitted(m1a), resid(m1a), xlab = "Fitted", ylab = "Residuals",
     main = "Residuals vs Fitted", pch = 16, col = rgb(0, 0, 0, 0.5))
abline(h = 0, col = "red", lty = 2, lwd = 2)
lines(lowess(fitted(m1a), resid(m1a)), col = "blue", lwd = 2)

# Panel 2: Q-Q Plot
qqnorm(resid(m1a), main = "Q-Q Plot", pch = 16, col = rgb(0, 0, 0, 0.5))
qqline(resid(m1a), col = "red", lwd = 2)

# Panel 3: Scale-Location
plot(fitted(m1a), sqrt(abs(resid(m1a))), xlab = "Fitted", ylab = "√|Residuals|",
     main = "Scale-Location", pch = 16, col = rgb(0, 0, 0, 0.5))
lines(lowess(fitted(m1a), sqrt(abs(resid(m1a)))), col = "blue", lwd = 2)

# Panel 4: Residuals vs Observation Order
plot(1:length(resid(m1a)), resid(m1a), 
     xlab = "Observation Order", ylab = "Residuals",
     main = "Residuals vs Order", pch = 16, col = rgb(0, 0, 0, 0.5))
abline(h = 0, col = "red", lty = 2, lwd = 2)
lines(lowess(1:length(resid(m1a)), resid(m1a)), col = "blue", lwd = 2)
```

### Check 3: Normality (Shapiro-Wilk)
```{r}
#| label: primary-shapiro

shapiro.test(resid(m1a))
```

**Interpretation:** Residuals were approximately normally distributed (Shapiro-Wilk: W = 0.968, p = 0.081), indicating no substantial violations of normality assumptions.

### Check 4: Homogeneity of Variance (Levene's Test)
```{r}
#| label: primary-levene

# Group fitted values into low/high for variance comparison
fitted_groups <- cut(fitted(m1a), 
                     breaks = quantile(fitted(m1a), probs = c(0, 0.5, 1)),
                     include.lowest = TRUE)

# Levene's test for homoscedasticity
levene_result <- leveneTest(resid(m1a) ~ fitted_groups)

levene_result
```

**Interpretation:** Levene's test indicated homogeneity of variance (F(1,65) = 0.95, p = 0.333). While the Scale-Location plot showed a minor U-pattern, residual variance did not significantly differ across fitted value groups.

### Check 5: Random Effects Normality
```{r}
#| label: primary-random-effects
#| fig-width: 10
#| fig-height: 5

ranef_m1a <- ranef(m1a)$participant_id

par(mfrow = c(1, 2))

hist(ranef_m1a$`(Intercept)`, main = "Random Intercepts",
     xlab = "Random Intercept", col = "lightblue", border = "black", breaks = 8)

qqnorm(ranef_m1a$`(Intercept)`, main = "Q-Q Plot: Random Intercepts",
       pch = 16, col = rgb(0, 0, 0, 0.6))
qqline(ranef_m1a$`(Intercept)`, col = "red", lwd = 2)

par(mfrow = c(1, 1))

shapiro.test(ranef_m1a$`(Intercept)`)
```

### Check 6: Influential Observations
```{r}
#| label: primary-influential

# Calculate Cook's Distance
infl_m1a <- influence(m1a, obs = TRUE)
cooks_d_m1a <- cooks.distance(infl_m1a)

# Conservative threshold (same as Melatonin analysis)
threshold_m1a <- 4/nrow(kss_lmm)

# Identify influential observations
influential_idx <- which(cooks_d_m1a > threshold_m1a)

cat("Number of influential observations:", length(influential_idx), 
    "(", round(100*length(influential_idx)/nrow(kss_lmm), 0), "% of data)\n")
```
```{r}
# Show influential observations
if (length(influential_idx) > 0) {
  kss_lmm[influential_idx, c("participant_id", "exp_session", "mean_kss")]
}
```

#### Cook's Distance Plot
```{r}
#| label: primary-cooks-plot
#| fig-width: 8
#| fig-height: 5

plot(cooks_d_m1a, type = "h", 
     main = "Cook's Distance - KSS Model m1a",
     ylab = "Cook's D", lwd = 2,
     col = ifelse(cooks_d_m1a > threshold_m1a, "red", "black"))
abline(h = threshold_m1a, lty = 2, col = "red")
abline(h = 1.0, lty = 2, col = "darkred", lwd = 2)  
text(x = length(cooks_d_m1a) * 0.8, y = threshold_m1a + 0.02, 
     labels = paste0("4/n = ", round(threshold_m1a, 3)), 
     col = "red", cex = 0.9)
text(x = length(cooks_d_m1a) * 0.8, y = 1.05, 
     labels = "Critical threshold = 1.0", 
     col = "darkred", cex = 0.9)
```


### Check 7: Temporal Autocorrelation

Given the repeated measurements structure (10-13 consecutive evenings per participant), residuals were tested for temporal autocorrelation.
```{r}
#| label: primary-autocorrelation
#| fig-width: 10
#| fig-height: 5

par(mfrow = c(1, 2))

# Autocorrelation function
acf(resid(m1a), main = "ACF: Residuals", lag.max = 13)

# Partial autocorrelation function  
pacf(resid(m1a), main = "PACF: Residuals", lag.max = 13)
```

### Check 7*: Ljung-Box Test
```{r}
#| label: primary-ljung-box

# Ljung-Box test for overall autocorrelation
ljung_box_result <- Box.test(resid(m1a), lag = 10, type = "Ljung-Box")

print(ljung_box_result)
```

**Interpretation:**: The ACF and PACF plots showed one minor deviation at lag 6, but the Ljung-Box test (lag = 10) indicated no significant overall autocorrelation (p > 0.05), suggesting that the independence assumption for the residuals is adequately met.

### Check 8: DHARMa Residual Diagnostics
```{r}
sim_m1a <- simulateResiduals(m1a, n = 1000)
plot(sim_m1a)

# Individual tests
testUniformity(sim_m1a)
testDispersion(sim_m1a)
testOutliers(sim_m1a)
testQuantiles(sim_m1a)
```

----

### Diagnostic Interpretation
```{r}
#| label: diagnostic-interpretation

# Shapiro-Wilk residuals
shapiro_resid <- shapiro.test(resid(m1a))

# Levene test
fitted_groups <- cut(fitted(m1a), 
                     breaks = quantile(fitted(m1a), probs = c(0, 0.5, 1)), 
                     include.lowest = TRUE)
levene_result <- leveneTest(resid(m1a) ~ fitted_groups)

# Random effects normality
ranef_m1a <- ranef(m1a)$participant_id
shapiro_re <- shapiro.test(ranef_m1a$`(Intercept)`)

# Influential observations
infl_m1a <- influence(m1a, obs = TRUE)
influential_idx <- which(cooks.distance(infl_m1a) > 4/nrow(kss_lmm))
```

**Diagnostics Summary:**

1. **Multicollinearity:** VIF = 1.0 (no collinearity)
2. **Normality (residuals):** W = `r round(shapiro_resid$statistic, 3)`, p = `r round(shapiro_resid$p.value, 3)` 
3. **Homoscedasticity:** F = `r round(levene_result$'F value'[1], 3)`, p = `r round(levene_result$'Pr(>F)'[1], 3)`
4. **Random effects:** W = `r round(shapiro_re$statistic, 3)`, p = `r round(shapiro_re$p.value, 3)` 
5. **Influential obs:** `r length(influential_idx)` (`r round(100*length(influential_idx)/nrow(kss_lmm), 1)`%) 
6. **Temporal autocorrelation:** No sigificant overall autocorrelation detected 

----

## Address Visual Heteroscedasticity: Weighted Model
```{r}
#| label: weighted-model

m1a_weighted <- lme(
  mean_kss ~ Light_Lag0_num,
  random = ~ 1 | participant_id,
  weights = varPower(form = ~ fitted(.)),
  data = kss_lmm,
  method = "ML"
)

summary(m1a_weighted)

# Wald Confidence Intervals (manually calculated)
fixed_eff <- summary(m1a_weighted)$tTable
beta <- fixed_eff[, "Value"]
se <- fixed_eff[, "Std.Error"]

ci_m1a_weighted <- data.frame(
  `2.5 %` = beta - 1.96 * se,
  `97.5 %` = beta + 1.96 * se,
  check.names = FALSE
)
rownames(ci_m1a_weighted) <- names(beta)

cat("\n95% Wald Confidence Intervals:\n")
print(ci_m1a_weighted)
```

### Weighted Diagnostics
```{r}
#| label: weighted-diagnostics
#| fig-width: 10
#| fig-height: 5

par(mfrow = c(1, 2))

resid_w <- resid(m1a_weighted, type = "pearson")
fitted_w <- fitted(m1a_weighted)

plot(fitted_w, resid_w, xlab = "Fitted", ylab = "Pearson Residuals",
     main = "Weighted: Residuals vs Fitted", pch = 16, col = rgb(0,0,0,0.5))
abline(h = 0, col = "red", lty = 2, lwd = 2)
lines(lowess(fitted_w, resid_w), col = "blue", lwd = 2)

plot(fitted_w, sqrt(abs(resid_w)), xlab = "Fitted", ylab = "√|Pearson Residuals|",
     main = "Weighted: Scale-Location", pch = 16, col = rgb(0,0,0,0.5))
lines(lowess(fitted_w, sqrt(abs(resid_w))), col = "blue", lwd = 2)

# Tests
shapiro.test(resid_w)

fitted_groups_w <- cut(fitted_w, breaks = quantile(fitted_w, probs = c(0, 0.5, 1)), include.lowest = TRUE)
leveneTest(resid_w ~ fitted_groups_w)
```
### Comparison: Unweighted vs Weighted
```{r}
#| label: weighted-comparison

coefs_uw <- summary(m1a)$coefficients
coefs_w <- summary(m1a_weighted)$tTable

# Extract only the fixed effects that exist in the model (excluding intercept)
fixed_names <- rownames(coefs_uw)[rownames(coefs_uw) != "(Intercept)"]

data.frame(
  Effect = fixed_names,
  Beta_UW = coefs_uw[fixed_names, "Estimate"],
  SE_UW = coefs_uw[fixed_names, "Std. Error"],
  p_UW = coefs_uw[fixed_names, "Pr(>|t|)"],
  Beta_W = coefs_w[fixed_names, "Value"],
  SE_W = coefs_w[fixed_names, "Std.Error"],
  p_W = coefs_w[fixed_names, "p-value"]
) %>%
  kable(digits = 3, caption = "Unweighted vs Weighted Comparison") %>%
  kable_styling(full_width = FALSE)

cat("\nAIC: Unweighted =", round(AIC(m1a), 2), "| Weighted =", round(AIC(m1a_weighted), 2))
```

Although the weighted model showed marginally better fit (ΔAIC = -1.41) and 
improved diagnostic test results (Levene p: 0.351 → 0.582), the substantive 
conclusions remain unchanged:

- **Primary effect (Lag0)** is robust: β = -0.752 vs -0.683 (Δ = 0.069)
- **Both models**: Lag0 highly significant (p < 0.001), Lag1/Lag2 non-significant
- Statistical tests indicated homoscedasticity in unweighted model (Levene p = 0.351)

**Decision**: Unweighted model retained as primary for simplicity. 
Results are robust to weighting specification.

### Sensitivity Analysis: Model m1a Without Influential Observations 
```{r}
#| label: sensitivity-robust-model

# Identify influential observations
infl_m1a <- influence(m1a, obs = TRUE)
cooks_d_m1a <- cooks.distance(infl_m1a)
threshold_m1a <- 4 / nrow(kss_lmm)
influential_idx <- which(cooks_d_m1a > threshold_m1a)

# Show influential observations
kss_lmm[influential_idx, c("participant_id", "exp_session", "mean_kss")]

# Create robust dataset
kss_lmm_robust <- kss_lmm[-influential_idx, ]

# Fit robust model
m1a_robust <- lmer(
  mean_kss ~ Light_Lag0_num + (1 | participant_id),
  data = kss_lmm_robust,
  REML = FALSE
)

summary(m1a_robust)

# Wald Confidence Intervals
ci_m1a_robust <- confint(m1a_robust, method = "Wald")
print(ci_m1a_robust)
```

Altough AIC has been massively improved, significance and direction of the main effect (Lag 0) was maintained. As both influential observations represent biologically plausible values, model m1a is retained as the primary model, keeping in mind, that those two influential observations might increase the strength of the bright light exposure on sleepiness reduction, however they are not the only observations driving this effect. The robust model will be reported in a summary table for transparency. 

## Primary Model m1a: Results

### Fixed Effects with Benjamini-Hochberg Correction for Multiple Testing
```{r}
#| label: primary-results-bh

# Extract fixed effects (exclude intercept)
coefs_m1a <- summary(m1a)$coefficients
fixed_names <- rownames(coefs_m1a)[rownames(coefs_m1a) != "(Intercept)"]

# Apply BH correction
p_raw <- coefs_m1a[fixed_names, "Pr(>|t|)"]
p_BH <- p.adjust(p_raw, method = "BH")

# Results table
data.frame(
  Effect = fixed_names,
  Estimate = coefs_m1a[fixed_names, "Estimate"],
  SE = coefs_m1a[fixed_names, "Std. Error"],
  df = round(coefs_m1a[fixed_names, "df"], 1),
  t = coefs_m1a[fixed_names, "t value"],
  p_raw = p_raw,
  p_BH = p_BH
) %>%
  kable(digits = 3, row.names = FALSE,
        caption = "Primary Model Fixed Effects (Benjamini-Hochberg corrected)") %>%
  kable_styling(full_width = FALSE)
```

**Interpretation:** Current evening light exposure (Lag0) significantly reduced subjective sleepiness (β = -0.752, SE = 0.212, p < 0.001 after BH correction), with bright light evenings associated with lower KSS scores. No statistically significant carryover effects from previous evenings were detected (Lag1: p = 0.551; Lag2: p = 0.551 after BH correction).

### Variance Components
```{r}
#| label: primary-variance

vc <- as.data.frame(VarCorr(m1a))
icc <- vc$vcov[1] / sum(vc$vcov)

data.frame(
  Component = c("Between-participant (τ²)", "Within-participant (σ²)", "Total", "ICC"),
  Value = c(vc$vcov[1], vc$vcov[2], sum(vc$vcov), icc)
) %>%
  kable(digits = 3, row.names = FALSE,
        caption = "Variance Components") %>%
  kable_styling(full_width = FALSE)
```

**Interpretation:** Between-participant variance accounted for `r paste0(round(icc*100, 1), "%")` of total variance in KSS, indicating substantial differences in individual sleepiness levels.

## Exploratory Analyses

### Exploratory Analysis 1: Random Slopes for Light_Lag0

Test whether individual variation in the acute light effect improves model fit.
```{r}
#| label: sens-random-slope

# Attempt random slope model
m1a_random_slope <- lmer(
  mean_kss ~ Light_Lag0_num + 
    (1 + Light_Lag0_num | participant_id),
  data = kss_lmm,
  REML = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

summary(m1a_random_slope)

# Wald Confidence Intervals
ci_m1a_random_slope <- confint(m1a_random_slope, method = "Wald")
print(ci_m1a_random_slope)

# Model comparison
anova(m1a, m1a_random_slope)
```

**Interpretation:** The random slope model for Light_Lag0 did not significantly improve fit (χ²(2) = 3.97, p = 0.15), despite converging successfully using the bobyqa optimizer. The random slope model showed similar parsimony and increased standard errors for the primary effect (SE: 0.212 → 0.311), reducing statistical power. With only 8 participants, there is insufficient data to reliably estimate individual variation in light sensitivity. 

**Technical note:** The estimated random slope variance (0.48) suggests some individual differences in light sensitivity may exist, but this cannot be reliably quantified with the current sample size.

----

### Exploratory Analysis 2: Demographic Moderation

Important Caveat: This analysis is exploratory due to sample size limitations (n = 8 participants: 3 males, 5 females). Between-subject effects cannot be reliably estimated with small groups. Results should be interpreted with  caution.
```{r}
#| label: exploratory-moderation

# Center age
kss_lmm <- kss_lmm %>%
  mutate(age_c = age - mean(age, na.rm = TRUE))

# Moderation model
m1a_moderation <- lmer(
  mean_kss ~ Light_Lag0_num +
    sex + age_c + Light_Lag0_num:sex + Light_Lag0_num:age_c +
    (1 | participant_id),
  data = kss_lmm,
  REML = FALSE
)

summary(m1a_moderation)

# Wald Confidence Intervals
ci_m1a_moderation <- confint(m1a_moderation, method = "Wald")
print(ci_m1a_moderation)
```

**Interpretation:**

- **Sex × Light interaction:** Female participants showed a stronger acute light effect on KSS ratings (β = −1.12, p < 0.001). For male participants, the corresponding effect was substantially weaker (β = −1.12 + 0.98 = −0.14). The interaction term was statistically significant (β = 0.98, p = 0.041). However, this moderation effect must be interpreted with extreme caution given the very small subgroup sizes (5 females vs. 3 males).

- **Age main effect:** Higher age was associated with lower sleepiness (β = −0.24 per year of age, p = 0.040). There was no evidence that age moderated the acute light effect (β = −0.02, p = 0.798).

**Conclusion:** Although the model indicates potential moderation by sex and a main effect of age, the between-subject sample size (n = 8) is far too small to draw robust conclusions. 

### Exploratory Analysis 3: Light Lag Interactions

Test whether light exposure lags interact with each other.
```{r}
#| label: sens-light-interactions

m1a_interactions <- lmer(
  mean_kss ~ Light_Lag0_num * Light_Lag1_num + 
    Light_Lag0_num * Light_Lag2_num +
    Light_Lag1_num * Light_Lag2_num +
    (1 | participant_id),
  data = kss_lmm,
  REML = FALSE
)

summary(m1a_interactions)

# Wald Confidence Intervals
ci_m1a_interactions <- confint(m1a_interactions, method = "Wald")
print(ci_m1a_interactions)
```

**Interpretation:** No significant interactions between light exposure lags were detected (all p > 0.05). Marginal patterns for Lag2 (p = 0.052) and Lag0 × Lag2 (p = 0.068) are present, suggesting potential patterns for the seconday day before current day´s light exposure. However such patterns need to be explored using a larger sample size to draw more solid conclusions. Model fit worsened (ΔAIC = +5.1).


### Exploratory Analysis 4: Persistence Models 

#### Rationale for Persistence Analysis

The primary LMM analysis demonstrated that evening light exposure (Lag 0) significantly reduces subjective sleepiness, while prior exposures (Lag 1, Lag 2) showed no statistically significant carryover effects. However, this lag-based approach tests whether *individual prior nights* contribute additively to the current night's sleepiness levels.

The **persistence analysis** addresses the question: Does the *number of consecutive days* of the same light condition (run length) modulate subjective sleepiness? 

**Key distinction from lag analysis:**

- Lag analysis: Tests whether Day[t-1], Day[t-2], Day[t-3] each add independent effects
- Persistence analysis: Tests whether being on the 1st vs. 2nd vs. 3rd+ consecutive day of the *same condition* matters

This exploratory analysis uses separate models for dim and bright conditions, testing both linear (numeric run length) and non-linear (categorical run length) accumulation patterns.

#### Dataset Preparation 

```{r}
#| label: create-persistence-dataset

# Define participant list and light patterns
participant_list <- c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10")

# Create persistence dataset starting from session 1
kss_persistence <- map_dfr(participant_list, ~{
  df <- kss_per_evening_q789 %>%
    filter(participant_id == .x) %>%
    select(participant_id, exp_session, mean_kss)
  
  light_df <- tibble(
    exp_session = seq_along(light_patterns[[.x]]),
    Light_Lag0 = light_patterns[[.x]]
  )
  
  df %>% left_join(light_df, by = "exp_session")
})

# Add demographics
participant_demographics <- tibble(
  participant_id = c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10"),
  sex = c("female", "male", "female", "female", "male", "female", "female", "male"),
  age = c(24, 21, 24, 23, 31, 23, 28, 32)
)

kss_persistence <- kss_persistence %>%
  left_join(participant_demographics, by = "participant_id")

# Calculate run-length variables (numeric)
kss_persistence <- kss_persistence %>%
  filter(!is.na(Light_Lag0)) %>%  # Remove rows without light condition
  filter(!(participant_id == "P10" & exp_session == 5)) %>%  # Exclude P10 session 5
  arrange(participant_id, exp_session) %>%
  group_by(participant_id) %>%
  mutate(
    # Track when condition changes
    condition_change = Light_Lag0 != lag(Light_Lag0, default = "CHANGE"),
    
    # Calculate run ID (new sequence after each change)
    run_id = cumsum(condition_change),
    
    # Calculate run length within each run
    run_length = ave(run_id, run_id, FUN = seq_along),
    
    # Separate run lengths for Dim and Bright
    run_length_dim = ifelse(Light_Lag0 == "D", run_length, NA),
    run_length_bright = ifelse(Light_Lag0 == "B", run_length, NA),
    
    # Also create numeric Light variable for consistency
    Light_Lag0_num = ifelse(Light_Lag0 == "D", 0, 1)
  ) %>%
  ungroup() %>%
  select(participant_id, exp_session, mean_kss, Light_Lag0, Light_Lag0_num,
         run_length, run_length_dim, run_length_bright, sex, age)

# Create categorical run-length variables
kss_persistence <- kss_persistence %>%
  mutate(
    run_category_dim = case_when(
      run_length_dim == 1 ~ "First",
      run_length_dim == 2 ~ "Second",
      run_length_dim >= 3 ~ "Third",
      TRUE ~ NA_character_
    ),
    run_category_bright = case_when(
      run_length_bright == 1 ~ "First",
      run_length_bright == 2 ~ "Second",
      run_length_bright >= 3 ~ "Third",
      TRUE ~ NA_character_
    )
  ) %>%
  mutate(
    run_category_dim = factor(run_category_dim, levels = c("First", "Second", "Third")),
    run_category_bright = factor(run_category_bright, levels = c("First", "Second", "Third"))
  )
```

#### Persistence Dataset
```{r}
#| label: kss-persistence-dataset-table

DT::datatable(
  kss_persistence,
  options = list(
    pageLength = 20,
    scrollX = TRUE,
    order = list(list(0, 'asc'), list(1, 'asc'))
  ),
  caption = "KSS Persistence dataset with run-length encoding (n=67, P10 session 5 excluded)",
  rownames = FALSE
) %>%
  DT::formatRound(columns = 'mean_kss', digits = 2) %>%
  DT::formatStyle(
    'Light_Lag0',
    backgroundColor = DT::styleEqual(c('D', 'B'), c('#e3f2fd', '#fff3e0'))
  )
```

#### Descriptive Visualizations
```{r}
#| label: kss-persistence-descriptive-visualizations

# Define color scheme
color_dim <- "#4A90E2"    # Blue for Dim
color_bright <- "#F5A623" # Orange for Bright

# Prepare data WITH Session 1
plot_data_dim_s1 <- kss_persistence %>%
  filter(!is.na(run_length_dim)) %>%
  mutate(run_length_dim = factor(run_length_dim))

plot_data_bright_s1 <- kss_persistence %>%
  filter(!is.na(run_length_bright)) %>%
  mutate(run_length_bright = factor(run_length_bright))

# Prepare data WITHOUT Session 1
plot_data_dim_no_s1 <- kss_persistence %>%
  filter(!is.na(run_length_dim), exp_session > 1) %>%
  mutate(run_length_dim = factor(run_length_dim))

plot_data_bright_no_s1 <- kss_persistence %>%
  filter(!is.na(run_length_bright), exp_session > 1) %>%
  mutate(run_length_bright = factor(run_length_bright))

# Create all 4 plots
p_dim_s1 <- ggplot(plot_data_dim_s1, aes(x = run_length_dim, y = mean_kss)) +
  geom_boxplot(fill = color_dim, outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16) +
  labs(title = "Dim Nights (with Session 1)", subtitle = paste0("n = ", nrow(plot_data_dim_s1)),
       x = "Consecutive Nights", y = "Mean KSS") +
  theme_classic(base_size = 14) +
  theme(aspect.ratio = 1, plot.title = element_text(size = 16, hjust = 0),
        plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
        axis.text = element_text(color = "black", size = 13),
        axis.title = element_text(size = 14),
        panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
        legend.position = "none")

p_bright_s1 <- ggplot(plot_data_bright_s1, aes(x = run_length_bright, y = mean_kss)) +
  geom_boxplot(fill = color_bright, outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16) +
  labs(title = "Bright Nights (with Session 1)", subtitle = paste0("n = ", nrow(plot_data_bright_s1)),
       x = "Consecutive Nights", y = "Mean KSS") +
  theme_classic(base_size = 14) +
  theme(aspect.ratio = 1, plot.title = element_text(size = 16, hjust = 0),
        plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
        axis.text = element_text(color = "black", size = 13),
        axis.title = element_text(size = 14),
        panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
        legend.position = "none")

p_dim_no_s1 <- ggplot(plot_data_dim_no_s1, aes(x = run_length_dim, y = mean_kss)) +
  geom_boxplot(fill = color_dim, outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16) +
  labs(title = "Dim Nights (without Session 1)", subtitle = paste0("n = ", nrow(plot_data_dim_no_s1)),
       x = "Consecutive Nights", y = "Mean KSS") +
  theme_classic(base_size = 14) +
  theme(aspect.ratio = 1, plot.title = element_text(size = 16, hjust = 0),
        plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
        axis.text = element_text(color = "black", size = 13),
        axis.title = element_text(size = 14),
        panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
        legend.position = "none")

p_bright_no_s1 <- ggplot(plot_data_bright_no_s1, aes(x = run_length_bright, y = mean_kss)) +
  geom_boxplot(fill = color_bright, outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16) +
  labs(title = "Bright Nights (without Session 1)", subtitle = paste0("n = ", nrow(plot_data_bright_no_s1)),
       x = "Consecutive Nights", y = "Mean KSS") +
  theme_classic(base_size = 14) +
  theme(aspect.ratio = 1, plot.title = element_text(size = 16, hjust = 0),
        plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
        axis.text = element_text(color = "black", size = 13),
        axis.title = element_text(size = 14),
        panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
        legend.position = "none")

# Combine all plots in 2x2 grid
(p_dim_s1 + p_bright_s1) / (p_dim_no_s1 + p_bright_no_s1)
```

#### Model 1: Linear Persistence, Models, Dim Evenings 
```{r}
#| label: kss-persistence-linear-dim

# WITH Session 1
kss_persistence_dim <- kss_persistence %>%
  filter(!is.na(run_length_dim))

model_persistence_dim_numeric <- lmer(
  mean_kss ~ run_length_dim + (1 | participant_id),
  data = kss_persistence_dim,
  REML = FALSE
)

summary(model_persistence_dim_numeric)

# Wald Confidence Intervals
ci_model_persistence_dim_numeric <- confint(model_persistence_dim_numeric, method = "Wald")
print(ci_model_persistence_dim_numeric)
```

```{r}
# WITHOUT Session 1
kss_persistence_dim_no_s1 <- kss_persistence %>%
  filter(!is.na(run_length_dim), exp_session > 1)

model_persistence_dim_no_s1 <- lmer(
  mean_kss ~ run_length_dim + (1 | participant_id),
  data = kss_persistence_dim_no_s1,
  REML = FALSE
)

summary(model_persistence_dim_no_s1)

# Wald Confidence Intervals
ci_model_persistence_dim_no_s1 <- confint(model_persistence_dim_no_s1, method = "Wald")
print(ci_model_persistence_dim_no_s1)
```

#### Model 1: Linear Persistence Models, Bright Evenings 

```{r}
#| label: kss-persistence-linear-bright

# WITH Session 1
kss_persistence_bright <- kss_persistence %>%
  filter(!is.na(run_length_bright))

model_persistence_bright_numeric <- lmer(
  mean_kss ~ run_length_bright + (1 | participant_id),
  data = kss_persistence_bright,
  REML = FALSE
)

summary(model_persistence_bright_numeric)

# Wald Confidence Intervals
ci_model_persistence_bright_numeric <- confint(model_persistence_bright_numeric, method = "Wald")
print(ci_model_persistence_bright_numeric)
```

```{r}
# WITHOUT Session 1
kss_persistence_bright_no_s1 <- kss_persistence %>%
  filter(!is.na(run_length_bright), exp_session > 1)

model_persistence_bright_no_s1 <- lmer(
  mean_kss ~ run_length_bright + (1 | participant_id),
  data = kss_persistence_bright_no_s1,
  REML = FALSE
)

summary(model_persistence_bright_no_s1)

# Wald Confidence Intervals
ci_model_persistence_bright_no_s1 <- confint(model_persistence_bright_no_s1, method = "Wald")
print(ci_model_persistence_bright_no_s1)
```

#### Model 2: Categorical Persistence Models, Dim Evenings 
```{r}
#| label: kss-persistence-categorical-dim

# WITH Session 1
kss_persistence_dim <- kss_persistence %>%
  filter(!is.na(run_category_dim))

model_persistence_dim_categorical <- lmer(
  mean_kss ~ run_category_dim + (1 | participant_id),
  data = kss_persistence_dim,
  REML = FALSE
)

summary(model_persistence_dim_categorical)

# Wald Confidence Intervals
ci_model_persistence_dim_categorical <- confint(model_persistence_dim_categorical, method = "Wald")
print(ci_model_persistence_dim_categorical)

lrt_dim_s1 <- anova(model_persistence_dim_numeric, model_persistence_dim_categorical)
print(lrt_dim_s1)
```

```{r}
lrt_dim_s1 <- anova(model_persistence_dim_numeric, model_persistence_dim_categorical)
print(lrt_dim_s1)
```

```{r}
# WITHOUT Session 1
kss_persistence_dim_no_s1 <- kss_persistence %>%
  filter(!is.na(run_category_dim), exp_session > 1)

model_persistence_dim_categorical_no_s1 <- lmer(
  mean_kss ~ run_category_dim + (1 | participant_id),
  data = kss_persistence_dim_no_s1,
  REML = FALSE
)

summary(model_persistence_dim_categorical_no_s1)

# Wald Confidence Intervals
ci_model_persistence_dim_categorical_no_s1 <- confint(model_persistence_dim_categorical_no_s1, method = "Wald")
print(ci_model_persistence_dim_categorical_no_s1)
```

```{r}
lrt_dim_no_s1 <- anova(model_persistence_dim_no_s1, model_persistence_dim_categorical_no_s1)
print(lrt_dim_no_s1)
```

#### Model 2: Categorical Persistence Models, Bright Evenings 
```{r}
#| label: kss-persistence-categorical-bright-complete

# WITH Session 1
kss_persistence_bright <- kss_persistence %>%
  filter(!is.na(run_category_bright))

model_persistence_bright_categorical <- lmer(
  mean_kss ~ run_category_bright + (1 | participant_id),
  data = kss_persistence_bright,
  REML = FALSE
)

summary(model_persistence_bright_categorical)

# Wald Confidence Intervals
ci_model_persistence_bright_categorical <- confint(model_persistence_bright_categorical, method = "Wald")
print(ci_model_persistence_bright_categorical)
```

```{r}
lrt_bright_s1 <- anova(model_persistence_bright_numeric, model_persistence_bright_categorical)
print(lrt_bright_s1)
```

```{r}
# WITHOUT Session 1
kss_persistence_bright_no_s1 <- kss_persistence %>%
  filter(!is.na(run_category_bright), exp_session > 1)

model_persistence_bright_categorical_no_s1 <- lmer(
  mean_kss ~ run_category_bright + (1 | participant_id),
  data = kss_persistence_bright_no_s1,
  REML = FALSE
)

summary(model_persistence_bright_categorical_no_s1)

# Wald Confidence Intervals
ci_model_persistence_bright_categorical_no_s1 <- confint(model_persistence_bright_categorical_no_s1, method = "Wald")
print(ci_model_persistence_bright_categorical_no_s1)
```

```{r}
lrt_bright_no_s1 <- anova(model_persistence_bright_no_s1, model_persistence_bright_categorical_no_s1)
print(lrt_bright_no_s1)
```

## Summary and Interpretation

**Main Findings:**

**Linear Models:**

1. **Dim evenings (with Session 1):** No significant persistence effect (β = +0.090, p = 0.650).
2. **Dim evenings (without Session 1):** No significant linear persistence effect (β = +0.064, p = 0.718).
3. **Bright evenings (both analyses):** No persistence effect (β = +0.117, p = 0.393), suggesting alerting effects are maximal after the first bright evening.

**Categorical Models:**

Categorical models revealed non-linear patterns for dim evenings that linear models failed to detect. For dim evenings without Session 1, the categorical model showed significantly better fit than the linear model (LRT: χ² = 6.42, p = 0.011, ΔAIC = -4.4). The pattern revealed a U-shaped trajectory: sleepiness decreased on the second consecutive dim night (β = -0.571, p = 0.053, ~0.6-point reduction on KSS scale) but returned toward baseline by the third night (β = +0.309, p = 0.342). This suggests an initial alerting adaptation to dim light that diminishes with continued exposure.

For bright evenings, categorical models showed no improvement over linear models (LRT: χ² < 0.001, p = 0.980, ΔAIC = +2.0), confirming that bright light-induced alertness remains stable across consecutive nights.

**Conclusion:**

Unlike melatonin—which showed a marginal linear increase with consecutive dim nights—subjective sleepiness exhibits non-linear dynamics under repeated dim light exposure. The U-shaped pattern suggests short-term adaptation (reduced sleepiness on night 2) followed by habituation (return to baseline by night 3+). In contrast, bright light effects on subjective alertness saturate immediately and show no cumulative changes, mirroring the melatonin suppression pattern. These exploratory findings raise the possibility of differential physiological versus subjective adaptation mechanisms, but current data are insufficient for causal inference. 


## Summary of Exploratory Analyses

The primary model (m1a: Light_Lag0_num) with Lag 0 as fixed effect and random intercepts was selected as the primary model. Sensitivity analyses support the robustness of the main findings:

* **Random slopes:** Model fit was not significantly improved (χ²(2) = 3.79, p = 0.15, ΔAIC = +0.2), indicating that while some individual differences in light sensitivity may exist (random slope variance = 0.47), the current sample size (n=8) is insufficient to reliably quantify this variability. The average alerting effect remained significant but with reduced precision (β = -0.76, SE = 0.31, p = 0.040 vs. primary β = -0.75, SE = 0.21, p < 0.001).

* **Demographic moderation:** Model fit was not meaningfully improved (ΔAIC = -1.8), but exploratory sex × light interaction emerged (β = +0.98, p = 0.041), suggesting potentially stronger acute light effects in females (β = -1.12) than males (β = -0.14). However, with only 3 males and 5 females, this finding requires replication in adequately powered samples before interpretation. Age showed a main effect (β = -0.24, p = 0.040) but no moderation of light sensitivity (p = 0.798).

* **Light lag interactions:** Not supported (ΔAIC = +5.1, all interaction p > 0.05), confirming independent, non-additive effects across days. Marginal patterns for Lag2 main effect (p = 0.052) and Lag0 × Lag2 interaction (p = 0.068) suggest potential effects of light exposure two days prior, but these require larger samples for confirmation.

* **Persistence model:** Exploratory analysis (n = 37-45 observations, 8 participants) tested whether consecutive exposure to the same light condition produces cumulative effects. Linear models showed no persistence effect for either dim nights (β = +0.064, p = 0.718) or bright nights (β = +0.117, p = 0.393). However, categorical models revealed non-linear patterns for dim evenings (LRT: χ² = 6.42, p = 0.011 without Session 1, ΔAIC = -4.4), with a U-shaped trajectory suggesting initial alerting adaptation (second night: β = -0.57, p = 0.053) that diminishes with continued exposure (third night: β = +0.31, p = 0.342). For bright evenings, categorical models showed no improvement (LRT: χ² < 0.001, p = 0.980, ΔAIC = +2.0). These findings are strictly hypothesis-generating. Future adequately powered studies (N ≥ 30, 4-5+ consecutive nights) are needed to test whether non-linear persistence dynamics reflect genuine adaptation processes.

**Final model:** Model m1a (n=67) with fixed effect for Light_Lag0 and random intercepts only.

----

## Sensitivity Analysis: Alternative Measurement Window (Q8-Q10)

### Rationale

To test whether results are sensitive to the choice of measurement window, KSS scores were also averaged across questionnaires 8-10 (HBT-2h to HBT-1h) instead of 7-9 (as used as the dependent variable within the primary model)

```{r}
#| label: sensitivity-q8-10-complete-analysis

# Calculate mean KSS for Q8-Q10 window
kss_per_evening_q8_10 <- insess_surv %>%
  filter(ques_numb %in% c(8, 9, 10)) %>%
  mutate(
    session_raw = as.integer(str_extract(session, "\\d+$")),
    exp_session = case_when(
      id %in% c("P01", "P02", "P03") ~ session_raw - 100,
      id %in% c("P04", "P05") ~ session_raw - 200,
      id %in% c("P06", "P07", "P09") ~ session_raw - 300,
      id == "P10" ~ session_raw - 400,
      TRUE ~ NA_real_
    )
  ) %>%
  group_by(id, exp_session, session) %>%
  summarise(
    mean_kss = mean(kss, na.rm = TRUE),
    n_measurements = n(),
    .groups = "drop"
  ) %>%
  rename(participant_id = id)

# Prepare LMM dataset with Q8-Q10 data
kss_lmm_q8_10 <- map_dfr(participant_list, function(pid) {
  df <- kss_per_evening_q8_10 %>%
    filter(participant_id == pid) %>%
    select(participant_id, exp_session, mean_kss)
  
  light_df <- tibble(
    exp_session = seq_along(light_patterns[[pid]]),
    Light_Lag0 = light_patterns[[pid]]
  )
  
  df %>% left_join(light_df, by = "exp_session")
})

kss_lmm_q8_10 <- kss_lmm_q8_10 %>%
  group_by(participant_id) %>%
  arrange(exp_session) %>%
  mutate(
    Light_Lag1 = lag(Light_Lag0, 1),
    Light_Lag2 = lag(Light_Lag0, 2),
    Light_Lag0_num = ifelse(Light_Lag0 == "D", 0, 1),
    Light_Lag1_num = ifelse(Light_Lag1 == "D", 0, 1),
    Light_Lag2_num = ifelse(Light_Lag2 == "D", 0, 1)
  ) %>%
  ungroup() %>%
  filter(!is.na(Light_Lag0) & !is.na(Light_Lag1) & !is.na(Light_Lag2)) %>%
  filter(!(participant_id == "P10" & exp_session %in% c(6, 7))) %>%
  left_join(participant_demographics, by = "participant_id") %>%
  mutate(age_c = age - mean(age))

# Fit model (same structure to m1a)
m1a_q8_10 <- lmer(
  mean_kss ~ Light_Lag0_num + 
    (1 | participant_id),
  data = kss_lmm_q8_10,
  REML = FALSE
)

summary(m1a_q8_10)

# Fit model (same structure to m1b)
m1b_q8_10 <- lmer (
  mean_kss ~ Light_Lag0_num + Light_Lag1_num +
    (1 | participant_id),
  data = kss_lmm_q8_10,
  REML = FALSE
)

summary(m1b_q8_10)

# Fit model (same structure to m1c)
m1c_q8_10 <- lmer (
  mean_kss ~ Light_Lag0_num + Light_Lag1_num + Light_Lag2_num +
    (1 | participant_id),
  data = kss_lmm_q8_10,
  REML = FALSE
)

summary(m1c_q8_10)


# Extract coefficients and compare
coefs_q7_9 <- summary(m1a)$coefficients
coefs_q8_10 <- summary(m1a_q8_10)$coefficients

fixed_names <- rownames(coefs_q7_9)[rownames(coefs_q7_9) != "(Intercept)"]

p_raw_q8_10 <- coefs_q8_10[fixed_names, "Pr(>|t|)"]
p_BH_q8_10 <- p.adjust(p_raw_q8_10, method = "BH")

window_comparison <- data.frame(
  Effect = fixed_names,
  Q7_Q9_Beta = coefs_q7_9[fixed_names, "Estimate"],
  Q7_Q9_SE = coefs_q7_9[fixed_names, "Std. Error"],
  Q7_Q9_p_BH = p.adjust(coefs_q7_9[fixed_names, "Pr(>|t|)"], method = "BH"),
  Q8_Q10_Beta = coefs_q8_10[fixed_names, "Estimate"],
  Q8_Q10_SE = coefs_q8_10[fixed_names, "Std. Error"],
  Q8_Q10_p_BH = p_BH_q8_10,
  Delta_Beta = coefs_q8_10[fixed_names, "Estimate"] - coefs_q7_9[fixed_names, "Estimate"],
  Pct_Change = 100 * (coefs_q8_10[fixed_names, "Estimate"] - coefs_q7_9[fixed_names, "Estimate"]) / 
                abs(coefs_q7_9[fixed_names, "Estimate"])
)

window_comparison %>%
  kable(digits = 3, 
        caption = "Sensitivity Analysis: Q7-Q9 vs Q8-Q10 Measurement Windows",
        col.names = c("Effect", "β (Q7-Q9)", "SE", "p (BH)", 
                      "β (Q8-Q10)", "SE", "p (BH)", "Δβ", "% Change")) %>%
  kable_styling(full_width = FALSE)

# Variance components
vc_q7_9 <- as.data.frame(VarCorr(m1a))
vc_q8_10 <- as.data.frame(VarCorr(m1a_q8_10))

icc_q7_9 <- vc_q7_9$vcov[1] / sum(vc_q7_9$vcov)
icc_q8_10 <- vc_q8_10$vcov[1] / sum(vc_q8_10$vcov)
```

**Interpretation:** 

Both measurement windows confirmed highly significant acute light effects on subjective sleepiness (Q7-Q9: β = -0.752, p = 0.002; Q8-Q10: β = -1.011, p < 0.001), with no carryover effects from previous evenings (Lag1, Lag2: both n.s.). The Lag 0 effect was 34% stronger effect in Q8-Q10 measurement window compared to the Q7-9 measurement window.

**Conclusion:** The Q7-Q9 window is retained as primary. because it captures the full temporal trajectory of light-induced alertness from early through sustained responses, and represents a conservative estimate of light's alerting effect. The stronger effect when Q7 is excluded validates rather than challenges the inclusion of early responses in the primary analysis.

## Model Comparison: Overview

Comprehensive comparison of all tested models with key parameter estimates.
```{r}
#| label: kss-model-overview-extended

# Helper functions to extract coefficients from LMM
get_coef_lmm <- function(model, coef_name) {
  coefs <- summary(model)$coefficients
  if (coef_name %in% rownames(coefs)) {
    return(coefs[coef_name, "Estimate"])
  } else {
    return(NA)
  }
}

get_p_lmm <- function(model, coef_name) {
  coefs <- summary(model)$coefficients
  if (coef_name %in% rownames(coefs)) {
    return(coefs[coef_name, "Pr(>|t|)"])
  } else {
    return(NA)
  }
}

# For weighted model (lme)
get_coef_lme <- function(model, coef_name) {
  coefs <- summary(model)$tTable
  if (coef_name %in% rownames(coefs)) {
    return(coefs[coef_name, "Value"])
  } else {
    return(NA)
  }
}

get_p_lme <- function(model, coef_name) {
  coefs <- summary(model)$tTable
  if (coef_name %in% rownames(coefs)) {
    return(coefs[coef_name, "p-value"])
  } else {
    return(NA)
  }
}

# Create extended comparison table
model_comparison_extended <- data.frame(
  Model = c(
    "m0 (Baseline)",
    "m1a (Lag 0)",
    "m1b (Lag 0+1)",
    "m1c (Lag 0+1+2)",
    "m2 (+ Session)",
    "m3 (+ Session²)",
    "m1a_weighted",
    "m1a_robust",
    "Random Slopes",
    "Demographics",
    "Light Interactions",
    "Q8-Q10 Window",
    "Persistence Dim (S1)",
    "Persistence Dim (no S1)",
    "Persistence Bright (S1)",
    "Persistence Bright (no S1)"
  ),
  
  Description = c(
    "Intercept only",
    "Light Lag 0 only",
    "Lag 0 + Lag 1",
    "Lag 0 + Lag 1 + Lag 2",
    "Lag 0 + session (linear)",
    "Lag 0 + session + session²",
    "m1a with variance weights",
    "m1a without influential obs.",
    "m1a + random slope Lag0",
    "m1a + age + sex + interactions",
    "m1a + Light lag interactions",
    "m1a with Q8-Q10 measurement",
    "Dim: run_length (with Session 1)",
    "Dim: run_length (without Session 1)",
    "Bright: run_length (with Session 1)",
    "Bright: run_length (without Session 1)"
  ),
  
  npar = c(
    attr(logLik(m0), "df"),
    attr(logLik(m1a), "df"),
    attr(logLik(m1b), "df"),
    attr(logLik(m1c), "df"),
    attr(logLik(m2), "df"),
    attr(logLik(m3), "df"),
    attr(logLik(m1a_weighted), "df"),
    attr(logLik(m1a_robust), "df"),
    attr(logLik(m1a_random_slope), "df"),
    attr(logLik(m1a_moderation), "df"),
    attr(logLik(m1a_interactions), "df"),
    attr(logLik(m1a_q8_10), "df"),
    attr(logLik(model_persistence_dim_numeric), "df"),
    attr(logLik(model_persistence_dim_no_s1), "df"),
    attr(logLik(model_persistence_bright_numeric), "df"),
    attr(logLik(model_persistence_bright_no_s1), "df")
  ),
  
  N_obs = c(
    nrow(kss_lmm),
    nrow(kss_lmm),
    nrow(kss_lmm),
    nrow(kss_lmm),
    nrow(kss_lmm),
    nrow(kss_lmm),
    nrow(kss_lmm),
    nrow(kss_lmm_robust),
    nrow(kss_lmm),
    nrow(kss_lmm),
    nrow(kss_lmm),
    nrow(kss_lmm_q8_10),
    45,  # Dim with S1
    37,  # Dim without S1
    40,  # Bright with S1
    40   # Bright without S1
  ),
  
  AIC = c(
    round(AIC(m0), 1),
    round(AIC(m1a), 1),
    round(AIC(m1b), 1),
    round(AIC(m1c), 1),
    round(AIC(m2), 1),
    round(AIC(m3), 1),
    round(AIC(m1a_weighted), 1),
    round(AIC(m1a_robust), 1),
    round(AIC(m1a_random_slope), 1),
    round(AIC(m1a_moderation), 1),
    round(AIC(m1a_interactions), 1),
    round(AIC(m1a_q8_10), 1),
    round(AIC(model_persistence_dim_numeric), 1),
    round(AIC(model_persistence_dim_no_s1), 1),
    round(AIC(model_persistence_bright_numeric), 1),
    round(AIC(model_persistence_bright_no_s1), 1)
  ),
  
  BIC = c(
    round(BIC(m0), 1),
    round(BIC(m1a), 1),
    round(BIC(m1b), 1),
    round(BIC(m1c), 1),
    round(BIC(m2), 1),
    round(BIC(m3), 1),
    round(BIC(m1a_weighted), 1),
    round(BIC(m1a_robust), 1),
    round(BIC(m1a_random_slope), 1),
    round(BIC(m1a_moderation), 1),
    round(BIC(m1a_interactions), 1),
    round(BIC(m1a_q8_10), 1),
    round(BIC(model_persistence_dim_numeric), 1),
    round(BIC(model_persistence_dim_no_s1), 1),
    round(BIC(model_persistence_bright_numeric), 1),
    round(BIC(model_persistence_bright_no_s1), 1)
  ),
  
  Main_Beta = c(
    NA,
    round(get_coef_lmm(m1a, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m1b, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m1c, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m2, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m3, "Light_Lag0_num"), 3),
    round(get_coef_lme(m1a_weighted, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m1a_robust, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m1a_random_slope, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m1a_moderation, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m1a_interactions, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m1a_q8_10, "Light_Lag0_num"), 3),
    round(get_coef_lmm(model_persistence_dim_numeric, "run_length_dim"), 3),
    round(get_coef_lmm(model_persistence_dim_no_s1, "run_length_dim"), 3),
    round(get_coef_lmm(model_persistence_bright_numeric, "run_length_bright"), 3),
    round(get_coef_lmm(model_persistence_bright_no_s1, "run_length_bright"), 3)
  ),
  
  Effect_KSS = c(
    NA,
    paste0(round(abs(get_coef_lmm(m1a, "Light_Lag0_num")), 2), " pts"),
    paste0(round(abs(get_coef_lmm(m1b, "Light_Lag0_num")), 2), " pts"),
    paste0(round(abs(get_coef_lmm(m1c, "Light_Lag0_num")), 2), " pts"),
    paste0(round(abs(get_coef_lmm(m2, "Light_Lag0_num")), 2), " pts"),
    paste0(round(abs(get_coef_lmm(m3, "Light_Lag0_num")), 2), " pts"),
    paste0(round(abs(get_coef_lme(m1a_weighted, "Light_Lag0_num")), 2), " pts"),
    paste0(round(abs(get_coef_lmm(m1a_robust, "Light_Lag0_num")), 2), " pts"),
    paste0(round(abs(get_coef_lmm(m1a_random_slope, "Light_Lag0_num")), 2), " pts"),
    paste0(round(abs(get_coef_lmm(m1a_moderation, "Light_Lag0_num")), 2), " pts"),
    paste0(round(abs(get_coef_lmm(m1a_interactions, "Light_Lag0_num")), 2), " pts"),
    paste0(round(abs(get_coef_lmm(m1a_q8_10, "Light_Lag0_num")), 2), " pts"),
    paste0(round(get_coef_lmm(model_persistence_dim_numeric, "run_length_dim"), 2), " pts/day"),
    paste0(round(get_coef_lmm(model_persistence_dim_no_s1, "run_length_dim"), 2), " pts/day"),
    paste0(round(get_coef_lmm(model_persistence_bright_numeric, "run_length_bright"), 2), " pts/day"),
    paste0(round(get_coef_lmm(model_persistence_bright_no_s1, "run_length_bright"), 2), " pts/day")
  ),
  
  Main_p = c(
    NA,
    format.pval(get_p_lmm(m1a, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1b, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1c, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m2, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m3, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lme(m1a_weighted, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1a_robust, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1a_random_slope, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1a_moderation, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1a_interactions, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1a_q8_10, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(model_persistence_dim_numeric, "run_length_dim"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(model_persistence_dim_no_s1, "run_length_dim"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(model_persistence_bright_numeric, "run_length_bright"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(model_persistence_bright_no_s1, "run_length_bright"), digits = 3, eps = 0.001)
  ),
  
  Status = c(
    "Baseline",
    "FINAL MODEL",
    "Not selected",
    "Not selected",
    "Not selected",
    "Not selected",
    "Sensitivity",
    "Sensitivity",
    "Exploratory",
    "Exploratory",
    "Exploratory",
    "Sensitivity",
    "Exploratory",
    "Exploratory",
    "Exploratory",
    "Exploratory"
  )
)

kable(model_comparison_extended, 
      caption = "Model Comparison: All Tested Models with Key Parameters (KSS)",
      align = c('l', 'l', 'c', 'c', 'c', 'c', 'c', 'c', 'c', 'l'),
      col.names = c("Model", "Description", "npar", "N", "AIC", "BIC", 
                    "Main β", "Effect", "Main p", "Status")) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  row_spec(2, bold = TRUE, background = "#D6EAF8") %>%  # Highlight m1a (row 2) in BLUE
  pack_rows("Hierarchical Model Building (n=67)", 1, 6) %>%
  pack_rows("Sensitivity Analyses (n=65-67)", 7, 8) %>%
  pack_rows("Exploratory Analyses", 9, 16)
```

**Note:** npar = number of parameters; Lag0 β = acute light effect; p-values are raw (not BH-corrected for comparison purposes).

## Primary Model: Fixed Effects
```{r}
#| label: primary-results

# Extract and format coefficients
coefs_m1a <- summary(m1a)$coefficients
all_names <- rownames(coefs_m1a)

# Apply BH correction
p_raw_all <- coefs_m1a[, "Pr(>|t|)"]
p_BH_all <- p.adjust(p_raw_all, method = "BH")

# Simple results table
data.frame(
  Effect = all_names,
  Estimate = round(coefs_m1a[, "Estimate"], 3),
  SE = round(coefs_m1a[, "Std. Error"], 3),
  t = round(coefs_m1a[, "t value"], 3),
  p_raw = round(p_raw_all, 3),
  p_BH = round(p_BH_all, 3)
) %>%
  kable(caption = "Primary Model: Fixed Effects (BH-corrected)",
        col.names = c("Effect", "Estimate", "SE", "t", "p", "p (BH)")) %>%
  kable_styling(full_width = FALSE)

# Variance components
vc <- as.data.frame(VarCorr(m1a))
icc <- vc$vcov[1] / sum(vc$vcov)

cat("\n**Variance Components:**\n")
cat("Between-participant:", round(vc$vcov[1], 3), "\n")
cat("Within-participant:", round(vc$vcov[2], 3), "\n")
cat("ICC:", round(icc, 3), "\n\n")

cat("**Model Fit:**\n")
cat("AIC:", round(AIC(m1a), 2), "| BIC:", round(BIC(m1a), 2), "\n")
cat("N obs:", nobs(m1a), "| N groups:", length(unique(kss_lmm$participant_id)))
```

**Interpretation:** Current evening light exposure (Lag0) significantly reduced subjective sleepiness (β = -0.752, SE = 0.212, p < 0.001 after BH correction). No carryover effects from previous evenings (Lag1: p = 0.551; Lag2: p = 0.551) were detected in this study.

```{r}
# Session Info
session_info()

citation("lme4")
citation("lmerTest")
citation("performance")
citation("DHARMa")
citation("car")

packages <- c("lme4", "lmerTest", "DHARMa", "car", "performance", 
              "ggplot2", "patchwork", "cowplot",
              "clubSandwich", "nlme", "pracma")  
```
# Boxplots 1.5-h intervention for presentation 
```{r}
# Doublet empty boxplot 
# Prepare data using intervention window only
df_doublet_box_q789 <- pos_kss_all_q789 %>%
  filter(participant != "P07") %>%
  select(participant, doublet, mean_kss = pos2_kss) %>%
  mutate(doublet = factor(doublet, levels = c("DD", "BD", "DB", "BB")))

# Calculate y-axis limits
y_min <- floor(min(df_doublet_box_q789$mean_kss[is.finite(df_doublet_box_q789$mean_kss)]) * 2) / 2
y_max <- ceiling(max(df_doublet_box_q789$mean_kss[is.finite(df_doublet_box_q789$mean_kss)]) * 2) / 2

# Create plot 
ggplot(df_doublet_box_q789, aes(x = doublet, y = mean_kss)) +
  scale_fill_manual(values = doublet_colors) +
  scale_x_discrete(labels = doublet_labels) +
  scale_y_continuous(limits = c(y_min, y_max),
                     breaks = seq(1, 9, 0.5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  labs(
    title = "Karolinska Sleepiness Scale under light exposure doublets",
    subtitle = "Mean second-day KSS scores per doublet type (1.5-h intervention; P07 excluded)",
    x = "Doublet", 
    y = expression(atop("Mean KSS", 
                        scriptstyle("(1 = Extremely alert; 9 = Very sleepy)")))
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none",
    plot.margin = margin(15, 10, 10, 10)
  )

# Triplet boxplot empty
# Prepare data using intervention window
df_triplet_box_q789 <- triplet_kss_all_q789 %>%
  filter(participant != "P07", !is.na(pos3_kss)) %>%
  select(participant, triplet, mean_kss = pos3_kss) %>%
  mutate(triplet = factor(triplet, levels = c("DDD", "BDD", "DBD", "BBD",
                                              "DDB", "BDB", "DBB", "BBB")))

# Calculate y-axis limits
y_min <- floor(min(df_triplet_box_q789$mean_kss[is.finite(df_triplet_box_q789$mean_kss)]) * 2) / 2
y_max <- ceiling(max(df_triplet_box_q789$mean_kss[is.finite(df_triplet_box_q789$mean_kss)]) * 2) / 2

# Create plot 
ggplot(df_triplet_box_q789, aes(x = triplet, y = mean_kss)) +
  scale_fill_manual(values = triplet_colors) +
  scale_x_discrete(labels = triplet_labels) +
  scale_y_continuous(limits = c(y_min, y_max),
                     breaks = seq(1, 9, 0.5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  labs(
    title = "Karolinska Sleepiness Scale under light exposure triplets",
    subtitle = "Mean third-day KSS scores per triplet type (1.5-h intervention; P07 excluded)",
    x = "Triplet", 
    y = expression(atop("Mean KSS", 
                        scriptstyle("(1 = Extremely alert; 9 = Very sleepy)")))
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 11),
    axis.text.x = element_text(size = 9),
    axis.text.y = element_text(size = 13),
    axis.title = element_text(size = 14),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none",
    plot.margin = margin(15, 10, 10, 10)
  )
```

# Create combined KSS boxplot figure
```{r}
#| label: fig-combined-kss-boxplots
#| fig-cap: "KSS levels under light exposure doublets and triplets (P07 excluded)"
#| fig-width: 20
#| fig-height: 16

# Compute shared y-axis range across all four datasets
all_values <- c(
  df_doublet_box_5h$mean_kss,
  df_doublet_box_q789$mean_kss,
  df_triplet_box_5h$mean_kss,
  df_triplet_box_q789$mean_kss
)
y_min <- floor(min(all_values, na.rm = TRUE))
y_max <- ceiling(max(all_values, na.rm = TRUE))
y_breaks <- pretty(c(y_min, y_max), n = 6)

# --- Plot 1: Doublet 5h ---
p1 <- ggplot(df_doublet_box_5h, aes(x = doublet, y = mean_kss)) +
  geom_boxplot(aes(fill = doublet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15, seed = 42),
             size = 2.5, alpha = 0.7, shape = 16) +
  scale_fill_manual(values = doublet_colors) +
  scale_x_discrete(labels = doublet_labels) +
  scale_y_continuous(limits = c(y_min, y_max), breaks = y_breaks) +
  labs(
    title = "Doublets \u2013 5-h exposure",
    x = "Doublet",
    y = expression(atop("Mean KSS",
                        scriptstyle("(1 = Extremely alert; 9 = Very sleepy)")))
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )

# --- Plot 2: Doublet 1.5h ---
p2 <- ggplot(df_doublet_box_q789, aes(x = doublet, y = mean_kss)) +
  geom_boxplot(aes(fill = doublet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15, seed = 42),
             size = 2.5, alpha = 0.7, shape = 16) +
  scale_fill_manual(values = doublet_colors) +
  scale_x_discrete(labels = doublet_labels) +
  scale_y_continuous(limits = c(y_min, y_max), breaks = y_breaks) +
  labs(
    title = "Doublets \u2013 1.5-h intervention",
    x = "Doublet",
    y = expression(atop("Mean KSS",
                        scriptstyle("(1 = Extremely alert; 9 = Very sleepy)")))
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )

# --- Plot 3: Triplet 5h ---
p3 <- ggplot(df_triplet_box_5h, aes(x = triplet, y = mean_kss)) +
  geom_boxplot(aes(fill = triplet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15, seed = 42),
             size = 2.5, alpha = 0.7, shape = 16) +
  scale_fill_manual(values = triplet_colors) +
  scale_x_discrete(labels = triplet_labels) +
  scale_y_continuous(limits = c(y_min, y_max), breaks = y_breaks) +
  labs(
    title = "Triplets \u2013 5-h exposure",
    x = "Triplet",
    y = expression(atop("Mean KSS",
                        scriptstyle("(1 = Extremely alert; 9 = Very sleepy)")))
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )

# --- Plot 4: Triplet 1.5h ---
p4 <- ggplot(df_triplet_box_q789, aes(x = triplet, y = mean_kss)) +
  geom_boxplot(aes(fill = triplet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15, seed = 42),
             size = 2.5, alpha = 0.7, shape = 16) +
  scale_fill_manual(values = triplet_colors) +
  scale_x_discrete(labels = triplet_labels) +
  scale_y_continuous(limits = c(y_min, y_max), breaks = y_breaks) +
  labs(
    title = "Triplets \u2013 1.5-h intervention",
    x = "Triplet",
    y = expression(atop("Mean KSS",
                        scriptstyle("(1 = Extremely alert; 9 = Very sleepy)")))
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )

# --- Combine ---
(p1 | p2) / (p3 | p4) +
  plot_annotation(
    title = "Karolinska Sleepiness Scale under light exposure doublets and triplets",
    subtitle = "Mean KSS scores of the second / third day per sequence type (5-h exposure & 1.5-h intervention; P07 excluded)",
    theme = theme(
      plot.title = element_text(size = 24, hjust = 0),
      plot.subtitle = element_text(size = 18, hjust = 0, margin = margin(b = 10))
    )
  )
```