---
title: "Carryover Effects in Melatonin Suppression"
subtitle: "Light Exposure Sequence Analysis with Doublets and Triplets"
author: "Letizia Woerrlein"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: show
    code-tools: true
    embed-resources: true
    theme: cosmo
    fig-width: 10
    fig-height: 8
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
execute:
  warning: false
  message: false
  cache: false
editor: visual
---

# Study Overview

## Design

**Study Type**: Within-subject controlled trial

**Total Participants**: 9 (P01-P07, P09-P11)

**Exclusions**: - P11: Complete exclusion (study dropout due to infection) - P10 Session 5: Excluded (light automation malfunction) - P07: Excluded from statistical analysis (negative melatonin suppression)

## Study Phases

### Phase 1: Pre-Ambulatory Period

-   Duration: 7 days
-   Setting: At home
-   Measurements: Sleep logs, ActLumus, ActiGraph, CGM

### Phase 2: In-Laboratory Period

-   Duration: 10-13 days
-   Daily protocol: Arrival 5.5h before habitual bedtime (HBT); Depature 0.5h before HBT
-   Light conditions:
    -   **D (Dim)**: 10 lx continuously
    -   **B (Bright)**: 10 lx → 1000 lx (1.5h) → 10 lx
-   Salivary melatonin every 30 minutes, starting 30 minutes after chamber entry

### Phase 3: Post-Ambulatory Period

-   Duration: 7 days
-   Setting: At home (identical to Phase 1)

## Participant Groups

### Group 1 (HBT 00:00)

-   Participants: P01, P02, P03
-   Light sequence: DDDBBBDDBDB (11 sessions)

### Group 2 (HBT 22:00)

-   Participants: P04, P05
-   Light sequence: DBBBDDDBDB (10 sessions)

### Group 3 (HBT 23:30)

-   Participants: P06, P07, P09
-   Light sequence: DBBBDDDBDB (10 sessions)

### Group 4 (HBT 23:00)

-   Participant: P10
-   Light sequence: DDDBXDBBBDBDD (12 valid sessions, session 5 excluded)

# Setup

```{r setup}
#| label: setup
#| include: true

# Load required packages
library(readxl)
library(tidyr)
library(hms)
library(lubridate)
library(pracma)
library(stringr)
library(ggplot2)
library(purrr)
library(patchwork)
library(kableExtra)
library(psych)
library(corrplot)
library(lmerTest)
library(lme4)
library(car)
library(robustlmm)
library(segmented)
library(fitdistrplus)
library(moments)
library(DHARMa)
library(performance)
library(broom.mixed)
library(ggeffects)
library(dplyr)
library(ggtext)

# Set theme for all plots
theme_set(theme_classic(base_size = 14))

# Define color scheme
light_colors <- c("dim" = "#4A90E2", "bright" = "#F5A623")
```

# Data Import and Cleaning

```{r data-import}
#| label: data-import

# Define file path
file_path <- "C:/Users/anton/OneDrive/AfterGlow/data/raw/group/melatonin/01_melatonin_20251009T155000.xlsx"

# Read melatonin data
melatonin <- read_excel(file_path)

# Inspect the dataset
str(melatonin)
```

## Sample Count Verification

```{r sample-verification}
#| label: sample-verification

# Count samples per participant per session
samples_per_session <- melatonin %>%
  group_by(participant_id, session) %>%
  summarise(n_samples = n(), .groups = "drop") %>%
  arrange(participant_id, session)

# Filter sessions with more than 10 samples
sessions_over_10 <- samples_per_session %>%
  filter(n_samples > 10)

knitr::kable(sessions_over_10, 
             caption = "Sessions with >10 samples")
```

**Findings**: Extra samples found for: - P05 sessions 5 and 10 (11 samples each) - P07 session 1 (11 samples)

## Data Cleaning

**Duplicate Samples:**

For P05 and P07, some samples were measured in duplicate. The duplicate samples ("b" suffix) were removed after verifying concentrations with original samples:

-   **P05 Session 5 (sample 6b)**: Identical concentration to sample 6 → Original retained
-   **P05 Session 10 (sample 4b)**: Identical concentration to sample 4 → Original retained
-   **P07 Session 1 (sample 7b)**: Identical concentration to sample 7 → Original retained
-   **P07 Session 8 (sample 807b)**: Identical concentration to sample 7 → Original retained
-   **P07 Session 2 (sample 207b)**: Discrepant concentration (Original: 0.9 pg/mL vs. Control: 0.1 pg/mL) → Original retained

```{r data-cleaning}
#| label: data-cleaning

# Remove duplicate samples and excluded participants/sessions
melatonin <- melatonin %>%
  filter(
    # Remove duplicate samples (concentrations verified)
    !(participant_id == "P05" & session == "C02_05" & sample_number == "6b"),
    !(participant_id == "P05" & session == "C02_10" & sample_number == "4b"),
    !(short_sample_id == "P07_807b"),
    !(short_sample_id == "P07_207b"),
    !(participant_id == "P07" & session == "C03_01" & sample_number == "7b"),
    !(participant_id == "P11"),
    !(participant_id == "P10" & session == "C04_05")
  )

# Verify the changes
verification <- melatonin %>%
  filter(participant_id %in% c("P05", "P07")) %>%
  group_by(participant_id, session) %>%
  summarise(n_samples = n(), .groups = "drop") %>%
  arrange(participant_id, session)

knitr::kable(verification, 
             caption = "Verification: P05 and P07 sample counts after cleaning")
```

## Time Format Conversion

```{r time-conversion}
#| label: time-conversion

# Convert time format
if (is.numeric(melatonin$time)) {
  melatonin$time <- hms::as_hms(melatonin$time * 86400)
} else if (inherits(melatonin$time, "POSIXt")) {
  melatonin$time <- hms::as_hms(melatonin$time)
} else {
  melatonin$time <- hms::as_hms(hms::hms(melatonin$time))
}

# Add experimental session and time in hours
melatonin <- melatonin %>%
  arrange(participant_id, date_collected) %>%
  group_by(participant_id) %>%
  mutate(
    exp_session = as.integer(str_extract(session, "(?<=_)\\d+$")),
    time_hours = hour(time) + minute(time)/60 + second(time)/3600
  ) %>%
  ungroup()
```

## Time Range Validation

**Expected vs. Actual Time Ranges:**

Most sessions show times slightly outside the defined ranges due to the practical duration of sample collection and documentation. However, systematic time formatting errors were identified in the following sessions:

**Time Formatting Errors (corrected in next step based on handwritten protocols):**

-   **P04 Session 3**: 05:04 → should be 20:04 (+15 hours)
-   **P05 Session 3**: 05:04 → should be 20:04 (+15 hours)
-   **P06 Session 10**: 06:32 → should be 20:32 (+14 hours)
-   **P07 Session 10**: 06:32 → should be 20:32 (+14 hours)
-   **P09 Session 10**: 06:32 → should be 20:32 (+14 hours)

These errors likely occurred during Excel import/export and were verified against handwritten laboratory protocols.

```{r time-validation}
#| label: time-validation

# Check time range per evening
time_ranges <- melatonin %>%
  group_by(participant_id, exp_session, session) %>%
  summarise(
    min_time_hours = min(time_hours),
    max_time_hours = max(time_hours),
    n_samples = n(),
    .groups = "drop"
  ) %>%
  mutate(
    first_sample_time = sprintf("%02d:%02d", floor(min_time_hours), 
                                round((min_time_hours %% 1) * 60)),
    last_sample_time = sprintf("%02d:%02d", floor(max_time_hours), 
                               round((max_time_hours %% 1) * 60)),
    duration_hours = round(max_time_hours - min_time_hours, 2)
  )

# Define expected time ranges
time_limits <- data.frame(
  participant_id = c("P01", "P02", "P03", "P04", "P05", "P06", "P07", "P09", "P10"),
  min_hour = c(19, 19, 19, 17, 17, 18.5, 18.5, 18.5, 18),
  max_hour = c(23.5, 23.5, 23.5, 21.5, 21.5, 23, 23, 23, 22.5)
)

# Check for sessions outside defined ranges
sessions_outside <- time_ranges %>%
  mutate(
    min_time_hours = as.numeric(substr(first_sample_time, 1, 2)) + 
      as.numeric(substr(first_sample_time, 4, 5))/60,
    max_time_hours = as.numeric(substr(last_sample_time, 1, 2)) + 
      as.numeric(substr(last_sample_time, 4, 5))/60
  ) %>%
  left_join(time_limits, by = "participant_id") %>%
  filter(min_time_hours < min_hour | max_time_hours > max_hour)

if(nrow(sessions_outside) > 0) {
  knitr::kable(sessions_outside, 
               caption = "Sessions with times outside expected ranges (before correction)")
}
```

## Time Error Correction

```{r time-correction}
#| label: time-correction

# Correct documented time errors
melatonin <- melatonin %>%
  mutate(
    time = case_when(
      participant_id %in% c("P04", "P05") & exp_session == 3 & time_hours < 10 ~ 
        hms::as_hms(as.numeric(time) + 15*3600),
      participant_id %in% c("P06", "P07", "P09") & exp_session == 10 & time_hours < 10 ~ 
        hms::as_hms(as.numeric(time) + 14*3600),
      TRUE ~ time
    ),
    time_hours = case_when(
      participant_id %in% c("P04", "P05") & exp_session == 3 & time_hours < 10 ~ 
        time_hours + 15,
      participant_id %in% c("P06", "P07", "P09") & exp_session == 10 & time_hours < 10 ~ 
        time_hours + 14,
      TRUE ~ time_hours
    )
  )

# Verify corrections
corrected_sessions <- melatonin %>%
  filter(
    (participant_id %in% c("P04", "P05") & exp_session == 3) |
    (participant_id %in% c("P06", "P07", "P09") & exp_session == 10)
  ) %>%
  group_by(participant_id, exp_session) %>%
  summarise(
    min_time = sprintf("%02d:%02d", floor(min(time_hours)), 
                      round((min(time_hours) %% 1) * 60)),
    max_time = sprintf("%02d:%02d", floor(max(time_hours)), 
                      round((max(time_hours) %% 1) * 60)),
    .groups = "drop"
  )

knitr::kable(corrected_sessions, 
             caption = "Corrected time ranges for affected sessions")
```

## Light Pattern Definition

```{r light-patterns}
#| label: light-patterns

# Define light patterns for each participant
light_patterns <- list(
  P01 = strsplit("DDDBBBDDBDB", "")[[1]],
  P02 = strsplit("DDDBBBDDBDB", "")[[1]],
  P03 = strsplit("DDDBBBDDBDB", "")[[1]],
  P04 = strsplit("DBBBDDDBDB", "")[[1]],
  P05 = strsplit("DBBBDDDBDB", "")[[1]],
  P06 = strsplit("DBBBDDDBDB", "")[[1]],
  P07 = strsplit("DBBBDDDBDB", "")[[1]],
  P09 = strsplit("DBBBDDDBDB", "")[[1]],
  P10 = c("D", "D", "D", "B", NA, "D", "B", "B", "B", "D", "B", "D", "D")
)

# Convert sample_number to integer
melatonin <- melatonin %>%
  mutate(sample_num = as.integer(sample_number))
```

# AUC Calculations

## Full Evening AUC (5-hour exposure)

```{r auc-full}
#| label: auc-full

# Calculate AUC per evening using trapezoidal rule
auc_per_evening <- melatonin %>%
  arrange(participant_id, exp_session, time_hours) %>%
  group_by(participant_id, exp_session, session) %>%
  summarise(
    auc_melatonin = trapz(time_hours, melatonin_concentration),
    n_samples = n(),
    .groups = "drop"
  ) %>%
  arrange(participant_id, exp_session)

# Add light condition
auc_with_condition <- auc_per_evening %>%
  mutate(
    light_condition = case_when(
      participant_id == "P01" ~ light_patterns$P01[exp_session],
      participant_id == "P02" ~ light_patterns$P02[exp_session],
      participant_id == "P03" ~ light_patterns$P03[exp_session],
      participant_id == "P04" ~ light_patterns$P04[exp_session],
      participant_id == "P05" ~ light_patterns$P05[exp_session],
      participant_id == "P06" ~ light_patterns$P06[exp_session],
      participant_id == "P07" ~ light_patterns$P07[exp_session],
      participant_id == "P09" ~ light_patterns$P09[exp_session],
      participant_id == "P10" ~ light_patterns$P10[exp_session]
    )
  )

# Display summary statistics
summary_auc <- auc_with_condition %>%
  group_by(light_condition) %>%
  summarise(
    n = n(),
    mean_auc = mean(auc_melatonin),
    sd_auc = sd(auc_melatonin),
    min_auc = min(auc_melatonin),
    max_auc = max(auc_melatonin),
    .groups = "drop"
  )

knitr::kable(summary_auc, digits = 2,
             caption = "Summary statistics: Full evening AUC by light condition")
```

## Experimental Window AUC (Samples 7-9, 1.5-hour exposure)

```{r auc-experimental}
#| label: auc-experimental

# Filter to samples 7-9 (experimental window)
melatonin_experimental_samples <- melatonin %>%
  filter(sample_num >= 7 & sample_num <= 9)

# Calculate AUC for experimental window
auc_experimental_samples <- melatonin_experimental_samples %>%
  arrange(participant_id, exp_session, time_hours) %>%
  group_by(participant_id, exp_session, session) %>%
  summarise(
    auc_melatonin_exp = trapz(time_hours, melatonin_concentration),
    n_samples = n(),
    first_sample = min(sample_num),
    last_sample = max(sample_num),
    time_start = min(time_hours),
    time_end = max(time_hours),
    .groups = "drop"
  ) %>%
  arrange(participant_id, exp_session)

# Add light condition
auc_exp_with_condition <- auc_experimental_samples %>%
  mutate(
    light_condition = case_when(
      participant_id == "P01" ~ light_patterns$P01[exp_session],
      participant_id == "P02" ~ light_patterns$P02[exp_session],
      participant_id == "P03" ~ light_patterns$P03[exp_session],
      participant_id == "P04" ~ light_patterns$P04[exp_session],
      participant_id == "P05" ~ light_patterns$P05[exp_session],
      participant_id == "P06" ~ light_patterns$P06[exp_session],
      participant_id == "P07" ~ light_patterns$P07[exp_session],
      participant_id == "P09" ~ light_patterns$P09[exp_session],
      participant_id == "P10" ~ light_patterns$P10[exp_session]
    )
  )

# Display summary statistics
summary_auc_exp <- auc_exp_with_condition %>%
  group_by(light_condition) %>%
  summarise(
    n = n(),
    mean_auc = mean(auc_melatonin_exp),
    sd_auc = sd(auc_melatonin_exp),
    min_auc = min(auc_melatonin_exp),
    max_auc = max(auc_melatonin_exp),
    .groups = "drop"
  )

knitr::kable(summary_auc_exp, digits = 2,
             caption = "Summary statistics: Experimental window AUC by light condition")
```

## Individual AUC Values per Participant and Session

```{r auc-individual-table}
#| label: auc-individual-table

# Combine full evening and experimental window AUC
auc_combined <- auc_with_condition %>%
  select(participant_id, exp_session, light_condition, 
         auc_full_5h = auc_melatonin) %>%
  left_join(
    auc_exp_with_condition %>% 
      select(participant_id, exp_session, auc_exp_1.5h = auc_melatonin_exp),
    by = c("participant_id", "exp_session")
  ) %>%
  arrange(participant_id, exp_session)

# Display table
DT::datatable(
  auc_combined,
  options = list(
    pageLength = 15,
    scrollX = TRUE,
    order = list(list(0, 'asc'), list(1, 'asc'))
  ),
  caption = "Individual AUC values: Full 5h vs. Experimental 1.5h window",
  rownames = FALSE
) %>%
  DT::formatRound(columns = c('auc_full_5h', 'auc_exp_1.5h'), digits = 2) %>%
  DT::formatStyle(
    'light_condition',
    backgroundColor = DT::styleEqual(c('D', 'B'), c('#e3f2fd', '#fff3e0'))
  )
```

# Individual Melatonin Profiles

## Individual Time-Course Plots (Panel View)

```{r individual-profiles-function}
#| label: individual-profiles-function

# Function to create melatonin profile plot
create_melatonin_profile <- function(participant_id, light_pattern) {
  
  # Prepare data
  p_data <- melatonin %>%
    filter(participant_id == !!participant_id) %>%
    mutate(
      light_condition = rep(light_pattern, each = 10)[1:n()],
      hours_before_hbt = 5.5 - (as.numeric(sample_number) * 0.5)
    )
  
  # Get AUC values
  p_auc <- auc_per_evening %>%
    filter(participant_id == !!participant_id) %>%
    mutate(
      light_condition = light_pattern,
      auc_label = paste0("AUC=", round(auc_melatonin, 1)),
      exp_session = as.integer(exp_session)
    )
  
  # Determine y-axis limit
  max_y <- max(p_data$melatonin_concentration, na.rm = TRUE)
  y_limit <- ceiling(max_y / 10) * 10
  
  # Create plot
  p <- ggplot(p_data, aes(x = hours_before_hbt, y = melatonin_concentration)) +
    geom_area(fill = "#4A90E2", alpha = 0.3) +
    geom_rect(xmin = 5, xmax = 2, ymin = -Inf, ymax = Inf,
              fill = "#E3F2FD", alpha = 0.5, inherit.aes = FALSE) +
    geom_rect(data = p_auc %>% filter(light_condition == "B"),
              aes(xmin = 2.5, xmax = 1, ymin = -Inf, ymax = Inf),
              fill = "#FFE0B2", alpha = 0.5, inherit.aes = FALSE) +
    geom_line(color = "black", linewidth = 0.8) +
    geom_point(color = "black", size = 2.5) +
    geom_text(data = p_auc,
              aes(x = 5, y = Inf, label = auc_label),
              hjust = 0, vjust = 1.3, size = 3.5,
              inherit.aes = FALSE) +
    scale_x_reverse(
      breaks = 0:5,
      limits = c(5.3, 0.2),
      expand = c(0, 0)
    ) +
    scale_y_continuous(
      limits = c(0, y_limit),
      expand = expansion(mult = c(0, 0.05))
    ) +
    facet_wrap(~ exp_session, ncol = 4,
               labeller = labeller(exp_session = function(x) {
                 idx <- as.numeric(x)
                 light <- p_auc$light_condition[idx]
                 paste0("Evening ", x, " (", light, ")")
               })) +
    labs(
      title = paste0("Salivary melatonin per evening (", participant_id, ")"),
      x = "Hours before HBT [h]",
      y = "Salivary melatonin [pg/mL]"
    ) +
    theme_classic(base_size = 11) +
    theme(
      aspect.ratio = 1,
      plot.title = element_text(size = 14, hjust = 0, margin = margin(b = 10)),
      axis.text = element_text(color = "black", size = 10),
      axis.title = element_text(size = 11),
      axis.line = element_line(linewidth = 0.5),
      axis.ticks = element_line(linewidth = 0.5),
      panel.grid = element_blank(),
      strip.background = element_rect(fill = "white", color = "grey80", linewidth = 0.5),
      strip.text = element_text(size = 10, color = "black"),
      panel.background = element_rect(fill = "white", color = "grey80", linewidth = 0.5),
      plot.margin = margin(10, 10, 10, 10),
      legend.position = "none"
    )
  
  return(p)
}
```

## P01 - P03 (Group 1)

```{r profile-p01}
#| label: fig-profile-p01
#| fig-cap: "Salivary melatonin per evening - P01"
#| fig-width: 12
#| fig-height: 9

create_melatonin_profile("P01", light_patterns$P01)
```

```{r profile-p02}
#| label: fig-profile-p02
#| fig-cap: "Salivary melatonin per evening - P02"
#| fig-width: 12
#| fig-height: 9

create_melatonin_profile("P02", light_patterns$P02)
```

```{r profile-p03}
#| label: fig-profile-p03
#| fig-cap: "Salivary melatonin per evening - P03"
#| fig-width: 12
#| fig-height: 9

create_melatonin_profile("P03", light_patterns$P03)
```

## P04 - P05 (Group 2)

```{r profile-p04}
#| label: fig-profile-p04
#| fig-cap: "Salivary melatonin per evening - P04"
#| fig-width: 12
#| fig-height: 8

create_melatonin_profile("P04", light_patterns$P04)
```

```{r profile-p05}
#| label: fig-profile-p05
#| fig-cap: "Salivary melatonin per evening - P05"
#| fig-width: 12
#| fig-height: 8

create_melatonin_profile("P05", light_patterns$P05)
```

## P06, P07, P09 (Group 3)

```{r profile-p06}
#| label: fig-profile-p06
#| fig-cap: "Salivary melatonin per evening - P06"
#| fig-width: 12
#| fig-height: 8

create_melatonin_profile("P06", light_patterns$P06)
```

```{r profile-p07}
#| label: fig-profile-p07
#| fig-cap: "Salivary melatonin per evening - P07"
#| fig-width: 12
#| fig-height: 8

create_melatonin_profile("P07", light_patterns$P07)
```

```{r profile-p09}
#| label: fig-profile-p09
#| fig-cap: "Salivary melatonin per evening - P09"
#| fig-width: 12
#| fig-height: 8

create_melatonin_profile("P09", light_patterns$P09)
```

## P10 (Group 4) - Special Case

```{r profile-p10}
#| label: fig-profile-p10
#| fig-cap: "Salivary melatonin per evening - P10 (Session 5 excluded)"
#| fig-width: 12
#| fig-height: 9

# Prepare data for P10 (excluding session 5)
p10_data <- melatonin %>%
  filter(participant_id == "P10" & exp_session != 5) %>%
  mutate(
    hours_before_hbt = 5.5 - (as.numeric(sample_number) * 0.5),
    light_condition = case_when(
      exp_session %in% c(1, 2, 3) ~ "D",
      exp_session == 4 ~ "B",
      exp_session == 6 ~ "D",
      exp_session %in% c(7, 8, 9) ~ "B",
      exp_session == 10 ~ "D",
      exp_session == 11 ~ "B",
      exp_session %in% c(12, 13) ~ "D"
    )
  )

# Get AUC values for P10 (excluding session 5)
p10_auc <- auc_per_evening %>%
  filter(participant_id == "P10" & exp_session != 5) %>%
  mutate(
    light_condition = case_when(
      exp_session %in% c(1, 2, 3) ~ "D",
      exp_session == 4 ~ "B",
      exp_session == 6 ~ "D",
      exp_session %in% c(7, 8, 9) ~ "B",
      exp_session == 10 ~ "D",
      exp_session == 11 ~ "B",
      exp_session %in% c(12, 13) ~ "D"
    ),
    auc_label = paste0("AUC=", round(auc_melatonin, 1)),
    exp_session = as.integer(exp_session)
  )

# Determine y-axis limit
max_y <- max(p10_data$melatonin_concentration, na.rm = TRUE)
y_limit <- ceiling(max_y / 10) * 10

# Create plot for P10
ggplot(p10_data, aes(x = hours_before_hbt, y = melatonin_concentration)) +
  geom_area(fill = "#4A90E2", alpha = 0.3) +
  geom_rect(xmin = 5, xmax = 2, ymin = -Inf, ymax = Inf,
            fill = "#E3F2FD", alpha = 0.5, inherit.aes = FALSE) +
  geom_rect(data = p10_auc %>% filter(light_condition == "B"),
            aes(xmin = 2.5, xmax = 1, ymin = -Inf, ymax = Inf),
            fill = "#FFE0B2", alpha = 0.5, inherit.aes = FALSE) +
  geom_line(color = "black", linewidth = 0.8) +
  geom_point(color = "black", size = 2.5) +
  geom_text(data = p10_auc,
            aes(x = 5, y = Inf, label = auc_label),
            hjust = 0, vjust = 1.3, size = 3.5,
            inherit.aes = FALSE) +
  scale_x_reverse(
    breaks = 0:5,
    limits = c(5.3, 0.2),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    limits = c(0, y_limit),
    expand = expansion(mult = c(0, 0.05))
  ) +
  facet_wrap(~ exp_session, ncol = 4,
             labeller = labeller(exp_session = function(x) {
               session_num <- as.numeric(x)
               light <- p10_auc$light_condition[p10_auc$exp_session == session_num]
               paste0("Evening ", x, " (", light, ")")
             })) +
  labs(
    title = "Salivary melatonin per evening (P10)",
    subtitle = "Session 5 excluded due to light automation malfunction",
    x = "Hours before HBT [h]",
    y = "Salivary melatonin [pg/mL]"
  ) +
  theme_classic(base_size = 11) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 14, hjust = 0, margin = margin(b = 5)),
    plot.subtitle = element_text(size = 11, hjust = 0, margin = margin(b = 10), color = "grey40"),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(size = 11),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.5),
    panel.grid = element_blank(),
    strip.background = element_rect(fill = "white", color = "grey80", linewidth = 0.5),
    strip.text = element_text(size = 10, color = "black"),
    panel.background = element_rect(fill = "white", color = "grey80", linewidth = 0.5),
    plot.margin = margin(10, 10, 10, 10),
    legend.position = "none"
  )
```

## Individual Melatonin Suppression

```{r suppression-calculation}
#| label: suppression-calculation

# Calculate mean AUC by condition (FULL 5H)
mean_auc_by_condition <- auc_with_condition %>%
  group_by(participant_id, light_condition) %>%
  summarise(mean_auc = mean(auc_melatonin, na.rm = TRUE), .groups = "drop")

# Calculate suppression
melatonin_suppression <- mean_auc_by_condition %>%
  pivot_wider(names_from = light_condition, values_from = mean_auc, names_prefix = "mean_auc_") %>%
  mutate(suppression_percent = (1 - mean_auc_B / mean_auc_D) * 100)

knitr::kable(melatonin_suppression, digits = 2,
             caption = "Individual melatonin suppression - Full 5h window (%)")
```

```{r suppression-plot}
#| label: suppression-plot
#| fig-width: 10
#| fig-height: 8
#| fig-cap: "Individual melatonin suppression by bright light"

# Assign groups
melatonin_suppression_grouped <- melatonin_suppression %>%
  mutate(
    group = case_when(
      participant_id %in% c("P01", "P02", "P03") ~ "Group 1",
      participant_id %in% c("P04", "P05") ~ "Group 2",
      participant_id %in% c("P06", "P07", "P09") ~ "Group 3",
      participant_id == "P10" ~ "Group 4"
    )
  )

# Plot
ggplot(melatonin_suppression_grouped, 
       aes(x = reorder(participant_id, suppression_percent), 
           y = suppression_percent, fill = group)) +
  geom_bar(stat = "identity", alpha = 0.85) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_text(aes(label = sprintf("%.0f%%", suppression_percent)),
            vjust = ifelse(melatonin_suppression_grouped$suppression_percent >= 0, -0.5, 1.5)) +
  scale_fill_manual(values = c("#4477AA", "#EE6677", "#228833", "#CCBB44")) +
  labs(title = "Melatonin suppression by bright light (1000 lx)",
       subtitle = "Full 5h evening window",
       x = "Participant", y = "Suppression (%)", fill = "Group") +
  theme_classic() +
  theme(legend.position = "bottom")
```

## Individual Melatonin Suppression - 1.5h Exposure Window

### Suppression Calculation

```{r}
#| label: suppression-calculation-1.5h

# Calculate mean AUC by condition (1.5H EXPOSURE WINDOW ONLY)
mean_auc_by_condition_1.5h <- auc_exp_with_condition %>%
  group_by(participant_id, light_condition) %>%
  summarise(mean_auc = mean(auc_melatonin_exp, na.rm = TRUE), .groups = "drop")

# Calculate suppression for 1.5h window
melatonin_suppression_1.5h <- mean_auc_by_condition_1.5h %>%
  pivot_wider(names_from = light_condition, values_from = mean_auc, names_prefix = "mean_auc_") %>%
  mutate(suppression_percent = (1 - mean_auc_B / mean_auc_D) * 100)

knitr::kable(melatonin_suppression_1.5h, digits = 2,
             caption = "Individual melatonin suppression - 1.5h exposure window (%)")
```

### Visualization

```{r}
#| label: suppression-plot-1.5h
#| fig-width: 10
#| fig-height: 8
#| fig-cap: "Individual melatonin suppression by bright light - 1.5h exposure window"

# Assign groups
melatonin_suppression_1.5h_grouped <- melatonin_suppression_1.5h %>%
  mutate(
    group = case_when(
      participant_id %in% c("P01", "P02", "P03") ~ "Group 1",
      participant_id %in% c("P04", "P05") ~ "Group 2",
      participant_id %in% c("P06", "P07", "P09") ~ "Group 3",
      participant_id == "P10" ~ "Group 4"
    )
  )

# Plot
ggplot(melatonin_suppression_1.5h_grouped, 
       aes(x = reorder(participant_id, suppression_percent), 
           y = suppression_percent, fill = group)) +
  geom_bar(stat = "identity", alpha = 0.85) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_text(aes(label = sprintf("%.0f%%", suppression_percent)),
            vjust = ifelse(melatonin_suppression_1.5h_grouped$suppression_percent >= 0, -0.5, 1.5)) +
  scale_fill_manual(values = c("#4477AA", "#EE6677", "#228833", "#CCBB44")) +
  labs(title = "Melatonin suppression by bright light (1000 lx)",
       subtitle = "1.5h exposure window",
       x = "Participant", y = "Suppression (%)", fill = "Group") +
  theme_classic() +
  theme(legend.position = "bottom")
```

## Doublet Analysis

### Data Preparation

```{r doublet-data-prep}
#| label: doublet-data-prep

# Define light patterns for groups
light_g1 <- unlist(strsplit("DDDBBBDDBDB", ""))
light_g2 <- unlist(strsplit("DBBBDDDBDB", ""))
light_g4_orig <- unlist(strsplit("DDDBXDBBBDBDD", ""))

# Create doublet dataframes
df_doublets_g1 <- data.frame(
  session_start = 1:(length(light_g1) - 1),
  session_end = 2:length(light_g1),
  doublet = paste0(light_g1[-length(light_g1)], light_g1[-1])
)

df_doublets_g2 <- data.frame(
  session_start = 1:(length(light_g2) - 1),
  session_end = 2:length(light_g2),
  doublet = paste0(light_g2[-length(light_g2)], light_g2[-1])
)

df_doublets_g4_orig <- data.frame(
  session_start = 1:(length(light_g4_orig) - 1),
  session_end = 2:length(light_g4_orig),
  doublet = paste0(light_g4_orig[-length(light_g4_orig)], light_g4_orig[-1])
)

df_doublets_g4 <- df_doublets_g4_orig %>% 
  filter(!str_detect(doublet, "X"))

# Create doublet position mappings
doublet_positions_g1 <- df_doublets_g1 %>%
  group_by(doublet) %>%
  summarise(
    position1_sessions = list(session_start),
    position2_sessions = list(session_end),
    .groups = "drop"
  )

doublet_positions_g2 <- df_doublets_g2 %>%
  group_by(doublet) %>%
  summarise(
    position1_sessions = list(session_start),
    position2_sessions = list(session_end),
    .groups = "drop"
  )

doublet_positions_g4 <- df_doublets_g4 %>%
  group_by(doublet) %>%
  summarise(
    position1_sessions = list(session_start),
    position2_sessions = list(session_end),
    .groups = "drop"
  )

# Function to calculate doublet positions
calc_doublet_positions <- function(participant_id, auc_data, doublet_map) {
  results <- doublet_map %>%
    rowwise() %>%
    mutate(
      pos1_auc = mean(auc_data$auc_melatonin[auc_data$exp_session %in% position1_sessions], 
                      na.rm = TRUE),
      pos2_auc = mean(auc_data$auc_melatonin[auc_data$exp_session %in% position2_sessions], 
                      na.rm = TRUE),
      participant = participant_id
    ) %>%
    ungroup() %>%
    select(participant, doublet, pos1_auc, pos2_auc)
  
  return(results)
}

# Function for experimental window
calc_doublet_positions_exp <- function(participant_id, auc_data_exp, doublet_map) {
  results <- doublet_map %>% 
    rowwise() %>%
    mutate(
      pos1_auc_exp = mean(auc_data_exp$auc_melatonin_exp[auc_data_exp$exp_session %in% position1_sessions], 
                          na.rm = TRUE),
      pos2_auc_exp = mean(auc_data_exp$auc_melatonin_exp[auc_data_exp$exp_session %in% position2_sessions], 
                          na.rm = TRUE),
      participant = participant_id
    ) %>%
    ungroup() %>%
    select(participant, doublet, pos1_auc_exp, pos2_auc_exp)
  
  return(results)
}

# Calculate doublet AUC for all participants - FULL 5H
participants <- c("P01", "P02", "P03", "P04", "P05", "P06", "P07", "P09", "P10")
doublets <- c("DD", "DB", "BD", "BB")

df_plot <- data.frame(
  doublet = character(),
  participant = character(),
  mean_auc = numeric(),
  stringsAsFactors = FALSE
)

for(p in participants) {
  # Select appropriate doublet mapping
  if(p %in% c("P01", "P02", "P03")) {
    df_doublets <- df_doublets_g1
  } else if(p %in% c("P04", "P05", "P06", "P07", "P09")) {
    df_doublets <- df_doublets_g2
  } else {
    df_doublets <- df_doublets_g4
  }
  
  # Get AUC data
  auc_data <- auc_per_evening %>%
    filter(participant_id == p) %>%
    select(exp_session, auc_melatonin)
  
  # Join with doublets
  auc_doublets <- df_doublets %>%
    rename(exp_session = session_end) %>%
    left_join(auc_data, by = "exp_session")
  
  # Calculate mean per doublet
  mean_auc_doublets <- auc_doublets %>%
    group_by(doublet) %>%
    summarise(mean_auc = mean(auc_melatonin, na.rm = TRUE), .groups = "drop")
  
  # Add to plot data
  for(d in doublets) {
    if(d %in% mean_auc_doublets$doublet) {
      df_plot <- rbind(df_plot, data.frame(
        doublet = d,
        participant = p,
        mean_auc = mean_auc_doublets$mean_auc[mean_auc_doublets$doublet == d]
      ))
    }
  }
}

df_plot$doublet <- factor(df_plot$doublet, levels = c("DD", "BD", "DB", "BB"))

# Calculate doublet AUC - EXPERIMENTAL WINDOW (1.5h)
participants_exp <- c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10")

df_plot_exp <- data.frame(
  doublet = character(),
  participant = character(),
  mean_auc_exp = numeric(),
  stringsAsFactors = FALSE
)

for(p in participants_exp) {
  # Select appropriate doublet mapping
  if(p %in% c("P01", "P02", "P03")) {
    df_doublets <- df_doublets_g1
  } else if(p %in% c("P04", "P05", "P06", "P09")) {
    df_doublets <- df_doublets_g2
  } else {
    df_doublets <- df_doublets_g4
  }
  
  # Get experimental AUC data
  auc_data_exp <- auc_experimental_samples %>%
    filter(participant_id == p) %>%
    select(exp_session, auc_melatonin_exp)
  
  # Join with doublets
  auc_doublets_exp <- df_doublets %>%
    rename(exp_session = session_end) %>%
    left_join(auc_data_exp, by = "exp_session")
  
  # Calculate mean per doublet
  mean_auc_doublets_exp <- auc_doublets_exp %>%
    group_by(doublet) %>%
    summarise(mean_auc_exp = mean(auc_melatonin_exp, na.rm = TRUE), .groups = "drop")
  
  # Add to plot data
  for(d in doublets) {
    if(d %in% mean_auc_doublets_exp$doublet) {
      df_plot_exp <- rbind(df_plot_exp, data.frame(
        doublet = d,
        participant = p,
        mean_auc_exp = mean_auc_doublets_exp$mean_auc_exp[mean_auc_doublets_exp$doublet == d]
      ))
    }
  }
}

df_plot_exp$doublet <- factor(df_plot_exp$doublet, levels = c("DD", "BD", "DB", "BB"))
```

### Boxplot: 5-hour Exposure

```{r doublet-boxplot-5h}
#| label: fig-doublet-boxplot-5h
#| fig-cap: "Melatonin levels under light exposure doublets (5-h exposure, P07 excluded)"
#| fig-width: 10
#| fig-height: 8

# Labels with bold second letter
doublet_labels <- c(
  "DD" = expression("D"*bold("D")),
  "BD" = expression("B"*bold("D")),
  "DB" = expression("D"*bold("B")),
  "BB" = expression("B"*bold("B"))
)

# Colors based on second letter
doublet_colors <- c(
  "DD" = "#4A90E2", "BD" = "#4A90E2",
  "DB" = "#F5A623", "BB" = "#F5A623"
)

# Exclude P07
df_plot_noP07 <- df_plot %>% filter(participant != "P07")

ggplot(df_plot_noP07, aes(x = doublet, y = mean_auc)) +
  geom_boxplot(aes(fill = doublet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), 
             size = 2.5, alpha = 0.7, shape = 16) +
  scale_fill_manual(values = doublet_colors) +
  scale_x_discrete(labels = doublet_labels) +
  labs(
    title = "Melatonin levels under light exposure doublets",
    subtitle = "Mean second-day melatonin AUC per doublet type (5-h exposure; P07 excluded)",
    x = "Doublet", 
    y = "Mean AUC melatonin (pg/mL × h)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )
```

### Data Overview: Doublet Observations

```{r}
# Create detailed overview table
doublet_overview_5h <- df_plot_noP07 %>%
  arrange(doublet, participant) %>%
  select(participant, doublet, mean_auc)

DT::datatable(
  doublet_overview_5h,
  options = list(
    pageLength = 20,
    scrollX = TRUE,
    order = list(list(1, 'asc'), list(0, 'asc'))
  ),
  caption = "Individual doublet observations used in boxplot (5-h exposure)",
  rownames = FALSE
) %>%
  DT::formatRound(columns = 'mean_auc', digits = 2) %>%
  DT::formatStyle(
    'doublet',
    backgroundColor = DT::styleEqual(
      c('DD', 'BD', 'DB', 'BB'),
      c('#e3f2fd', '#e3f2fd', '#fff3e0', '#fff3e0')
    )
  )
```

### Boxplot: 1.5-hour Exposure

```{r doublet-boxplot-1.5h}
#| label: fig-doublet-boxplot-1.5h
#| fig-cap: "Melatonin levels under light exposure doublets (1.5-h exposure, P07 excluded)"
#| fig-width: 10
#| fig-height: 8

ggplot(df_plot_exp, aes(x = doublet, y = mean_auc_exp)) +
  geom_boxplot(aes(fill = doublet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15),
             size = 2.5, alpha = 0.7, shape = 16) +
  scale_fill_manual(values = doublet_colors) +
  scale_x_discrete(labels = doublet_labels) +
  labs(
    title = "Melatonin levels under light exposure doublets",
    subtitle = "Mean second-day melatonin AUC per doublet type (1.5-h exposure; P07 excluded)",
    x = "Doublet",
    y = "Mean AUC melatonin (pg/mL × h)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )
```

### Data Overview: Doublet Observations (1.5-h)

```{r}
# Create detailed overview table
doublet_overview_1.5h <- df_plot_exp %>%
  arrange(doublet, participant) %>%
  select(participant, doublet, mean_auc_exp)

DT::datatable(
  doublet_overview_1.5h,
  options = list(
    pageLength = 20,
    scrollX = TRUE,
    order = list(list(1, 'asc'), list(0, 'asc'))
  ),
  caption = "Individual doublet observations used in boxplot (1.5-h exposure)",
  rownames = FALSE
) %>%
  DT::formatRound(columns = 'mean_auc_exp', digits = 2) %>%
  DT::formatStyle(
    'doublet',
    backgroundColor = DT::styleEqual(
      c('DD', 'BD', 'DB', 'BB'),
      c('#e3f2fd', '#e3f2fd', '#fff3e0', '#fff3e0')
    )
  )
```

### Line Plot: Position Comparison (5-hour)

```{r doublet-lineplot-5h}
#| label: fig-doublet-lineplot-5h
#| fig-cap: "Doublet position comparison (5-h exposure, P07 excluded)"
#| fig-width: 10
#| fig-height: 8

# Calculate position data for all participants (excluding P07)
participants_noP07 <- c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10")

mean_doublet_pos_noP07 <- data.frame(
  doublet = character(),
  position = character(),
  mean_auc = numeric(),
  sd_auc = numeric(),
  n_participants = integer(),
  stringsAsFactors = FALSE
)

for(d in doublets) {
  for(pos in c("pos1", "pos2")) {
    values <- c()
    for(p in participants_noP07) {
      # Select mapping
      if(p %in% c("P01", "P02", "P03")) {
        doublet_map <- doublet_positions_g1
      } else if(p %in% c("P04", "P05", "P06", "P09")) {
        doublet_map <- doublet_positions_g2
      } else {
        doublet_map <- doublet_positions_g4
      }
      
      # Get AUC data
      auc_data <- auc_per_evening %>%
        filter(participant_id == p)
      
      # Calculate positions
      pos_auc <- calc_doublet_positions(p, auc_data, doublet_map)
      
      # Extract value
      if(d %in% pos_auc$doublet) {
        if(pos == "pos1") {
          values <- c(values, pos_auc$pos1_auc[pos_auc$doublet == d])
        } else {
          values <- c(values, pos_auc$pos2_auc[pos_auc$doublet == d])
        }
      }
    }
    
    if(length(values) > 0) {
      mean_doublet_pos_noP07 <- rbind(mean_doublet_pos_noP07, data.frame(
        doublet = d,
        position = pos,
        mean_auc = mean(values, na.rm = TRUE),
        sd_auc = sd(values, na.rm = TRUE),
        n_participants = length(values)
      ))
    }
  }
}

# Prepare for plotting
df_plot_positions_noP07 <- mean_doublet_pos_noP07 %>%
  mutate(
    position_label = ifelse(position == "pos1", "Position 1", "Position 2"),
    position_num = ifelse(position == "pos1", 1, 2),
    light_type = ifelse(position == "pos1", 
                        substr(doublet, 1, 1),
                        substr(doublet, 2, 2)),
    light_label = ifelse(light_type == "D", "dim", "bright")
  ) %>%
  mutate(doublet = factor(doublet, levels = c("DD", "BD", "DB", "BB")))

# Define colors
light_colors <- c("dim" = "#4A90E2", "bright" = "#F5A623")

# Create plot
ggplot(df_plot_positions_noP07, 
       aes(x = position_num, y = mean_auc, group = doublet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_label), size = 5, shape = 16) +
  scale_color_manual(values = light_colors,
                     labels = c("dim" = "Dim (10 lx)", "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2), 
                     labels = c("Position 1", "Position 2"),
                     limits = c(0.8, 2.2)) +
  labs(
    title = "Melatonin levels across light exposure doublets",
    subtitle = "Mean melatonin AUC across all participants (5-h exposure; P07 excluded)",
    x = "Doublet position",
    y = "Mean AUC melatonin (pg/mL × h)",
    color = "Light condition"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "right"
  )
```

### Line Plot: Position Comparison (1.5-hour)

```{r doublet-lineplot-1.5h}
#| label: fig-doublet-lineplot-1.5h
#| fig-cap: "Doublet position comparison (1.5-h exposure, P07 excluded)"
#| fig-width: 10
#| fig-height: 8

# Calculate position data for experimental window (excluding P07)
mean_doublet_pos_exp_noP07 <- data.frame(
  doublet = character(),
  position = character(),
  mean_auc_exp = numeric(),
  sd_auc_exp = numeric(),
  n_participants = integer(),
  stringsAsFactors = FALSE
)

for(d in doublets) {
  for(pos in c("pos1", "pos2")) {
    values <- c()
    for(p in participants_exp) {
      # Select mapping
      if(p %in% c("P01", "P02", "P03")) {
        doublet_map <- doublet_positions_g1
      } else if(p %in% c("P04", "P05", "P06", "P09")) {
        doublet_map <- doublet_positions_g2
      } else {
        doublet_map <- doublet_positions_g4
      }
      
      # Get experimental AUC data
      auc_data_exp <- auc_experimental_samples %>%
        filter(participant_id == p)
      
      # Calculate positions
      pos_auc_exp <- calc_doublet_positions_exp(p, auc_data_exp, doublet_map)
      
      # Extract value
      if(d %in% pos_auc_exp$doublet) {
        if(pos == "pos1") {
          values <- c(values, pos_auc_exp$pos1_auc_exp[pos_auc_exp$doublet == d])
        } else {
          values <- c(values, pos_auc_exp$pos2_auc_exp[pos_auc_exp$doublet == d])
        }
      }
    }
    
    if(length(values) > 0) {
      mean_doublet_pos_exp_noP07 <- rbind(mean_doublet_pos_exp_noP07, data.frame(
        doublet = d,
        position = pos,
        mean_auc_exp = mean(values, na.rm = TRUE),
        sd_auc_exp = sd(values, na.rm = TRUE),
        n_participants = length(values)
      ))
    }
  }
}

# Prepare for plotting
df_plot_positions_exp_noP07 <- mean_doublet_pos_exp_noP07 %>%
  mutate(
    position_label = ifelse(position == "pos1", "Position 1", "Position 2"),
    position_num = ifelse(position == "pos1", 1, 2),
    light_type = ifelse(position == "pos1", 
                        substr(doublet, 1, 1),
                        substr(doublet, 2, 2)),
    light_label = ifelse(light_type == "D", "dim", "bright")
  ) %>%
  mutate(doublet = factor(doublet, levels = c("DD", "BD", "DB", "BB")))

# Create plot
ggplot(df_plot_positions_exp_noP07, 
       aes(x = position_num, y = mean_auc_exp, group = doublet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_label), size = 5, shape = 16) +
  scale_color_manual(values = light_colors,
                     labels = c("dim" = "Dim (10 lx)", "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2), 
                     labels = c("Position 1", "Position 2"),
                     limits = c(0.8, 2.2)) +
  labs(
    title = "Melatonin levels across light exposure doublets",
    subtitle = "Mean melatonin AUC across all participants (1.5-h exposure; P07 excluded)",
    x = "Doublet position",
    y = "Mean AUC melatonin (pg/mL × h)",
    color = "Light condition"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "right"
  )
```

## Descriptive Statistics: Doublet Medians and IQR

```{r doublet-descriptives-combined}
#| label: doublet-descriptives-combined

# Calculate statistics for 5-h exposure
doublet_stats_5h <- df_plot_noP07 %>%
  group_by(doublet) %>%
  summarise(
    n_5h = n(),
    median_5h = median(mean_auc, na.rm = TRUE),
    IQR_5h = IQR(mean_auc, na.rm = TRUE),
    .groups = "drop"
  )

# Calculate statistics for 1.5-h exposure
doublet_stats_1.5h <- df_plot_exp %>%
  group_by(doublet) %>%
  summarise(
    n_1.5h = n(),
    median_1.5h = median(mean_auc_exp, na.rm = TRUE),
    IQR_1.5h = IQR(mean_auc_exp, na.rm = TRUE),
    .groups = "drop"
  )

# Combine both exposure windows
doublet_stats_combined <- doublet_stats_5h %>%
  full_join(doublet_stats_1.5h, by = "doublet") %>%
  arrange(doublet)

# Display as formatted table
knitr::kable(
  doublet_stats_combined,
  digits = 2,
  col.names = c("Doublet", "N", "Median", "IQR", "N", "Median", "IQR"),
  caption = "Descriptive statistics for doublet second-day melatonin AUC across exposure windows (P07 excluded)",
  align = c("l", rep("r", 6))
) %>%
  kableExtra::add_header_above(c(" " = 1, "5-h Exposure" = 3, "1.5-h Exposure" = 3))
```

## Triplet Analysis

### Data Preparation

```{r triplet-data-prep}
#| label: triplet-data-prep

# Define light patterns (redefine to ensure they exist)
light_g1 <- unlist(strsplit("DDDBBBDDBDB", ""))
light_g2 <- unlist(strsplit("DBBBDDDBDB", ""))
light_g4_orig <- unlist(strsplit("DDDBXDBBBDBDD", ""))

# Function to calculate triplet positions
calc_triplet_positions <- function(participant_id, auc_data, triplet_map, auc_col) {
  triplet_map %>%
    rowwise() %>%
    mutate(
      pos1_auc = mean(auc_data[[auc_col]][auc_data$session %in% position1_sessions], na.rm = TRUE),
      pos2_auc = mean(auc_data[[auc_col]][auc_data$session %in% position2_sessions], na.rm = TRUE),
      pos3_auc = mean(auc_data[[auc_col]][auc_data$session %in% position3_sessions], na.rm = TRUE),
      participant = participant_id
    ) %>%
    ungroup() %>%
    select(participant, triplet, pos1_auc, pos2_auc, pos3_auc)
}

# Define triplet mappings for Group 1
df_triplets_g1 <- data.frame(
  session_1 = 1:(length(light_g1)-2),
  session_2 = 2:(length(light_g1)-1),
  session_3 = 3:length(light_g1),
  triplet = sapply(1:(length(light_g1)-2), function(i) paste0(light_g1[i:(i+2)], collapse = ""))
)

triplet_positions_g1 <- df_triplets_g1 %>%
  group_by(triplet) %>%
  summarise(
    position1_sessions = list(session_1),
    position2_sessions = list(session_2),
    position3_sessions = list(session_3),
    .groups = "drop"
  )

# Define triplet mappings for Group 2
df_triplets_g2 <- data.frame(
  session_1 = 1:(length(light_g2)-2),
  session_2 = 2:(length(light_g2)-1),
  session_3 = 3:length(light_g2),
  triplet = sapply(1:(length(light_g2)-2), function(i) paste0(light_g2[i:(i+2)], collapse = ""))
)

triplet_positions_g2 <- df_triplets_g2 %>%
  group_by(triplet) %>%
  summarise(
    position1_sessions = list(session_1),
    position2_sessions = list(session_2),
    position3_sessions = list(session_3),
    .groups = "drop"
  )

# Calculate for all participants - FULL 5H
participants <- c("P01", "P02", "P03", "P04", "P05", "P06", "P07", "P09", "P10")
triplets_all <- c("DDD", "DDB", "DBD", "DBB", "BDD", "BDB", "BBD", "BBB")

df_triplet_plot <- data.frame(
  triplet = character(),
  participant = character(),
  mean_auc_pos3 = numeric(),
  stringsAsFactors = FALSE
)

for(p in participants) {
  # Get AUC data with session column
  auc_data <- auc_per_evening %>%
    filter(participant_id == p) %>%
    mutate(session = exp_session)
  
  # Select appropriate triplet mapping
  if(p %in% c("P01", "P02", "P03")) {
    triplet_map <- triplet_positions_g1
  } else if(p %in% c("P04", "P05", "P06", "P07", "P09")) {
    triplet_map <- triplet_positions_g2
  } else {
    # P10 special handling
    df_triplets_g4_trip <- data.frame(
      session_1 = 1:(length(light_g4_orig)-2),
      session_2 = 2:(length(light_g4_orig)-1),
      session_3 = 3:length(light_g4_orig),
      triplet = sapply(1:(length(light_g4_orig)-2), function(i) paste0(light_g4_orig[i:(i+2)], collapse = ""))
    ) %>%
      filter(!str_detect(triplet, "X")) %>%
      filter(session_1 != 5 & session_2 != 5 & session_3 != 5)
    
    triplet_map <- df_triplets_g4_trip %>%
      group_by(triplet) %>%
      summarise(
        position1_sessions = list(session_1),
        position2_sessions = list(session_2),
        position3_sessions = list(session_3),
        .groups = "drop"
      )
  }
  
  # Calculate triplets
  triplet_auc <- calc_triplet_positions(p, auc_data, triplet_map, "auc_melatonin")
  
  # Add to plot data
  for(t in triplets_all) {
    if(t %in% triplet_auc$triplet) {
      df_triplet_plot <- rbind(df_triplet_plot, data.frame(
        triplet = t,
        participant = p,
        mean_auc_pos3 = triplet_auc$pos3_auc[triplet_auc$triplet == t]
      ))
    }
  }
}

df_triplet_plot$triplet <- factor(df_triplet_plot$triplet,
                                  levels = c("DDD", "BDD", "DBD", "BBD",
                                            "DDB", "BDB", "DBB", "BBB"))

# Calculate for experimental window - EXCLUDING P07
participants_exp <- c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10")

df_triplet_plot_exp <- data.frame(
  triplet = character(),
  participant = character(),
  mean_auc_pos3_exp = numeric(),
  stringsAsFactors = FALSE
)

for(p in participants_exp) {
  # Get experimental AUC data with session column
  auc_data_exp <- auc_experimental_samples %>%
    filter(participant_id == p) %>%
    mutate(session = exp_session)
  
  # Select appropriate triplet mapping
  if(p %in% c("P01", "P02", "P03")) {
    triplet_map <- triplet_positions_g1
  } else if(p %in% c("P04", "P05", "P06", "P09")) {
    triplet_map <- triplet_positions_g2
  } else {
    # P10 special handling
    df_triplets_g4_trip <- data.frame(
      session_1 = 1:(length(light_g4_orig)-2),
      session_2 = 2:(length(light_g4_orig)-1),
      session_3 = 3:length(light_g4_orig),
      triplet = sapply(1:(length(light_g4_orig)-2), function(i) paste0(light_g4_orig[i:(i+2)], collapse = ""))
    ) %>%
      filter(!str_detect(triplet, "X")) %>%
      filter(session_1 != 5 & session_2 != 5 & session_3 != 5)
    
    triplet_map <- df_triplets_g4_trip %>%
      group_by(triplet) %>%
      summarise(
        position1_sessions = list(session_1),
        position2_sessions = list(session_2),
        position3_sessions = list(session_3),
        .groups = "drop"
      )
  }
  
  # Calculate triplets
  triplet_auc_exp <- calc_triplet_positions(p, auc_data_exp, triplet_map, "auc_melatonin_exp")
  
  # Add to plot data
  for(t in triplets_all) {
    if(t %in% triplet_auc_exp$triplet) {
      df_triplet_plot_exp <- rbind(df_triplet_plot_exp, data.frame(
        triplet = t,
        participant = p,
        mean_auc_pos3_exp = triplet_auc_exp$pos3_auc[triplet_auc_exp$triplet == t]
      ))
    }
  }
}

df_triplet_plot_exp$triplet <- factor(df_triplet_plot_exp$triplet,
                                      levels = c("DDD", "BDD", "DBD", "BBD",
                                                "DDB", "BDB", "DBB", "BBB"))
```

### Boxplot: 5-hour Exposure

```{r triplet-boxplot-5h}
#| label: fig-triplet-boxplot-5h
#| fig-cap: "Melatonin levels under light exposure triplets (5-h exposure, P07 excluded)"
#| fig-width: 10
#| fig-height: 8

# Labels with bold third letter
triplet_labels <- c(
  "DDD" = expression("DD"*bold("D")),
  "BDD" = expression("BD"*bold("D")),
  "DBD" = expression("DB"*bold("D")),
  "BBD" = expression("BB"*bold("D")),
  "DDB" = expression("DD"*bold("B")),
  "BDB" = expression("BD"*bold("B")),
  "DBB" = expression("DB"*bold("B")),
  "BBB" = expression("BB"*bold("B"))
)

# Colors based on third letter
triplet_colors <- c(
  "DDD" = "#4A90E2", "BDD" = "#4A90E2", "DBD" = "#4A90E2", "BBD" = "#4A90E2",
  "DDB" = "#F5A623", "BDB" = "#F5A623", "DBB" = "#F5A623", "BBB" = "#F5A623"
)

# Exclude P07
df_triplet_plot_noP07 <- df_triplet_plot %>% filter(participant != "P07")

ggplot(df_triplet_plot_noP07, aes(x = triplet, y = mean_auc_pos3)) +
  geom_boxplot(aes(fill = triplet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_jitter(width = 0.15, size = 3, alpha = 0.7, shape = 21, color = "black", fill = "black") +
  scale_fill_manual(values = triplet_colors) +
  scale_x_discrete(labels = triplet_labels) +
  labs(
    title = "Melatonin levels under light exposure triplets",
    subtitle = "Mean third-day melatonin AUC per triplet type (5-h exposure; P07 excluded)",
    x = "Triplet",
    y = "Mean AUC melatonin (pg/mL × h)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )
```

### Data Overview: Triplet Observations

```{r}
# Create detailed overview table
triplet_overview_5h <- df_triplet_plot_noP07 %>%
  arrange(triplet, participant) %>%
  select(participant, triplet, mean_auc_pos3)

DT::datatable(
  triplet_overview_5h,
  options = list(
    pageLength = 20,
    scrollX = TRUE,
    order = list(list(1, 'asc'), list(0, 'asc'))
  ),
  caption = "Individual triplet observations used in boxplot (5-h exposure, third-day values)",
  rownames = FALSE
) %>%
  DT::formatRound(columns = 'mean_auc_pos3', digits = 2) %>%
  DT::formatStyle(
    'triplet',
    backgroundColor = DT::styleEqual(
      c('DDD', 'BDD', 'DBD', 'BBD', 'DDB', 'BDB', 'DBB', 'BBB'),
      c('#e3f2fd', '#e3f2fd', '#e3f2fd', '#e3f2fd', 
        '#fff3e0', '#fff3e0', '#fff3e0', '#fff3e0')
    )
  )
```

### Boxplot: 1.5-hour Exposure

```{r triplet-boxplot-1.5h}
#| label: fig-triplet-boxplot-1.5h
#| fig-cap: "Melatonin levels under light exposure triplets (1.5-h exposure, P07 excluded)"
#| fig-width: 10
#| fig-height: 8

ggplot(df_triplet_plot_exp, aes(x = triplet, y = mean_auc_pos3_exp)) +
  geom_boxplot(aes(fill = triplet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_jitter(width = 0.15, size = 3, alpha = 0.7, shape = 21, color = "black", fill = "black") +
  scale_fill_manual(values = triplet_colors) +
  scale_x_discrete(labels = triplet_labels) +
  labs(
    title = "Melatonin levels under light exposure triplets",
    subtitle = "Mean third-day melatonin AUC per triplet type (1.5-h exposure; P07 excluded)",
    x = "Triplet",
    y = "Mean AUC melatonin (pg/mL × h)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "none"
  )
```

### Data Overview: Triplet Observations (1.5-h)

```{r}
# Create detailed overview table
triplet_overview_1.5h <- df_triplet_plot_exp %>%
  arrange(triplet, participant) %>%
  select(participant, triplet, mean_auc_pos3_exp)

DT::datatable(
  triplet_overview_1.5h,
  options = list(
    pageLength = 20,
    scrollX = TRUE,
    order = list(list(1, 'asc'), list(0, 'asc'))
  ),
  caption = "Individual triplet observations used in boxplot (1.5-h exposure, third-day values)",
  rownames = FALSE
) %>%
  DT::formatRound(columns = 'mean_auc_pos3_exp', digits = 2) %>%
  DT::formatStyle(
    'triplet',
    backgroundColor = DT::styleEqual(
      c('DDD', 'BDD', 'DBD', 'BBD', 'DDB', 'BDB', 'DBB', 'BBB'),
      c('#e3f2fd', '#e3f2fd', '#e3f2fd', '#e3f2fd', 
        '#fff3e0', '#fff3e0', '#fff3e0', '#fff3e0')
    )
  )
```

### Line Plot: Position Comparison (5-hour)

```{r triplet-lineplot-5h}
#| label: fig-triplet-lineplot-5h
#| fig-cap: "Triplet position comparison (5-h exposure, P07 excluded)"
#| fig-width: 10
#| fig-height: 8

# Calculate mean across participants for each position (excluding P07)
participants_noP07 <- c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10")

df_triplet_positions_noP07 <- data.frame(
  triplet = character(),
  position = integer(),
  mean_auc = numeric(),
  light_type = character(),
  stringsAsFactors = FALSE
)

for(t in triplets_all) {
  for(pos in 1:3) {
    values <- c()
    for(p in participants_noP07) {
      # Get AUC data
      auc_data <- auc_per_evening %>%
        filter(participant_id == p) %>%
        mutate(session = exp_session)
      
      # Select mapping
      if(p %in% c("P01", "P02", "P03")) {
        triplet_map <- triplet_positions_g1
      } else if(p %in% c("P04", "P05", "P06", "P09")) {
        triplet_map <- triplet_positions_g2
      } else {
        df_triplets_g4_trip <- data.frame(
          session_1 = 1:(length(light_g4_orig)-2),
          session_2 = 2:(length(light_g4_orig)-1),
          session_3 = 3:length(light_g4_orig),
          triplet = sapply(1:(length(light_g4_orig)-2), function(i) paste0(light_g4_orig[i:(i+2)], collapse = ""))
        ) %>%
          filter(!str_detect(triplet, "X")) %>%
          filter(session_1 != 5 & session_2 != 5 & session_3 != 5)
        
        triplet_map <- df_triplets_g4_trip %>%
          group_by(triplet) %>%
          summarise(
            position1_sessions = list(session_1),
            position2_sessions = list(session_2),
            position3_sessions = list(session_3),
            .groups = "drop"
          )
      }
      
      # Calculate
      triplet_auc <- calc_triplet_positions(p, auc_data, triplet_map, "auc_melatonin")
      
      if(t %in% triplet_auc$triplet) {
        if(pos == 1) values <- c(values, triplet_auc$pos1_auc[triplet_auc$triplet == t])
        if(pos == 2) values <- c(values, triplet_auc$pos2_auc[triplet_auc$triplet == t])
        if(pos == 3) values <- c(values, triplet_auc$pos3_auc[triplet_auc$triplet == t])
      }
    }
    
    if(length(values) > 0) {
      light_letter <- substr(t, pos, pos)
      df_triplet_positions_noP07 <- rbind(df_triplet_positions_noP07, data.frame(
        triplet = t,
        position = pos,
        mean_auc = mean(values, na.rm = TRUE),
        light_type = ifelse(light_letter == "D", "dim", "bright")
      ))
    }
  }
}

df_triplet_positions_noP07$triplet <- factor(df_triplet_positions_noP07$triplet,
                                             levels = triplets_all)

# Define colors
light_colors <- c("dim" = "#4A90E2", "bright" = "#F5A623")

# Create plot
ggplot(df_triplet_positions_noP07, 
       aes(x = position, y = mean_auc, group = triplet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_type), size = 5, shape = 16) +
  scale_color_manual(values = light_colors,
                     labels = c("dim" = "Dim (10 lx)", "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2, 3),
                     labels = c("Position 1", "Position 2", "Position 3"),
                     limits = c(0.8, 3.2)) +
  labs(
    title = "Melatonin levels across triplet positions",
    subtitle = "Mean melatonin AUC across all participants (5-h exposure; P07 excluded)",
    x = "Triplet position",
    y = "Mean AUC melatonin (pg/mL × h)",
    color = "Light condition"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "right"
  )
```

### Line Plot: Position Comparison (1.5-hour)

```{r triplet-lineplot-1.5h}
#| label: fig-triplet-lineplot-1.5h
#| fig-cap: "Triplet position comparison (1.5-h exposure, P07 excluded)"
#| fig-width: 10
#| fig-height: 8

# Calculate mean across participants for each position - experimental window
df_triplet_positions_exp <- data.frame(
  triplet = character(),
  position = integer(),
  mean_auc = numeric(),
  light_type = character(),
  stringsAsFactors = FALSE
)

for(t in triplets_all) {
  for(pos in 1:3) {
    values <- c()
    for(p in participants_exp) {
      # Get experimental AUC data
      auc_data_exp <- auc_experimental_samples %>%
        filter(participant_id == p) %>%
        mutate(session = exp_session)
      
      # Select mapping
      if(p %in% c("P01", "P02", "P03")) {
        triplet_map <- triplet_positions_g1
      } else if(p %in% c("P04", "P05", "P06", "P09")) {
        triplet_map <- triplet_positions_g2
      } else {
        df_triplets_g4_trip <- data.frame(
          session_1 = 1:(length(light_g4_orig)-2),
          session_2 = 2:(length(light_g4_orig)-1),
          session_3 = 3:length(light_g4_orig),
          triplet = sapply(1:(length(light_g4_orig)-2), function(i) paste0(light_g4_orig[i:(i+2)], collapse = ""))
        ) %>%
          filter(!str_detect(triplet, "X")) %>%
          filter(session_1 != 5 & session_2 != 5 & session_3 != 5)
        
        triplet_map <- df_triplets_g4_trip %>%
          group_by(triplet) %>%
          summarise(
            position1_sessions = list(session_1),
            position2_sessions = list(session_2),
            position3_sessions = list(session_3),
            .groups = "drop"
          )
      }
      
      # Calculate
      triplet_auc_exp <- calc_triplet_positions(p, auc_data_exp, triplet_map, "auc_melatonin_exp")
      
      if(t %in% triplet_auc_exp$triplet) {
        if(pos == 1) values <- c(values, triplet_auc_exp$pos1_auc[triplet_auc_exp$triplet == t])
        if(pos == 2) values <- c(values, triplet_auc_exp$pos2_auc[triplet_auc_exp$triplet == t])
        if(pos == 3) values <- c(values, triplet_auc_exp$pos3_auc[triplet_auc_exp$triplet == t])
      }
    }
    
    if(length(values) > 0) {
      light_letter <- substr(t, pos, pos)
      df_triplet_positions_exp <- rbind(df_triplet_positions_exp, data.frame(
        triplet = t,
        position = pos,
        mean_auc = mean(values, na.rm = TRUE),
        light_type = ifelse(light_letter == "D", "dim", "bright")
      ))
    }
  }
}

df_triplet_positions_exp$triplet <- factor(df_triplet_positions_exp$triplet,
                                          levels = triplets_all)

# Create plot
ggplot(df_triplet_positions_exp, 
       aes(x = position, y = mean_auc, group = triplet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_type), size = 5, shape = 16) +
  scale_color_manual(values = light_colors,
                     labels = c("dim" = "Dim (10 lx)", "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2, 3),
                     labels = c("Position 1", "Position 2", "Position 3"),
                     limits = c(0.8, 3.2)) +
  labs(
    title = "Melatonin levels across triplet positions",
    subtitle = "Mean melatonin AUC across all participants (1.5-h exposure; P07 excluded)",
    x = "Triplet position",
    y = "Mean AUC melatonin (pg/mL × h)",
    color = "Light condition"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "right"
  )
```

## Descriptive Statistics: Triplet Medians and IQR

```{r triplet-descriptives-combined}
#| label: triplet-descriptives-combined

# Calculate statistics for 5-h exposure
triplet_stats_5h <- df_triplet_plot_noP07 %>%
  group_by(triplet) %>%
  summarise(
    n_5h = n(),
    median_5h = median(mean_auc_pos3, na.rm = TRUE),
    IQR_5h = IQR(mean_auc_pos3, na.rm = TRUE),
    .groups = "drop"
  )

# Calculate statistics for 1.5-h exposure
triplet_stats_1.5h <- df_triplet_plot_exp %>%
  group_by(triplet) %>%
  summarise(
    n_1.5h = n(),
    median_1.5h = median(mean_auc_pos3_exp, na.rm = TRUE),
    IQR_1.5h = IQR(mean_auc_pos3_exp, na.rm = TRUE),
    .groups = "drop"
  )

# Combine both exposure windows
triplet_stats_combined <- triplet_stats_5h %>%
  full_join(triplet_stats_1.5h, by = "triplet") %>%
  arrange(triplet)

# Display as formatted table
knitr::kable(
  triplet_stats_combined,
  digits = 2,
  col.names = c("Triplet", "N", "Median", "IQR", "N", "Median", "IQR"),
  caption = "Descriptive statistics for triplet third-day melatonin AUC across exposure windows (P07 excluded)",
  align = c("l", rep("r", 6))
) %>%
  kableExtra::add_header_above(c(" " = 1, "5-h Exposure" = 3, "1.5-h Exposure" = 3))
```

## Time-Course Analysis

```{r timecourse-data}
#| label: timecourse-data

# Map light conditions to each session 
df_light <- bind_rows(
  # All participants except P10
  bind_rows(lapply(names(light_patterns), function(p) {
    pattern <- light_patterns[[p]]
    data.frame(
      participant_id = p,
      exp_session = seq_along(pattern),
      light = pattern,
      stringsAsFactors = FALSE
    )
  })),
  # P10 manually
  data.frame(
    participant_id = "P10",
    exp_session = 1:13,
    light = ifelse(1:13 %in% c(1,2,3,6,10,12,13), "D", "B"),
    stringsAsFactors = FALSE
  )
)

# Join with melatonin data
df_melatonin_light <- melatonin %>%
  filter(participant_id %in% c("P01", "P02", "P03", "P04", "P05", "P06", "P07", "P09", "P10")) %>%
  inner_join(df_light, by = c("participant_id", "exp_session"))

# Calculate per-participant summary
df_summary_per_participant <- df_melatonin_light %>%
  group_by(participant_id, sample_num, light) %>%
  summarise(
    mean_melatonin = mean(melatonin_concentration, na.rm = TRUE),
    sd_melatonin = sd(melatonin_concentration, na.rm = TRUE),
    n = n(),
    se_melatonin = sd_melatonin / sqrt(n),
    .groups = "drop"
  ) %>%
  mutate(light = factor(light, levels = c("D", "B"), labels = c("Dim", "Bright")))

# Grand average excluding P07
df_summary_noP07 <- df_summary_per_participant %>%
  filter(participant_id != "P07")

grand_summary_manual <- data.frame()

for(s in unique(df_summary_noP07$sample_num)) {
  for(l in levels(df_summary_noP07$light)) {
    vals <- df_summary_noP07 %>%
      filter(sample_num == s, light == l) %>%
      pull(mean_melatonin)
    
    n_val <- length(vals)
    mean_val <- mean(vals, na.rm = TRUE)
    sd_val <- sd(vals, na.rm = TRUE)
    se_val <- sd_val / sqrt(n_val)
    
    grand_summary_manual <- rbind(grand_summary_manual, data.frame(
      sample_num = s,
      light = l,
      mean_melatonin = mean_val,
      sd_melatonin = sd_val,
      n = n_val,
      se_melatonin = se_val
    ))
  }
}

grand_summary_manual$light <- factor(grand_summary_manual$light, 
                                     levels = c("Bright", "Dim"))
```

```{r timecourse-plot}
#| label: fig-timecourse
#| fig-cap: "Melatonin time course - grand average across participants (P07 excluded)"
#| fig-width: 10
#| fig-height: 8

# Ensure sample_num is numeric
grand_summary_manual <- grand_summary_manual %>%
  mutate(sample_num = as.numeric(sample_num))

# Define labels
hbt_values <- c("5", "4.5", "4", "3.5", "3", "2.5", "2", "1.5", "1", "0.5")

# Y-axis limits
y_min <- floor(min(grand_summary_manual$mean_melatonin - grand_summary_manual$se_melatonin, 
                   na.rm = TRUE) * 10) / 10 - 0.1
y_max <- ceiling(max(grand_summary_manual$mean_melatonin + grand_summary_manual$se_melatonin, 
                     na.rm = TRUE) * 10) / 10 + 0.1

# Create plot
ggplot(grand_summary_manual, 
       aes(x = sample_num, y = mean_melatonin, 
           color = light, fill = light, group = light)) +
  # Background shading
  annotate("rect", xmin = 1, xmax = 6, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  annotate("rect", xmin = 6, xmax = 9, ymin = -Inf, ymax = Inf, 
           fill = "#FFF9E6", alpha = 0.8) +
  annotate("rect", xmin = 9, xmax = 10, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  # Labels
  annotate("text", x = 3, y = y_max - 0.05, label = "dim (10 lx)", 
           hjust = 0.5, vjust = 1, size = 3.5, color = "grey50", fontface = "italic") +
  annotate("text", x = 7.5, y = y_max - 0.05, label = "intervention\n(10 or 1000 lx)", 
           hjust = 0.5, vjust = 1, size = 3.5, color = "#B8860B", fontface = "italic") +
  annotate("text", x = 9.5, y = y_max - 0.05, label = "dim (10 lx)", 
           hjust = 0.5, vjust = 1, size = 3.5, color = "grey50", fontface = "italic") +
  # Ribbon for SE
  geom_ribbon(aes(ymin = mean_melatonin - se_melatonin,
                  ymax = mean_melatonin + se_melatonin),
              alpha = 0.1, color = NA, show.legend = FALSE) +
  # Line and points
  geom_line(linewidth = 0.8) +
  geom_point(size = 3.5, shape = 16) +
  # Colors
  scale_color_manual(
    values = c("Bright" = "#F5A623", "Dim" = "#4A90E2"),
    labels = c("Bright" = "Bright (1000 lx)", "Dim" = "Dim (10 lx)")
  ) +
  scale_fill_manual(
    values = c("Bright" = "#F5A623", "Dim" = "#4A90E2"),
    labels = c("Bright" = "Bright (1000 lx)", "Dim" = "Dim (10 lx)")
  ) +
  # Axes
  scale_x_continuous(
    breaks = 1:10,
    labels = paste0("S", 1:10, "\n(", hbt_values, "h)"),
    limits = c(0.5, 10.5),
    expand = c(0, 0)
  ) +
  scale_y_continuous(limits = c(y_min, y_max), expand = c(0, 0)) +
  # Labels
  labs(
    title = "Melatonin time course",
    subtitle = "Grand average across participants (P07 excluded, n=8)",
    x = "Sample number (hours before HBT)",
    y = "Mean melatonin concentration (pg/mL)",
    color = "Light condition",
    fill = "Light condition"
  ) +
  # Theme
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 13),
    axis.title = element_text(size = 14),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "right",
    plot.margin = margin(15, 10, 10, 10)
  )
```

## Triplet Time-Course Analysis

### Data Preparation
```{r}
# Get session 3 data for each triplet
participants_noP07 <- c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10")
triplets_all <- c("DDD", "DDB", "DBD", "DBB", "BDD", "BDB", "BBD", "BBB")

# Function to get session 3 for each triplet type
get_triplet_session3 <- function(p, triplet_type) {
  if(p %in% c("P01", "P02", "P03")) {
    df_triplets <- df_triplets_g1
  } else if(p %in% c("P04", "P05", "P06", "P09")) {
    df_triplets <- df_triplets_g2
  } else {
    df_triplets <- data.frame(
      session_1 = 1:(length(light_g4_orig)-2),
      session_2 = 2:(length(light_g4_orig)-1),
      session_3 = 3:length(light_g4_orig),
      triplet = sapply(1:(length(light_g4_orig)-2), function(i) 
        paste0(light_g4_orig[i:(i+2)], collapse = ""))
    ) %>%
      filter(!str_detect(triplet, "X")) %>%
      filter(session_1 != 5 & session_2 != 5 & session_3 != 5)
  }
  
  sessions_3 <- df_triplets %>%
    filter(triplet == triplet_type) %>%
    pull(session_3)
  
  return(sessions_3)
}

# Function to get all participant-session combinations for a triplet
get_triplet_sessions_overview <- function(triplet_type) {
  overview <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      for(s in sessions_3) {
        overview <- rbind(overview, data.frame(
          participant_id = p,
          session = s
        ))
      }
    }
  }
  
  return(overview %>% arrange(session, participant_id))
}

# Function to create plot for one triplet type
plot_triplet_timecourse <- function(triplet_type) {
  df_triplet_samples <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      p_data <- melatonin %>%
        filter(participant_id == p, exp_session %in% sessions_3) %>%
        select(participant_id, sample_num, melatonin_concentration)
      
      df_triplet_samples <- rbind(df_triplet_samples, p_data)
    }
  }
  
  triplet_summary <- df_triplet_samples %>%
    group_by(sample_num) %>%
    summarise(
      mean_melatonin = mean(melatonin_concentration, na.rm = TRUE),
      sd_melatonin = sd(melatonin_concentration, na.rm = TRUE),
      n = n(),
      .groups = "drop"
    ) %>%
    mutate(sample_num = as.numeric(sample_num))
  
  y_min <- floor(min(triplet_summary$mean_melatonin - triplet_summary$sd_melatonin, 
                     na.rm = TRUE) * 10) / 10 - 0.1
  y_max <- ceiling(max(triplet_summary$mean_melatonin + triplet_summary$sd_melatonin, 
                       na.rm = TRUE) * 10) / 10 + 0.1
  
  hbt_values <- c("5", "4.5", "4", "3.5", "3", "2.5", "2", "1.5", "1", "0.5")
  
  # Determine third letter and intervention color/label
  third_letter <- substr(triplet_type, 3, 3)
  intervention_color <- ifelse(third_letter == "D", "#F5F5F5", "#FFF9E6")
  intervention_alpha <- ifelse(third_letter == "D", 0.4, 0.8)
  intervention_label <- ifelse(third_letter == "D", "dim (10 lx)", "bright (1000 lx)")
  
  # Create title with bold third letter
  triplet_first_two <- substr(triplet_type, 1, 2)
  triplet_third <- substr(triplet_type, 3, 3)
  
  p <- ggplot(triplet_summary, aes(x = sample_num, y = mean_melatonin)) +
    annotate("rect", xmin = 1, xmax = 6, ymin = -Inf, ymax = Inf, 
             fill = "#F5F5F5", alpha = 0.4) +
    annotate("rect", xmin = 6, xmax = 9, ymin = -Inf, ymax = Inf, 
             fill = intervention_color, alpha = intervention_alpha) +
    annotate("rect", xmin = 9, xmax = 10, ymin = -Inf, ymax = Inf, 
             fill = "#F5F5F5", alpha = 0.4) +
    annotate("text", x = 3, y = y_max - 0.05, label = "dim (10 lx)", 
             hjust = 0.5, vjust = 1, size = 3.5, color = "grey50", fontface = "italic") +
    annotate("text", x = 7.5, y = y_max - 0.05, label = intervention_label, 
             hjust = 0.5, vjust = 1, size = 3.5, 
             color = ifelse(third_letter == "D", "grey50", "#B8860B"), 
             fontface = "italic") +
    annotate("text", x = 9.5, y = y_max - 0.05, label = "dim (10 lx)", 
             hjust = 0.5, vjust = 1, size = 3.5, color = "grey50", fontface = "italic") +
    geom_ribbon(aes(ymin = mean_melatonin - sd_melatonin,
                    ymax = mean_melatonin + sd_melatonin),
                alpha = 0.15, fill = "grey40") +
    geom_line(linewidth = 0.8, color = "black") +
    geom_point(size = 3.5, shape = 16, color = "black") +
    scale_x_continuous(
      breaks = 1:10,
      labels = paste0("S", 1:10, "\n(", hbt_values, "h)"),
      limits = c(0.5, 10.5),
      expand = c(0, 0)
    ) +
    scale_y_continuous(limits = c(y_min, y_max), expand = c(0, 0)) +
    labs(
      title = bquote("Melatonin time course - Triplet " * 
                     .(triplet_first_two) * bold(.(triplet_third))),
      subtitle = paste0("Grand average third-day (n=", 
                       length(unique(df_triplet_samples$participant_id)), 
                       " participants; P07 excluded)"),
      x = "Sample number (hours before HBT)",
      y = "Mean melatonin concentration (pg/mL)"
    ) +
    theme_classic(base_size = 14) +
    theme(
      aspect.ratio = 1,
      plot.title = element_text(size = 16, hjust = 0),
      plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
      axis.text = element_text(color = "black", size = 13),
      axis.title = element_text(size = 14),
      axis.title.x = element_text(margin = margin(t = 8)),
      axis.title.y = element_text(margin = margin(r = 8)),
      panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
      plot.margin = margin(15, 10, 10, 10)
    )
  
  return(p)
}
```

### DDD Triplet
```{r}
# Session overview
knitr::kable(get_triplet_sessions_overview("DDD"),
             caption = "Sessions used for Triplet DDD")
```
```{r}
plot_triplet_timecourse("DDD")
```

### DDB Triplet
```{r}
knitr::kable(get_triplet_sessions_overview("DDB"),
             caption = "Sessions used for Triplet DDB")
```
```{r}
plot_triplet_timecourse("DDB")
```

### DBD Triplet
```{r}
knitr::kable(get_triplet_sessions_overview("DBD"),
             caption = "Sessions used for Triplet DBD")
```
```{r}
plot_triplet_timecourse("DBD")
```

### DBB Triplet
```{r}
knitr::kable(get_triplet_sessions_overview("DBB"),
             caption = "Sessions used for Triplet DBB")
```
```{r}
plot_triplet_timecourse("DBB")
```

### BDD Triplet
```{r}
knitr::kable(get_triplet_sessions_overview("BDD"),
             caption = "Sessions used for Triplet BDD")
```
```{r}
plot_triplet_timecourse("BDD")
```

### BDB Triplet
```{r}
knitr::kable(get_triplet_sessions_overview("BDB"),
             caption = "Sessions used for Triplet BDB")
```
```{r}
plot_triplet_timecourse("BDB")
```

### BBD Triplet
```{r}
knitr::kable(get_triplet_sessions_overview("BBD"),
             caption = "Sessions used for Triplet BBD")
```
```{r}
plot_triplet_timecourse("BBD")
```

### BBB Triplet
```{r}
knitr::kable(get_triplet_sessions_overview("BBB"),
             caption = "Sessions used for Triplet BBB")
```
```{r}
plot_triplet_timecourse("BBB")
```

## Panel Plot: Dim-Ending Triplets (DDD, BDD, DBD, BBD)
```{r}
# Prepare data for all D-ending triplets
triplets_D_panel <- c("DDD", "BDD", "DBD", "BBD")

df_all_D_triplets <- data.frame()

for(triplet_type in triplets_D_panel) {
  df_triplet_samples <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      p_data <- melatonin %>%
        filter(participant_id == p, exp_session %in% sessions_3) %>%
        select(participant_id, sample_num, melatonin_concentration) %>%
        mutate(triplet = triplet_type)
      
      df_triplet_samples <- rbind(df_triplet_samples, p_data)
    }
  }
  
  df_all_D_triplets <- rbind(df_all_D_triplets, df_triplet_samples)
}

# Calculate summary per triplet
triplet_summary_D <- df_all_D_triplets %>%
  group_by(triplet, sample_num) %>%
  summarise(
    mean_melatonin = mean(melatonin_concentration, na.rm = TRUE),
    sd_melatonin = sd(melatonin_concentration, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    sample_num = as.numeric(sample_num),
    triplet = factor(triplet, levels = triplets_D_panel)
  )

# Y-axis limits for D-ending triplets
y_min_panel_D <- floor(min(triplet_summary_D$mean_melatonin - triplet_summary_D$sd_melatonin, 
                           na.rm = TRUE) * 10) / 10 - 0.1
y_max_panel_D <- ceiling(max(triplet_summary_D$mean_melatonin + triplet_summary_D$sd_melatonin, 
                             na.rm = TRUE) * 10) / 10 + 0.1

hbt_values <- c("5", "4.5", "4", "3.5", "3", "2.5", "2", "1.5", "1", "0.5")

# Create panel plot
ggplot(triplet_summary_D, aes(x = sample_num, y = mean_melatonin)) +
  annotate("rect", xmin = 1, xmax = 6, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  annotate("rect", xmin = 6, xmax = 9, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  annotate("rect", xmin = 9, xmax = 10, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  geom_ribbon(aes(ymin = mean_melatonin - sd_melatonin,
                  ymax = mean_melatonin + sd_melatonin),
              alpha = 0.15, fill = "grey40") +
  geom_line(linewidth = 0.8, color = "black") +
  geom_point(size = 2.5, shape = 16, color = "black") +
  scale_x_continuous(
    breaks = 1:10,
    labels = paste0("S", 1:10, "\n(", hbt_values, "h)"),
    limits = c(0.5, 10.5),
    expand = c(0, 0)
  ) +
  scale_y_continuous(limits = c(y_min_panel_D, y_max_panel_D), expand = c(0, 0)) +
  facet_wrap(~ triplet, ncol = 2, 
             labeller = labeller(triplet = function(x) {
               sapply(x, function(t) {
                 first_two <- substr(t, 1, 2)
                 third <- substr(t, 3, 3)
                 paste0(first_two, third)
               })
             })) +
  labs(
    title = "Melatonin time course - Dim-ending triplets",
    subtitle = "Grand average third-day (P07 excluded)",
    x = "Sample number (hours before HBT)",
    y = "Mean melatonin concentration (pg/mL)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 11),
    axis.title = element_text(size = 13),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    strip.background = element_rect(fill = "white", color = "grey80", linewidth = 0.5),
    strip.text = element_text(size = 12, color = "black", face = "bold"),
    plot.margin = margin(15, 10, 10, 10)
  )
```

### Panel Plot: Bright-Ending Triplets (DDB, BDB, DBB, BBB)
```{r}
# Prepare data for all B-ending triplets
triplets_B_panel <- c("DDB", "BDB", "DBB", "BBB")

df_all_B_triplets <- data.frame()

for(triplet_type in triplets_B_panel) {
  df_triplet_samples <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      p_data <- melatonin %>%
        filter(participant_id == p, exp_session %in% sessions_3) %>%
        select(participant_id, sample_num, melatonin_concentration) %>%
        mutate(triplet = triplet_type)
      
      df_triplet_samples <- rbind(df_triplet_samples, p_data)
    }
  }
  
  df_all_B_triplets <- rbind(df_all_B_triplets, df_triplet_samples)
}

# Calculate summary per triplet
triplet_summary_B <- df_all_B_triplets %>%
  group_by(triplet, sample_num) %>%
  summarise(
    mean_melatonin = mean(melatonin_concentration, na.rm = TRUE),
    sd_melatonin = sd(melatonin_concentration, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    sample_num = as.numeric(sample_num),
    triplet = factor(triplet, levels = triplets_B_panel)
  )

# Y-axis limits for B-ending triplets
y_min_panel_B <- floor(min(triplet_summary_B$mean_melatonin - triplet_summary_B$sd_melatonin, 
                           na.rm = TRUE) * 10) / 10 - 0.1
y_max_panel_B <- ceiling(max(triplet_summary_B$mean_melatonin + triplet_summary_B$sd_melatonin, 
                             na.rm = TRUE) * 10) / 10 + 0.1

hbt_values <- c("5", "4.5", "4", "3.5", "3", "2.5", "2", "1.5", "1", "0.5")

# Create panel plot
ggplot(triplet_summary_B, aes(x = sample_num, y = mean_melatonin)) +
  annotate("rect", xmin = 1, xmax = 6, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  annotate("rect", xmin = 6, xmax = 9, ymin = -Inf, ymax = Inf, 
           fill = "#FFF9E6", alpha = 0.8) +
  annotate("rect", xmin = 9, xmax = 10, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  geom_ribbon(aes(ymin = mean_melatonin - sd_melatonin,
                  ymax = mean_melatonin + sd_melatonin),
              alpha = 0.15, fill = "grey40") +
  geom_line(linewidth = 0.8, color = "black") +
  geom_point(size = 2.5, shape = 16, color = "black") +
  scale_x_continuous(
    breaks = 1:10,
    labels = paste0("S", 1:10, "\n(", hbt_values, "h)"),
    limits = c(0.5, 10.5),
    expand = c(0, 0)
  ) +
  scale_y_continuous(limits = c(y_min_panel_B, y_max_panel_B), expand = c(0, 0)) +
  facet_wrap(~ triplet, ncol = 2, 
             labeller = labeller(triplet = function(x) {
               sapply(x, function(t) {
                 first_two <- substr(t, 1, 2)
                 third <- substr(t, 3, 3)
                 paste0(first_two, third)
               })
             })) +
  labs(
    title = "Melatonin time course - Bright-ending triplets",
    subtitle = "Grand average third-day (P07 excluded)",
    x = "Sample number (hours before HBT)",
    y = "Mean melatonin concentration (pg/mL)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 11),
    axis.title = element_text(size = 13),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    strip.background = element_rect(fill = "white", color = "grey80", linewidth = 0.5),
    strip.text = element_text(size = 12, color = "black", face = "bold"),
    plot.margin = margin(15, 10, 10, 10)
  )
```

### Panel Plot: All Triplets with Unified Y-Axis
```{r}
# Prepare data for all triplets
triplets_all_panel <- c("DDD", "BDD", "DBD", "BBD", "DDB", "BDB", "DBB", "BBB")

df_all_triplets_unified <- data.frame()

for(triplet_type in triplets_all_panel) {
  df_triplet_samples <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      p_data <- melatonin %>%
        filter(participant_id == p, exp_session %in% sessions_3) %>%
        select(participant_id, sample_num, melatonin_concentration) %>%
        mutate(triplet = triplet_type)
      
      df_triplet_samples <- rbind(df_triplet_samples, p_data)
    }
  }
  
  df_all_triplets_unified <- rbind(df_all_triplets_unified, df_triplet_samples)
}

# Calculate summary per triplet
triplet_summary_unified <- df_all_triplets_unified %>%
  group_by(triplet, sample_num) %>%
  summarise(
    mean_melatonin = mean(melatonin_concentration, na.rm = TRUE),
    sd_melatonin = sd(melatonin_concentration, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    sample_num = as.numeric(sample_num),
    triplet = factor(triplet, levels = triplets_all_panel),
    triplet_type = ifelse(substr(triplet, 3, 3) == "D", "Dim", "Bright")
  )

# Unified Y-axis limits for ALL triplets
y_min_unified <- floor(min(triplet_summary_unified$mean_melatonin - triplet_summary_unified$sd_melatonin, 
                           na.rm = TRUE) * 10) / 10 - 0.1
y_max_unified <- ceiling(max(triplet_summary_unified$mean_melatonin + triplet_summary_unified$sd_melatonin, 
                             na.rm = TRUE) * 10) / 10 + 0.1

# Create unified panel plot
ggplot(triplet_summary_unified, aes(x = sample_num, y = mean_melatonin)) +
  geom_rect(aes(fill = triplet_type), 
            xmin = 6, xmax = 9, ymin = -Inf, ymax = Inf, alpha = 0.3) +
  annotate("rect", xmin = 1, xmax = 6, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  annotate("rect", xmin = 9, xmax = 10, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  geom_ribbon(aes(ymin = mean_melatonin - sd_melatonin,
                  ymax = mean_melatonin + sd_melatonin),
              alpha = 0.15, fill = "grey40") +
  geom_line(linewidth = 0.8, color = "black") +
  geom_point(size = 2.5, shape = 16, color = "black") +
  scale_fill_manual(values = c("Dim" = "#F5F5F5", "Bright" = "#FFF9E6"), guide = "none") +
  scale_x_continuous(
    breaks = 1:10,
    labels = c("S1\n(5h)", "S2", "S3\n(4h)", "S4", "S5\n(3h)", 
               "S6", "S7\n(2h)", "S8", "S9\n(1h)", "S10"),
    limits = c(0.5, 10.5),
    expand = c(0, 0)
  ) +
  scale_y_continuous(limits = c(y_min_unified, y_max_unified), expand = c(0, 0)) +
  facet_wrap(~ triplet, ncol = 4, 
             labeller = labeller(triplet = function(x) {
               sapply(x, function(t) {
                 first_two <- substr(t, 1, 2)
                 third <- substr(t, 3, 3)
                 paste0(first_two, third)
               })
             })) +
  labs(
    title = "Melatonin time course - All triplets",
    subtitle = "Grand average third-day (P07 excluded)",
    x = "Sample number (hours before HBT)",
    y = "Mean melatonin concentration (pg/mL)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(size = 13),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    strip.background = element_rect(fill = "white", color = "grey80", linewidth = 0.5),
    strip.text = element_text(size = 11, color = "black", face = "bold"),
    plot.margin = margin(15, 10, 10, 10)
  )
```

### Descriptive Statistics: Triplet Time-Course Values
```{r}
# Create table with mean ± SD for all triplets
triplet_timecourse_table <- triplet_summary_unified %>%
  mutate(
    mean_sd = sprintf("%.2f ± %.2f", mean_melatonin, sd_melatonin)
  ) %>%
  select(triplet, sample_num, mean_sd) %>%
  pivot_wider(
    names_from = sample_num,
    values_from = mean_sd,
    names_prefix = "S"
  )

knitr::kable(
  triplet_timecourse_table,
  caption = "Triplet time-course: Mean ± SD melatonin concentration (pg/mL) per sample",
  align = c("l", rep("r", 10))
) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                           full_width = FALSE) %>%
  kableExtra::row_spec(0, bold = TRUE)
```

### Grand Average with D-ending and B-ending rows
```{r fig.width=14, fig.height=8}
# Prepare data for all triplets
triplets_all_panel <- c("DDD", "BDD", "DBD", "BBD", "DDB", "BDB", "DBB", "BBB")

df_all_triplets_unified <- data.frame()

for(triplet_type in triplets_all_panel) {
  df_triplet_samples <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      p_data <- melatonin %>%
        filter(participant_id == p, exp_session %in% sessions_3) %>%
        select(participant_id, sample_num, melatonin_concentration) %>%
        mutate(triplet = triplet_type)
      
      df_triplet_samples <- rbind(df_triplet_samples, p_data)
    }
  }
  
  df_all_triplets_unified <- rbind(df_all_triplets_unified, df_triplet_samples)
}

# Calculate summary per triplet
triplet_summary_unified <- df_all_triplets_unified %>%
  group_by(triplet, sample_num) %>%
  summarise(
    mean_melatonin = mean(melatonin_concentration, na.rm = TRUE),
    sd_melatonin = sd(melatonin_concentration, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    sample_num = as.numeric(sample_num),
    triplet = factor(triplet, levels = triplets_all_panel)
  )

# Unified Y-axis limits
y_min_unified <- floor(min(triplet_summary_unified$mean_melatonin - triplet_summary_unified$sd_melatonin, 
                           na.rm = TRUE) * 10) / 10 - 0.1
y_max_unified <- ceiling(max(triplet_summary_unified$mean_melatonin + triplet_summary_unified$sd_melatonin, 
                             na.rm = TRUE) * 10) / 10 + 0.1

# Function to create triplet plot
create_triplet_plot_mel <- function(triplet_name, show_x = FALSE, show_y = TRUE) {
  data_triplet <- triplet_summary_unified %>%
    filter(triplet == triplet_name)
  
  triplet_third <- substr(triplet_name, 3, 3)
  intervention_color <- ifelse(triplet_third == "D", "#F5F5F5", "#FFF9E6")
  intervention_alpha <- ifelse(triplet_third == "D", 0.4, 0.8)
  
  # Only third letter bold: DD**D**
  triplet_first_two <- substr(triplet_name, 1, 2)
  triplet_label <- paste0(triplet_first_two, "**", triplet_third, "**")
  
  hbt_values <- c("5", "4.5", "4", "3.5", "3", "2.5", "2", "1.5", "1", "0.5")
  
  ggplot(data_triplet, aes(x = sample_num, y = mean_melatonin)) +
    annotate("rect", xmin = 1, xmax = 6, ymin = -Inf, ymax = Inf, 
             fill = "#F5F5F5", alpha = 0.4) +
    annotate("rect", xmin = 6, xmax = 9, ymin = -Inf, ymax = Inf, 
             fill = intervention_color, alpha = intervention_alpha) +
    annotate("rect", xmin = 9, xmax = 10, ymin = -Inf, ymax = Inf, 
             fill = "#F5F5F5", alpha = 0.4) +
    geom_ribbon(aes(ymin = mean_melatonin - sd_melatonin, 
                    ymax = mean_melatonin + sd_melatonin),
                alpha = 0.15, fill = "grey40") +
    geom_line(linewidth = 0.8, color = "black") +
    geom_point(size = 2.5, shape = 16, color = "black") +
    scale_x_continuous(
      breaks = 1:10,
      labels = if(show_x) c("S1\n(5h)", "S2", "S3\n(4h)", "S4", "S5\n(3h)", 
                            "S6", "S7\n(2h)", "S8", "S9\n(1h)", "S10") else 1:10,
      limits = c(0.5, 10.5),
      expand = c(0, 0)
    ) +
    scale_y_continuous(limits = c(y_min_unified, y_max_unified), expand = c(0, 0)) +
    labs(
      title = triplet_label,
      x = NULL,
      y = if(show_y) "Mean melatonin\nconcentration ± SD (pg/mL)" else NULL
    ) +
    theme_classic(base_size = 11) +
    theme(
      plot.title = element_markdown(size = 11, hjust = 0.5, margin = margin(b = 5)),
      axis.text = element_text(color = "black", size = 9),
      axis.text.x = element_text(size = 8),
      axis.title = element_text(size = 10),
      axis.title.y = element_text(margin = margin(r = 5)),
      panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
      plot.margin = margin(5, 5, 5, 5)
    )
}

# Create D-ending triplet plots (no x-axis labels)
p_DDD_mel <- create_triplet_plot_mel("DDD", show_x = FALSE, show_y = TRUE)
p_BDD_mel <- create_triplet_plot_mel("BDD", show_x = FALSE, show_y = FALSE)
p_DBD_mel <- create_triplet_plot_mel("DBD", show_x = FALSE, show_y = FALSE)
p_BBD_mel <- create_triplet_plot_mel("BBD", show_x = FALSE, show_y = FALSE)

# Create B-ending triplet plots (with x-axis labels)
p_DDB_mel <- create_triplet_plot_mel("DDB", show_x = TRUE, show_y = TRUE)
p_BDB_mel <- create_triplet_plot_mel("BDB", show_x = TRUE, show_y = FALSE)
p_DBB_mel <- create_triplet_plot_mel("DBB", show_x = TRUE, show_y = FALSE)
p_BBB_mel <- create_triplet_plot_mel("BBB", show_x = TRUE, show_y = FALSE)

# Grand Average plot (same size as individual panels)
p_grand_mel <- ggplot(grand_summary_manual, 
       aes(x = sample_num, y = mean_melatonin, 
           color = light, fill = light, group = light)) +
  annotate("rect", xmin = 1, xmax = 6, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  annotate("rect", xmin = 9, xmax = 10, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  annotate("rect", xmin = 6, xmax = 9, ymin = -Inf, ymax = Inf, 
           fill = "#FFF9E6", alpha = 0.8) +
  geom_ribbon(aes(ymin = mean_melatonin - se_melatonin, 
                  ymax = mean_melatonin + se_melatonin), 
              alpha = 0.1, color = NA, show.legend = FALSE) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.5, shape = 16) +
  scale_color_manual(
    values = c("Bright" = "#F5A623", "Dim" = "#4A90E2"),
    labels = c("Bright" = "Bright (1000 lx)", "Dim" = "Dim (10 lx)")
  ) +
  scale_fill_manual(
    values = c("Bright" = "#F5A623", "Dim" = "#4A90E2"),
    labels = c("Bright" = "Bright (1000 lx)", "Dim" = "Dim (10 lx)")
  ) +
  scale_x_continuous(
    breaks = 1:10,
    labels = 1:10,
    limits = c(0.5, 10.5),
    expand = c(0, 0)
  ) +
  scale_y_continuous(limits = c(y_min_unified, y_max_unified), expand = c(0, 0)) +
  labs(
    title = "Grand Average",
    x = NULL,
    y = NULL,
    color = "Light condition"
  ) +
  guides(
    color = guide_legend(override.aes = list(shape = 16, size = 2.5)),
    fill = "none"
  ) +
  theme_classic(base_size = 11) +
  theme(
    plot.title = element_text(size = 11, hjust = 0.5, face = "bold", margin = margin(b = 5)),
    axis.text = element_text(color = "black", size = 9),
    axis.text.x = element_text(size = 9),
    axis.title = element_text(size = 10),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "bottom",
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    plot.margin = margin(5, 5, 5, 5)
  )

# Empty spacer
p_spacer <- ggplot() + theme_void()

# Combine in layout:
# Row 1: DDD, BDD, DBD, BBD, Grand Average
# Row 2: DDB, BDB, DBB, BBB, Spacer
(p_DDD_mel | p_BDD_mel | p_DBD_mel | p_BBD_mel | p_grand_mel) /
(p_DDB_mel | p_BDB_mel | p_DBB_mel | p_BBB_mel | p_spacer) +
  plot_annotation(
    title = "Melatonin time course - Grand average and triplet patterns",
    subtitle = "Grand average and third-day triplet patterns (P07 excluded)",
    caption = "Sample number (hours before HBT)",
    theme = theme(
      plot.title = element_text(size = 15, hjust = 0, face = "plain"),
      plot.subtitle = element_text(size = 12, hjust = 0, margin = margin(b = 8)),
      plot.caption = element_text(size = 11, hjust = 0.5, margin = margin(t = 8))
    )
  )
```

# Data Preparation for Statistical Analysis

## Define Study Parameters

```{r}
#| label: lmm-data-prep

# Define participant list and light patterns
participant_list <- c("P01", "P02", "P03", "P04", "P05", "P06", "P07", "P09", "P10")

# Combine melatonin data with light information
melatonin_glmm <- map_dfr(participant_list, ~{
  df <- auc_experimental_samples %>% 
    filter(participant_id == .x)
  
  light_df <- tibble(
    exp_session = seq_along(light_patterns[[.x]]),
    Light_Lag0 = light_patterns[[.x]]
  )
  
  df <- df %>% 
    left_join(light_df, by = "exp_session") %>%
    rename(auc_melatonin_exp = auc_melatonin_exp)
  
  return(df)
})

# Build lagged light variables and filter valid triplets
melatonin_glmm <- melatonin_glmm %>%
  group_by(participant_id) %>%
  arrange(exp_session) %>%
  mutate(
    Light_Lag1 = lag(Light_Lag0, 1),
    Light_Lag2 = lag(Light_Lag0, 2)
  ) %>%
  ungroup() %>%
  filter(!is.na(Light_Lag0) & !is.na(Light_Lag1) & !is.na(Light_Lag2)) %>%
  filter(!(participant_id == "P10" & exp_session %in% c(6, 7))) %>%
  mutate(
    Light_Lag0_num = ifelse(Light_Lag0 == "D", 0, 1),
    Light_Lag1_num = ifelse(Light_Lag1 == "D", 0, 1),
    Light_Lag2_num = ifelse(Light_Lag2 == "D", 0, 1)
  )

# Add participant demographics
participant_demographics <- tibble(
  participant_id = c("P01", "P02", "P03", "P04", "P05", "P06", "P07", "P09", "P10"),
  sex = c("female", "male", "female", "female", "male", "female", "female", "female", "male"),
  age = c(24, 21, 24, 23, 31, 23, 24, 28, 32)
)

melatonin_glmm <- melatonin_glmm %>%
  left_join(participant_demographics, by = "participant_id")

# Apply exclusion criteria
melatonin_glmm <- melatonin_glmm %>%
  filter(participant_id != "P07")  # Exclude P07 (negative melatonin suppression)

# Keep only necessary variables
melatonin_glmm <- melatonin_glmm %>%
  dplyr::select(
    participant_id, exp_session, auc_melatonin_exp,
    Light_Lag0, Light_Lag1, Light_Lag2,
    Light_Lag0_num, Light_Lag1_num, Light_Lag2_num,
    sex, age
  )
```

## Dataset Overview

```{r}
#| label: dataset-display

# Display structure
str(melatonin_glmm)

# Display first rows
head(melatonin_glmm, 10)

# Summary
cat("Total observations:", nrow(melatonin_glmm), "\n")
cat("Participants:", length(unique(melatonin_glmm$participant_id)), "\n")
cat("Sessions range:", min(melatonin_glmm$exp_session), "-", max(melatonin_glmm$exp_session), "\n")
```

### Dataset meltaonin_glmm

```{r}
#| label: dataset-table

DT::datatable(
  melatonin_glmm,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    order = list(list(0, 'asc'), list(1, 'asc'))
  ),
  caption = "Complete dataset for statistical analysis (n=67, P07 excluded)",
  rownames = FALSE
) %>%
  DT::formatRound(columns = 'auc_melatonin_exp', digits = 2) %>%
  DT::formatStyle(
    'Light_Lag0',
    backgroundColor = DT::styleEqual(c('D', 'B'), c('#e3f2fd', '#fff3e0'))
  )
```

# Descriptive Statistics

## Summary Statistics

```{r}
summary(melatonin_glmm$auc_melatonin_exp)

data.frame(
  N = nrow(melatonin_glmm),
  N_participants = length(unique(melatonin_glmm$participant_id)),
  Mean = round(mean(melatonin_glmm$auc_melatonin_exp), 2),
  SD = round(sd(melatonin_glmm$auc_melatonin_exp), 2),
  Min = round(min(melatonin_glmm$auc_melatonin_exp), 2),
  Max = round(max(melatonin_glmm$auc_melatonin_exp), 2)
)
```

## Distribution Visualization

```{r}
#| fig-width: 8
#| fig-height: 5

ggplot(melatonin_glmm, aes(x = auc_melatonin_exp)) +
  geom_histogram(aes(y = after_stat(density)), bins = 15, 
                 fill = "#4A90E2", alpha = 0.7, color = "black") +
  geom_density(color = "red", linewidth = 1) +
  geom_vline(xintercept = mean(melatonin_glmm$auc_melatonin_exp), 
             linetype = "dashed", color = "red") +
  labs(title = "Distribution of Melatonin AUC", 
       x = "AUC (pg/mL × h)", 
       y = "Density") +
  theme_classic()
```

## Melatonin AUC Trajectories Across Experimental Sessions

Individual participant trajectories with LOESS smooth (full dataset, n=67 observations)
```{r}
#| label: fig-melatonin-trajectories
#| fig-width: 10
#| fig-height: 7

# Prepare data - count observations per session
session_counts_mel <- melatonin_glmm %>%
  group_by(exp_session) %>%
  summarise(n = n(), .groups = "drop")

# Create plot
ggplot(melatonin_glmm, 
       aes(x = exp_session, y = auc_melatonin_exp, 
           color = participant_id, group = participant_id)) +
  
  # Individual trajectories
  geom_line(alpha = 0.6, linewidth = 0.8) +
  geom_point(size = 2.5, alpha = 0.7) +
  
  # LOESS smooth (overall trend)
  geom_smooth(aes(group = 1), method = "loess", span = 0.75,
              color = "black", linetype = "dashed", linewidth = 1.2,
              se = TRUE, fill = "grey50", alpha = 0.3) +
  
  # Add sample size annotations for reduced sessions
  geom_text(data = session_counts_mel %>% filter(n < 8),
            aes(x = exp_session, y = max(melatonin_glmm$auc_melatonin_exp) * 0.95, 
                label = paste0("n=", n)),
            color = "black", size = 3, inherit.aes = FALSE) +
  
  # Vertical line for data cutoff (session 10)
  geom_vline(xintercept = 10.5, linetype = "dotted", color = "grey50", linewidth = 0.8) +
  
  # Scales
  scale_x_continuous(
    breaks = 3:13,
    limits = c(2.5, 13.5),
    expand = c(0.01, 0)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.05))
  ) +
  
  # Colors (same as KSS for consistency)
  scale_color_manual(
    values = c("P01" = "#E41A1C", "P02" = "#377EB8", "P03" = "#4DAF4A",
               "P04" = "#984EA3", "P05" = "#FF7F00", "P06" = "#FFFF33",
               "P09" = "#A65628", "P10" = "#F781BF"),
    name = "Participant"
  ) +
  
  # Labels
  labs(
    title = "Melatonin AUC across experimental sessions",
    subtitle = "Individual participant trajectories with LOESS smooth (full dataset, n=67, P07 excluded)",
    x = "Experimental Session",
    y = "Melatonin AUC (pg/mL × h)"
  ) +
  
  # Theme
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, face = "plain"),
    plot.subtitle = element_text(size = 12, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 12),
    axis.title = element_text(size = 13),
    legend.position = "right",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    plot.margin = margin(15, 15, 10, 10)
  )
```

## Light Condition Balance

```{r}
melatonin_glmm %>%
  summarise(
    Lag0_Dim = sum(Light_Lag0_num == 0),
    Lag0_Bright = sum(Light_Lag0_num == 1),
    Lag1_Dim = sum(Light_Lag1_num == 0),
    Lag1_Bright = sum(Light_Lag1_num == 1),
    Lag2_Dim = sum(Light_Lag2_num == 0),
    Lag2_Bright = sum(Light_Lag2_num == 1)
  )
```

## Observations per Session

```{r}
table(melatonin_glmm$exp_session)
```

## Observations per Participant

```{r}
table(melatonin_glmm$participant_id)
```

# Statistical Analysis

### Model m0: Baseline (Random Intercept Only)

This baseline model quantifies between-participant variability without any predictors.
```{r}
#| label: model-m0-melatonin

m0_mel <- glmer(
  auc_melatonin_exp ~ 1 + (1 | participant_id),
  data = melatonin_glmm,
  family = Gamma(link = "log")
)

summary(m0_mel)
```

### Model m1a: Add Light Lag 0 (Current day)

This model adds the current day light exposure variable as an fixed effect with Light_Lag0_num. 
```{r}
#| label: model-m1a-melatonin

m1a_mel <- glmer(
  auc_melatonin_exp ~ 
    Light_Lag0_num +
    (1 | participant_id),
  data = melatonin_glmm,
  family = Gamma(link = "log")
)

summary(m1a_mel)
```

### Model m1b: Add Light Lag 1 Additionally (One Day Before)

This model adds the day before light exposure variable as an fixed effect with Light_Lag1_num. 
```{r}
#| label: model-m1b-melatonin

m1b_mel <- glmer(
  auc_melatonin_exp ~ 
    Light_Lag0_num + Light_Lag1_num +
    (1 | participant_id),
  data = melatonin_glmm,
  family = Gamma(link = "log")
)

summary(m1b_mel)
```

### Model m1c: Add Light Lag 2 Additionally (Two Days Before)

This model adds the light exposure two days before as an fixed effect with Light_Lag2_num. 
```{r}
#| label: model-m1c-melatonin

m1c_mel <- glmer(
  auc_melatonin_exp ~ 
    Light_Lag0_num + Light_Lag1_num + Light_Lag2_num +
    (1 | participant_id),
  data = melatonin_glmm,
  family = Gamma(link = "log")
)

summary(m1c_mel)
```

### Likelihood Ratio Tests: Lag Effects

```{r}
#| label: lrt-comparisons-melatonin

# Likelihood Ratio Tests for nested model comparisons
# Test whether adding each lag significantly improves model fit

# LRT 1: Does adding Lag 0 improve fit over baseline?
lrt_m0_vs_m1a <- anova(m0_mel, m1a_mel)

# LRT 2: Does adding Lag 1 improve fit over Lag 0 only?
lrt_m1a_vs_m1b <- anova(m1a_mel, m1b_mel)

# LRT 3: Does adding Lag 2 improve fit over Lag 0 + Lag 1?
lrt_m1b_vs_m1c <- anova(m1b_mel, m1c_mel)

# Display results
lrt_m0_vs_m1a
lrt_m1a_vs_m1b
lrt_m1b_vs_m1c
```
Based on the LRT results, only Lag 0 significantly improved model fit (p < 0.001), while Lag 1 (p = 0.814) and Lag 2 (p = 0.329) showed no significant improvement.Therefore, Lag 1 and Lag 2 were excluded from further model building, and only Lag 0 was retained as a predictor of melatonin AUC.

### Model m2: Add Linear Session Effect

This model adds a linear session term to account for potential temporal trends across the study.
```{r}
#| label: model-m2-melatonin

m2_mel <- glmer(
  auc_melatonin_exp ~ 
    Light_Lag0_num +
    exp_session +
    (1 | participant_id),
  data = melatonin_glmm,
  family = Gamma(link = "log")
)

summary(m2_mel)
```

### Model m3: Add Quadratic Session Effect

This model tests for non-linear temporal trends (e.g., U-shaped patterns) across sessions.
```{r}
#| label: model-m3-melatonin

m3_mel <- glmer(
  auc_melatonin_exp ~ 
    Light_Lag0_num +
    exp_session + I(exp_session^2) +
    (1 | participant_id),
  data = melatonin_glmm,
  family = Gamma(link = "log")
)

summary(m3_mel)
```

### Likelihood Ratio Tests: Session Effects
```{r}
#| label: lrt-session-effects-melatonin

# LRT 1: Does adding exp_session (linear) improve fit over Lag 0 only?
lrt_m1a_vs_m2 <- anova(m1a_mel, m2_mel)

# LRT 2: Does adding quadratic term improve fit over linear only?
lrt_m2_vs_m3 <- anova(m2_mel, m3_mel)

lrt_m1a_vs_m2
lrt_m2_vs_m3
```
Based on the LRT results, model fit was significantly improved by adding 
linear exp_session (p = 0.031), but not by adding the quadratic term 
(p = 0.182). Therefore, analysis will be proceeded with m2 (Lag 0 + linear session effect) 
to test for interaction effects in the next step.

### Model m4: Test Light × Session Interaction

This model tests whether the acute light suppression effect (Light_Lag0) changes across sessions (habituation or sensitization).
```{r}
#| label: model-m4-melatonin

m4_mel <- glmer(
  auc_melatonin_exp ~ 
    Light_Lag0_num +
    exp_session +
    Light_Lag0_num:exp_session +
    (1 | participant_id),
  data = melatonin_glmm,
  family = Gamma(link = "log")
)

summary(m4_mel)
```

### Likelihood Ratio Test: Light × Session Interaction
```{r}
#| label: lrt-interaction-melatonin

# LRT: Does adding the interaction improve fit over main effects only?
lrt_m2_vs_m4 <- anova(m2_mel, m4_mel)
lrt_m2_vs_m4
```
Based on the LRT results, model fit was not significantly improved by adding the interaction between Light_Lag0_num and exp_session (p = 0.707). Therefore, model m2 (main effects only) was retained for the full dataset analysis. However, comprehensive diagnostics revealed problems (see Section 8.2), leading to sensitivity analysis with influential observations excluded.

## Model SelectionSummary (full dataset) (model m2)

Based on the likelihood ratio tests and hierarchical model building strategy:

**Step 1: Testing Lag Effects**
- **m0 vs m1a**: Adding Lag 0 significantly improved fit over baseline (LRT: χ²(1) = 30.37, p < 0.001) 
- **m1a vs m1b**: Adding Lag 1 did not improve fit over Lag 0 only (LRT: χ²(1) = 0.06, p = 0.814) 
- **m1b vs m1c**: Adding Lag 2 did not improve fit over Lag 0 + Lag 1 (LRT: χ²(1) = 0.95, p = 0.329) 
- Note: Lag 2 was tested for completeness despite non-significant Lag 1
- **Decision**: Proceed with m1a (Light_Lag0_num only)

**Step 2: Testing Session Effects**
- **m1a vs m2**: Adding linear session effect significantly improved fit (LRT: χ²(1) = 4.66, p = 0.031) 
- **m2 vs m3**: Adding quadratic session term did not improve fit (LRT: χ²(1) = 1.78, p = 0.182) ✗
- **Decision**: Proceed with m2 (Light_Lag0_num + exp_session)

**Step 3: Testing Interaction**
- **m2 vs m4**: Adding Light × Session interaction did not improve fit (LRT: χ²(1) = 0.14, p = 0.707) 
- **Decision**: Retain m2 as final model

**Final Model (m2)**: 
- Predictors: Light_Lag0_num + exp_session
- Random effect: (1 | participant_id)
- AIC: 224.21
- Interpretation: Melatonin AUC is influenced by current evening light exposure and shows a linear habituation effect across sessions, with no evidence of light sensitization/habituation.

**Final Model:** Light_Lag0 + exp_session + (1 | participant_id)

### Final Basic Model Results with BH Correction (full dataset) (model m2)
```{r}
#| label: final-model-bh-melatonin

# Extract coefficients from m2
m2_coef <- summary(m2_mel)$coefficients
m2_fixed <- rownames(m2_coef)[rownames(m2_coef) != "(Intercept)"]

# Apply BH correction
m2_p_raw <- m2_coef[m2_fixed, "Pr(>|z|)"]
m2_p_bh <- p.adjust(m2_p_raw, method = "BH")

# Create results table
m2_results <- data.frame(
  Effect = m2_fixed,
  Estimate = round(m2_coef[m2_fixed, "Estimate"], 3),
  SE = round(m2_coef[m2_fixed, "Std. Error"], 3),
  t = round(m2_coef[m2_fixed, "t value"], 3),  # << GEÄNDERT: "t value" statt "z value"
  p_raw = format.pval(m2_p_raw, digits = 3, eps = 0.001),
  p_BH = format.pval(m2_p_bh, digits = 3, eps = 0.001)
)

# Add suppression percentage for Light_Lag0
m2_results <- m2_results %>%
  mutate(
    Suppression_pct = ifelse(
      grepl("Light_Lag0", Effect),
      paste0(round((1 - exp(Estimate)) * 100, 1), "%"),
      NA
    )
  )

kable(m2_results, 
      caption = "Final model (m2) results with Benjamini-Hochberg correction") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Model Diagnostics

### Check 1: Positive, Non-Zero Values

Gamma distribution requires strictly positive values.
```{r}
#| label: check-positive-values

min(melatonin_glmm$auc_melatonin_exp)
sum(melatonin_glmm$auc_melatonin_exp <= 0)
```

**Result:** All values are positive (min = `r round(min(melatonin_glmm$auc_melatonin_exp), 3)`), no zeros.

### Check 2: Right-Skewed Distribution
```{r}
#| label: check-skewness

skew_value <- skewness(melatonin_glmm$auc_melatonin_exp)
```

**Skewness:** `r round(skew_value, 2)`

**Interpretation:** Strongly right-skewed (skewness > 1), appropriate for Gamma distribution.

#### Histogram with Density Overlay
```{r}
#| label: fig-histogram-gamma
#| fig-width: 8
#| fig-height: 5

ggplot(melatonin_glmm, aes(x = auc_melatonin_exp)) +
  geom_histogram(aes(y = after_stat(density)), bins = 15, 
                 fill = "#4A90E2", alpha = 0.7, color = "black") +
  geom_density(color = "red", linewidth = 1) +
  labs(title = "Distribution of Melatonin AUC (n=67, P07 excluded)",
       x = "AUC (pg/mL × h)", 
       y = "Density") +
  theme_classic(base_size = 13)
```

#### Gamma Distribution Fit
```{r}
#| label: fig-gamma-fit
#| fig-width: 10
#| fig-height: 8

fit_gamma <- fitdist(melatonin_glmm$auc_melatonin_exp, "gamma")
plot(fit_gamma)
```

**Interpretation:** Good fit to Gamma distribution suggested.

### Check 3: DHARMa Residual Diagnostics
```{r}
#| label: dharma-diagnostics-m2
#| fig-width: 10
#| fig-height: 5

sim_m2 <- simulateResiduals(m2_mel, n = 1000)
plot(sim_m2)
```

**Interpretation:** Problematic quantile deviation detected. 

#### Individual DHARMa Tests
```{r}
#| label: dharma-tests-individual

testUniformity(sim_m2)
testDispersion(sim_m2)
testOutliers(sim_m2)
testQuantiles(sim_m2)
```

**Results:**

- **Uniformity test:** p = `r round(testUniformity(sim_m2, plot = FALSE)$p.value, 3)`
- **Dispersion test:** p = `r round(testDispersion(sim_m2, plot = FALSE)$p.value, 3)`
- **Outlier test:** p = `r round(testOutliers(sim_m2, plot = FALSE)$p.value, 3)`
- **Quantile test:** p = `r round(testQuantiles(sim_m2, plot = FALSE)$p.value, 3)`

**Conclusion:** Problematic deviations found in DHARMA quantile tests. Therefore, test model without influential observations.

### Check 4: Comprehensive Diagnostics
```{r}
#| label: check-model-comprehensive
#| fig-width: 12
#| fig-height: 10

check_model(m2_mel)
```

### Check 5: Multicollinearity (VIF)
```{r}
#| label: vif-check

vif(m2_mel)
```

**Interpretation:** All VIF < 2 → No multicollinearity issues.

### Check 6: Influential Observations
```{r}
#| label: influential-observations

infl_m2 <- influence(m2_mel)
cooks_d_m2 <- cooks.distance(infl_m2)
threshold_m2 <- 4/nrow(melatonin_glmm)

influential_idx <- which(cooks_d_m2 > threshold_m2)
```

**Number of influential observations:** `r length(influential_idx)` (`r round(100*length(influential_idx)/nrow(melatonin_glmm), 1)`% of data)
```{r}
#| label: influential-obs-table

melatonin_glmm[influential_idx, c("participant_id", "exp_session", "auc_melatonin_exp")]
```

#### Cook's Distance Plot
```{r}
#| label: fig-cooks-distance
#| fig-width: 10
#| fig-height: 5

plot(cooks_d_m2, type = "h", main = "Cook's Distance - Melatonin Model m2",
     ylab = "Cook's D", lwd = 2,
     col = ifelse(cooks_d_m2 > threshold_m2, "red", "black"))
abline(h = threshold_m2, lty = 2, col = "red")
```

**Interpretation:** `r length(influential_idx)` influential observations identified. All values are biologically plausible (no statistical outliers per DHARMa).

#### Sensitivity Analysis: Exclude Influential Observations
```{r}
#| label: robust-model

# Remove influential observations
melatonin_glmm_robust <- melatonin_glmm[-influential_idx, ]

# Refit model
Melatonin_GLMM_robust <- glmer(
  auc_melatonin_exp ~ 
    Light_Lag0_num + exp_session +
    (1 | participant_id),
  data = melatonin_glmm_robust,
  family = Gamma(link = "log")
)

summary(Melatonin_GLMM_robust)

# Check diagnostics
sim_robust <- simulateResiduals(Melatonin_GLMM_robust, n = 1000)
plot(sim_robust)
testQuantiles(sim_robust)
```
Robustness check: Model m2_mel_robust (n = 63, excluding 4 influential observations)
Results:
 - Light effect remains highly significant and stronger (β = -1.49, p < 0.001)
 - Session effect becomes non-significant after removing influential observations (β = -0.03, p = 0.315)
- Decision: Conduct comprehensive hierarchical model building on the 
  robust dataset (excluding 4 influential observations) due to:
  1. Massively improved AIC/BIC (m2: 224.2 → m2_robust: 186.6)
  2. Significant quantile deviations in full model (p = 0.023)
  3. Need to re-evaluate session effect significance
 
## Sensitivity Analysis: Hierarchical Model Building on Robust Dataset

### Rationale

Given the identification of 4 influential observations with Cook's D > 4/n (see Check 6 above), a sensitivity analysis by repeating the entire hierarchical model building strategy on the robust dataset was performed. 

1. Assess whether model selection decisions are stable after removing influential cases
2. Evaluate whether model fit improves
3. Determine the most appropriate final model for subsequent analyses

**Note:** The robust dataset `melatonin_glmm_robust` (n = `r nrow(melatonin_glmm_robust)`) was created in Check 6 by excluding the 4 influential observations identified via Cook's Distance analysis.

---

### Step 1: Testing Lag Effects (Robust Dataset)

#### Baseline Model (m0_robust)
```{r}
#| label: m0-robust

m0_robust <- glmer(
  auc_melatonin_exp ~ 1 + (1 | participant_id),
  data = melatonin_glmm_robust,
  family = Gamma(link = "log")
)

summary(m0_robust)
```

#### Model with Lag 0 only (m1a_robust)
```{r}
#| label: m1a-robust

m1a_robust <- glmer(
  auc_melatonin_exp ~ 
    Light_Lag0_num + 
    (1 | participant_id),
  data = melatonin_glmm_robust,
  family = Gamma(link = "log")
)

summary(m1a_robust)
```

**LRT: m0_robust vs m1a_robust**
```{r}
#| label: lrt-m0-m1a-robust

anova(m0_robust, m1a_robust)
```

**Interpretation:** 
- χ²(1) = `r round(anova(m0_robust, m1a_robust)$Chisq[2], 2)`
- p = `r format.pval(anova(m0_robust, m1a_robust)$"Pr(>Chisq)"[2], digits = 3, eps = 0.001)`
- Adding Lag 0 significantly improves fit over baseline model

---

#### Model with Lag 0 + Lag 1 (m1b_robust)
```{r}
#| label: m1b-robust

m1b_robust <- glmer(
  auc_melatonin_exp ~ 
    Light_Lag0_num + Light_Lag1_num +
    (1 | participant_id),
  data = melatonin_glmm_robust,
  family = Gamma(link = "log")
)

summary(m1b_robust)
```

**LRT: m1a_robust vs m1b_robust**
```{r}
#| label: lrt-m1a-m1b-robust

anova(m1a_robust, m1b_robust)
```

**Interpretation:** 
- χ²(1) = `r round(anova(m1a_robust, m1b_robust)$Chisq[2], 2)`
- p = `r format.pval(anova(m1a_robust, m1b_robust)$"Pr(>Chisq)"[2], digits = 3, eps = 0.001)`
- Adding Lag 1 does not improve fit over Lag 0 only model 

---

#### Model with Lag 0 + Lag 1 + Lag 2 (m1c_robust)
```{r}
#| label: m1c-robust

m1c_robust <- glmer(
  auc_melatonin_exp ~ 
    Light_Lag0_num + Light_Lag1_num + Light_Lag2_num +
    (1 | participant_id),
  data = melatonin_glmm_robust,
  family = Gamma(link = "log")
)

summary(m1c_robust)
```

**LRT: m1b_robust vs m1c_robust**
```{r}
#| label: lrt-m1b-m1c-robust

anova(m1b_robust, m1c_robust)
```

**Interpretation:** 
- χ²(1) = `r round(anova(m1b_robust, m1c_robust)$Chisq[2], 2)`
- p = `r format.pval(anova(m1b_robust, m1c_robust)$"Pr(>Chisq)"[2], digits = 3, eps = 0.001)`
- Adding Lag 2 does not improve fit

**Decision Step 1 (Robust Dataset):** Proceed with m1a_robust (Light_Lag0_num only)

---

### Step 2: Testing Session Effects (Robust Dataset)

#### Model with Linear Session Effect (m2_robust)
```{r}
#| label: m2-robust

m2_robust <- glmer(
  auc_melatonin_exp ~ 
    Light_Lag0_num + exp_session +
    (1 | participant_id),
  data = melatonin_glmm_robust,
  family = Gamma(link = "log")
)

summary(m2_robust)
```

**LRT: m1a_robust vs m2_robust**
```{r}
#| label: lrt-m1a-m2-robust

anova(m1a_robust, m2_robust)
```

**Interpretation:** 
- χ²(1) = `r round(anova(m1a_robust, m2_robust)$Chisq[2], 2)`
- p = `r format.pval(anova(m1a_robust, m2_robust)$"Pr(>Chisq)"[2], digits = 3, eps = 0.001)`
- Adding linear session effect does not improve fit over Lag 0 model only

---

#### Model with Quadratic Session Effect (m3_robust)
```{r}
#| label: m3-robust

m3_robust <- glmer(
  auc_melatonin_exp ~ 
    Light_Lag0_num + exp_session + I(exp_session^2) +
    (1 | participant_id),
  data = melatonin_glmm_robust,
  family = Gamma(link = "log")
)

summary(m3_robust)
```

**LRT: m2_robust vs m3_robust**
```{r}
#| label: lrt-m2-m3-robust

anova(m2_robust, m3_robust)
```

**Interpretation:** 
- χ²(1) = `r round(anova(m2_robust, m3_robust)$Chisq[2], 2)`
- p = `r format.pval(anova(m2_robust, m3_robust)$"Pr(>Chisq)"[2], digits = 3, eps = 0.001)`
- Adding quadratic term does not improve fit

**Decision Step 2 (Robust Dataset):** Proceed with m1a_robust (Light_Lag0_num)

**Decision:** Retain m1a_robust as final model

**Summary Robust Model m1a**
```{r}
#| label: m1a_robust

summary(m1a_robust)
```

### Check 1 Robust Model: Positive, Non-Zero Values

Gamma distribution requires strictly positive values.
```{r}
#| label: check-positive-values-robust

min(melatonin_glmm_robust$auc_melatonin_exp)
sum(melatonin_glmm_robust$auc_melatonin_exp <= 0)
```

**Result:** All values are positive (min = `r round(min(melatonin_glmm_robust$auc_melatonin_exp), 3)`), no zeros.

---

### Check 2 Robust Model: Right-Skewed Distribution
```{r}
#| label: check-skewness-robust

skew_value_robust <- skewness(melatonin_glmm_robust$auc_melatonin_exp)
```

**Skewness:** `r round(skew_value_robust, 2)`

**Interpretation:** `r if(skew_value_robust > 1) {"Strongly right-skewed (skewness > 1), appropriate for Gamma distribution."} else {"Right-skewed, appropriate for Gamma distribution."}`

#### Histogram with Density Overlay
```{r}
#| label: fig-histogram-gamma-robust
#| fig-cap: "Distribution of Melatonin AUC in robust dataset (n=63)"
#| fig-width: 8
#| fig-height: 5

ggplot(melatonin_glmm_robust, aes(x = auc_melatonin_exp)) +
  geom_histogram(aes(y = after_stat(density)), bins = 15, 
                 fill = "#50C878", alpha = 0.7, color = "black") +
  geom_density(color = "red", linewidth = 1) +
  labs(title = "Distribution of Melatonin AUC (n=63, robust dataset)",
       x = "AUC (pg/mL × h)", 
       y = "Density") +
  theme_classic(base_size = 13)
```

#### Gamma Distribution Fit
```{r}
#| label: fig-gamma-fit-robust
#| fig-cap: "Gamma distribution fit for robust dataset"
#| fig-width: 10
#| fig-height: 8

fit_gamma_robust <- fitdist(melatonin_glmm_robust$auc_melatonin_exp, "gamma")
plot(fit_gamma_robust)
```

**Interpretation:** Good fit to Gamma distribution suggested.

---

### Check 3 Robust Model: DHARMa Residual Diagnostics
```{r}
#| label: dharma-diagnostics-m1a-robust
#| fig-cap: "DHARMa residual diagnostics for m1a_robust"
#| fig-width: 10
#| fig-height: 5

sim_m1a_robust <- simulateResiduals(m1a_robust, n = 1000)
plot(sim_m1a_robust)
```

#### Individual DHARMa Tests
```{r}
#| label: dharma-tests-individual-robust

testUniformity(sim_m1a_robust)
testDispersion(sim_m1a_robust)
testOutliers(sim_m1a_robust)
testQuantiles(sim_m1a_robust)
```

**Results:**

- **Uniformity test:** p = `r round(testUniformity(sim_m1a_robust, plot = FALSE)$p.value, 3)`
- **Dispersion test:** p = `r round(testDispersion(sim_m1a_robust, plot = FALSE)$p.value, 3)`
- **Outlier test:** p = `r round(testOutliers(sim_m1a_robust, plot = FALSE)$p.value, 3)`
- **Quantile test:** p = `r if(is.na(testQuantiles(sim_m1a_robust, plot = FALSE)$p.value)) {"NA (test failed)"} else {round(testQuantiles(sim_m1a_robust, plot = FALSE)$p.value, 3)}`

---

### Check 4 Robust Model: Comprehensive Diagnostics
```{r}
#| label: check-model-comprehensive-robust
#| fig-cap: "Comprehensive model diagnostics for m1a_robust"
#| fig-width: 12
#| fig-height: 10

check_model(m1a_robust)
```

**Interpretation:** Visual inspection of residual patterns, linearity, 
homogeneity of variance, and normality of random effects.

---

### Check 5 Robust Model: Multicollinearity (VIF)
```{r}
#| label: vif-check-robust

# VIF only applicable for models with multiple fixed effects
# m1a_robust has only one predictor (Light_Lag0_num), so VIF not needed
cat("Model m1a_robust contains only one predictor (Light_Lag0_num).\n")
cat("Multicollinearity check not applicable.\n")
```

**Note:** VIF is not applicable as m1a_robust contains only a single 
fixed effect predictor (Light_Lag0_num). Multicollinearity can only 
occur with multiple predictors.

### Final Model Results with BH Correction (m1a_robust)
```{r}
#| label: final-model-bh-m1a-robust
#| tbl-cap: "Fixed effects for final robust model with Benjamini-Hochberg correction"

# Extract coefficients
m1a_coef <- summary(m1a_robust)$coefficients
m1a_fixed <- rownames(m1a_coef)[rownames(m1a_coef) != "(Intercept)"]

# BH correction
m1a_p_raw <- m1a_coef[m1a_fixed, "Pr(>|z|)"]
m1a_p_bh <- p.adjust(m1a_p_raw, method = "BH")

# Create results table
m1a_results <- data.frame(
  Effect = m1a_fixed,
  Estimate = round(m1a_coef[m1a_fixed, "Estimate"], 3),
  SE = round(m1a_coef[m1a_fixed, "Std. Error"], 3),
  t = round(m1a_coef[m1a_fixed, "t value"], 3),  # Wie bei m2
  p_raw = format.pval(m1a_p_raw, digits = 3, eps = 0.001),
  p_BH = format.pval(m1a_p_bh, digits = 3, eps = 0.001),
  Suppression = paste0(round((1 - exp(m1a_coef[m1a_fixed, "Estimate"])) * 100, 1), "%")
)

kable(m1a_results, 
      col.names = c("Effect", "β (log scale)", "SE", "t", "p", "p (BH-adj)", "Suppression"),
      align = c("l", "r", "r", "r", "r", "r", "c")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

**Note:** Only one predictor, therefore BH correction does not change p-value.
## Model Diagnostics for Final Robust Model (m1a_robust)

All diagnostics are performed on the robust dataset (n = 63) after 
exclusion of 4 influential observations identified in the full dataset 
analysis.

## Comparison: Full Dataset Model (m2) vs. Robust Dataset Model (m1a_robust)

To evaluate the impact of excluding influential observations, 
the final models from both datasets will be compared.

**Methodological note:** Models are fitted to datasets of different sizes 
(m2: n=67, m1a_robust: n=63). While direct AIC/BIC comparison across datasets 
has limitations, the observed improvement magnitude (ΔAIC ≈ -39) substantially 
exceeds sample size effects. Within-dataset comparison (m1a_robust vs. m2_robust) 
confirms model superiority.

---

### Model Specifications
```{r}
#| label: tbl-model-specs-comparison
#| tbl-cap: "Model specifications comparison"

data.frame(
  Aspect = c("Formula", "N observations", "N parameters", "Dataset"),
  m2_Full = c(
    "Light_Lag0 + exp_session",
    "67",
    "5",
    "Full"
  ),
  m1a_Robust = c(
    "Light_Lag0 only",
    "63",
    "4",
    "4 influential obs excluded"
  )
) %>%
  kable(col.names = c("", "m2 (Full)", "m1a_robust")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

---

### Model Fit Statistics
```{r}
#| label: tbl-fit-comparison
#| tbl-cap: "Model fit comparison"

data.frame(
  Metric = c("AIC", "BIC"),
  m2 = c(AIC(m2_mel), BIC(m2_mel)),
  m1a = c(AIC(m1a_robust), BIC(m1a_robust)),
  Delta = c(
    AIC(m1a_robust) - AIC(m2_mel),
    BIC(m1a_robust) - BIC(m2_mel)
  )
) %>%
  mutate(across(where(is.numeric), ~round(.x, 1))) %>%
  kable(col.names = c("Metric", "m2", "m1a_robust", "Δ")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

**Result:** Robust model shows substantially better fit (ΔAIC = `r round(AIC(m1a_robust) - AIC(m2_mel), 1)`, ΔBIC = `r round(BIC(m1a_robust) - BIC(m2_mel), 1)`).

---

### Fixed Effects Comparison
```{r}
#| label: tbl-effects-comparison
#| tbl-cap: "Fixed effects estimates comparison"

coef_m2 <- summary(m2_mel)$coefficients
coef_m1a <- summary(m1a_robust)$coefficients

data.frame(
  Effect = c("Light_Lag0_num", "exp_session"),
  Beta_m2 = c(
    coef_m2["Light_Lag0_num", "Estimate"],
    coef_m2["exp_session", "Estimate"]
  ),
  p_m2 = c(
    coef_m2["Light_Lag0_num", "Pr(>|z|)"],
    coef_m2["exp_session", "Pr(>|z|)"]
  ),
  Beta_m1a = c(
    coef_m1a["Light_Lag0_num", "Estimate"],
    NA
  ),
  p_m1a = c(
    coef_m1a["Light_Lag0_num", "Pr(>|z|)"],
    NA
  ),
  Suppression = c(
    paste0(round((1 - exp(coef_m2["Light_Lag0_num", "Estimate"])) * 100, 1), "% / ",
           round((1 - exp(coef_m1a["Light_Lag0_num", "Estimate"])) * 100, 1), "%"),
    "—"
  )
) %>%
  mutate(
    across(starts_with("Beta"), ~round(.x, 3)),
    across(starts_with("p"), ~format.pval(.x, digits = 3, eps = 0.001))
  ) %>%
  kable(col.names = c("Effect", "β (m2)", "p", "β (m1a)", "p", "Suppression")) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  row_spec(1, bold = TRUE, background = "#D3F4FF")
```

**Key findings:**
- **Light effect:** Highly consistent across models (both p < 0.001), stronger in robust model
- **Session effect:** Significant in full model (p = 0.031), not included in robust model (LRT p = 0.316)

---

### Diagnostic Performance
```{r}
#| label: tbl-diagnostics-comparison
#| tbl-cap: "DHARMa diagnostic tests comparison"

data.frame(
  Test = c("Uniformity (KS)", "Dispersion", "Outliers", "Quantiles"),
  m2_Full = c(
    format.pval(testUniformity(sim_m2, plot = FALSE)$p.value, digits = 3),
    format.pval(testDispersion(sim_m2, plot = FALSE)$p.value, digits = 3),
    format.pval(testOutliers(sim_m2, plot = FALSE)$p.value, digits = 3),
    "0.023*"
  ),
  m1a_Robust = c(
    format.pval(testUniformity(sim_m1a_robust, plot = FALSE)$p.value, digits = 3),
    format.pval(testDispersion(sim_m1a_robust, plot = FALSE)$p.value, digits = 3),
    format.pval(testOutliers(sim_m1a_robust, plot = FALSE)$p.value, digits = 3),
    "NULL"
  )
) %>%
  kable(col.names = c("Test", "m2 (p-value)", "m1a_robust")) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  add_footnote("* Significant quantile deviations in full model", notation = "none")
```

**Result:** Full model showed significant quantile deviations; robust model passed all applicable tests.

---

### Effect Stability
```{r}
#| label: fig-effect-stability
#| fig-cap: "Light effect estimates with 95% confidence intervals"
#| fig-width: 6
#| fig-height: 4

# Extract estimates and CIs
light_m2 <- fixef(m2_mel)["Light_Lag0_num"]
se_m2 <- sqrt(vcov(m2_mel)["Light_Lag0_num", "Light_Lag0_num"])

light_m1a <- fixef(m1a_robust)["Light_Lag0_num"]
se_m1a <- sqrt(vcov(m1a_robust)["Light_Lag0_num", "Light_Lag0_num"])

data.frame(
  Model = factor(c("m2 (Full)", "m1a_robust"), 
                 levels = c("m2 (Full)", "m1a_robust")),
  Estimate = c(light_m2, light_m1a),
  SE = c(se_m2, se_m1a)
) %>%
  ggplot(aes(x = Model, y = Estimate)) +
  geom_point(size = 3, color = "#4A90E2") +
  geom_errorbar(aes(ymin = Estimate - 1.96*SE, ymax = Estimate + 1.96*SE), 
                width = 0.15, linewidth = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red", alpha = 0.5) +
  labs(x = NULL, y = "β (log scale)") +
  theme_classic(base_size = 12)
```

---

### Conclusion

**Final model selected: m1a_robust**

**Justification:**
1. Superior model fit (ΔAIC = -38.6, ΔBIC = -40.7)
2. Good diagnostic performance (all tests passed)
3. Stable primary effect (Light: β = -1.48, p < 0.001, 77.2% suppression)
4. Session effect not justified in robust dataset (LRT p = 0.316)
5. Adheres to parsimony (4 vs. 5 parameters)

**Implication:** All subsequent exploratory analyses based on m1a_robust.

### Final Model Selection

Following diagnostic-driven model selection principles, **m1a_robust** 
is selected as the final model.

---

## Exploratory Analyses

### Exploratory Analysis 1: Random Slopes for Light_Lag0

This model tests whether participants show **individual differences** in their light suppression response by allowing the Light_Lag0 effect to vary across participants.

**Rationale:** The final model assumes that all participants have the same light suppression effect (fixed effect only). This analysis tests whether allowing individual variation in light sensitivity improves model fit and whether the average effect remains significant.
```{r}
#| label: sensitivity-random-slopes

# Fit model with random slope for Light_Lag0
m_random_slopes <- glmer(
  auc_melatonin_exp ~ 
    Light_Lag0_num + exp_session +
    (1 + Light_Lag0_num | participant_id),
  data = melatonin_glmm_robust,
  family = Gamma(link = "log"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

summary(m_random_slopes)
```

**Model equation:** log(μᵢⱼ) = (β₀ + u₀ᵢ) + (β₁ + u₁ᵢ)×Light_Lag0_num + β₂×exp_session

where:
- u₀ᵢ ~ N(0, τ₀²) = random intercept for participant i
- u₁ᵢ ~ N(0, τ₁²) = random slope for Light_Lag0_num for participant i
- Correlation between u₀ᵢ and u₁ᵢ was estimated

**Interpretation:**

Model fit was substantially improved by adding random slopes (AIC = 176 vs. 189.8, **ΔAIC = -13.8**), indicating that **meaningful individual differences** exist in light suppression response. Key findings were as follows:

1. **Random slope variance (τ₁² = 0.37, SD = 0.61):** Substantial variation was observed across participants in their melatonin suppression response to bright light, with some individuals being more light-sensitive than others.

2. **Strong negative correlation (r = -0.78):** Participants with **higher baseline melatonin levels** (higher intercept) exhibited **stronger suppression** (more negative slope). This suggests that individuals with robust melatonin production are also more responsive to light-induced suppression.

3. **Average effect remained robust:** Despite individual variation, the **average light suppression effect remained highly significant** (β = -1.48, p < 0.001, 77.3% suppression) and was nearly identical to the fixed slope model (77.5% suppression). This confirms that the primary finding is not dependent on the assumption of uniform light sensitivity across participants.

4. **Residual variance was reduced:** Additional variance was explained by the random slope (residual variance: 0.419 → 0.339, -19% reduction), indicating that individual differences in light sensitivity account for meaningful within-person variability.

5. **Session effect remained non-significant:** The exp_session effect was not significant in the random slopes model (β = -0.03, p = 0.250), consistent with findings from the fixed slopes model after influential observations were removed.

**Conclusion:** While substantial melatonin suppression in response to bright light was observed across all participants, the magnitude of suppression varied between individuals. Participants with higher baseline melatonin production showed stronger light-induced suppression, consistent with a "ceiling effect" where greater secretion allows for greater suppression. The primary finding of robust light suppression (~77%) remains valid as an average population effect.

**Note:** With n=8 participants, the random slope variance estimate should be interpreted cautiously. However, the substantial improvement in model fit (ΔAIC = -13.8) provides strong evidence that non-trivial individual differences exist.

### Exploratory Analysis 2: Age and Sex Moderation
```{r}
#| label: sensitivity-demographics-moderation

# Center age
melatonin_glmm_robust <- melatonin_glmm_robust %>%
  mutate(age_c = age - mean(age))

# Moderation model 
m_demographics_moderation <- glmer(
  auc_melatonin_exp ~ 
    Light_Lag0_num + exp_session +
    sex + age_c + 
    Light_Lag0_num:sex + Light_Lag0_num:age_c +
    (1 | participant_id),
  data = melatonin_glmm_robust,
  family = Gamma(link = "log"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

summary(m_demographics_moderation)
```

**Interpretation:**

Model fit was not improved by adding demographic moderators (AIC = 189.7 vs. 189.8, **ΔAIC = -0.1**), indicating that age and sex do not substantially explain melatonin variability beyond what is captured by random intercepts. Key findings were as follows:

1. **Main effects were non-significant:** Neither sex (β = -0.011, p = 0.987) nor age (β = 0.062, p = 0.474) showed significant effects on baseline melatonin levels. This suggests that similar average melatonin production exists across demographic groups in this sample.

2. **Exploratory sex interaction (p = 0.037):** Males showed potentially **stronger light suppression** (86.5%) compared to females (68.6%), with a significant interaction term (β = -0.85, p = 0.037). However, this finding requires **critical interpretation**:
   - **Sample composition:** Only **3 males vs. 5 females** (n=8 total)
   - **Statistical power:** The analysis was underpowered for between-subject effects
   - **Model parsimony:** Nearly identical AIC (ΔAIC = -0.1) suggests no meaningful improvement despite added complexity

3. **No age moderation was observed:** The Age × Light interaction was non-significant (β = 0.041, p = 0.442), indicating that light sensitivity does not systematically vary with age in this young-adult sample (ages 21-32).

4. **Average effect remained robust:** The light suppression effect remained highly significant (β = -1.16, p < 0.001) after controlling for demographics, confirming that the primary finding is not confounded by age or sex.

**Conclusion:** With only n=8 participants, demographic effects cannot be reliably estimated. The exploratory sex interaction is **hypothesis-generating only** and requires replication in larger samples before substantive conclusions can be drawn. The final model without demographics was retained based on superior parsimony (nearly identical AIC despite added complexity), avoidance of overfitting, and the study's primary focus on within-subject light effects rather than between-subject demographic differences.

**Recommendation:** Future studies with larger samples (n ≥ 30 per sex) should investigate potential sex differences in circadian light sensitivity, which have been reported in some (but not all) prior literature.

### Exploratory Analysis 3: Light Lag Interactions

This model tests whether light exposure on different evenings interact with each other (e.g., does the effect of current light depend on previous light exposure?).

**Rationale:** The final model assumes independent, additive effects of Light_Lag0, Light_Lag1, and Light_Lag2. This analysis tests whether these lags show interactive (synergistic or antagonistic) effects.
```{r}
#| label: sensitivity-light-interactions

# Fit model with all pairwise light lag interactions
m_light_interactions <- glmer(
  auc_melatonin_exp ~ 
    Light_Lag0_num * Light_Lag1_num + 
    Light_Lag0_num * Light_Lag2_num +
    Light_Lag1_num * Light_Lag2_num +
    exp_session +
    (1 | participant_id),
  data = melatonin_glmm_robust,
  family = Gamma(link = "log"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

summary(m_light_interactions)
```

**Model equation:** 
log(μᵢⱼ) = β₀ + β₁×Lag0 + β₂×Lag1 + β₃×Lag2 + β₄×Session + β₅×(Lag0×Lag1) + β₆×(Lag0×Lag2) + β₇×(Lag1×Lag2) + uᵢ

**Interpretation of interactions:**
- **β₅ (Lag0 × Lag1):** Does yesterday's light modify today's light effect?
- **β₆ (Lag0 × Lag2):** Does 2-days-ago light modify today's light effect?
- **β₇ (Lag1 × Lag2):** Do yesterday's and 2-days-ago light interact?

**Interpretation:**

The light lag interaction model did **not** improve fit (AIC = 193.3 vs. 189.8, **ΔAIC = +3.5**).

1. **All interactions non-significant:**
   - **Lag0 × Lag1:** β = 0.31, p = 0.521 (yesterday's light does not modify today's response)
   - **Lag0 × Lag2:** β = 0.46, p = 0.203 (2-days-ago light does not modify today's response)  
   - **Lag1 × Lag2:** β = -0.02, p = 0.960 (no interaction between prior exposures)

2. **No evidence for habituation or sensitization:** The absence of significant Lag0 × Lag1/Lag2 interactions suggests that **prior bright light exposure does not alter the magnitude of response to current light in this population**. This suggests:

3. **Independent daily effects confirmed:** Each evening's light exposure acts **independently** on the melatonin system, with current exposure (Lag0) being the sole significant predictor (β = -1.84, p < 0.001, 84.1% suppression when prior exposures were dim). Prior exposures (Lag1, Lag2) showed no main effects and did not moderate the acute response.

**Conclusion:** The final model with additive (non-interactive) light lag effects is preferred based on parsimony (superior AIC), statistical evidence (no significant interactions), and biological plausibility (each evening's light acts independently). This finding does not support the hypothesis that there are carryover effects of light induced melatonin suppression from previous days on the current day. 

### Exploratory Analysis 4: Persistence Model 

#### Rationale for Persistence Analysis

The primary GLMM analysis demonstrated that evening light exposure (Lag 0) significantly suppresses melatonin, while prior exposures (Lag 1, Lag 2) showed no carryover effects. However, this lag-based approach tests whether *individual prior nights* contribute additively to the current night's melatonin levels.

The **persistence analysis** addresses the question: Does the *number of consecutive days* of the same light condition (run length) modulate melatonin responses? This tests whether:

1. **Dim nights:** Multiple consecutive dark evenings produce cumulative recovery of melatonin (i.e., does melatonin continue to increase with each additional dark night?)

2. **Bright nights:** Multiple consecutive bright evenings produce cumulative suppression (i.e., does melatonin continue to decrease with each additional bright night?)

**Key distinction from lag analysis:**

- Lag analysis: Tests whether Day[t-1], Day[t-2], Day[t-3] each add independent effects
- Persistence analysis: Tests whether being on the 1st vs. 2nd vs. 3rd+ consecutive day of the *same condition* matters

This exploratory analysis uses separate models for dim and bright conditions, testing both linear (numeric run length) and non-linear (categorical run length) accumulation patterns.

#### Dataset Preparation 
```{r}
#| label: create-persistence-dataset

# Define participant list and light patterns
participant_list <- c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10")  # P07 excluded

# Create persistence dataset starting from session 1
melatonin_persistence <- map_dfr(participant_list, ~{
  df <- auc_experimental_samples %>% 
    filter(participant_id == .x)
  
  light_df <- tibble(
    exp_session = seq_along(light_patterns[[.x]]),
    Light_Lag0 = light_patterns[[.x]]
  )
  
  df <- df %>% 
    left_join(light_df, by = "exp_session") %>%
    rename(auc_melatonin_exp = auc_melatonin_exp)
  
  return(df)
})

# Add demographics
participant_demographics <- tibble(
  participant_id = c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10"),
  sex = c("female", "male", "female", "female", "male", "female", "female", "male"),
  age = c(24, 21, 24, 23, 31, 23, 28, 32)
)

melatonin_persistence <- melatonin_persistence %>%
  left_join(participant_demographics, by = "participant_id")

# Calculate run-length variables (numeric)
melatonin_persistence <- melatonin_persistence %>%
  filter(!is.na(Light_Lag0)) %>%  # Remove rows without light condition
  filter(!(participant_id == "P10" & exp_session == 5)) %>%  # Exclude P10 session 5
  arrange(participant_id, exp_session) %>%
  group_by(participant_id) %>%
  mutate(
    # Track when condition changes
    condition_change = Light_Lag0 != lag(Light_Lag0, default = "CHANGE"),
    
    # Calculate run ID (new sequence after each change)
    run_id = cumsum(condition_change),
    
    # Calculate run length within each run
    run_length = ave(run_id, run_id, FUN = seq_along),
    
    # Separate run lengths for Dim and Bright
    run_length_dim = ifelse(Light_Lag0 == "D", run_length, NA),
    run_length_bright = ifelse(Light_Lag0 == "B", run_length, NA),
    
    # Also create numeric Light variable for consistency
    Light_Lag0_num = ifelse(Light_Lag0 == "D", 0, 1)
  ) %>%
  ungroup() %>%
  select(participant_id, exp_session, auc_melatonin_exp, Light_Lag0, Light_Lag0_num,
         run_length, run_length_dim, run_length_bright, sex, age)

# Create categorical run-length variables
melatonin_persistence <- melatonin_persistence %>%
  mutate(
    run_category_dim = case_when(
      run_length_dim == 1 ~ "First",
      run_length_dim == 2 ~ "Second",
      run_length_dim >= 3 ~ "Third",
      TRUE ~ NA_character_
    ),
    run_category_bright = case_when(
      run_length_bright == 1 ~ "First",
      run_length_bright == 2 ~ "Second",
      run_length_bright >= 3 ~ "Third",
      TRUE ~ NA_character_
    )
  ) %>%
  mutate(
    run_category_dim = factor(run_category_dim, levels = c("First", "Second", "Third")),
    run_category_bright = factor(run_category_bright, levels = c("First", "Second", "Third"))
  )
```

#### Persistence Dataset
```{r}
#| label: display-persistence-dataset

DT::datatable(
  melatonin_persistence,
  options = list(
    pageLength = 20,
    scrollX = TRUE,
    order = list(list(0, 'asc'), list(1, 'asc'))
  ),
  caption = "Persistence dataset with run-length encoding (n=85, P07 excluded, all sessions)",
  rownames = FALSE
) %>%
  DT::formatRound(columns = 'auc_melatonin_exp', digits = 2) %>%
  DT::formatStyle(
    'Light_Lag0',
    backgroundColor = DT::styleEqual(c('D', 'B'), c('#e3f2fd', '#fff3e0'))
  )
```

#### Descriptive Visualizations
```{r}
#| label: fig-persistence-boxplots-all
#| fig-width: 12
#| fig-height: 12

# Define color scheme
color_dim <- "#4A90E2"    # Blue for Dim
color_bright <- "#F5A623" # Orange for Bright

# Prepare data WITH Session 1
plot_data_dim_s1 <- melatonin_persistence %>%
  filter(!is.na(run_length_dim)) %>%
  mutate(run_length_dim = factor(run_length_dim))

plot_data_bright_s1 <- melatonin_persistence %>%
  filter(!is.na(run_length_bright)) %>%
  mutate(run_length_bright = factor(run_length_bright))

# Prepare data WITHOUT Session 1
plot_data_dim_no_s1 <- melatonin_persistence %>%
  filter(!is.na(run_length_dim), exp_session > 1) %>%
  mutate(run_length_dim = factor(run_length_dim))

plot_data_bright_no_s1 <- melatonin_persistence %>%
  filter(!is.na(run_length_bright), exp_session > 1) %>%
  mutate(run_length_bright = factor(run_length_bright))

# Create all 4 plots
p_dim_s1 <- ggplot(plot_data_dim_s1, aes(x = run_length_dim, y = auc_melatonin_exp)) +
  geom_boxplot(fill = color_dim, outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16) +
  labs(title = "Dim Nights (with Session 1)", subtitle = paste0("n = ", nrow(plot_data_dim_s1)),
       x = "Consecutive Nights", y = "Melatonin AUC (pg/mL × h)") +
  theme_classic(base_size = 14) +
  theme(aspect.ratio = 1, plot.title = element_text(size = 16, hjust = 0),
        plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
        axis.text = element_text(color = "black", size = 13),
        axis.title = element_text(size = 14),
        panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
        legend.position = "none")

p_bright_s1 <- ggplot(plot_data_bright_s1, aes(x = run_length_bright, y = auc_melatonin_exp)) +
  geom_boxplot(fill = color_bright, outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16) +
  labs(title = "Bright Nights (with Session 1)", subtitle = paste0("n = ", nrow(plot_data_bright_s1)),
       x = "Consecutive Nights", y = "Melatonin AUC (pg/mL × h)") +
  theme_classic(base_size = 14) +
  theme(aspect.ratio = 1, plot.title = element_text(size = 16, hjust = 0),
        plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
        axis.text = element_text(color = "black", size = 13),
        axis.title = element_text(size = 14),
        panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
        legend.position = "none")

p_dim_no_s1 <- ggplot(plot_data_dim_no_s1, aes(x = run_length_dim, y = auc_melatonin_exp)) +
  geom_boxplot(fill = color_dim, outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16) +
  labs(title = "Dim Nights (without Session 1)", subtitle = paste0("n = ", nrow(plot_data_dim_no_s1)),
       x = "Consecutive Nights", y = "Melatonin AUC (pg/mL × h)") +
  theme_classic(base_size = 14) +
  theme(aspect.ratio = 1, plot.title = element_text(size = 16, hjust = 0),
        plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
        axis.text = element_text(color = "black", size = 13),
        axis.title = element_text(size = 14),
        panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
        legend.position = "none")

p_bright_no_s1 <- ggplot(plot_data_bright_no_s1, aes(x = run_length_bright, y = auc_melatonin_exp)) +
  geom_boxplot(fill = color_bright, outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16) +
  labs(title = "Bright Nights (without Session 1)", subtitle = paste0("n = ", nrow(plot_data_bright_no_s1)),
       x = "Consecutive Nights", y = "Melatonin AUC (pg/mL × h)") +
  theme_classic(base_size = 14) +
  theme(aspect.ratio = 1, plot.title = element_text(size = 16, hjust = 0),
        plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
        axis.text = element_text(color = "black", size = 13),
        axis.title = element_text(size = 14),
        panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
        legend.position = "none")

# Combine all plots in 2x2 grid
(p_dim_s1 + p_bright_s1) / (p_dim_no_s1 + p_bright_no_s1)
```

#### Model 1: Linear Persistence Models, Dim Evenings 
```{r}
#| label: model-persistence-dim-linear

# WITH Session 1
melatonin_persistence_dim <- melatonin_persistence %>%
  filter(!is.na(run_length_dim))

model_persistence_dim_numeric <- glmer(
  auc_melatonin_exp ~ run_length_dim + (1 | participant_id),
  data = melatonin_persistence_dim,
  family = Gamma(link = "log")
)

summary(model_persistence_dim_numeric)

# WITHOUT Session 1
melatonin_persistence_dim_no_s1 <- melatonin_persistence %>%
  filter(!is.na(run_length_dim), exp_session > 1)

model_persistence_dim_no_s1 <- glmer(
  auc_melatonin_exp ~ run_length_dim + (1 | participant_id),
  data = melatonin_persistence_dim_no_s1,
  family = Gamma(link = "log")
)

summary(model_persistence_dim_no_s1)
```

#### Model 1: Linear Persistence Models, Bright Evenings
```{r}
#| label: model-persistence-bright-linear

# WITH Session 1
melatonin_persistence_bright <- melatonin_persistence %>%
  filter(!is.na(run_length_bright))

model_persistence_bright_numeric <- glmer(
  auc_melatonin_exp ~ run_length_bright + (1 | participant_id),
  data = melatonin_persistence_bright,
  family = Gamma(link = "log")
)

summary(model_persistence_bright_numeric)

# WITHOUT Session 1
melatonin_persistence_bright_no_s1 <- melatonin_persistence %>%
  filter(!is.na(run_length_bright), exp_session > 1)

model_persistence_bright_no_s1 <- glmer(
  auc_melatonin_exp ~ run_length_bright + (1 | participant_id),
  data = melatonin_persistence_bright_no_s1,
  family = Gamma(link = "log")
)

summary(model_persistence_bright_no_s1)
```

#### Model 2: Categorical Persistence Models, Dim Evenings 
```{r}
#| label: model-persistence-dim-categorical

# WITH Session 1
melatonin_persistence_dim <- melatonin_persistence %>%
  filter(!is.na(run_category_dim))

model_persistence_dim_categorical <- glmer(
  auc_melatonin_exp ~ run_category_dim + (1 | participant_id),
  data = melatonin_persistence_dim,
  family = Gamma(link = "log")
)

summary(model_persistence_dim_categorical)

lrt_dim_s1 <- anova(model_persistence_dim_numeric, model_persistence_dim_categorical)
print(lrt_dim_s1)

# WITHOUT Session 1
melatonin_persistence_dim_no_s1 <- melatonin_persistence %>%
  filter(!is.na(run_category_dim), exp_session > 1)

model_persistence_dim_categorical_no_s1 <- glmer(
  auc_melatonin_exp ~ run_category_dim + (1 | participant_id),
  data = melatonin_persistence_dim_no_s1,
  family = Gamma(link = "log")
)

summary(model_persistence_dim_categorical_no_s1)

lrt_dim_no_s1 <- anova(model_persistence_dim_no_s1, model_persistence_dim_categorical_no_s1)
print(lrt_dim_no_s1)
```

#### Model 2: Categorical Persistence Models, Bright Evenings 
```{r}
#| label: model-persistence-bright-categorical

# WITH Session 1
melatonin_persistence_bright <- melatonin_persistence %>%
  filter(!is.na(run_category_bright))

model_persistence_bright_categorical <- glmer(
  auc_melatonin_exp ~ run_category_bright + (1 | participant_id),
  data = melatonin_persistence_bright,
  family = Gamma(link = "log")
)

summary(model_persistence_bright_categorical)

lrt_bright_s1 <- anova(model_persistence_bright_numeric, model_persistence_bright_categorical)
print(lrt_bright_s1)

# WITHOUT Session 1
melatonin_persistence_bright_no_s1 <- melatonin_persistence %>%
  filter(!is.na(run_category_bright), exp_session > 1)

model_persistence_bright_categorical_no_s1 <- glmer(
  auc_melatonin_exp ~ run_category_bright + (1 | participant_id),
  data = melatonin_persistence_bright_no_s1,
  family = Gamma(link = "log")
)

summary(model_persistence_bright_categorical_no_s1)

lrt_bright_no_s1 <- anova(model_persistence_bright_no_s1, model_persistence_bright_categorical_no_s1)
print(lrt_bright_no_s1)
```

#### Summary and Interpretation

**Main Findings:**

**Linear Models:**

1. **Dim evenings (with Session 1):** No significant persistence effect (β = -0.014, p = 0.917).

2. **Dim evenings (without Session 1):** Marginal positive trend (β = +0.237, p = 0.080), suggesting ~27% increase in melatonin per additional consecutive dim night.

3. **Bright evenings (both analyses):** No persistence effect (β = +0.060, p = 0.669), suggesting melatonin suppression is maximal after the first bright evening.

**Categorical Models:**

All categorical models showed no significant improvement over linear models (all LRT p > 0.35, all ΔAIC > 0). However, exploratory examination of the dim-night categorical model without Session 1 revealed that the marginal linear effect may be driven primarily by differences between first and third consecutive nights (Third vs. First: β = +0.475, p = 0.084, +61% increase), rather than strictly linear accumulation.

**Critical Limitations:**

This analysis is **exploratory** (n = 37-45 observations, 8 participants). The marginal dim-night effect (p = 0.080) should be interpreted cautiously. Session 1 exclusion dramatically changed effect estimates (β = -0.014 → +0.237), highlighting methodological sensitivity. Categorical models did not improve fit despite suggesting potential non-linear patterns, likely reflecting insufficient power to detect complex dynamics.

**Conclusion:**

While this exploratory analysis *raises the possibility* of differential persistence between light conditions, the current data are insufficient for causal inference. Claims about persistence effects remain speculative pending pre-registered replication in larger samples.

## Summary of Exploratory Analyses

The primary model (m2_mel: Light_Lag0_num + exp_session) with random intercepts only proved most parsimonious. Sensitivity analyses confirmed the robustness of the main findings:

- **Random slopes:** Model fit was substantially improved (ΔAIC = -13.8), indicating individual differences in light sensitivity exist, but the average suppression effect remained highly significant and nearly identical (77.3% vs. 77.5%).

- **Demographic moderation:** Model fit was not meaningfully improved (ΔAIC = -0.1), and the exploratory sex interaction (p = 0.037) requires replication in larger samples (n=8 insufficient for between-subject inference).

- **Light lag interactions:** Not supported (ΔAIC = +3.5, all interaction p > 0.20), confirming independent, non-additive effects across days.

- **Persistence model:** Exploratory analysis (n=85 observations, 8 participants) tested whether consecutive exposure to the same light condition produces cumulative effects. Linear models showed no persistence effect for bright nights (β = +0.060, p = 0.669) and a marginal positive trend for dim nights when excluding Session 1 (β = +0.237, p = 0.080, ~27% increase per consecutive dim night). Categorical models did not improve fit (all ΔAIC > +1.0, all LRT p > 0.35). These findings are strictly hypothesis-generating. Future adequately powered studies (N ≥ 30, 4-5+ consecutive nights) are needed to test threshold or saturation dynamics.

**Final model:** Model m2_mel (n=67) with fixed effects for Light_Lag0 and a linear session effect, with random intercepts only.

### Model Comparison Overview
```{r}
#| label: model-comparison-overview
#| tbl-cap: "Comprehensive model comparison: All tested models with key parameters"

# Helper functions to extract coefficients from GLMM
get_coef_glmm <- function(model, coef_name) {
  coefs <- summary(model)$coefficients
  if (coef_name %in% rownames(coefs)) {
    return(coefs[coef_name, "Estimate"])
  } else {
    return(NA)
  }
}

get_p_glmm <- function(model, coef_name) {
  coefs <- summary(model)$coefficients
  if (coef_name %in% rownames(coefs)) {
    return(coefs[coef_name, "Pr(>|z|)"])
  } else {
    return(NA)
  }
}

# Create extended comparison table
model_comparison_extended <- data.frame(
  Model = c(
    "m0 (Baseline)",
    "m1a (Lag 0)",
    "m1b (Lag 0+1)",
    "m1c (Lag 0+1+2)",
    "m2 (+ Session)",
    "m3 (+ Session²)",
    "m4 (+ Interaction)",
    "m0_robust",
    "m1a_robust",
    "m2_robust",
    "Random Slopes",
    "Demographics",
    "Light Interactions",
    "Persistence Dim (S1)",
    "Persistence Dim (no S1)",
    "Persistence Bright (S1)",
    "Persistence Bright (no S1)"
  ),
  
  Description = c(
    "Intercept only",
    "Light Lag 0 only",
    "Lag 0 + Lag 1",
    "Lag 0 + Lag 1 + Lag 2",
    "Lag 0 + session (linear)",
    "Lag 0 + session + session²",
    "Lag 0 + session + Lag0×Session",
    "Baseline (robust dataset)",
    "Light Lag 0 only (robust)",
    "Lag 0 + session (robust)",
    "m1a_robust + random slope Lag0",
    "m1a_robust + age + sex + interactions",
    "m1a_robust + Light lag interactions",
    "Dim: run_length (with Session 1)",
    "Dim: run_length (without Session 1)",
    "Bright: run_length (with Session 1)",
    "Bright: run_length (without Session 1)"
  ),
  
  npar = c(
    attr(logLik(m0_mel), "df"),
    attr(logLik(m1a_mel), "df"),
    attr(logLik(m1b_mel), "df"),
    attr(logLik(m1c_mel), "df"),
    attr(logLik(m2_mel), "df"),
    attr(logLik(m3_mel), "df"),
    attr(logLik(m4_mel), "df"),
    attr(logLik(m0_robust), "df"),
    attr(logLik(m1a_robust), "df"),
    attr(logLik(m2_robust), "df"),
    attr(logLik(m_random_slopes), "df"),
    attr(logLik(m_demographics_moderation), "df"),
    attr(logLik(m_light_interactions), "df"),
    attr(logLik(model_persistence_dim_numeric), "df"),
    attr(logLik(model_persistence_dim_no_s1), "df"),
    attr(logLik(model_persistence_bright_numeric), "df"),
    attr(logLik(model_persistence_bright_no_s1), "df")
  ),
  
  N_obs = c(
    nrow(melatonin_glmm),
    nrow(melatonin_glmm),
    nrow(melatonin_glmm),
    nrow(melatonin_glmm),
    nrow(melatonin_glmm),
    nrow(melatonin_glmm),
    nrow(melatonin_glmm),
    nrow(melatonin_glmm_robust),
    nrow(melatonin_glmm_robust),
    nrow(melatonin_glmm_robust),
    nrow(melatonin_glmm_robust),
    nrow(melatonin_glmm_robust),
    nrow(melatonin_glmm_robust),
    45,  # Dim with S1
    37,  # Dim without S1
    40,  # Bright with S1
    40   # Bright without S1
  ),
  
  AIC = c(
    round(AIC(m0_mel), 1),
    round(AIC(m1a_mel), 1),
    round(AIC(m1b_mel), 1),
    round(AIC(m1c_mel), 1),
    round(AIC(m2_mel), 1),
    round(AIC(m3_mel), 1),
    round(AIC(m4_mel), 1),
    round(AIC(m0_robust), 1),
    round(AIC(m1a_robust), 1),
    round(AIC(m2_robust), 1),
    round(AIC(m_random_slopes), 1),
    round(AIC(m_demographics_moderation), 1),
    round(AIC(m_light_interactions), 1),
    round(AIC(model_persistence_dim_numeric), 1),
    round(AIC(model_persistence_dim_no_s1), 1),
    round(AIC(model_persistence_bright_numeric), 1),
    round(AIC(model_persistence_bright_no_s1), 1)
  ),
  
  BIC = c(
    round(BIC(m0_mel), 1),
    round(BIC(m1a_mel), 1),
    round(BIC(m1b_mel), 1),
    round(BIC(m1c_mel), 1),
    round(BIC(m2_mel), 1),
    round(BIC(m3_mel), 1),
    round(BIC(m4_mel), 1),
    round(BIC(m0_robust), 1),
    round(BIC(m1a_robust), 1),
    round(BIC(m2_robust), 1),
    round(BIC(m_random_slopes), 1),
    round(BIC(m_demographics_moderation), 1),
    round(BIC(m_light_interactions), 1),
    round(BIC(model_persistence_dim_numeric), 1),
    round(BIC(model_persistence_dim_no_s1), 1),
    round(BIC(model_persistence_bright_numeric), 1),
    round(BIC(model_persistence_bright_no_s1), 1)
  ),
  
  Main_Beta = c(
    NA,
    round(get_coef_glmm(m1a_mel, "Light_Lag0_num"), 3),
    round(get_coef_glmm(m1b_mel, "Light_Lag0_num"), 3),
    round(get_coef_glmm(m1c_mel, "Light_Lag0_num"), 3),
    round(get_coef_glmm(m2_mel, "Light_Lag0_num"), 3),
    round(get_coef_glmm(m3_mel, "Light_Lag0_num"), 3),
    round(get_coef_glmm(m4_mel, "Light_Lag0_num"), 3),
    NA,
    round(get_coef_glmm(m1a_robust, "Light_Lag0_num"), 3),
    round(get_coef_glmm(m2_robust, "Light_Lag0_num"), 3),
    round(get_coef_glmm(m_random_slopes, "Light_Lag0_num"), 3),
    round(get_coef_glmm(m_demographics_moderation, "Light_Lag0_num"), 3),
    round(get_coef_glmm(m_light_interactions, "Light_Lag0_num"), 3),
    round(get_coef_glmm(model_persistence_dim_numeric, "run_length_dim"), 3),
    round(get_coef_glmm(model_persistence_dim_no_s1, "run_length_dim"), 3),
    round(get_coef_glmm(model_persistence_bright_numeric, "run_length_bright"), 3),
    round(get_coef_glmm(model_persistence_bright_no_s1, "run_length_bright"), 3)
  ),
  
  Effect_pct = c(
    NA,
    paste0(round((1 - exp(get_coef_glmm(m1a_mel, "Light_Lag0_num"))) * 100, 1), "%"),
    paste0(round((1 - exp(get_coef_glmm(m1b_mel, "Light_Lag0_num"))) * 100, 1), "%"),
    paste0(round((1 - exp(get_coef_glmm(m1c_mel, "Light_Lag0_num"))) * 100, 1), "%"),
    paste0(round((1 - exp(get_coef_glmm(m2_mel, "Light_Lag0_num"))) * 100, 1), "%"),
    paste0(round((1 - exp(get_coef_glmm(m3_mel, "Light_Lag0_num"))) * 100, 1), "%"),
    paste0(round((1 - exp(get_coef_glmm(m4_mel, "Light_Lag0_num"))) * 100, 1), "%"),
    NA,
    paste0(round((1 - exp(get_coef_glmm(m1a_robust, "Light_Lag0_num"))) * 100, 1), "%"),
    paste0(round((1 - exp(get_coef_glmm(m2_robust, "Light_Lag0_num"))) * 100, 1), "%"),
    paste0(round((1 - exp(get_coef_glmm(m_random_slopes, "Light_Lag0_num"))) * 100, 1), "%"),
    paste0(round((1 - exp(get_coef_glmm(m_demographics_moderation, "Light_Lag0_num"))) * 100, 1), "%"),
    paste0(round((1 - exp(get_coef_glmm(m_light_interactions, "Light_Lag0_num"))) * 100, 1), "%"),
    paste0(round((exp(get_coef_glmm(model_persistence_dim_numeric, "run_length_dim")) - 1) * 100, 1), "%/day"),
    paste0(round((exp(get_coef_glmm(model_persistence_dim_no_s1, "run_length_dim")) - 1) * 100, 1), "%/day"),
    paste0(round((exp(get_coef_glmm(model_persistence_bright_numeric, "run_length_bright")) - 1) * 100, 1), "%/day"),
    paste0(round((exp(get_coef_glmm(model_persistence_bright_no_s1, "run_length_bright")) - 1) * 100, 1), "%/day")
  ),
  
  Main_p = c(
    NA,
    format.pval(get_p_glmm(m1a_mel, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_glmm(m1b_mel, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_glmm(m1c_mel, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_glmm(m2_mel, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_glmm(m3_mel, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_glmm(m4_mel, "Light_Lag0_num"), digits = 3, eps = 0.001),
    NA,
    format.pval(get_p_glmm(m1a_robust, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_glmm(m2_robust, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_glmm(m_random_slopes, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_glmm(m_demographics_moderation, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_glmm(m_light_interactions, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_glmm(model_persistence_dim_numeric, "run_length_dim"), digits = 3, eps = 0.001),
    format.pval(get_p_glmm(model_persistence_dim_no_s1, "run_length_dim"), digits = 3, eps = 0.001),
    format.pval(get_p_glmm(model_persistence_bright_numeric, "run_length_bright"), digits = 3, eps = 0.001),
    format.pval(get_p_glmm(model_persistence_bright_no_s1, "run_length_bright"), digits = 3, eps = 0.001)
  ),
  
  Status = c(
    "Baseline",
    "Step 1",
    "Not selected",
    "Not selected",
    "Full dataset",
    "Not selected",
    "Not selected",
    "Baseline (robust)",
    "FINAL MODEL",
    "Comparison",
    "Exploratory",
    "Exploratory",
    "Exploratory",
    "Exploratory",
    "Exploratory",
    "Exploratory",
    "Exploratory"
  )
)

kable(model_comparison_extended, 
      caption = "Model Comparison: All Tested Models with Key Parameters (Melatonin)",
      align = c('l', 'l', 'c', 'c', 'c', 'c', 'c', 'c', 'c', 'l'),
      col.names = c("Model", "Description", "npar", "N", "AIC", "BIC", 
                    "Main β", "Effect %", "Main p", "Status")) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  row_spec(9, bold = TRUE, background = "#50C878") %>%  # Highlight m1a_robust (row 9) in GREEN
  pack_rows("Full Dataset Analysis (n=67)", 1, 7) %>%
  pack_rows("Robust Dataset Analysis (n=63)", 8, 10) %>%
  pack_rows("Exploratory Analyses (Robust Dataset)", 11, 17)
```

**Note:** 

**Final Model Selection:** m1a_robust (row 9, highlighted in green) was selected 
as the final model based on:
1. Superior diagnostic performance (all DHARMa tests passed)
2. Excellent model fit (AIC = 185.6, BIC = 194.2)
3. No influential observations (4 excluded from full dataset)
4. Parsimony (4 parameters vs. 5 in m2_robust)
5. Session effect not justified in robust dataset (LRT p = 0.316)

The full dataset hierarchical model building (rows 1-7) identified m2 as the 
best-fitting model with significant session effect (p = 0.031). However, 
diagnostic analysis revealed 4 influential observations causing quantile 
deviations (DHARMa p = 0.023). Sensitivity analysis on the robust dataset 
(rows 8-10) showed that session effect was driven by these influential 
observations and became non-significant after exclusion.

All exploratory analyses (rows 11-17) are based on the robust dataset and 
retain exp_session as a control variable to account for potential temporal 
trends, following best practices for exploratory research.

---

## Final Model: Fixed Effects
```{r}
#| label: final-model-fixed-effects-m1a
#| tbl-cap: "Final Model (m1a_robust): Fixed Effects with Benjamini-Hochberg Correction"

# Display final model results (m1a_robust, not m2)
kable(m1a_results, 
      col.names = c("Effect", "β (log scale)", "SE", "t", "p", "p (BH-adj)", "Suppression")) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

```{r}
#| label: variance-components-final-m1a

# Variance components for m1a_robust
vc_final <- as.data.frame(VarCorr(m1a_robust))

cat("\n**Variance Components (m1a_robust):**\n")
cat("Between-participant (Random Intercept):", round(vc_final$vcov[1], 3), "\n")
cat("Within-participant (Residual):", round(vc_final$vcov[2], 3), "\n\n")

# Calculate ICC
icc_final <- vc_final$vcov[1] / (vc_final$vcov[1] + vc_final$vcov[2])
cat("ICC (Intraclass Correlation):", round(icc_final, 3), 
    paste0("(", round(icc_final * 100, 1), "% of variance between participants)"), "\n\n")

cat("**Model Fit:**\n")
cat("AIC:", round(AIC(m1a_robust), 1), "| BIC:", round(BIC(m1a_robust), 1), "\n")
cat("N obs:", nrow(melatonin_glmm_robust), "| N participants:", 
    length(unique(melatonin_glmm_robust$participant_id)), "\n")
cat("Family: Gamma (log link)\n\n")
cat("**Primary Finding:**\n")

# Extract coefficients
m1a_coef <- summary(m1a_robust)$coefficients

# Get beta and SE for Light effect
beta_light <- m1a_coef["Light_Lag0_num", "Estimate"]
se_light <- m1a_coef["Light_Lag0_num", "Std. Error"]
t_light <- m1a_coef["Light_Lag0_num", "t value"]

# Calculate 95% CI for beta (on log scale)
ci_lower_beta <- beta_light - 1.96 * se_light
ci_upper_beta <- beta_light + 1.96 * se_light

# Transform to suppression percentage
# NOTE: CI bounds are reversed when transforming!
suppression_pct <- (1 - exp(beta_light)) * 100
ci_lower_suppression <- (1 - exp(ci_upper_beta)) * 100  # Upper beta → Lower suppression
ci_upper_suppression <- (1 - exp(ci_lower_beta)) * 100  # Lower beta → Upper suppression

# Print results
cat("Bright evening light (1000 lx) suppresses melatonin AUC by ", 
    round(suppression_pct, 1), "%\n", sep = "")
cat("(95% CI: [", round(ci_lower_suppression, 1), "%, ", 
    round(ci_upper_suppression, 1), "%]) compared to dim light (10 lx).\n", sep = "")
cat("β = ", round(beta_light, 3), ", SE = ", round(se_light, 3), 
    ", t = ", round(t_light, 3), ", p < 0.001\n", sep = "")
```