---
title: "Psychomotor Vigilance Test (PVT) analysis"
subtitle: "Carryover effects in melatonin suppression"
author: "Letizia Woerrlein"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: show
    code-tools: true
    code-line-numbers: true
    embed-resources: true
    theme: cosmo
    df-print: paged
execute:
  warning: false
  message: false
  cache: false
---

# Study Overview

## Background

The Psychomotor Vigilance Test (PVT) is a sustained-attention task that measures behavioral alertness and response speed. In this study, a 10-minute auditory PVT task was applied every 60 minutes (please note that due to time conflicts with depature times from the laboratory, the last two PVTs had only 30 minutes temporal distance. It was examined how different evening light exposure patterns (dim vs. bright) affect PVT performance across multiple consecutive days.

## Study Design Summary

- Participants: 9 total (P01-P07, P09-P11)
  - P11: Excluded (study dropout)
  - P07: Excluded from final analysis
  - Final sample: n = 8 participants

- Light Conditions:
  - D (Dim): 5 hours at 10 lx
  - B (Bright): 3h at 10 lx → 1.5h at 1000 lx → 0.5h at 10 lx

- PVT Testing:
  - 6 tests per evening session
  - Tests administered every 60 minutes (first 4 tests), then every 30 minutes (last 2 tests)
  - Primary analysis: Mean of trials 4 & 5 (during the 1.5h dim vs. bright intervention period)

## Key Analysis Features

- Lag analysis: Examining effects of current day (Lag 0), previous day (Lag 1), and two days prior (Lag 2)
- Primary outcome: Mean response speed (1/RT in milliseconds)
- Higher response speed = better performance (faster reactions)
- Approach: Convert simple reaction time into response speed for each recording of every session and PVT trail. Then build the average of all response speeds per PVT trail to produce one mean response speed for every PVT trail. Subsequently (dependent on which analysis step is taken), build the mean either for all PVT trails on a given session day (5h-exposure window) or create the average of trail 4 and 5 for every experimental evening to represent the light intervention window. 

---

# Setup

## Load Required Packages
```{r setup}
library(tidyverse)
library(gridExtra)
library(grid)
library(ggplot2)
library(cowplot)
library(lme4)
library(lmerTest) 
library(performance)
library(DHARMa)
library(car)
library(knitr)
library(kableExtra)
library(clubSandwich)
library(patchwork)
library(ggtext)

knitr::opts_knit$set(root.dir = normalizePath("C:/Users/anton/OneDrive/AfterGlow"))
```

---

# Data Import and Processing

## Step 1: Import PVT Data from RTF Files
```{r data-import}
# Use absolute path 
data_path <- "C:/Users/anton/OneDrive/AfterGlow/data/raw"
all_files <- list.files(data_path, pattern = "\\.(RTF|rtf)$", recursive = TRUE, full.names = TRUE)
cat("Found", length(all_files), "RTF files\n")

# Filter for PVT files
all_files <- all_files[grepl("pvt", all_files, ignore.case = TRUE)]
cat("Found", length(all_files), "PVT RTF files after filtering\n\n")

pvt_data_list <- list()
problem_lines_detailed <- data.frame()

for (i in seq_along(all_files)) {
  filepath <- all_files[i]
  
  participant <- str_extract(filepath, "P[0-9]{2}")
  session <- as.numeric(str_extract(str_extract(filepath, "([0-9]{2})_expsession"), "[0-9]{2}"))
  trial <- as.numeric(str_extract(str_extract(filepath, "pvt-([0-9]{2})_"), "[0-9]{2}"))
  
  raw_lines <- readLines(filepath, warn = FALSE)
  
  # ========== SPECIAL PARSER FOR P10 SESSION 10 ========== (different data format due to issues with data saving)
  if (participant == "P10" && session == 10) {
    rt_values <- c()
    et_values <- c()
    
    # Parse lines with format: \par    RT,  ET
    for (line in raw_lines) {
      # Remove \par and extract numbers separated by comma
      clean_line <- gsub("\\\\par", "", line)
      clean_line <- trimws(clean_line)
      
      # Try to extract two numbers separated by comma
      if (grepl("^[0-9]+,\\s*[0-9.]+$", clean_line)) {
        parts <- str_split(clean_line, ",")[[1]]
        if (length(parts) == 2) {
          rt <- as.numeric(trimws(parts[1]))
          et <- as.numeric(trimws(parts[2]))
          if (!is.na(rt) && !is.na(et)) {
            rt_values <- c(rt_values, rt)
            et_values <- c(et_values, et)
          }
        }
      }
    }
    
    if (length(rt_values) > 0 && length(rt_values) == length(et_values)) {
      pvt_data_list[[i]] <- data.frame(
        participant, session, trial,
        trial_within_test = 1:length(rt_values),
        RT = rt_values, ET = et_values, filepath
      )
    }
    
    if (i %% 50 == 0) cat("Processed", i, "of", length(all_files), "files\n")
    next  # Skip normal parsing
  }
  # ========== END SPECIAL PARSER ==========
  
  # Regular parsing for all other files
  data_start <- which(grepl("DATA IS RT\\(ms\\) & ET\\(sec\\):", raw_lines))
  
  if (length(data_start) == 0) next
  
  data_lines <- raw_lines[(data_start + 1):length(raw_lines)]
  rt_values <- c()
  et_values <- c()
  
  for (line_idx in seq_along(data_lines)) {
    numbers <- as.numeric(unlist(str_extract_all(data_lines[line_idx], "[0-9.]+")))
    numbers <- numbers[!is.na(numbers)]
    
    if (length(numbers) == 2) {
      rt_values <- c(rt_values, numbers[1])
      et_values <- c(et_values, numbers[2])
    } else if (length(numbers) == 1) {
      problem_lines_detailed <- rbind(problem_lines_detailed, data.frame(
        file = basename(filepath), participant, session, trial,
        line_number = data_start + line_idx, values_in_line = 1,
        line_content = trimws(data_lines[line_idx]), action = "Skipped"
      ))
    }
  }
  
  if (length(rt_values) == 0 || length(rt_values) != length(et_values)) next
  
  pvt_data_list[[i]] <- data.frame(
    participant, session, trial,
    trial_within_test = 1:length(rt_values),
    RT = rt_values, ET = et_values, filepath
  )
  
  if (i %% 50 == 0) cat("Processed", i, "of", length(all_files), "files\n")
}

pvt_data <- bind_rows(pvt_data_list[!sapply(pvt_data_list, is.null)])
cat("\nTotal rows imported:", nrow(pvt_data), "\n")

if (nrow(pvt_data) > 0) {
  cat("Participants:", paste(unique(pvt_data$participant), collapse = ", "), "\n")
}

# Verification: Check P10 sessions
cat("\n=== P10 SESSION CHECK ===\n")
p10_sessions_imported <- pvt_data %>%
  filter(participant == "P10") %>%
  group_by(session) %>%
  summarise(
    n_trials = n_distinct(trial),
    n_rows = n(),
    .groups = "drop"
  ) %>%
  arrange(session)

cat("P10 Sessions imported:\n")
print(p10_sessions_imported, n = 20)

# Check if Session 9 and 10 are different
if (10 %in% p10_sessions_imported$session && 9 %in% p10_sessions_imported$session) {
  s9_sample <- pvt_data %>%
    filter(participant == "P10", session == 9, trial == 1) %>%
    head(5) %>%
    pull(RT)
  
  s10_sample <- pvt_data %>%
    filter(participant == "P10", session == 10, trial == 1) %>%
    head(5) %>%
    pull(RT)
  
  cat("\nSession 9, Trial 1, First 5 RTs:", paste(s9_sample, collapse = ", "), "\n")
  cat("Session 10, Trial 1, First 5 RTs:", paste(s10_sample, collapse = ", "), "\n")
  
  if (identical(s9_sample, s10_sample)) {
    cat("WARNING: Session 9 and 10 appear identical!\n")
  } else {
    cat("SUCCESS: Session 9 and 10 are different!\n")
  }
}
```

## Step 2: Data Validation (Spot Check)
```{r data-validation}
if (nrow(pvt_data) > 0) {
  set.seed(42)
  random_checks <- pvt_data %>%
    group_by(participant) %>%
    slice_sample(n = 1) %>%
    ungroup() %>%
    slice_sample(n = 5)
  
  verification_results <- list()
  
  for (i in 1:nrow(random_checks)) {
    check <- random_checks[i, ]
    
    raw_lines <- readLines(check$filepath, warn = FALSE)
    
    # Extract participant and session for special handling
    participant <- check$participant
    session <- check$session
    
    rt_values <- c()
    et_values <- c()
    
    # ========== SPECIAL PARSER FOR P10 SESSION 10 ==========
    if (participant == "P10" && session == 10) {
      for (line in raw_lines) {
        clean_line <- gsub("\\\\par", "", line)
        clean_line <- trimws(clean_line)
        
        if (grepl("^[0-9]+,\\s*[0-9.]+$", clean_line)) {
          parts <- str_split(clean_line, ",")[[1]]
          if (length(parts) == 2) {
            rt <- as.numeric(trimws(parts[1]))
            et <- as.numeric(trimws(parts[2]))
            if (!is.na(rt) && !is.na(et)) {
              rt_values <- c(rt_values, rt)
              et_values <- c(et_values, et)
            }
          }
        }
      }
    } else {
      # Regular parsing for all other files
      data_start <- which(grepl("DATA IS RT\\(ms\\) & ET\\(sec\\):", raw_lines))
      if (length(data_start) > 0) {
        data_lines <- raw_lines[(data_start + 1):length(raw_lines)]
        
        for (line in data_lines) {
          numbers <- as.numeric(unlist(str_extract_all(line, "[0-9.]+")))
          numbers <- numbers[!is.na(numbers)]
          if (length(numbers) == 2) {
            rt_values <- c(rt_values, numbers[1])
            et_values <- c(et_values, numbers[2])
          }
        }
      }
    }
    # ========== END SPECIAL PARSER ==========
    
    if (check$trial_within_test <= length(rt_values)) {
      file_rt <- rt_values[check$trial_within_test]
      file_et <- et_values[check$trial_within_test]
      status <- ifelse(abs(check$RT - file_rt) < 0.01 && abs(check$ET - file_et) < 0.01, "MATCH", "MISMATCH")
    } else {
      status <- "NOT FOUND"
    }
    
    verification_results[[i]] <- list(participant = check$participant, session = check$session, 
                                      trial = check$trial, status = status)
  }
  
  cat("\n Validation:", sum(sapply(verification_results, function(x) x$status == "MATCH")), 
      "of 5 matches\n")
} else {
  cat("\n No data found - skipping validation\n")
}
```

## Step 3: Data Cleaning

Reaction times < 100ms are described as false starts, reaction times >= 500ms are identified as attention lapses and extreme outliers of reaction times above 30000ms are identified as outliers and will be exlcuded from analysis. Afterwards, response speed will be calculated from simple reacction times using this formula: 1000 / RT.

```{r data-cleaning}
pvt_clean <- pvt_data %>%
  filter(!is.na(RT), RT > 0) %>%
  mutate(
    false_start = ifelse(RT < 100, 1, 0),
    lapse = ifelse(RT >= 500, 1, 0),
    extreme_outlier = ifelse(RT > 30000, 1, 0),
    response_speed = 1000 / RT
  )

extreme_outliers <- pvt_clean %>% filter(RT > 30000)

if (nrow(extreme_outliers) > 0) {
  print(extreme_outliers %>%
          select(participant, session, trial, RT) %>%
          mutate(RT_sec = round(RT/1000, 1)) %>%
          arrange(desc(RT)))
  
  extreme_outliers %>%
    group_by(trial) %>%
    summarise(n = n()) %>%
    print()
}

pvt_clean <- pvt_clean %>%
  filter(RT <= 30000)
```

## Step 4: Calculate Summary Metrics

Following Basner & Dinges (2011) guidelines for PVT analysis.

```{r summary-metrics}
pvt_summary <- pvt_clean %>%
  group_by(participant, session, trial) %>%
  summarise(
    mean_response_speed = mean(response_speed[RT >= 100], na.rm = TRUE),
    n_lapses = sum(lapse, na.rm = TRUE),
    slowest_10pct_speed = mean(response_speed[RT >= quantile(RT, 0.9, na.rm = TRUE)], na.rm = TRUE),
    fastest_10pct_speed = mean(response_speed[RT <= quantile(RT, 0.1, na.rm = TRUE)], na.rm = TRUE),
    n_false_starts = sum(false_start, na.rm = TRUE),
    n_trials = n(),
    .groups = "drop"
  )

DT::datatable(
  pvt_summary,
  options = list(
    pageLength = 20,
    scrollX = TRUE,
    scrollY = "400px",
    order = list(list(0, 'asc'), list(1, 'asc'), list(2, 'asc'))
  ),
  caption = "PVT Summary: Mean response speed and performance metrics",
  rownames = FALSE,
  filter = 'top'
) %>%
  DT::formatStyle(
    'participant',
    backgroundColor = DT::styleEqual(
      unique(pvt_summary$participant),
      rep(c('#f0f0f0', '#ffffff'), length.out = length(unique(pvt_summary$participant)))
    )
  )
```

## Step 5: Calculate Session-Level Means

```{r session-means-table}
pvt_session_means <- pvt_summary %>%
  filter(!(participant == "P10" & session == 5)) %>%
  group_by(participant, session) %>%
  summarise(mean_response_speed = mean(mean_response_speed, na.rm = TRUE), .groups = "drop")

pvt_p01 <- pvt_session_means %>% filter(participant == "P01")
pvt_p02 <- pvt_session_means %>% filter(participant == "P02")
pvt_p03 <- pvt_session_means %>% filter(participant == "P03")
pvt_p04 <- pvt_session_means %>% filter(participant == "P04")
pvt_p05 <- pvt_session_means %>% filter(participant == "P05")
pvt_p06 <- pvt_session_means %>% filter(participant == "P06")
pvt_p07 <- pvt_session_means %>% filter(participant == "P07")
pvt_p09 <- pvt_session_means %>% filter(participant == "P09")
pvt_p10 <- pvt_session_means %>% filter(participant == "P10")
```

```{r session-means}
pvt_session_means <- pvt_summary %>%
  filter(!(participant == "P10" & session == 5)) %>%
  group_by(participant, session) %>%
  summarise(mean_response_speed = mean(mean_response_speed, na.rm = TRUE), .groups = "drop")

# Display as interactive table
DT::datatable(
  pvt_session_means,
  options = list(
    pageLength = 20,
    scrollX = TRUE,
    scrollY = "400px",
    order = list(list(0, 'asc'), list(1, 'asc'))
  ),
  caption = "PVT Session Means: Mean response speed per participant and session (averaged across all 6 trials; P10 Session 5 excluded)",
  rownames = FALSE,
  filter = 'top'
) %>%
  DT::formatRound(columns = 'mean_response_speed', digits = 3) %>%
  DT::formatStyle(
    'participant',
    backgroundColor = DT::styleEqual(
      unique(pvt_session_means$participant),
      c('#E8F4F8', '#FFF9E6', '#E8F8F5', '#F9E8F8', '#FFF3E0', 
        '#E3F2FD', '#FCE4EC', '#F1F8E9', '#FFF8E1')
    )
  )

# Create individual participant dataframes
pvt_p01 <- pvt_session_means %>% filter(participant == "P01")
pvt_p02 <- pvt_session_means %>% filter(participant == "P02")
pvt_p03 <- pvt_session_means %>% filter(participant == "P03")
pvt_p04 <- pvt_session_means %>% filter(participant == "P04")
pvt_p05 <- pvt_session_means %>% filter(participant == "P05")
pvt_p06 <- pvt_session_means %>% filter(participant == "P06")
pvt_p07 <- pvt_session_means %>% filter(participant == "P07")
pvt_p09 <- pvt_session_means %>% filter(participant == "P09")
pvt_p10 <- pvt_session_means %>% filter(participant == "P10")
```

---

# Doublet and Triplet Pattern Definitions

## Group 1: DDDBBBDDBDB (11 sessions)

Participants: P01, P02, P03

```{r patterns-group1}
doublet_positions_g1_pvt <- tibble(
  doublet = c("DD", "DB", "BD", "BB"),
  position1_sessions = list(
    c(1, 2, 7),
    c(3, 8, 10),
    c(6, 9),
    c(4, 5)
  ),
  position2_sessions = list(
    c(2, 3, 8),
    c(4, 9, 11),
    c(7, 10),
    c(5, 6)
  )
)

triplet_positions_g1_pvt <- tibble(
  triplet = c("DDD", "DDB", "DBB", "BBB", "BBD", "BDD", "DBD", "BDB"),
  position1_sessions = list(
    c(1),         # DDD: 1-2-3
    c(2, 7),      # DDB: 2-3-4, 7-8-9
    c(3),         # DBB: 3-4-5
    c(4),         # BBB: 4-5-6
    c(5),         # BBD: 5-6-7
    c(6),         # BDD: 6-7-8
    c(8),         # DBD: 8-9-10
    c(9)          # BDB: 9-10-11
  ),
  position2_sessions = list(
    c(2),
    c(3, 8),
    c(4),
    c(5),
    c(6),
    c(7),
    c(9),
    c(10)
  ),
  position3_sessions = list(
    c(3),
    c(4, 9),
    c(5),
    c(6),
    c(7),
    c(8),
    c(10),
    c(11)
  )
)
```

## Group 2 & 3: DBBBDDDBDB (10 sessions)

Participants: P04, P05, P06, P07, P09

```{r patterns-group2-3}
doublet_positions_g2_g3_pvt <- tibble(
  doublet = c("DD", "DB", "BD", "BB"),
  position1_sessions = list(
    c(5, 6),
    c(1, 7, 9),
    c(4, 8),
    c(2, 3)
  ),
  position2_sessions = list(
    c(6, 7),
    c(2, 8, 10),
    c(5, 9),
    c(3, 4)
  )
)

triplet_positions_g2_g3_pvt <- tibble(
  triplet = c("DDD", "DDB", "DBB", "BBB", "BBD", "BDD", "DBD", "BDB"),
  position1_sessions = list(
    c(5),         # DDD: 5-6-7
    c(6),         # DDB: 6-7-8
    c(1),         # DBB: 1-2-3
    c(2),         # BBB: 2-3-4
    c(3),         # BBD: 3-4-5
    c(4),         # BDD: 4-5-6
    c(7),         # DBD: 7-8-9
    c(8)          # BDB: 8-9-10
  ),
  position2_sessions = list(
    c(6),
    c(7),
    c(2),
    c(3),
    c(4),
    c(5),
    c(8),
    c(9)
  ),
  position3_sessions = list(
    c(7),
    c(8),
    c(3),
    c(4),
    c(5),
    c(6),
    c(9),
    c(10)
  )
)
```

## Group 4: DDDBDBBBDBDD (P10 only, session 5 excluded)

```{r patterns-group4}
doublet_positions_g4_pvt <- tibble(
  doublet = c("DD", "DB", "BD", "BB"),
  position1_sessions = list(
    c(1, 2, 12),
    c(3, 6, 10),
    c(9, 11),
    c(7, 8)
  ),
  position2_sessions = list(
    c(2, 3, 13),
    c(4, 7, 11),
    c(10, 12),
    c(8, 9)
  )
)

triplet_positions_g4_pvt <- tibble(
  triplet = c("DDD", "DDB", "DBB", "BBB", "BBD", "BDD", "DBD", "BDB"),
  position1_sessions = list(
    c(1),         # DDD: 1-2-3
    c(2),         # DDB: 2-3-4
    c(6),         # DBB: 6-7-8
    c(7),         # BBB: 7-8-9
    c(8),         # BBD: 8-9-10
    c(11),        # BDD: 11-12-13
    c(10),        # DBD: 10-11-12
    c(9)          # BDB: 9-10-11
  ),
  position2_sessions = list(
    c(2),
    c(3),
    c(7),
    c(8),
    c(9),
    c(12),
    c(11),
    c(10)
  ),
  position3_sessions = list(
    c(3),
    c(4),
    c(8),
    c(9),
    c(10),
    c(13),
    c(12),
    c(11)
  )
)
```

---

# Pattern Analysis Functions

## Calculate Doublet Positions

```{r function-doublet}
calc_doublet_positions_pvt <- function(participant_id, pvt_data, doublet_map) {
  results <- doublet_map %>%
    rowwise() %>%
    mutate(
      pos1_speed = mean(pvt_data$mean_response_speed[pvt_data$session %in% position1_sessions], 
                        na.rm = TRUE),
      pos2_speed = mean(pvt_data$mean_response_speed[pvt_data$session %in% position2_sessions], 
                        na.rm = TRUE),
      participant = participant_id
    ) %>%
    ungroup() %>%
    select(participant, doublet, pos1_speed, pos2_speed)
  
  return(results)
}
```

## Calculate Triplet Positions

```{r function-triplet}
calc_triplet_positions_pvt <- function(participant_id, pvt_data, triplet_map) {
  results <- triplet_map %>%
    rowwise() %>%
    mutate(
      pos1_speed = if(length(position1_sessions) > 0) {
        mean(pvt_data$mean_response_speed[pvt_data$session %in% position1_sessions], na.rm = TRUE)
      } else { NA_real_ },
      pos2_speed = if(length(position2_sessions) > 0) {
        mean(pvt_data$mean_response_speed[pvt_data$session %in% position2_sessions], na.rm = TRUE)
      } else { NA_real_ },
      pos3_speed = if(length(position3_sessions) > 0) {
        mean(pvt_data$mean_response_speed[pvt_data$session %in% position3_sessions], na.rm = TRUE)
      } else { NA_real_ },
      participant = participant_id
    ) %>%
    ungroup() %>%
    select(participant, triplet, pos1_speed, pos2_speed, pos3_speed)
  
  return(results)
}
```

---

# Calculate Pattern Positions for All Participants

## Group 1 (P01, P02, P03)

```{r calc-positions-g1}
pos_pvt_doublet_p01 <- calc_doublet_positions_pvt("P01", pvt_p01, doublet_positions_g1_pvt)
pos_pvt_doublet_p02 <- calc_doublet_positions_pvt("P02", pvt_p02, doublet_positions_g1_pvt)
pos_pvt_doublet_p03 <- calc_doublet_positions_pvt("P03", pvt_p03, doublet_positions_g1_pvt)

pos_pvt_triplet_p01 <- calc_triplet_positions_pvt("P01", pvt_p01, triplet_positions_g1_pvt)
pos_pvt_triplet_p02 <- calc_triplet_positions_pvt("P02", pvt_p02, triplet_positions_g1_pvt)
pos_pvt_triplet_p03 <- calc_triplet_positions_pvt("P03", pvt_p03, triplet_positions_g1_pvt)
```

## Group 2 & 3 (P04, P05, P06, P07, P09)

```{r calc-positions-g2-3}
pos_pvt_doublet_p04 <- calc_doublet_positions_pvt("P04", pvt_p04, doublet_positions_g2_g3_pvt)
pos_pvt_doublet_p05 <- calc_doublet_positions_pvt("P05", pvt_p05, doublet_positions_g2_g3_pvt)
pos_pvt_doublet_p06 <- calc_doublet_positions_pvt("P06", pvt_p06, doublet_positions_g2_g3_pvt)
pos_pvt_doublet_p07 <- calc_doublet_positions_pvt("P07", pvt_p07, doublet_positions_g2_g3_pvt)
pos_pvt_doublet_p09 <- calc_doublet_positions_pvt("P09", pvt_p09, doublet_positions_g2_g3_pvt)

pos_pvt_triplet_p04 <- calc_triplet_positions_pvt("P04", pvt_p04, triplet_positions_g2_g3_pvt)
pos_pvt_triplet_p05 <- calc_triplet_positions_pvt("P05", pvt_p05, triplet_positions_g2_g3_pvt)
pos_pvt_triplet_p06 <- calc_triplet_positions_pvt("P06", pvt_p06, triplet_positions_g2_g3_pvt)
pos_pvt_triplet_p07 <- calc_triplet_positions_pvt("P07", pvt_p07, triplet_positions_g2_g3_pvt)
pos_pvt_triplet_p09 <- calc_triplet_positions_pvt("P09", pvt_p09, triplet_positions_g2_g3_pvt)
```

## Group 4 (P10)

```{r calc-positions-g4}
pos_pvt_doublet_p10 <- calc_doublet_positions_pvt("P10", pvt_p10, doublet_positions_g4_pvt)
pos_pvt_triplet_p10 <- calc_triplet_positions_pvt("P10", pvt_p10, triplet_positions_g4_pvt)
```

## Combine All Positions

```{r combine-positions}
all_doublet_positions_pvt <- bind_rows(
  pos_pvt_doublet_p01, pos_pvt_doublet_p02, pos_pvt_doublet_p03,
  pos_pvt_doublet_p04, pos_pvt_doublet_p05, pos_pvt_doublet_p06,
  pos_pvt_doublet_p07, pos_pvt_doublet_p09, pos_pvt_doublet_p10
)

all_triplet_positions_pvt <- bind_rows(
  pos_pvt_triplet_p01, pos_pvt_triplet_p02, pos_pvt_triplet_p03,
  pos_pvt_triplet_p04, pos_pvt_triplet_p05, pos_pvt_triplet_p06,
  pos_pvt_triplet_p07, pos_pvt_triplet_p09, pos_pvt_triplet_p10
)
```

## Verification: P10 Triplets

```{r verify-p10}
cat("\n=== VERIFICATION: P10 TRIPLETS ===\n")
cat("P10 Pattern: DDDBDBBBDBDD (Sessions 1-4, 6-13)\n")
cat("Expected triplets (consecutive only):\n")
cat("  1-2-3: DDD\n")
cat("  2-3-4: DDB\n")
cat("  6-7-8: DBB\n")
cat("  7-8-9: BBB\n")
cat("  8-9-10: BBD\n")
cat("  9-10-11: BDB\n")
cat("  10-11-12: DBD\n")
cat("  11-12-13: BDD\n\n")
print(pos_pvt_triplet_p10)
```

---

# Visualization Setup

## Design Configuration

```{r viz-config}
light_colors <- c("dim" = "#4A90E2", "bright" = "#F5A623")

doublet_colors_pvt <- c(
  "DD" = "#4A90E2", "BD" = "#4A90E2",
  "DB" = "#F5A623", "BB" = "#F5A623"
)

triplet_colors_pvt <- c(
  "DDD" = "#4A90E2", "BDD" = "#4A90E2", "DBD" = "#4A90E2", "BBD" = "#4A90E2",
  "DDB" = "#F5A623", "BDB" = "#F5A623", "DBB" = "#F5A623", "BBB" = "#F5A623"
)

doublet_labels_pvt <- c(
  "DD" = expression("D"*bold("D")),
  "BD" = expression("B"*bold("D")),
  "DB" = expression("D"*bold("B")),
  "BB" = expression("B"*bold("B"))
)

triplet_labels_pvt <- c(
  "DDD" = expression("DD"*bold("D")),
  "BDD" = expression("BD"*bold("D")),
  "DBD" = expression("DB"*bold("D")),
  "BBD" = expression("BB"*bold("D")),
  "DDB" = expression("DD"*bold("B")),
  "BDB" = expression("BD"*bold("B")),
  "DBB" = expression("DB"*bold("B")),
  "BBB" = expression("BB"*bold("B"))
)
```

---

# Results: Full 5-Hour Chamber Exposure

## Prepare Data (Excluding P07)

```{r prep-data-5h}
participant_doublet_means_noP07 <- all_doublet_positions_pvt %>%
  filter(participant != "P07") %>%
  select(participant, doublet, mean_speed = pos2_speed) %>%
  mutate(doublet = factor(doublet, levels = c("DD", "BD", "DB", "BB")))

participant_triplet_means_noP07 <- all_triplet_positions_pvt %>%
  filter(participant != "P07", !is.na(pos3_speed)) %>%
  select(participant, triplet, mean_speed = pos3_speed) %>%
  mutate(triplet = factor(triplet, levels = c("DDD", "BDD", "DBD", "BBD",
                                              "DDB", "BDB", "DBB", "BBB")))

mean_doublet_positions <- all_doublet_positions_pvt %>%
  filter(participant != "P07") %>%
  group_by(doublet) %>%
  summarise(
    n_participants = n(),
    pos1_mean = mean(pos1_speed, na.rm = TRUE),
    pos2_mean = mean(pos2_speed, na.rm = TRUE),
    .groups = "drop"
  )

df_plot_doublet_positions <- mean_doublet_positions %>%
  pivot_longer(cols = c(pos1_mean, pos2_mean), names_to = "position", values_to = "mean_speed") %>%
  mutate(
    position_num = ifelse(position == "pos1_mean", 1, 2),
    light_type = ifelse(position == "pos1_mean", substr(doublet, 1, 1), substr(doublet, 2, 2)),
    light_label = ifelse(light_type == "D", "dim", "bright"),
    doublet = factor(doublet, levels = c("DD", "BD", "DB", "BB"))
  )

mean_triplet_positions <- all_triplet_positions_pvt %>%
  filter(participant != "P07") %>%
  group_by(triplet) %>%
  summarise(
    n_participants = n(),
    pos1_mean = mean(pos1_speed, na.rm = TRUE),
    pos2_mean = mean(pos2_speed, na.rm = TRUE),
    pos3_mean = mean(pos3_speed, na.rm = TRUE),
    .groups = "drop"
  )

df_plot_triplet_positions <- mean_triplet_positions %>%
  pivot_longer(cols = c(pos1_mean, pos2_mean, pos3_mean), names_to = "position", values_to = "mean_speed") %>%
  mutate(
    position_num = case_when(
      position == "pos1_mean" ~ 1,
      position == "pos2_mean" ~ 2,
      position == "pos3_mean" ~ 3
    ),
    light_type = substr(triplet, position_num, position_num),
    light_label = ifelse(light_type == "D", "dim", "bright"),
    triplet = factor(triplet, levels = c("DDD", "BDD", "DBD", "BBD", "DDB", "BDB", "DBB", "BBB"))
  ) %>%
  filter(!is.na(mean_speed))
```

## Y-Axis Limits

```{r y-limits-5h}
all_doublet_speeds <- participant_doublet_means_noP07$mean_speed
y_min_doublet <- floor(min(all_doublet_speeds, na.rm = TRUE) * 20) / 20
y_max_doublet <- ceiling(max(all_doublet_speeds, na.rm = TRUE) * 20) / 20

all_triplet_speeds <- participant_triplet_means_noP07$mean_speed
y_min_triplet <- floor(min(all_triplet_speeds, na.rm = TRUE) * 20) / 20
y_max_triplet <- ceiling(max(all_triplet_speeds, na.rm = TRUE) * 20) / 20

all_line_speeds <- c(df_plot_doublet_positions$mean_speed, df_plot_triplet_positions$mean_speed)
y_min_line <- floor(min(all_line_speeds, na.rm = TRUE) * 2) / 2
y_max_line <- ceiling(max(all_line_speeds, na.rm = TRUE) * 2) / 2
```

## Individual Plots

### Doublet Boxplot

```{r plot-doublet-box-5h, fig.width=8, fig.height=8}
pvt_doublet_box <- ggplot(participant_doublet_means_noP07, aes(x = doublet, y = mean_speed)) +
  geom_boxplot(aes(fill = doublet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16, stroke = 0) +
  scale_fill_manual(values = doublet_colors_pvt) +
  scale_x_discrete(labels = doublet_labels_pvt) +
  scale_y_continuous(limits = c(y_min_doublet, y_max_doublet),
                     breaks = seq(floor(y_min_doublet), ceiling(y_max_doublet), 0.5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  labs(title = "Psychomotor vigilance test (PVT) for light exposure doublets",
       subtitle = "Mean response speed by doublet type and participant (5-h light exposure; P07 excluded)",
       x = "Doublet", 
       y = "Mean response speed (1/s)") +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0, face = "plain"),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10), face = "plain"),
    axis.text = element_text(color = "black", size = 13, face = "plain"),
    axis.title = element_text(size = 14, face = "plain"),
    axis.title.x = element_text(face = "plain"),
    axis.title.y = element_text(margin = margin(r = 10), face = "plain"),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    plot.margin = margin(15, 10, 10, 10),
    legend.position = "none"
  )

print(pvt_doublet_box)
```

### Triplet Boxplot

```{r plot-triplet-box-5h, fig.width=8, fig.height=8}
pvt_triplet_box <- ggplot(participant_triplet_means_noP07, aes(x = triplet, y = mean_speed)) +
  geom_boxplot(aes(fill = triplet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16, stroke = 0) +
  scale_fill_manual(values = triplet_colors_pvt) +
  scale_x_discrete(labels = triplet_labels_pvt) +
  scale_y_continuous(limits = c(y_min_triplet, y_max_triplet),
                     breaks = seq(floor(y_min_triplet), ceiling(y_max_triplet), 0.5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  labs(title = "Psychomotor vigilance test (PVT) for light exposure triplets",
       subtitle = "Mean response speed by doublet type and participant (5-h light exposure; P07 excluded)",
       x = "Triplet", 
       y = "Mean response speed (1/s)") +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0, face = "plain"),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10), face = "plain"),
    axis.text = element_text(color = "black", size = 13, face = "plain"),
    axis.text.x = element_text(angle = 0, hjust = 0.5, face = "plain"),
    axis.title = element_text(size = 14, face = "plain"),
    axis.title.x = element_text(face = "plain"),
    axis.title.y = element_text(margin = margin(r = 10), face = "plain"),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    plot.margin = margin(15, 10, 10, 10),
    legend.position = "none"
  )

print(pvt_triplet_box)
```

### Doublet Line Plot

```{r plot-doublet-line-5h, fig.width=8, fig.height=8}
pvt_doublet_line <- ggplot(df_plot_doublet_positions, aes(x = position_num, y = mean_speed, group = doublet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_label), size = 3.5, shape = 16) +
  scale_color_manual(values = light_colors, labels = c("dim" = "Dim (10 lx)", "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2), labels = c("Position 1", "Position 2"), limits = c(0.8, 2.2)) +
  scale_y_continuous(limits = c(y_min_line, y_max_line),
                     breaks = seq(floor(y_min_line), ceiling(y_max_line), 0.5),
                     expand = expansion(mult = c(0.05, 0.05))) +
  labs(title = "Psychomotor vigilance test (PVT) response speed across doublet positions",
       subtitle = "Mean values across all participants (5-h light exposure; P07 excluded)",
       x = "Doublet position", 
       y = "Mean response speed (1/s)",
       color = "Light condition") +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0, face = "plain"),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10), face = "plain"),
    axis.text = element_text(color = "black", size = 13, face = "plain"),
    axis.title = element_text(size = 14, face = "plain"),
    axis.title.x = element_text(face = "plain"),
    axis.title.y = element_text(margin = margin(r = 10), face = "plain"),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    plot.margin = margin(15, 10, 10, 10),
    legend.position = "right",
    legend.title = element_text(size = 13, face = "plain"),
    legend.text = element_text(size = 12, face = "plain")
  )

print(pvt_doublet_line)
```

### Triplet Line Plot

```{r plot-triplet-line-5h, fig.width=8, fig.height=8}
pvt_triplet_line <- ggplot(df_plot_triplet_positions, aes(x = position_num, y = mean_speed, group = triplet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_label), size = 3.5, shape = 16) +
  scale_color_manual(values = light_colors, labels = c("dim" = "Dim (10 lx)", "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2, 3), labels = c("Position 1", "Position 2", "Position 3"), limits = c(0.8, 3.2)) +
  scale_y_continuous(limits = c(y_min_line, y_max_line),
                     breaks = seq(floor(y_min_line), ceiling(y_max_line), 0.5),
                     expand = expansion(mult = c(0.05, 0.05))) +
  labs(title = "Psychomotor vigilance test (PVT) response speed across triplet positions",
       subtitle = "Mean values across all participants (5-h light exposure; P07 excluded)",
       x = "Triplet position", 
       y = "Mean response speed (1/s)",
       color = "Light condition") +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0, face = "plain"),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10), face = "plain"),
    axis.text = element_text(color = "black", size = 13, face = "plain"),
    axis.title = element_text(size = 14, face = "plain"),
    axis.title.x = element_text(face = "plain"),
    axis.title.y = element_text(margin = margin(r = 10), face = "plain"),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    plot.margin = margin(15, 10, 10, 10),
    legend.position = "right",
    legend.title = element_text(size = 13, face = "plain"),
    legend.text = element_text(size = 12, face = "plain")
  )

print(pvt_triplet_line)
```

## Combined 4-Panel Figure (5-Hour Exposure)

```{r plot-combined-5h, fig.width=14, fig.height=14}
p1 <- ggplot(participant_doublet_means_noP07, aes(x = doublet, y = mean_speed)) +
  geom_boxplot(aes(fill = doublet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16, stroke = 0) +
  scale_fill_manual(values = doublet_colors_pvt) +
  scale_x_discrete(labels = doublet_labels_pvt) +
  scale_y_continuous(limits = c(y_min_doublet, y_max_doublet),
                     breaks = seq(floor(y_min_doublet), ceiling(y_max_doublet), 0.5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  labs(x = "Doublet", 
       y = "Mean response speed (1/s)") +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.margin = margin(10, 10, 10, 10),
    axis.text = element_text(color = "black", size = 13, face = "plain"),
    axis.title = element_text(size = 14, face = "plain"),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    legend.position = "none"
  ) +
  annotation_custom(grob = textGrob("A", x = 0.05, y = 0.95, gp = gpar(fontsize = 16, fontface = "bold")),
                    xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf)

p2 <- ggplot(df_plot_doublet_positions, aes(x = position_num, y = mean_speed, group = doublet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_label), size = 3.5, shape = 16) +
  scale_color_manual(values = light_colors, labels = c("dim" = "Dim (10 lx)", "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2), labels = c("Position 1", "Position 2"), limits = c(0.8, 2.2)) +
  scale_y_continuous(limits = c(y_min_line, y_max_line),
                     breaks = seq(floor(y_min_line), ceiling(y_max_line), 0.5),
                     expand = expansion(mult = c(0.05, 0.05))) +
  labs(x = "Doublet position", 
       y = "Mean response speed (1/s)",
       color = "Light condition") +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.margin = margin(10, 10, 10, 10),
    axis.text = element_text(color = "black", size = 13, face = "plain"),
    axis.title = element_text(size = 14, face = "plain"),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    legend.position = "none"
  ) +
  annotation_custom(grob = textGrob("B", x = 0.05, y = 0.95, gp = gpar(fontsize = 16, fontface = "bold")),
                    xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf)

p3 <- ggplot(participant_triplet_means_noP07, aes(x = triplet, y = mean_speed)) +
  geom_boxplot(aes(fill = triplet), outlier.shape = NA, width = 0.45, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.1), size = 2.5, alpha = 0.7, shape = 16, stroke = 0) +
  scale_fill_manual(values = triplet_colors_pvt) +
  scale_x_discrete(labels = triplet_labels_pvt) +
  scale_y_continuous(limits = c(y_min_triplet, y_max_triplet),
                     breaks = seq(floor(y_min_triplet), ceiling(y_max_triplet), 0.5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  labs(x = "Triplet", 
       y = "Mean response speed (1/s)") +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.margin = margin(10, 10, 10, 10),
    axis.text = element_text(color = "black", size = 11, face = "plain"),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 9),
    axis.text.y = element_text(size = 13),
    axis.title = element_text(size = 14, face = "plain"),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    legend.position = "none"
  ) +
  annotation_custom(grob = textGrob("C", x = 0.05, y = 0.95, gp = gpar(fontsize = 16, fontface = "bold")),
                    xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf)

p4 <- ggplot(df_plot_triplet_positions, aes(x = position_num, y = mean_speed, group = triplet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_label), size = 3.5, shape = 16) +
  scale_color_manual(values = light_colors, labels = c("dim" = "Dim (10 lx)", "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2, 3), labels = c("Position 1", "Position 2", "Position 3"), limits = c(0.8, 3.2)) +
  scale_y_continuous(limits = c(y_min_line, y_max_line),
                     breaks = seq(floor(y_min_line), ceiling(y_max_line), 0.5),
                     expand = expansion(mult = c(0.05, 0.05))) +
  labs(x = "Triplet position", 
       y = "Mean response speed (1/s)",
       color = "Light condition") +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.margin = margin(10, 10, 10, 10),
    axis.text = element_text(color = "black", size = 13, face = "plain"),
    axis.title = element_text(size = 14, face = "plain"),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    legend.position = "none"
  ) +
  annotation_custom(grob = textGrob("D", x = 0.05, y = 0.95, gp = gpar(fontsize = 16, fontface = "bold")),
                    xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf)

# Extract legend
p4_with_legend <- ggplot(df_plot_triplet_positions, aes(x = position_num, y = mean_speed, group = triplet)) +
  geom_point(aes(color = light_label), size = 3.5, shape = 16) +
  scale_color_manual(values = light_colors, labels = c("dim" = "Dim (10 lx)", "bright" = "Bright (1000 lx)")) +
  labs(color = "Light condition") +
  theme_classic(base_size = 14) +
  theme(
    legend.title = element_text(size = 13, face = "plain"),
    legend.text = element_text(size = 12, face = "plain")
  )

legend_grob <- cowplot::get_legend(p4_with_legend)

# Convert to grobs and align
g1 <- ggplotGrob(p1)
g2 <- ggplotGrob(p2)
g3 <- ggplotGrob(p3)
g4 <- ggplotGrob(p4)

maxWidth <- unit.pmax(g1$widths, g2$widths, g3$widths, g4$widths)
maxHeight <- unit.pmax(g1$heights, g2$heights, g3$heights, g4$heights)

g1$widths <- maxWidth
g2$widths <- maxWidth
g3$widths <- maxWidth
g4$widths <- maxWidth

g1$heights <- maxHeight
g2$heights <- maxHeight
g3$heights <- maxHeight
g4$heights <- maxHeight

title_grob <- textGrob(
  "Psychomotor vigilance test (PVT) response speed under doublet and triplet light patterns",
  gp = gpar(fontsize = 16, fontface = "plain"), hjust = 0, x = 0.01
)

subtitle_grob <- textGrob(
  "Mean response speed by pattern type and participant (5-h light exposure; P07 excluded)",
  gp = gpar(fontsize = 13, fontface = "plain"), hjust = 0, x = 0.01
)

plots_grid <- arrangeGrob(g1, g2, g3, g4, ncol = 2, nrow = 2)

pvt_combined_plot <- grid.arrange(
  title_grob, subtitle_grob,
  arrangeGrob(plots_grid, legend_grob, ncol = 2, widths = c(10, 1)),
  ncol = 1, heights = c(0.7, 0.6, 10)
)

print(pvt_combined_plot)
```

---

# Results: Trials 4 & 5 Only (1.5-Hour Light Intervention Window)

## Calculate Session Means for Trials 4 & 5

```{r session-means-t45}
pvt_session_means_t45 <- pvt_summary %>%
  filter(!(participant == "P10" & session == 5)) %>%
  filter(trial %in% c(4, 5)) %>%
  group_by(participant, session) %>%
  summarise(mean_response_speed = mean(mean_response_speed, na.rm = TRUE), .groups = "drop")

cat("Total sessions with Trial 4&5 data:", nrow(pvt_session_means_t45), "\n")
cat("Sessions per participant:\n")
print(pvt_session_means_t45 %>% count(participant))

pvt_p01_t45 <- pvt_session_means_t45 %>% filter(participant == "P01")
print(pvt_p01_t45)
pvt_p02_t45 <- pvt_session_means_t45 %>% filter(participant == "P02")
print(pvt_p02_t45)
pvt_p03_t45 <- pvt_session_means_t45 %>% filter(participant == "P03")
print(pvt_p03_t45)
pvt_p04_t45 <- pvt_session_means_t45 %>% filter(participant == "P04")
print(pvt_p04_t45)
pvt_p05_t45 <- pvt_session_means_t45 %>% filter(participant == "P05")
print(pvt_p05_t45)
pvt_p06_t45 <- pvt_session_means_t45 %>% filter(participant == "P06")
print(pvt_p06_t45)
pvt_p07_t45 <- pvt_session_means_t45 %>% filter(participant == "P07")
print(pvt_p07_t45)
pvt_p09_t45 <- pvt_session_means_t45 %>% filter(participant == "P09")
print(pvt_p09_t45)
pvt_p10_t45 <- pvt_session_means_t45 %>% filter(participant == "P10")
print(pvt_p10_t45)
```

## Calculate Pattern Positions for Trials 4 & 5

### Group 1

```{r calc-positions-t45-g1}
pos_pvt_doublet_p01_t45 <- calc_doublet_positions_pvt("P01", pvt_p01_t45, doublet_positions_g1_pvt)
pos_pvt_doublet_p02_t45 <- calc_doublet_positions_pvt("P02", pvt_p02_t45, doublet_positions_g1_pvt)
pos_pvt_doublet_p03_t45 <- calc_doublet_positions_pvt("P03", pvt_p03_t45, doublet_positions_g1_pvt)

pos_pvt_triplet_p01_t45 <- calc_triplet_positions_pvt("P01", pvt_p01_t45, triplet_positions_g1_pvt)
pos_pvt_triplet_p02_t45 <- calc_triplet_positions_pvt("P02", pvt_p02_t45, triplet_positions_g1_pvt)
pos_pvt_triplet_p03_t45 <- calc_triplet_positions_pvt("P03", pvt_p03_t45, triplet_positions_g1_pvt)
```

### Group 2 & 3

```{r calc-positions-t45-g2-3}
pos_pvt_doublet_p04_t45 <- calc_doublet_positions_pvt("P04", pvt_p04_t45, doublet_positions_g2_g3_pvt)
pos_pvt_doublet_p05_t45 <- calc_doublet_positions_pvt("P05", pvt_p05_t45, doublet_positions_g2_g3_pvt)
pos_pvt_doublet_p06_t45 <- calc_doublet_positions_pvt("P06", pvt_p06_t45, doublet_positions_g2_g3_pvt)
pos_pvt_doublet_p07_t45 <- calc_doublet_positions_pvt("P07", pvt_p07_t45, doublet_positions_g2_g3_pvt)
pos_pvt_doublet_p09_t45 <- calc_doublet_positions_pvt("P09", pvt_p09_t45, doublet_positions_g2_g3_pvt)

pos_pvt_triplet_p04_t45 <- calc_triplet_positions_pvt("P04", pvt_p04_t45, triplet_positions_g2_g3_pvt)
pos_pvt_triplet_p05_t45 <- calc_triplet_positions_pvt("P05", pvt_p05_t45, triplet_positions_g2_g3_pvt)
pos_pvt_triplet_p06_t45 <- calc_triplet_positions_pvt("P06", pvt_p06_t45, triplet_positions_g2_g3_pvt)
pos_pvt_triplet_p07_t45 <- calc_triplet_positions_pvt("P07", pvt_p07_t45, triplet_positions_g2_g3_pvt)
pos_pvt_triplet_p09_t45 <- calc_triplet_positions_pvt("P09", pvt_p09_t45, triplet_positions_g2_g3_pvt)
```

### Group 4

```{r calc-positions-t45-g4}
pos_pvt_doublet_p10_t45 <- calc_doublet_positions_pvt("P10", pvt_p10_t45, doublet_positions_g4_pvt)
pos_pvt_triplet_p10_t45 <- calc_triplet_positions_pvt("P10", pvt_p10_t45, triplet_positions_g4_pvt)
```

### Combine

```{r combine-positions-t45}
all_doublet_positions_pvt_t45 <- bind_rows(
  pos_pvt_doublet_p01_t45, pos_pvt_doublet_p02_t45, pos_pvt_doublet_p03_t45,
  pos_pvt_doublet_p04_t45, pos_pvt_doublet_p05_t45, pos_pvt_doublet_p06_t45,
  pos_pvt_doublet_p07_t45, pos_pvt_doublet_p09_t45, pos_pvt_doublet_p10_t45
)

all_triplet_positions_pvt_t45 <- bind_rows(
  pos_pvt_triplet_p01_t45, pos_pvt_triplet_p02_t45, pos_pvt_triplet_p03_t45,
  pos_pvt_triplet_p04_t45, pos_pvt_triplet_p05_t45, pos_pvt_triplet_p06_t45,
  pos_pvt_triplet_p07_t45, pos_pvt_triplet_p09_t45, pos_pvt_triplet_p10_t45
)
```

## Prepare Data for Plotting (Trials 4&5, Excluding P07)

```{r prep-data-t45}
participant_doublet_means_noP07_t45 <- all_doublet_positions_pvt_t45 %>%
  filter(participant != "P07") %>%
  select(participant, doublet, mean_speed = pos2_speed) %>%
  mutate(doublet = factor(doublet, levels = c("DD", "BD", "DB", "BB")))

participant_triplet_means_noP07_t45 <- all_triplet_positions_pvt_t45 %>%
  filter(participant != "P07", !is.na(pos3_speed)) %>%
  select(participant, triplet, mean_speed = pos3_speed) %>%
  mutate(triplet = factor(triplet, levels = c("DDD", "BDD", "DBD", "BBD",
                                              "DDB", "BDB", "DBB", "BBB")))

mean_doublet_positions_t45 <- all_doublet_positions_pvt_t45 %>%
  filter(participant != "P07") %>%
  group_by(doublet) %>%
  summarise(
    n_participants = n(),
    pos1_mean = mean(pos1_speed, na.rm = TRUE),
    pos2_mean = mean(pos2_speed, na.rm = TRUE),
    .groups = "drop"
  )

df_plot_doublet_positions_t45 <- mean_doublet_positions_t45 %>%
  pivot_longer(cols = c(pos1_mean, pos2_mean), names_to = "position", values_to = "mean_speed") %>%
  mutate(
    position_num = ifelse(position == "pos1_mean", 1, 2),
    light_type = ifelse(position == "pos1_mean", substr(doublet, 1, 1), substr(doublet, 2, 2)),
    light_label = ifelse(light_type == "D", "dim", "bright"),
    doublet = factor(doublet, levels = c("DD", "BD", "DB", "BB"))
  )

mean_triplet_positions_t45 <- all_triplet_positions_pvt_t45 %>%
  filter(participant != "P07") %>%
  group_by(triplet) %>%
  summarise(
    n_participants = n(),
    pos1_mean = mean(pos1_speed, na.rm = TRUE),
    pos2_mean = mean(pos2_speed, na.rm = TRUE),
    pos3_mean = mean(pos3_speed, na.rm = TRUE),
    .groups = "drop"
  )

df_plot_triplet_positions_t45 <- mean_triplet_positions_t45 %>%
  pivot_longer(cols = c(pos1_mean, pos2_mean, pos3_mean), names_to = "position", values_to = "mean_speed") %>%
  mutate(
    position_num = case_when(
      position == "pos1_mean" ~ 1,
      position == "pos2_mean" ~ 2,
      position == "pos3_mean" ~ 3
    ),
    light_type = substr(triplet, position_num, position_num),
    light_label = ifelse(light_type == "D", "dim", "bright"),
    triplet = factor(triplet, levels = c("DDD", "BDD", "DBD", "BBD", "DDB", "BDB", "DBB", "BBB"))
  ) %>%
  filter(!is.na(mean_speed))
```

## Y-Axis Limits (Trials 4&5)

```{r y-limits-t45}
all_doublet_speeds_t45 <- participant_doublet_means_noP07_t45$mean_speed
y_min_doublet_t45 <- floor(min(all_doublet_speeds_t45, na.rm = TRUE) * 20) / 20
y_max_doublet_t45 <- ceiling(max(all_doublet_speeds_t45, na.rm = TRUE) * 20) / 20

all_triplet_speeds_t45 <- participant_triplet_means_noP07_t45$mean_speed
y_min_triplet_t45 <- floor(min(all_triplet_speeds_t45, na.rm = TRUE) * 20) / 20
y_max_triplet_t45 <- ceiling(max(all_triplet_speeds_t45, na.rm = TRUE) * 20) / 20

all_line_speeds_t45 <- c(df_plot_doublet_positions_t45$mean_speed, df_plot_triplet_positions_t45$mean_speed)
y_min_line_t45 <- floor(min(all_line_speeds_t45, na.rm = TRUE) * 2) / 2
y_max_line_t45 <- ceiling(max(all_line_speeds_t45, na.rm = TRUE) * 2) / 2
```

## Individual Plots (Trials 4&5)

### Doublet Boxplot

```{r plot-doublet-box-t45, fig.width=8, fig.height=8}
pvt_doublet_box_t45 <- ggplot(participant_doublet_means_noP07_t45, aes(x = doublet, y = mean_speed)) +
  geom_boxplot(aes(fill = doublet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16, stroke = 0) +
  scale_fill_manual(values = doublet_colors_pvt) +
  scale_x_discrete(labels = doublet_labels_pvt) +
  scale_y_continuous(limits = c(y_min_doublet_t45, y_max_doublet_t45),
                     breaks = seq(floor(y_min_doublet_t45), ceiling(y_max_doublet_t45), 0.5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  labs(title = "Psychomotor vigilance test (PVT) for light exposure doublets",
       subtitle = "Mean response speed by doublet type and participant (1.5-h dim vs. bright period; P07 excluded)",
       x = "Doublet", 
       y = "Mean response speed (1/s)") +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0, face = "plain"),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10), face = "plain"),
    axis.text = element_text(color = "black", size = 13, face = "plain"),
    axis.title = element_text(size = 14, face = "plain"),
    axis.title.x = element_text(face = "plain"),
    axis.title.y = element_text(margin = margin(r = 10), face = "plain"),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    plot.margin = margin(15, 10, 10, 10),
    legend.position = "none"
  )

print(pvt_doublet_box_t45)
```

### Descriptive Statistics: Doublet Medians and IQR
```{r doublet-stats-pvt}
# Calculate statistics for 5-h exposure
doublet_stats_5h <- participant_doublet_means_noP07 %>%
  group_by(doublet) %>%
  summarise(
    n_5h = n(),
    median_5h = median(mean_speed, na.rm = TRUE),
    IQR_5h = IQR(mean_speed, na.rm = TRUE),
    .groups = "drop"
  )

# Calculate statistics for 1.5-h exposure
doublet_stats_1.5h <- participant_doublet_means_noP07_t45 %>%
  group_by(doublet) %>%
  summarise(
    n_1.5h = n(),
    median_1.5h = median(mean_speed, na.rm = TRUE),
    IQR_1.5h = IQR(mean_speed, na.rm = TRUE),
    .groups = "drop"
  )

# Combine both exposure windows
doublet_stats_combined <- doublet_stats_5h %>%
  full_join(doublet_stats_1.5h, by = "doublet") %>%
  arrange(doublet)

# Display as formatted table
knitr::kable(
  doublet_stats_combined,
  digits = 2,
  col.names = c("Doublet", "N", "Median", "IQR", "N", "Median", "IQR"),
  caption = "Descriptive statistics for doublet second-day PVT response speed across exposure windows (P07 excluded)",
  align = c("l", rep("r", 6))
) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  kableExtra::add_header_above(c(" " = 1, "5-h Exposure" = 3, "1.5-h Exposure" = 3)) %>%
  kableExtra::row_spec(0, bold = TRUE)
```

### Triplet Boxplot

```{r plot-triplet-box-t45, fig.width=8, fig.height=8}
pvt_triplet_box_t45 <- ggplot(participant_triplet_means_noP07_t45, aes(x = triplet, y = mean_speed)) +
  geom_boxplot(aes(fill = triplet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16, stroke = 0) +
  scale_fill_manual(values = triplet_colors_pvt) +
  scale_x_discrete(labels = triplet_labels_pvt) +
  scale_y_continuous(limits = c(y_min_triplet_t45, y_max_triplet_t45),
                     breaks = seq(floor(y_min_triplet_t45), ceiling(y_max_triplet_t45), 0.5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  labs(title = "Psychomotor vigilance test (PVT) for light exposure triplets",
       subtitle = "Mean response speed by triplet type and participant (1.5-h dim vs. bright period; P07 excluded)",
       x = "Triplet", 
       y = "Mean response speed (1/s)") +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0, face = "plain"),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10), face = "plain"),
    axis.text = element_text(color = "black", size = 13, face = "plain"),
    axis.text.x = element_text(angle = 0, hjust = 0.5, face = "plain"),
    axis.title = element_text(size = 14, face = "plain"),
    axis.title.x = element_text(face = "plain"),
    axis.title.y = element_text(margin = margin(r = 10), face = "plain"),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    plot.margin = margin(15, 10, 10, 10),
    legend.position = "none"
  )

print(pvt_triplet_box_t45)
```

### Descriptive Statistics: Triplet Medians and IQR
```{r triplet-stats-pvt}
# Calculate statistics for 5-h exposure
triplet_stats_5h <- participant_triplet_means_noP07 %>%
  group_by(triplet) %>%
  summarise(
    n_5h = n(),
    median_5h = median(mean_speed, na.rm = TRUE),
    IQR_5h = IQR(mean_speed, na.rm = TRUE),
    .groups = "drop"
  )

# Calculate statistics for 1.5-h exposure
triplet_stats_1.5h <- participant_triplet_means_noP07_t45 %>%
  group_by(triplet) %>%
  summarise(
    n_1.5h = n(),
    median_1.5h = median(mean_speed, na.rm = TRUE),
    IQR_1.5h = IQR(mean_speed, na.rm = TRUE),
    .groups = "drop"
  )

# Combine both exposure windows
triplet_stats_combined <- triplet_stats_5h %>%
  full_join(triplet_stats_1.5h, by = "triplet") %>%
  arrange(triplet)

# Display as formatted table
knitr::kable(
  triplet_stats_combined,
  digits = 2,
  col.names = c("Triplet", "N", "Median", "IQR", "N", "Median", "IQR"),
  caption = "Descriptive statistics for triplet third-day PVT response speed across exposure windows (P07 excluded)",
  align = c("l", rep("r", 6))
) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  kableExtra::add_header_above(c(" " = 1, "5-h Exposure" = 3, "1.5-h Exposure" = 3)) %>%
  kableExtra::row_spec(0, bold = TRUE)
```

## Detailed Data Overview: Triplet Boxplot
```{r triplet-data-detailed}
#| label: triplet-data-detailed
#| echo: true

# Show all data points used in the boxplot
detailed_triplet_data <- participant_triplet_means_noP07_t45 %>%
  select(participant, triplet, mean_speed) %>%
  arrange(triplet, participant)

# Display as formatted table
knitr::kable(
  detailed_triplet_data,
  digits = 3,
  col.names = c("Participant", "Triplet", "Mean Speed (1/s)"),
  caption = "Detailed data for triplet boxplot (1.5-h exposure, Trials 4&5, P07 excluded)",
  align = c("l", "l", "r")
) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  kableExtra::row_spec(0, bold = TRUE) %>%
  kableExtra::pack_rows(index = table(detailed_triplet_data$triplet))

# Summary statistics
summary_stats <- detailed_triplet_data %>%
  group_by(triplet) %>%
  summarise(
    N = n(),
    Participants = paste(sort(unique(participant)), collapse = ", "),
    Median = median(mean_speed),
    Mean = mean(mean_speed),
    SD = sd(mean_speed),
    Min = min(mean_speed),
    Max = max(mean_speed),
    .groups = "drop"
  )

knitr::kable(
  summary_stats,
  digits = 3,
  caption = "Summary statistics by triplet type",
  align = c("l", "r", "l", rep("r", 5))
) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  kableExtra::row_spec(0, bold = TRUE)
```

### Doublet Line Plot

```{r plot-doublet-line-t45, fig.width=8, fig.height=8}
pvt_doublet_line_t45 <- ggplot(df_plot_doublet_positions_t45, aes(x = position_num, y = mean_speed, group = doublet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_label), size = 3.5, shape = 16) +
  scale_color_manual(values = light_colors, labels = c("dim" = "Dim (10 lx)", "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2), labels = c("Position 1", "Position 2"), limits = c(0.8, 2.2)) +
  scale_y_continuous(limits = c(y_min_line_t45, y_max_line_t45),
                     breaks = seq(floor(y_min_line_t45), ceiling(y_max_line_t45), 0.5),
                     expand = expansion(mult = c(0.05, 0.05))) +
  labs(title = "Psychomotor vigilance test (PVT) response speed across doublet positions",
       subtitle = "Mean response speed by triplet type and participant (1.5-h dim vs. bright period; P07 excluded)",
       x = "Doublet position", 
       y = "Mean response speed (1/s)",
       color = "Light condition") +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0, face = "plain"),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10), face = "plain"),
    axis.text = element_text(color = "black", size = 13, face = "plain"),
    axis.title = element_text(size = 14, face = "plain"),
    axis.title.x = element_text(face = "plain"),
    axis.title.y = element_text(margin = margin(r = 10), face = "plain"),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    plot.margin = margin(15, 10, 10, 10),
    legend.position = "right",
    legend.title = element_text(size = 13, face = "plain"),
    legend.text = element_text(size = 12, face = "plain")
  )

print(pvt_doublet_line_t45)
```

### Triplet Line Plot

```{r plot-triplet-line-t45, fig.width=8, fig.height=8}
pvt_triplet_line_t45 <- ggplot(df_plot_triplet_positions_t45, aes(x = position_num, y = mean_speed, group = triplet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_label), size = 3.5, shape = 16) +
  scale_color_manual(values = light_colors, labels = c("dim" = "Dim (10 lx)", "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2, 3), labels = c("Position 1", "Position 2", "Position 3"), limits = c(0.8, 3.2)) +
  scale_y_continuous(limits = c(y_min_line_t45, y_max_line_t45),
                     breaks = seq(floor(y_min_line_t45), ceiling(y_max_line_t45), 0.5),
                     expand = expansion(mult = c(0.05, 0.05))) +
  labs(title = "Psychomotor vigilance test (PVT) response speed across triplet positions",
       subtitle = "Mean response speed by triplet type and participant (1.5-h dim vs. bright period; P07 excluded)",
       x = "Triplet position", 
       y = "Mean response speed (1/s)",
       color = "Light condition") +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0, face = "plain"),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10), face = "plain"),
    axis.text = element_text(color = "black", size = 13, face = "plain"),
    axis.title = element_text(size = 14, face = "plain"),
    axis.title.x = element_text(face = "plain"),
    axis.title.y = element_text(margin = margin(r = 10), face = "plain"),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    plot.margin = margin(15, 10, 10, 10),
    legend.position = "right",
    legend.title = element_text(size = 13, face = "plain"),
    legend.text = element_text(size = 12, face = "plain")
  )

print(pvt_triplet_line_t45)
```

## Descriptive Statistics (Trials 4&5)

```{r descriptive-stats-t45}
triplet_summary_t45 <- participant_triplet_means_noP07_t45 %>%
  group_by(triplet) %>%
  summarise(
    n = n(),
    median = median(mean_speed, na.rm = TRUE),
    Q1 = quantile(mean_speed, 0.25, na.rm = TRUE),
    Q3 = quantile(mean_speed, 0.75, na.rm = TRUE),
    IQR = IQR(mean_speed, na.rm = TRUE)
  ) %>%
  mutate(
    median = round(median, 3),
    Q1 = round(Q1, 3),
    Q3 = round(Q3, 3),
    IQR = round(IQR, 3)
  )

print(triplet_summary_t45)
```

## Combined 4-Panel Figure (Trials 4&5)

```{r plot-combined-t45, fig.width=14, fig.height=14}
p1_t45 <- ggplot(participant_doublet_means_noP07_t45, aes(x = doublet, y = mean_speed)) +
  geom_boxplot(aes(fill = doublet), outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16, stroke = 0) +
  scale_fill_manual(values = doublet_colors_pvt) +
  scale_x_discrete(labels = doublet_labels_pvt) +
  scale_y_continuous(limits = c(y_min_doublet_t45, y_max_doublet_t45),
                     breaks = seq(floor(y_min_doublet_t45), ceiling(y_max_doublet_t45), 0.5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  labs(x = "Doublet", 
       y = "Mean response speed (1/s)") +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.margin = margin(10, 10, 10, 10),
    axis.text = element_text(color = "black", size = 13, face = "plain"),
    axis.title = element_text(size = 14, face = "plain"),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    legend.position = "none"
  ) +
  annotation_custom(grob = textGrob("A", x = 0.05, y = 0.95, gp = gpar(fontsize = 16, fontface = "bold")),
                    xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf)

p2_t45 <- ggplot(df_plot_doublet_positions_t45, aes(x = position_num, y = mean_speed, group = doublet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_label), size = 3.5, shape = 16) +
  scale_color_manual(values = light_colors, labels = c("dim" = "Dim (10 lx)", "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2), labels = c("Position 1", "Position 2"), limits = c(0.8, 2.2)) +
  scale_y_continuous(limits = c(y_min_line_t45, y_max_line_t45),
                     breaks = seq(floor(y_min_line_t45), ceiling(y_max_line_t45), 0.5),
                     expand = expansion(mult = c(0.05, 0.05))) +
  labs(x = "Doublet position", 
       y = "Mean response speed (1/s)",
       color = "Light condition") +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.margin = margin(10, 10, 10, 10),
    axis.text = element_text(color = "black", size = 13, face = "plain"),
    axis.title = element_text(size = 14, face = "plain"),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    legend.position = "none"
  ) +
  annotation_custom(grob = textGrob("B", x = 0.05, y = 0.95, gp = gpar(fontsize = 16, fontface = "bold")),
                    xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf)

p3_t45 <- ggplot(participant_triplet_means_noP07_t45, aes(x = triplet, y = mean_speed)) +
  geom_boxplot(aes(fill = triplet), outlier.shape = NA, width = 0.45, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.1), size = 2.5, alpha = 0.7, shape = 16, stroke = 0) +
  scale_fill_manual(values = triplet_colors_pvt) +
  scale_x_discrete(labels = triplet_labels_pvt) +
  scale_y_continuous(limits = c(y_min_triplet_t45, y_max_triplet_t45),
                     breaks = seq(floor(y_min_triplet_t45), ceiling(y_max_triplet_t45), 0.5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  labs(x = "Triplet", 
       y = "Mean response speed (1/s)") +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.margin = margin(10, 10, 10, 10),
    axis.text = element_text(color = "black", size = 11, face = "plain"),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 9),
    axis.text.y = element_text(size = 13),
    axis.title = element_text(size = 14, face = "plain"),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    legend.position = "none"
  ) +
  annotation_custom(grob = textGrob("C", x = 0.05, y = 0.95, gp = gpar(fontsize = 16, fontface = "bold")),
                    xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf)

p4_t45 <- ggplot(df_plot_triplet_positions_t45, aes(x = position_num, y = mean_speed, group = triplet)) +
  geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = light_label), size = 3.5, shape = 16) +
  scale_color_manual(values = light_colors, labels = c("dim" = "Dim (10 lx)", "bright" = "Bright (1000 lx)")) +
  scale_x_continuous(breaks = c(1, 2, 3), labels = c("Position 1", "Position 2", "Position 3"), limits = c(0.8, 3.2)) +
  scale_y_continuous(limits = c(y_min_line_t45, y_max_line_t45),
                     breaks = seq(floor(y_min_line_t45), ceiling(y_max_line_t45), 0.5),
                     expand = expansion(mult = c(0.05, 0.05))) +
  labs(x = "Triplet position", 
       y = "Mean response speed (1/s)",
       color = "Light condition") +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.margin = margin(10, 10, 10, 10),
    axis.text = element_text(color = "black", size = 13, face = "plain"),
    axis.title = element_text(size = 14, face = "plain"),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    legend.position = "none"
  ) +
  annotation_custom(grob = textGrob("D", x = 0.05, y = 0.95, gp = gpar(fontsize = 16, fontface = "bold")),
                    xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf)

# Extract legend
p_legend_t45 <- ggplot(df_plot_triplet_positions_t45, aes(x = position_num, y = mean_speed, group = triplet)) +
  geom_point(aes(color = light_label), size = 3.5, shape = 16) +
  scale_color_manual(values = light_colors, labels = c("dim" = "Dim (10 lx)", "bright" = "Bright (1000 lx)")) +
  labs(color = "Light condition") +
  theme_classic(base_size = 14) +
  theme(
    legend.title = element_text(size = 13, face = "plain"),
    legend.text = element_text(size = 12, face = "plain")
  )

legend_grob_t45 <- cowplot::get_legend(p_legend_t45)

# Convert to grobs and align
g1_t45 <- ggplotGrob(p1_t45)
g2_t45 <- ggplotGrob(p2_t45)
g3_t45 <- ggplotGrob(p3_t45)
g4_t45 <- ggplotGrob(p4_t45)

maxWidth_t45 <- unit.pmax(g1_t45$widths, g2_t45$widths, g3_t45$widths, g4_t45$widths)
maxHeight_t45 <- unit.pmax(g1_t45$heights, g2_t45$heights, g3_t45$heights, g4_t45$heights)

g1_t45$widths <- maxWidth_t45
g2_t45$widths <- maxWidth_t45
g3_t45$widths <- maxWidth_t45
g4_t45$widths <- maxWidth_t45

g1_t45$heights <- maxHeight_t45
g2_t45$heights <- maxHeight_t45
g3_t45$heights <- maxHeight_t45
g4_t45$heights <- maxHeight_t45

title_grob_t45 <- textGrob(
  "Psychomotor vigilance test (PVT) response speed under doublet and triplet light patterns",
  gp = gpar(fontsize = 16, fontface = "plain"), hjust = 0, x = 0.01
)

subtitle_grob_t45 <- textGrob(
  "Mean response speed by pattern type and participant (1.5-h dim vs. bright period; P07 excluded)",
  gp = gpar(fontsize = 13, fontface = "plain"), hjust = 0, x = 0.01
)

plots_grid_t45 <- arrangeGrob(g1_t45, g2_t45, g3_t45, g4_t45, ncol = 2, nrow = 2)

pvt_combined_plot_t45 <- grid.arrange(
  title_grob_t45, subtitle_grob_t45,
  arrangeGrob(plots_grid_t45, legend_grob_t45, ncol = 2, widths = c(10, 1)),
  ncol = 1, heights = c(0.7, 0.6, 10)
)

print(pvt_combined_plot_t45)
```

## Individual PVT Response Speed Profiles

### Light Exposure Sequences
```{r define-light-patterns}
# Define light exposure patterns for each participant
light_patterns <- list(
  P01 = "DDDBBBDDBDB",    # Group 1: HBT 00:00, 11 sessions
  P02 = "DDDBBBDDBDB",    # Group 1: HBT 00:00, 11 sessions
  P03 = "DDDBBBDDBDB",    # Group 1: HBT 00:00, 11 sessions
  P04 = "DBBBDDDBDB",     # Group 2: HBT 22:00, 10 sessions
  P05 = "DBBBDDDBDB",     # Group 2: HBT 22:00, 10 sessions
  P06 = "DBBBDDDBDB",     # Group 3: HBT 23:30, 10 sessions
  P07 = "DBBBDDDBDB",     # Group 3: HBT 23:30, 10 sessions (excluded from analysis)
  P09 = "DBBBDDDBDB",     # Group 3: HBT 23:30, 10 sessions
  P10 = "DDDBXDBBBDBDD"   # Group 4: HBT 23:00, 13 sessions (X = session 5 excluded)
)

# Note: 
# D = Dim (10 lx)
# B = Bright (1000 lx)
# X = Excluded session (P10 session 5 - light automation malfunction)
```

### Individual Time-Course Plots (Panel View)
```{r pvt-profile-function}
# Function to create PVT response speed profile plot
create_pvt_profile <- function(participant_id, light_pattern) {
  
  # Use pvt_summary which already has mean_response_speed per trial
  p_data <- pvt_summary %>%
    filter(participant == participant_id) %>%
    select(session, trial, mean_response_speed) %>%
    mutate(light_condition = rep(strsplit(light_pattern, "")[[1]], each = 6)[1:n()])
  
  if (nrow(p_data) == 0) stop(paste("No data found for participant:", participant_id))
  
  # Add custom x-position with reduced spacing between trial 5 and 6
  p_data <- p_data %>%
    mutate(x_position = case_when(
      trial == 1 ~ 1.0,
      trial == 2 ~ 2.0,
      trial == 3 ~ 3.0,
      trial == 4 ~ 4.0,
      trial == 5 ~ 5.0,
      trial == 6 ~ 5.5  # Half distance (0.5 instead of 1.0)
    ))
  
  # Get session means (all 6 trials)
  p_means <- pvt_session_means %>%
    filter(participant == participant_id) %>%
    mutate(
      light_condition = strsplit(light_pattern, "")[[1]],
      mean_label = paste0("M=", round(mean_response_speed, 2))
    )
  
  # Create lookup for light conditions
  light_lookup <- setNames(p_means$light_condition, p_means$session)
  
  # Determine y-axis limits
  max_y <- max(p_data$mean_response_speed, na.rm = TRUE)
  min_y <- min(p_data$mean_response_speed, na.rm = TRUE)
  y_range <- max_y - min_y
  y_limit_upper <- ceiling((max_y + y_range * 0.1) * 10) / 10
  y_limit_lower <- floor((min_y - y_range * 0.1) * 10) / 10
  
  # Create plot
  p <- ggplot(p_data, aes(x = x_position, y = mean_response_speed)) +
    # Orange background for trials 4, 5, 6 in BRIGHT sessions only
    # Start at midpoint between trial 3 and 4, end at midpoint after trial 6
    geom_rect(data = p_data %>% filter(light_condition == "B", trial == 4) %>% 
                group_by(session) %>% slice(1),
              xmin = 4, xmax = 5.5, ymin = -Inf, ymax = Inf,
              fill = "#FFE0B2", alpha = 0.5, inherit.aes = FALSE) +
    geom_line(color = "black", linewidth = 0.8) +
    geom_point(color = "black", size = 2.5) +
   geom_text(data = p_means,
          aes(x = 0.5, y = Inf, label = mean_label),
          hjust = 0, vjust = 1.3, size = 3.5, inherit.aes = FALSE) +
    scale_x_continuous(
      breaks = c(1, 2, 3, 4, 5, 5.5),
      labels = c("1", "2", "3", "4", "5", "6"),
      limits = c(0.5, 6),
      expand = c(0, 0)
    ) +
    scale_y_continuous(limits = c(y_limit_lower, y_limit_upper), 
                       expand = expansion(mult = c(0, 0.05))) +
    facet_wrap(~ session, ncol = 4,
               labeller = as_labeller(function(x) {
                 paste0("Session ", x, " (", light_lookup[x], ")")
               })) +
    labs(
      title = paste0("PVT Response Speed per Session (", participant_id, ")"),
      x = "Trial Number",
      y = "Mean Response Speed (1/s)"
    ) +
    theme_classic(base_size = 11) +
    theme(
      aspect.ratio = 1,
      plot.title = element_text(size = 14, hjust = 0, margin = margin(b = 10)),
      axis.text = element_text(color = "black", size = 10),
      axis.title = element_text(size = 11),
      axis.line = element_line(linewidth = 0.5),
      axis.ticks = element_line(linewidth = 0.5),
      panel.grid = element_blank(),
      strip.background = element_rect(fill = "white", color = "grey80", linewidth = 0.5),
      strip.text = element_text(size = 10, color = "black"),
      panel.background = element_rect(fill = "white", color = "grey80", linewidth = 0.5),
      plot.margin = margin(10, 10, 10, 10),
      legend.position = "none"
    )
  
  return(p)
}
```

### P01 - P03 (Group 1)
```{r pvt-profile-p01}
#| fig-width: 12
#| fig-height: 9
create_pvt_profile("P01", light_patterns$P01)
```
```{r pvt-profile-p02}
#| fig-width: 12
#| fig-height: 9
create_pvt_profile("P02", light_patterns$P02)
```
```{r pvt-profile-p03}
#| fig-width: 12
#| fig-height: 9
create_pvt_profile("P03", light_patterns$P03)
```

### P04 - P05 (Group 2)
```{r pvt-profile-p04}
#| fig-width: 12
#| fig-height: 9
create_pvt_profile("P04", light_patterns$P04)
```
```{r pvt-profile-p05}
#| fig-width: 12
#| fig-height: 9
create_pvt_profile("P05", light_patterns$P05)
```

### P06, P07, P09 (Group 3)
```{r pvt-profile-p06}
#| fig-width: 12
#| fig-height: 9
create_pvt_profile("P06", light_patterns$P06)
```
```{r pvt-profile-p07}
#| fig-width: 12
#| fig-height: 9
# P07 - define pattern manually since it was excluded from analysis
create_pvt_profile("P07", "DBBBDDDBDB")
```
```{r pvt-profile-p09}
#| fig-width: 12
#| fig-height: 9
create_pvt_profile("P09", light_patterns$P09)
```

### P10 (Group 4) - Special Case
```{r pvt-profile-p10}
#| fig-width: 12
#| fig-height: 9

# P10 with session 5 excluded 
p10_data_special <- pvt_summary %>%
  filter(participant == "P10", session != 5) %>%
  select(session, trial, mean_response_speed) %>%
  mutate(
    # Add custom x-position with reduced spacing between trial 5 and 6
    x_position = case_when(
      trial == 1 ~ 1.0,
      trial == 2 ~ 2.0,
      trial == 3 ~ 3.0,
      trial == 4 ~ 4.0,
      trial == 5 ~ 5.0,
      trial == 6 ~ 5.5  # Half distance (0.5 instead of 1.0)
    ),
    light_condition = case_when(
      session %in% c(1, 2, 3) ~ "D",
      session == 4 ~ "B",
      session == 6 ~ "D",
      session %in% c(7, 8, 9) ~ "B",
      session == 10 ~ "D",
      session == 11 ~ "B",
      session %in% c(12, 13) ~ "D"
    )
  )

p10_means_special <- pvt_session_means %>%
  filter(participant == "P10", session != 5) %>%
  mutate(
    light_condition = case_when(
      session %in% c(1, 2, 3) ~ "D",
      session == 4 ~ "B",
      session == 6 ~ "D",
      session %in% c(7, 8, 9) ~ "B",
      session == 10 ~ "D",
      session == 11 ~ "B",
      session %in% c(12, 13) ~ "D"
    ),
    mean_label = paste0("M=", round(mean_response_speed, 2))
  )

light_lookup_p10 <- setNames(p10_means_special$light_condition, p10_means_special$session)

max_y <- max(p10_data_special$mean_response_speed, na.rm = TRUE)
min_y <- min(p10_data_special$mean_response_speed, na.rm = TRUE)
y_range <- max_y - min_y
y_limit_upper <- ceiling((max_y + y_range * 0.1) * 10) / 10
y_limit_lower <- floor((min_y - y_range * 0.1) * 10) / 10

ggplot(p10_data_special, aes(x = x_position, y = mean_response_speed)) +
  geom_rect(data = p10_data_special %>% filter(light_condition == "B", trial == 4) %>% 
              group_by(session) %>% slice(1),
            xmin = 4, xmax = 5.5, ymin = -Inf, ymax = Inf,
            fill = "#FFE0B2", alpha = 0.5, inherit.aes = FALSE) +
  geom_line(color = "black", linewidth = 0.8) +
  geom_point(color = "black", size = 2.5) +
  geom_text(data = p10_means_special,
            aes(x = 0.5, y = Inf, label = mean_label),
            hjust = 0, vjust = 1.3, size = 3.5, inherit.aes = FALSE) +
  scale_x_continuous(
    breaks = c(1, 2, 3, 4, 5, 5.5),
    labels = c("1", "2", "3", "4", "5", "6"),
    limits = c(0.5, 6),
    expand = c(0, 0)
  ) +
  scale_y_continuous(limits = c(y_limit_lower, y_limit_upper), 
                     expand = expansion(mult = c(0, 0.05))) +
  facet_wrap(~ session, ncol = 4,
             labeller = as_labeller(function(x) {
               paste0("Session ", x, " (", light_lookup_p10[x], ")")
             })) +
  labs(
    title = "PVT Response Speed per Session (P10)",
    subtitle = "Session 5 excluded due to light automation malfunction",
    x = "Trial Number",
    y = "Mean Response Speed (1/s)"
  ) +
  theme_classic(base_size = 11) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 14, hjust = 0, margin = margin(b = 5)),
    plot.subtitle = element_text(size = 11, hjust = 0, margin = margin(b = 10), color = "grey40"),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(size = 11),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.5),
    panel.grid = element_blank(),
    strip.background = element_rect(fill = "white", color = "grey80", linewidth = 0.5),
    strip.text = element_text(size = 10, color = "black"),
    panel.background = element_rect(fill = "white", color = "grey80", linewidth = 0.5),
    plot.margin = margin(10, 10, 10, 10),
    legend.position = "none"
  )
```
## Calculate Grand Average Time-Course
```{r calculate-grand-average}
# Map light conditions to pvt_summary
pvt_light <- pvt_summary %>%
  filter(!(participant == "P10" & session == 5)) %>%  # Exclude P10 Session 5
  left_join(
    bind_rows(
      # P01-P09
      map_dfr(c("P01", "P02", "P03", "P04", "P05", "P06", "P07", "P09"), ~{
        pattern <- strsplit(light_patterns[[.x]], "")[[1]]
        tibble(
          participant = .x,
          session = seq_along(pattern),
          light_condition = pattern
        )
      }),
      # P10 special
      tibble(
        participant = "P10",
        session = c(1:4, 6:13),
        light_condition = c("D","D","D","B","D","B","B","B","D","B","D","D")
      )
    ),
    by = c("participant", "session")
  )

# Calculate per-participant means per trial and light condition
pvt_participant_summary <- pvt_light %>%
  filter(participant != "P07") %>%  # Exclude P07
  group_by(participant, trial, light_condition) %>%
  summarise(
    mean_response_speed = mean(mean_response_speed, na.rm = TRUE),
    .groups = "drop"
  )

# Calculate grand average across participants
pvt_grand_summary <- pvt_participant_summary %>%
  group_by(trial, light_condition) %>%
  summarise(
    mean_speed = mean(mean_response_speed, na.rm = TRUE),
    sd_speed = sd(mean_response_speed, na.rm = TRUE),
    n = n(),
    se_speed = sd_speed / sqrt(n),
    .groups = "drop"
  ) %>%
  mutate(
    light_condition = factor(light_condition, 
                            levels = c("D", "B"), 
                            labels = c("Dim", "Bright"))
  )

# Display summary
print(pvt_grand_summary)
```

## Time Course Plot
```{r plot-grand-average-timecourse, fig.width=10, fig.height=8}
# Add time mapping
time_mapping <- tibble(
  trial = 1:6,
  hours_before_hbt = c(5.5, 4.5, 3.5, 2.5, 1.5, 1.0)
)

pvt_grand_summary <- pvt_grand_summary %>%
  left_join(time_mapping, by = "trial")

# Y-axis limits
y_min_trial <- floor(min(pvt_grand_summary$mean_speed - pvt_grand_summary$se_speed, na.rm = TRUE) * 20) / 20 - 0.05
y_max_trial <- ceiling(max(pvt_grand_summary$mean_speed + pvt_grand_summary$se_speed, na.rm = TRUE) * 20) / 20 + 0.15

light_colors <- c("Dim" = "#4A90E2", "Bright" = "#F5A623")

pvt_grand_avg_plot <- ggplot(pvt_grand_summary, 
                             aes(x = hours_before_hbt, y = mean_speed, 
                                 color = light_condition, fill = light_condition, group = light_condition)) +
  annotate("rect", xmin = 2.5, xmax = 5.5, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  annotate("rect", xmin = 1.0, xmax = 2.5, ymin = -Inf, ymax = Inf, 
           fill = "#FFF9E6", alpha = 0.8) +
  annotate("rect", xmin = 0.5, xmax = 1.0, ymin = -Inf, ymax = Inf, 
           fill = "#F5F5F5", alpha = 0.4) +
  annotate("text", x = 4.0, y = y_max_trial - 0.02, label = "dim (10 lx)", 
           hjust = 0.5, vjust = 1, size = 3, color = "grey50", fontface = "italic") +
  annotate("text", x = 1.75, y = y_max_trial - 0.02, label = "intervention\n(10 or 1000 lx)", 
           hjust = 0.5, vjust = 1, size = 3.5, color = "#B8860B", fontface = "italic") +
  annotate("text", x = 0.75, y = y_max_trial - 0.02, label = "dim (10 lx)", 
           hjust = 0.5, vjust = 1, size = 3, color = "grey50", fontface = "italic") +
  geom_ribbon(aes(ymin = mean_speed - se_speed, 
                  ymax = mean_speed + se_speed),
              alpha = 0.1, color = NA, show.legend = FALSE) +
  geom_line(linewidth = 0.6) +
  geom_point(size = 5, shape = 16) +
  scale_color_manual(
    values = light_colors,
    labels = c("Dim" = "Dim (10 lx)", "Bright" = "Bright (1000 lx)")
  ) +
  scale_fill_manual(
    values = light_colors,
    labels = c("Dim" = "Dim (10 lx)", "Bright" = "Bright (1000 lx)")
  ) +
  scale_x_reverse(
    breaks = c(5.5, 4.5, 3.5, 2.5, 1.5, 1.0),
    labels = c("PVT 1\n(5.5h)", "PVT 2\n(4.5h)", "PVT 3\n(3.5h)", 
               "PVT 4\n(2.5h)", "PVT 5\n(1.5h)", "PVT 6\n(1.0h)")
  ) +
  scale_y_continuous(
    limits = c(y_min_trial, y_max_trial),
    expand = c(0, 0)
  ) +
  labs(
    title = "Psychomotor vigilance test (PVT) response speed time course",
    subtitle = "Grand average across all participants (P07 excluded, n=8)",
    x = "PVT trial number (hours before HBT)",
    y = "Mean response speed (1/s)",
    color = "Light condition",
    fill = "Light condition"
  ) +
  theme_classic(base_size = 14) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = 16, hjust = 0, face = "plain"),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10), face = "plain"),
    axis.text = element_text(color = "black", size = 13, face = "plain"),
    axis.title = element_text(size = 14, face = "plain"),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.5),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "right",
    legend.title = element_text(size = 13, face = "plain"),
    legend.text = element_text(size = 12, face = "plain"),
    legend.key.size = unit(0.5, "cm"),
    plot.margin = margin(15, 10, 10, 10)
  )

print(pvt_grand_avg_plot)
```

## Triplet Time-Course Analysis

### Data Preparation
```{r triplet-data-prep}
# Define light patterns for groups
light_g1 <- strsplit("DDDBBBDDBDB", "")[[1]]  # Group 1 (P01, P02, P03)
light_g2 <- strsplit("DBBBDDDBDB", "")[[1]]   # Group 2 (P04, P05) and Group 3 (P06, P09)
light_g4_orig <- strsplit("DDDBXDBBBDBDD", "")[[1]]  # Group 4 (P10)

# Create triplet dataframes for each group
df_triplets_g1 <- data.frame(
  session_1 = 1:(length(light_g1)-2),
  session_2 = 2:(length(light_g1)-1),
  session_3 = 3:length(light_g1),
  triplet = sapply(1:(length(light_g1)-2), function(i) 
    paste0(light_g1[i:(i+2)], collapse = ""))
)

df_triplets_g2 <- data.frame(
  session_1 = 1:(length(light_g2)-2),
  session_2 = 2:(length(light_g2)-1),
  session_3 = 3:length(light_g2),
  triplet = sapply(1:(length(light_g2)-2), function(i) 
    paste0(light_g2[i:(i+2)], collapse = ""))
)

# Participant list
participants_noP07 <- c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10")
triplets_all <- c("DDD", "DDB", "DBD", "DBB", "BDD", "BDB", "BBD", "BBB")

# Function to get session 3 for each triplet type
get_triplet_session3 <- function(p, triplet_type) {
  if(p %in% c("P01", "P02", "P03")) {
    df_triplets <- df_triplets_g1
  } else if(p %in% c("P04", "P05", "P06", "P09")) {
    df_triplets <- df_triplets_g2
  } else {
    df_triplets <- data.frame(
      session_1 = 1:(length(light_g4_orig)-2),
      session_2 = 2:(length(light_g4_orig)-1),
      session_3 = 3:length(light_g4_orig),
      triplet = sapply(1:(length(light_g4_orig)-2), function(i) 
        paste0(light_g4_orig[i:(i+2)], collapse = ""))
    ) %>%
      filter(!str_detect(triplet, "X")) %>%
      filter(session_1 != 5 & session_2 != 5 & session_3 != 5)
  }
  
  sessions_3 <- df_triplets %>%
    filter(triplet == triplet_type) %>%
    pull(session_3)
  
  return(sessions_3)
}

# Function to get all participant-session combinations for a triplet
get_triplet_sessions_overview <- function(triplet_type) {
  overview <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      for(s in sessions_3) {
        overview <- rbind(overview, data.frame(
          participant_id = p,
          session = s
        ))
      }
    }
  }
  
  return(overview %>% arrange(session, participant_id))
}

# Function to create plot for one triplet type
plot_triplet_timecourse_pvt <- function(triplet_type) {
  df_triplet_trials <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      # Use pvt_summary instead of pvt_data
      p_data <- pvt_summary %>%
        filter(participant == p, session %in% sessions_3) %>%
        select(session, trial, mean_response_speed) %>%
        group_by(trial) %>%
        summarise(
          mean_response_speed = mean(mean_response_speed, na.rm = TRUE),
          .groups = "drop"
        ) %>%
        mutate(participant_id = p)
      
      df_triplet_trials <- rbind(df_triplet_trials, p_data)
    }
  }
  
  triplet_summary <- df_triplet_trials %>%
    group_by(trial) %>%
    summarise(
      mean_speed = mean(mean_response_speed, na.rm = TRUE),
      sd_speed = sd(mean_response_speed, na.rm = TRUE),
      n = n(),
      .groups = "drop"
    ) %>%
    # Add custom x-position with reduced spacing between trial 5 and 6
    mutate(x_position = case_when(
      trial == 1 ~ 1.0,
      trial == 2 ~ 2.0,
      trial == 3 ~ 3.0,
      trial == 4 ~ 4.0,
      trial == 5 ~ 5.0,
      trial == 6 ~ 5.5  # Half distance (0.5 instead of 1.0)
    ))
  
  y_min <- floor(min(triplet_summary$mean_speed - triplet_summary$sd_speed, 
                     na.rm = TRUE) * 10) / 10 - 0.1
  y_max <- ceiling(max(triplet_summary$mean_speed + triplet_summary$sd_speed, 
                       na.rm = TRUE) * 10) / 10 + 0.1
  
  # Determine third letter and color
  third_letter <- substr(triplet_type, 3, 3)
  trial_4_5_color <- ifelse(third_letter == "D", "#F5F5F5", "#FFE0B2")
  trial_4_5_alpha <- ifelse(third_letter == "D", 0.4, 0.5)
  
  # Create title with bold third letter
  triplet_first_two <- substr(triplet_type, 1, 2)
  triplet_third <- substr(triplet_type, 3, 3)
  
  p <- ggplot(triplet_summary, aes(x = x_position, y = mean_speed)) +
    # Orange/grey background from trial 4 to trial 6
    annotate("rect", xmin = 4, xmax = 5.5, ymin = -Inf, ymax = Inf, 
             fill = trial_4_5_color, alpha = trial_4_5_alpha) +
    geom_ribbon(aes(ymin = mean_speed - sd_speed,
                    ymax = mean_speed + sd_speed),
                alpha = 0.15, fill = "grey40") +
    geom_line(linewidth = 0.8, color = "black") +
    geom_point(size = 2.0, shape = 16, color = "black") +
    scale_x_continuous(
      breaks = c(1, 2, 3, 4, 5, 5.5),
      labels = c("1", "2", "3", "4", "5", "6"),
      limits = c(0.5, 6),
      expand = c(0, 0)
    ) +
    scale_y_continuous(limits = c(y_min, y_max), expand = c(0, 0)) +
    labs(
      title = bquote("PVT response speed - Triplet " * 
                     .(triplet_first_two) * bold(.(triplet_third))),
      subtitle = paste0("Grand average third-day (n=", 
                       length(unique(df_triplet_trials$participant_id)), 
                       " participants; P07 excluded)"),
      x = "Trial Number",
      y = "Mean Response Speed (1/s)"
    ) +
    theme_classic(base_size = 14) +
    theme(
      aspect.ratio = 1,
      plot.title = element_text(size = 16, hjust = 0),
      plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
      axis.text = element_text(color = "black", size = 13),
      axis.title = element_text(size = 14),
      panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
      plot.margin = margin(15, 10, 10, 10)
    )
  
  return(p)
}
```

### DDD Triplet
```{r triplet-ddd}
knitr::kable(get_triplet_sessions_overview("DDD"),
             caption = "Sessions used for Triplet DDD")
plot_triplet_timecourse_pvt("DDD")
```

### DDB Triplet
```{r triplet-ddb}
knitr::kable(get_triplet_sessions_overview("DDB"),
             caption = "Sessions used for Triplet DDB")
plot_triplet_timecourse_pvt("DDB")
```

### DBD Triplet
```{r triplet-dbd}
knitr::kable(get_triplet_sessions_overview("DBD"),
             caption = "Sessions used for Triplet DBD")
plot_triplet_timecourse_pvt("DBD")
```

### DBB Triplet
```{r triplet-dbb}
knitr::kable(get_triplet_sessions_overview("DBB"),
             caption = "Sessions used for Triplet DBB")
plot_triplet_timecourse_pvt("DBB")
```

### BDD Triplet
```{r triplet-bdd}
knitr::kable(get_triplet_sessions_overview("BDD"),
             caption = "Sessions used for Triplet BDD")
plot_triplet_timecourse_pvt("BDD")
```

### BDB Triplet
```{r triplet-bdb}
knitr::kable(get_triplet_sessions_overview("BDB"),
             caption = "Sessions used for Triplet BDB")
plot_triplet_timecourse_pvt("BDB")
```

### BBD Triplet
```{r triplet-bbd}
knitr::kable(get_triplet_sessions_overview("BBD"),
             caption = "Sessions used for Triplet BBD")
plot_triplet_timecourse_pvt("BBD")
```

### BBB Triplet
```{r triplet-bbb}
knitr::kable(get_triplet_sessions_overview("BBB"),
             caption = "Sessions used for Triplet BBB")
plot_triplet_timecourse_pvt("BBB")
```

### Panel Plot: Dim-Ending Triplets (DDD, BDD, DBD, BBD)
```{r panel-dim-triplets, fig.width=10, fig.height=10}
triplets_D_panel <- c("DDD", "BDD", "DBD", "BBD")

df_all_D_triplets <- data.frame()

for(triplet_type in triplets_D_panel) {
  df_triplet_trials <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      p_data <- pvt_summary %>%
        filter(participant == p, session %in% sessions_3) %>%
        select(session, trial, mean_response_speed) %>%
        group_by(trial) %>%
        summarise(mean_response_speed = mean(mean_response_speed, na.rm = TRUE), .groups = "drop") %>%
        mutate(triplet = triplet_type)
      
      df_triplet_trials <- rbind(df_triplet_trials, p_data)
    }
  }
  
  df_all_D_triplets <- rbind(df_all_D_triplets, df_triplet_trials)
}

triplet_summary_D <- df_all_D_triplets %>%
  group_by(triplet, trial) %>%
  summarise(
    mean_speed = mean(mean_response_speed, na.rm = TRUE),
    sd_speed = sd(mean_response_speed, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(triplet = factor(triplet, levels = triplets_D_panel)) %>%
  # Add custom x-position with reduced spacing between trial 5 and 6
  mutate(x_position = case_when(
    trial == 1 ~ 1.0,
    trial == 2 ~ 2.0,
    trial == 3 ~ 3.0,
    trial == 4 ~ 4.0,
    trial == 5 ~ 5.0,
    trial == 6 ~ 5.5  # Half distance (0.5 instead of 1.0)
  ))

y_min_panel_D <- floor(min(triplet_summary_D$mean_speed - triplet_summary_D$sd_speed, na.rm = TRUE) * 10) / 10 - 0.1
y_max_panel_D <- ceiling(max(triplet_summary_D$mean_speed + triplet_summary_D$sd_speed, na.rm = TRUE) * 10) / 10 + 0.1

ggplot(triplet_summary_D, aes(x = x_position, y = mean_speed)) +
  annotate("rect", xmin = 4, xmax = 5.5, ymin = -Inf, ymax = Inf, fill = "#F5F5F5", alpha = 0.4) +
  geom_ribbon(aes(ymin = mean_speed - sd_speed, ymax = mean_speed + sd_speed), alpha = 0.15, fill = "grey40") +
  geom_line(linewidth = 0.8, color = "black") +
  geom_point(size = 2.0, shape = 16, color = "black") +
  scale_x_continuous(
    breaks = c(1, 2, 3, 4, 5, 5.5),
    labels = c("1", "2", "3", "4", "5", "6"),
    limits = c(0.5, 6),
    expand = c(0, 0)
  ) +
  scale_y_continuous(limits = c(y_min_panel_D, y_max_panel_D), expand = c(0, 0)) +
  facet_wrap(~ triplet, ncol = 2) +
  labs(
    title = "PVT Response Speed - Dim-ending triplets",
    subtitle = "Grand average third-day (P07 excluded)",
    x = "Trial Number",
    y = "Mean Response Speed (1/s)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 11),
    axis.title = element_text(size = 13),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    strip.background = element_rect(fill = "white", color = "grey80", linewidth = 0.5),
    strip.text = element_text(size = 12, color = "black", face = "bold"),
    plot.margin = margin(15, 10, 10, 10)
  )
```

### Panel Plot: Bright-Ending Triplets (DDB, BDB, DBB, BBB)
```{r panel-bright-triplets, fig.width=10, fig.height=10}
triplets_B_panel <- c("DDB", "BDB", "DBB", "BBB")

df_all_B_triplets <- data.frame()

for(triplet_type in triplets_B_panel) {
  df_triplet_trials <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      p_data <- pvt_summary %>%
        filter(participant == p, session %in% sessions_3) %>%
        select(session, trial, mean_response_speed) %>%
        group_by(trial) %>%
        summarise(mean_response_speed = mean(mean_response_speed, na.rm = TRUE), .groups = "drop") %>%
        mutate(triplet = triplet_type)
      
      df_triplet_trials <- rbind(df_triplet_trials, p_data)
    }
  }
  
  df_all_B_triplets <- rbind(df_all_B_triplets, df_triplet_trials)
}

triplet_summary_B <- df_all_B_triplets %>%
  group_by(triplet, trial) %>%
  summarise(
    mean_speed = mean(mean_response_speed, na.rm = TRUE),
    sd_speed = sd(mean_response_speed, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(triplet = factor(triplet, levels = triplets_B_panel)) %>%
  # Add custom x-position with reduced spacing between trial 5 and 6
  mutate(x_position = case_when(
    trial == 1 ~ 1.0,
    trial == 2 ~ 2.0,
    trial == 3 ~ 3.0,
    trial == 4 ~ 4.0,
    trial == 5 ~ 5.0,
    trial == 6 ~ 5.5  # Half distance (0.5 instead of 1.0)
  ))

y_min_panel_B <- floor(min(triplet_summary_B$mean_speed - triplet_summary_B$sd_speed, na.rm = TRUE) * 10) / 10 - 0.1
y_max_panel_B <- ceiling(max(triplet_summary_B$mean_speed + triplet_summary_B$sd_speed, na.rm = TRUE) * 10) / 10 + 0.1

ggplot(triplet_summary_B, aes(x = x_position, y = mean_speed)) +
  annotate("rect", xmin = 4, xmax = 5.5, ymin = -Inf, ymax = Inf, fill = "#FFE0B2", alpha = 0.5) +
  geom_ribbon(aes(ymin = mean_speed - sd_speed, ymax = mean_speed + sd_speed), alpha = 0.15, fill = "grey40") +
  geom_line(linewidth = 0.8, color = "black") +
  geom_point(size = 2.0, shape = 16, color = "black") +
  scale_x_continuous(
    breaks = c(1, 2, 3, 4, 5, 5.5),
    labels = c("1", "2", "3", "4", "5", "6"),
    limits = c(0.5, 6),
    expand = c(0, 0)
  ) +
  scale_y_continuous(limits = c(y_min_panel_B, y_max_panel_B), expand = c(0, 0)) +
  facet_wrap(~ triplet, ncol = 2) +
  labs(
    title = "PVT Response Speed - Bright-ending triplets",
    subtitle = "Grand average third-day (P07 excluded)",
    x = "Trial Number",
    y = "Mean Response Speed (1/s)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, hjust = 0),
    plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 11),
    axis.title = element_text(size = 13),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    strip.background = element_rect(fill = "white", color = "grey80", linewidth = 0.5),
    strip.text = element_text(size = 12, color = "black", face = "bold"),
    plot.margin = margin(15, 10, 10, 10)
  )
```

### Panel Plot: All Triplets with Grand Average
```{r panel-all-triplets-2rows, fig.width=14, fig.height=7}
# Prepare data for all triplets
triplets_all_panel <- c("DDD", "BDD", "DBD", "BBD", "DDB", "BDB", "DBB", "BBB")
df_all_triplets_unified <- data.frame()

for(triplet_type in triplets_all_panel) {
  df_triplet_trials <- data.frame()
  
  for(p in participants_noP07) {
    sessions_3 <- get_triplet_session3(p, triplet_type)
    
    if(length(sessions_3) > 0) {
      p_data <- pvt_summary %>%
        filter(participant == p, session %in% sessions_3) %>%
        select(participant, trial, mean_response_speed) %>%
        mutate(triplet = triplet_type)
      
      df_triplet_trials <- rbind(df_triplet_trials, p_data)
    }
  }
  
  df_all_triplets_unified <- rbind(df_all_triplets_unified, df_triplet_trials)
}

# Calculate summary per triplet
triplet_summary_unified <- df_all_triplets_unified %>%
  group_by(triplet, trial) %>%
  summarise(
    mean_speed = mean(mean_response_speed, na.rm = TRUE),
    sd_speed = sd(mean_response_speed, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    trial = as.numeric(trial),
    triplet = factor(triplet, levels = triplets_all_panel)
  ) %>%
  # Add custom x-position with reduced spacing between trial 5 and 6
  mutate(x_position = case_when(
    trial == 1 ~ 1.0,
    trial == 2 ~ 2.0,
    trial == 3 ~ 3.0,
    trial == 4 ~ 4.0,
    trial == 5 ~ 5.0,
    trial == 6 ~ 5.5  # Half distance
  ))

# Unified Y-axis limits
y_min_unified <- floor(min(triplet_summary_unified$mean_speed - triplet_summary_unified$sd_speed, 
                           na.rm = TRUE) * 10) / 10 - 0.1
y_max_unified <- ceiling(max(triplet_summary_unified$mean_speed + triplet_summary_unified$sd_speed, 
                             na.rm = TRUE) * 10) / 10 + 0.1

# Function to create triplet plot
create_triplet_plot_pvt <- function(triplet_name, show_x = FALSE, show_y = TRUE) {
  data_triplet <- triplet_summary_unified %>%
    filter(triplet == triplet_name)
  
  triplet_third <- substr(triplet_name, 3, 3)
  intervention_color <- ifelse(triplet_third == "D", "#F5F5F5", "#FFE0B2")
  intervention_alpha <- ifelse(triplet_third == "D", 0.4, 0.5)
  
  # Only third letter bold
  triplet_first_two <- substr(triplet_name, 1, 2)
  triplet_label <- paste0(triplet_first_two, "**", triplet_third, "**")
  
  ggplot(data_triplet, aes(x = x_position, y = mean_speed)) +
    annotate("rect", xmin = 4, xmax = 5.5, ymin = -Inf, ymax = Inf, 
             fill = intervention_color, alpha = intervention_alpha) +
    geom_ribbon(aes(ymin = mean_speed - sd_speed, 
                    ymax = mean_speed + sd_speed),
                alpha = 0.15, fill = "grey40") +
    geom_line(linewidth = 0.8, color = "black") +
    geom_point(size = 2.5, shape = 16, color = "black") +
    scale_x_continuous(
      breaks = c(1, 2, 3, 4, 5, 5.5),
      labels = if(show_x) c("1", "2", "3", "4", "5", "6") else c("1", "2", "3", "4", "5", "6"),
      limits = c(0.5, 6),
      expand = c(0, 0)
    ) +
    scale_y_continuous(limits = c(y_min_unified, y_max_unified), expand = c(0, 0)) +
    labs(
      title = triplet_label,
      x = NULL,
      y = if(show_y) "Mean Response Speed\n± SD (1/s)" else NULL
    ) +
    theme_classic(base_size = 11) +
    theme(
      plot.title = element_markdown(size = 11, hjust = 0.5, margin = margin(b = 5)),
      axis.text = element_text(color = "black", size = 9),
      axis.text.x = element_text(size = 8),
      axis.title = element_text(size = 10),
      axis.title.y = element_text(margin = margin(r = 5)),
      panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
      plot.margin = margin(5, 5, 5, 5)
    )
}

# Create D-ending triplet plots (Row 1 - no x-axis labels)
p_DDD_pvt <- create_triplet_plot_pvt("DDD", show_x = FALSE, show_y = TRUE)
p_BDD_pvt <- create_triplet_plot_pvt("BDD", show_x = FALSE, show_y = FALSE)
p_DBD_pvt <- create_triplet_plot_pvt("DBD", show_x = FALSE, show_y = FALSE)
p_BBD_pvt <- create_triplet_plot_pvt("BBD", show_x = FALSE, show_y = FALSE)

# Create B-ending triplet plots (Row 2 - with x-axis labels)
p_DDB_pvt <- create_triplet_plot_pvt("DDB", show_x = TRUE, show_y = TRUE)
p_BDB_pvt <- create_triplet_plot_pvt("BDB", show_x = TRUE, show_y = FALSE)
p_DBB_pvt <- create_triplet_plot_pvt("DBB", show_x = TRUE, show_y = FALSE)
p_BBB_pvt <- create_triplet_plot_pvt("BBB", show_x = TRUE, show_y = FALSE)

# Grand Average plot (same size as individual panels)
# Add x_position to pvt_grand_summary
pvt_grand_summary_adjusted <- pvt_grand_summary %>%
  mutate(x_position = case_when(
    trial == 1 ~ 1.0,
    trial == 2 ~ 2.0,
    trial == 3 ~ 3.0,
    trial == 4 ~ 4.0,
    trial == 5 ~ 5.0,
    trial == 6 ~ 5.5
  ))

p_grand_pvt <- ggplot(pvt_grand_summary_adjusted, 
       aes(x = x_position, y = mean_speed, 
           color = light_condition, fill = light_condition, group = light_condition)) +
  annotate("rect", xmin = 4, xmax = 5.5, ymin = -Inf, ymax = Inf, 
           fill = "#FFE0B2", alpha = 0.5) +
  geom_ribbon(aes(ymin = mean_speed - se_speed, 
                  ymax = mean_speed + se_speed), 
              alpha = 0.1, color = NA, show.legend = FALSE) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.5, shape = 16) +
  scale_color_manual(
    values = c("Bright" = "#F5A623", "Dim" = "#4A90E2"),
    labels = c("Bright" = "Bright (1000 lx)", "Dim" = "Dim (10 lx)")
  ) +
  scale_fill_manual(
    values = c("Bright" = "#F5A623", "Dim" = "#4A90E2"),
    labels = c("Bright" = "Bright (1000 lx)", "Dim" = "Dim (10 lx)")
  ) +
  scale_x_continuous(
    breaks = c(1, 2, 3, 4, 5, 5.5),
    labels = c("1", "2", "3", "4", "5", "6"),
    limits = c(0.5, 6),
    expand = c(0, 0)
  ) +
  scale_y_continuous(limits = c(y_min_unified, y_max_unified), expand = c(0, 0)) +
  labs(
    title = "Grand Average",
    x = NULL,
    y = NULL,
    color = "Light condition"
  ) +
  guides(
    color = guide_legend(override.aes = list(shape = 16, size = 2.5)),
    fill = "none"
  ) +
  theme_classic(base_size = 11) +
  theme(
    plot.title = element_text(size = 11, hjust = 0.5, face = "bold", margin = margin(b = 5)),
    axis.text = element_text(color = "black", size = 9),
    axis.text.x = element_text(size = 9),
    axis.title = element_text(size = 10),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    legend.position = "bottom",
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    plot.margin = margin(5, 5, 5, 5)
  )

# Empty spacer
p_spacer <- ggplot() + theme_void()

# Combine in layout:
# Row 1: DDD, BDD, DBD, BBD, Grand Average
# Row 2: DDB, BDB, DBB, BBB, Spacer
(p_DDD_pvt | p_BDD_pvt | p_DBD_pvt | p_BBD_pvt | p_grand_pvt) /
(p_DDB_pvt | p_BDB_pvt | p_DBB_pvt | p_BBB_pvt | p_spacer) +
  plot_annotation(
    title = "PVT Response Speed - Grand average and triplet patterns",
    subtitle = "Grand average and third-day triplet patterns (P07 excluded)",
    caption = "Trial Number",
    theme = theme(
      plot.title = element_text(size = 15, hjust = 0, face = "plain"),
      plot.subtitle = element_text(size = 12, hjust = 0, margin = margin(b = 8)),
      plot.caption = element_text(size = 11, hjust = 0.5, margin = margin(t = 8))
    )
  )
```

### Descriptive Statistics: Triplet Time-Course Values
```{r triplet-stats-table}
if(!exists("triplet_summary_unified")) {
  triplets_all_panel <- c("DDD", "BDD", "DBD", "BBD", "DDB", "BDB", "DBB", "BBB")
  df_all_triplets_unified <- data.frame()
  
  for(triplet_type in triplets_all_panel) {
    df_triplet_trials <- data.frame()
    
    for(p in participants_noP07) {
      sessions_3 <- get_triplet_session3(p, triplet_type)
      
      if(length(sessions_3) > 0) {
        p_data <- pvt_summary %>%
          filter(participant == p, session %in% sessions_3) %>%
          select(participant, trial, mean_response_speed) %>%
          mutate(triplet = triplet_type)
        
        df_triplet_trials <- rbind(df_triplet_trials, p_data)
      }
    }
    
    df_all_triplets_unified <- rbind(df_all_triplets_unified, df_triplet_trials)
  }
  
  triplet_summary_unified <- df_all_triplets_unified %>%
    group_by(triplet, trial) %>%
    summarise(
      mean_speed = mean(mean_response_speed, na.rm = TRUE),
      sd_speed = sd(mean_response_speed, na.rm = TRUE),
      n = n(),
      .groups = "drop"
    ) %>%
    mutate(
      trial = as.numeric(trial),
      triplet = factor(triplet, levels = triplets_all_panel)
    )
}

# Create table
triplet_timecourse_table <- triplet_summary_unified %>%
  mutate(mean_sd = sprintf("%.2f ± %.2f", mean_speed, sd_speed)) %>%
  select(triplet, trial, mean_sd) %>%
  pivot_wider(names_from = trial, values_from = mean_sd, names_prefix = "T")

knitr::kable(
  triplet_timecourse_table,
  caption = "Triplet time-course: Mean ± SD response speed (1/s) per trial",
  align = c("l", rep("r", 6))
) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  kableExtra::row_spec(0, bold = TRUE)
```

---

# Statistical Analysis: Linear Mixed Models

## Data Preparation for LMM

### Define Study Parameters
```{r lmm-setup}
participant_list <- c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10")

light_patterns <- list(
  "P01" = "DDDBBBDDBDB",
  "P02" = "DDDBBBDDBDB",
  "P03" = "DDDBBBDDBDB",
  "P04" = "DBBBDDDBDB",
  "P05" = "DBBBDDDBDB",
  "P06" = "DBBBDDDBDB",
  "P09" = "DBBBDDDBDB",
  "P10" = "DDDBXDBBBDBDD"
)
```

### Combine PVT Data with Light Information
```{r lmm-data-prep}
pvt_lmm <- map_dfr(participant_list, ~{
  df <- pvt_session_means_t45 %>% 
    filter(participant == .x) %>%
    rename(
      participant_id = participant,
      exp_session = session
    )
  
  pattern <- light_patterns[[.x]]
  pattern_vec <- str_split(pattern, "")[[1]]
  
  if (.x == "P10") {
    valid_sessions <- which(pattern_vec != "X")
    light_df <- tibble(
      exp_session = valid_sessions,
      Light_Lag0 = pattern_vec[valid_sessions]
    )
  } else {
    light_df <- tibble(
      exp_session = seq_along(pattern_vec),
      Light_Lag0 = pattern_vec
    )
  }
  
  df <- df %>% 
    left_join(light_df, by = "exp_session")
  
  return(df)
})
```

### Build Lagged Light Variables
```{r build-lags}
pvt_lmm <- pvt_lmm %>%
  group_by(participant_id) %>%
  arrange(exp_session) %>%
  mutate(
    Light_Lag1 = lag(Light_Lag0, 1),
    Light_Lag2 = lag(Light_Lag0, 2)
  ) %>%
  ungroup() %>%
  filter(!is.na(Light_Lag0) & !is.na(Light_Lag1) & !is.na(Light_Lag2)) %>%
  filter(!(participant_id == "P10" & exp_session %in% c(6, 7))) %>%
  mutate(
    Light_Lag0_num = ifelse(Light_Lag0 == "D", 0, 1),
    Light_Lag1_num = ifelse(Light_Lag1 == "D", 0, 1),
    Light_Lag2_num = ifelse(Light_Lag2 == "D", 0, 1)
  )
```

### Add Demographics
```{r add-demographics}
participant_demographics <- tibble(
  participant_id = c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10"),
  sex = c("female", "male", "female", "female", "male", "female", "female", "male"),
  age = c(24, 21, 24, 23, 31, 23, 28, 32)
)

pvt_lmm <- pvt_lmm %>%
  left_join(participant_demographics, by = "participant_id")
```

### Final Selection
```{r final-selection}
pvt_lmm <- pvt_lmm %>%
  dplyr::select(
    participant_id, exp_session, mean_response_speed,
    Light_Lag0, Light_Lag1, Light_Lag2,
    Light_Lag0_num, Light_Lag1_num, Light_Lag2_num,
    sex, age
  )
```

## Dataset Overview
```{r lmm-overview}
str(pvt_lmm)

light_summary <- pvt_lmm %>%
  summarise(
    Lag0_Dim = sum(Light_Lag0_num == 0, na.rm = TRUE),
    Lag0_Bright = sum(Light_Lag0_num == 1, na.rm = TRUE),
    Lag1_Dim = sum(Light_Lag1_num == 0, na.rm = TRUE),
    Lag1_Bright = sum(Light_Lag1_num == 1, na.rm = TRUE),
    Lag2_Dim = sum(Light_Lag2_num == 0, na.rm = TRUE),
    Lag2_Bright = sum(Light_Lag2_num == 1, na.rm = TRUE)
  )

print(light_summary)

table(pvt_lmm$exp_session)
table(pvt_lmm$participant_id)
```

## Dataset for Statistical Analysis (pvt_lmm)
```{r pvt-lmm-dataset-table}
#| label: pvt-lmm-dataset-table
#| echo: true

DT::datatable(
  pvt_lmm,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    order = list(list(0, 'asc'), list(1, 'asc'))
  ),
  caption = paste0("Complete dataset for statistical analysis (n=", nrow(pvt_lmm), ", P07 excluded)"),
  rownames = FALSE
) %>%
  DT::formatRound(columns = 'mean_response_speed', digits = 2) %>%
  DT::formatStyle(
    'Light_Lag0',
    backgroundColor = DT::styleEqual(c('D', 'B'), c('#e3f2fd', '#fff3e0'))
  )
```

## Distribution Analysis
```{r distribution-analysis}
summary(pvt_lmm$mean_response_speed)

data.frame(
  N = nrow(pvt_lmm),
  N_participants = length(unique(pvt_lmm$participant_id)),
  Mean = round(mean(pvt_lmm$mean_response_speed, na.rm = TRUE), 3),
  SD = round(sd(pvt_lmm$mean_response_speed, na.rm = TRUE), 3),
  Min = round(min(pvt_lmm$mean_response_speed, na.rm = TRUE), 3),
  Max = round(max(pvt_lmm$mean_response_speed, na.rm = TRUE), 3)
)
```
```{r plot-distribution, fig.width=8, fig.height=6}
ggplot(pvt_lmm, aes(x = mean_response_speed)) +
  geom_histogram(aes(y = after_stat(density)), bins = 15, 
                 fill = "#4A90E2", alpha = 0.7, color = "black") +
  geom_density(color = "red", linewidth = 1) +
  geom_vline(xintercept = mean(pvt_lmm$mean_response_speed, na.rm = TRUE), 
             linetype = "dashed", color = "red") +
  labs(title = "Distribution of PVT Mean Response Speed (Trials 4&5)",
       subtitle = paste0("P07 excluded, n=", nrow(pvt_lmm), " observations"),
       x = "Mean response speed (1/s)", 
       y = "Density") +
  theme_classic(base_size = 14)
```

## PVT Trajectories Across Sessions
```{r plot-trajectories, fig.width=10, fig.height=8}
session_counts_pvt <- pvt_lmm %>%
  group_by(exp_session) %>%
  summarise(n = n(), .groups = "drop")

ggplot(pvt_lmm, 
       aes(x = exp_session, y = mean_response_speed, 
           color = participant_id, group = participant_id)) +
  geom_line(alpha = 0.6, linewidth = 0.8) +
  geom_point(size = 2.5, alpha = 0.7) +
  geom_smooth(aes(group = 1), method = "loess", span = 0.75,
              color = "black", linetype = "dashed", linewidth = 1.2,
              se = TRUE, fill = "grey50", alpha = 0.3) +
  geom_text(data = session_counts_pvt %>% filter(n < 8),
            aes(x = exp_session, y = max(pvt_lmm$mean_response_speed, na.rm = TRUE) * 0.98, 
                label = paste0("n=", n)),
            color = "black", size = 3, inherit.aes = FALSE) +
  geom_vline(xintercept = 10.5, linetype = "dotted", color = "grey50", linewidth = 0.8) +
  scale_x_continuous(
    breaks = 3:13,
    limits = c(2.5, 13.5),
    expand = c(0.01, 0)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.05))
  ) +
  scale_color_manual(
    values = c("P01" = "#E41A1C", "P02" = "#377EB8", "P03" = "#4DAF4A",
               "P04" = "#984EA3", "P05" = "#FF7F00", "P06" = "#FFFF33",
               "P09" = "#A65628", "P10" = "#F781BF"),
    name = "Participant"
  ) +
  labs(
    title = "PVT response speed across experimental sessions",
    subtitle = paste0("Individual participant trajectories with LOESS smooth (Trials 4&5, n=", 
                      nrow(pvt_lmm), ", P07 excluded)"),
    x = "Experimental Session",
    y = "Mean response speed (1/s)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(size = 16, face = "plain"),
    plot.subtitle = element_text(size = 12, margin = margin(b = 10)),
    axis.text = element_text(color = "black", size = 12),
    axis.title = element_text(size = 13),
    legend.position = "right",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3),
    plot.margin = margin(15, 15, 10, 10)
  )
```

---

# Model Building and Comparison

## Model m0: Baseline (Intercept Only)
```{r model-m0}
m0_pvt <- lmer(
  mean_response_speed ~ 1 + (1 | participant_id),
  data = pvt_lmm,
  REML = FALSE
)

summary(m0_pvt)

vc_m0 <- as.data.frame(VarCorr(m0_pvt))
icc_m0 <- vc_m0$vcov[1] / (vc_m0$vcov[1] + vc_m0$vcov[2])
```

## Model m1a: Add Current Day Light (Lag 0)
```{r model-m1a}
m1a_pvt <- lmer(
  mean_response_speed ~ 
    Light_Lag0_num +
    (1 | participant_id),
  data = pvt_lmm,
  REML = FALSE
)

summary(m1a_pvt)
```

## Model m1b: Add Previous Day Light (Lag 1)
```{r model-m1b}
m1b_pvt <- lmer(
  mean_response_speed ~ 
    Light_Lag0_num + Light_Lag1_num +
    (1 | participant_id),
  data = pvt_lmm,
  REML = FALSE
)

summary(m1b_pvt)
```

## Model m1c: Add Two Days Prior (Lag 2)
```{r model-m1c}
m1c_pvt <- lmer(
  mean_response_speed ~ 
    Light_Lag0_num + Light_Lag1_num + Light_Lag2_num +
    (1 | participant_id),
  data = pvt_lmm,
  REML = FALSE
)

summary(m1c_pvt)
```

## Model m2: Add Linear Session Term
```{r model-m2}
m2_pvt <- lmer(
  mean_response_speed ~ 
    Light_Lag0_num + Light_Lag1_num +
    exp_session +
    (1 | participant_id),
  data = pvt_lmm,
  REML = FALSE
)

summary(m2_pvt)
```

## Model m3: Add Quadratic Session Term
```{r model-m3}
m3_pvt <- lmer(
  mean_response_speed ~ 
    Light_Lag0_num + Light_Lag1_num +
    exp_session + I(exp_session^2) +
    (1 | participant_id),
  data = pvt_lmm,
  REML = FALSE
)

summary(m3_pvt)
```

## Likelihood Ratio Tests
```{r lrt-tests}
anova(m0_pvt, m1a_pvt)
anova(m1a_pvt, m1b_pvt)
anova(m1b_pvt, m1c_pvt)
anova(m1b_pvt, m2_pvt)
anova(m1b_pvt, m3_pvt)
anova(m2_pvt, m3_pvt)
```
## Model Selection Summary

Based on the likelihood ratio tests, the following conclusions can be drawn about the hierarchical model building process:

Based on LRT tests, model m1b (Lag 0 + Lag 1 + random intercept) is chosen as the final model for primary analysis

Key findings:

- Lag 0 (Current day light): Adding Lag 0 significantly improved fit over random intercept model only (m0 vs. m1a: χ²=8.19, p<0.01)

- Lag 1 (Previous day): Adding Lag 1 additionally significantly improve fit over model m1a (m1a vs. m1b: χ²=6.64, p<0.05)

- Lag 2 (Two days prior): Adding Lag 2 additionally did not significantly improve model fit over model m1b  (m1b vs. m1c: χ²=1.30, p=0.253)

- Linear session effect: Adding a linear session term did not significantly improve model fit over model m1b (m1b vs. m2: χ²=0.56, p=0.454)

- Quadratic session effect: Adding a quadratic session term did not significantly improve model fit over model m2 (m2 vs. m3: χ²=0.53, p=0.466)

**Final model selection**: Based on the LRT results, **model m1b** was selected as the final primary model, using Lag 0 and Lag 1 as fixed effects as well as participant ID as the random intercept.

This model includes:

- Random intercepts for participants

- Fixed effects for: Lag 0 (current evening light exposure), Lag 1 (previous evenings light exposure)

---

# Primary Model Diagnostics

## Check 1: Multicollinearity (VIF)
```{r check-vif, fig.width=8, fig.height=6}
vif_m1_pvt <- vif(m1b_pvt)

barplot(vif_m1_pvt, 
        main = "Variance Inflation Factors (PVT Primary Model)",
        ylab = "VIF",
        col = ifelse(vif_m1_pvt > 5, "red", 
                     ifelse(vif_m1_pvt > 2, "orange", "lightblue")),
        las = 2,
        ylim = c(0, max(vif_m1_pvt) * 1.2))
abline(h = 5, col = "red", lty = 2, lwd = 2)
abline(h = 10, col = "darkred", lty = 2, lwd = 2)
legend("topright", 
       legend = c("VIF < 2 (excellent)", "VIF 2-5 (acceptable)", "VIF > 5 (concern)"),
       fill = c("lightblue", "orange", "red"),
       cex = 0.8)

print(round(vif_m1_pvt, 2))
```

Interpretation Check 1 VIF: VIF values of 1.00 for both predictors indicate no multicollinearity 
(all values < 2, excellent).

## Check 2: Residual Plots
```{r check-residuals, fig.width=10, fig.height=10}
# First calculate Cook's D with influence.ME 
infl_m1_pvt_check2 <- influence(m1b_pvt, obs = TRUE)
cooks_d_pvt_check2 <- cooks.distance(infl_m1_pvt_check2)
threshold_pvt_check2 <- 4/nrow(pvt_lmm)

par(mfrow = c(2, 2))

plot(fitted(m1b_pvt), resid(m1b_pvt), 
     xlab = "Fitted", ylab = "Residuals",
     main = "Residuals vs Fitted", pch = 16, col = rgb(0, 0, 0, 0.5))
abline(h = 0, col = "red", lty = 2, lwd = 2)
lines(lowess(fitted(m1b_pvt), resid(m1b_pvt)), col = "blue", lwd = 2)

qqnorm(resid(m1b_pvt), main = "Q-Q Plot", pch = 16, col = rgb(0, 0, 0, 0.5))
qqline(resid(m1b_pvt), col = "red", lwd = 2)

plot(fitted(m1b_pvt), sqrt(abs(resid(m1b_pvt))), 
     xlab = "Fitted", ylab = "√|Residuals|",
     main = "Scale-Location", pch = 16, col = rgb(0, 0, 0, 0.5))
lines(lowess(fitted(m1b_pvt), sqrt(abs(resid(m1b_pvt)))), col = "blue", lwd = 2)

plot(cooks_d_pvt_check2, type = "h", main = "Cook's Distance",
     ylab = "Cook's D", lwd = 2,
     col = ifelse(cooks_d_pvt_check2 > threshold_pvt_check2, "red", "black"))
abline(h = threshold_pvt_check2, col = "red", lty = 2, lwd = 2)

par(mfrow = c(1, 1))
```

Interpretation Check 2 Residual Plots: 

- Residual vs. Fitted plot: The relationship between the predictors and dependent variable seems to be linear. 

- QQ Plot: Minor deviations are observed in the central region around zero, however normality assumption seems to be satisfied reasonably.

- Scale Location Plot: Visual inspection suggests the homoscedasticity assumption is largely met, though there is a slight suggestion of decreasing variance with higher fitted values. This deviation appears minor and unlikely to substantially affect inference.

- Cook´s Distance Plot: 2 influential observations exceeding the threshold of 4/n are shown. A robust model removing those observations using the model m1b structure will be tested as part of the sensitivity analyses.


## Check 3: Normality (Shapiro-Wilk)
```{r check-normality}
shapiro.test(resid(m1b_pvt))
```

Interpretation Check 3 Normality: Shapiro-Wilk test shows no evidence for significant deviation of normality assumption (p=0.7637)

## Check 4: Heteroskedasticity (Levene's Test)
```{r check-heteroskedasticity, fig.width=10, fig.height=5}
par(mfrow = c(1, 2))

plot(fitted(m1b_pvt), resid(m1b_pvt), 
     xlab = "Fitted", ylab = "Residuals",
     main = "Residuals vs Fitted", pch = 16, col = rgb(0, 0, 0, 0.5))
abline(h = 0, col = "red", lty = 2, lwd = 2)
lines(lowess(fitted(m1b_pvt), resid(m1b_pvt)), col = "blue", lwd = 2)

plot(fitted(m1b_pvt), sqrt(abs(resid(m1b_pvt))), 
     xlab = "Fitted", ylab = "√|Residuals|",
     main = "Scale-Location", pch = 16, col = rgb(0, 0, 0, 0.5))
lines(lowess(fitted(m1b_pvt), sqrt(abs(resid(m1b_pvt)))), col = "blue", lwd = 2)
abline(h = mean(sqrt(abs(resid(m1b_pvt)))), col = "red", lty = 2, lwd = 2)

par(mfrow = c(1, 1))

fitted_groups <- cut(fitted(m1b_pvt), 
                     breaks = quantile(fitted(m1b_pvt), probs = c(0, 0.5, 1)),
                     include.lowest = TRUE)
leveneTest(resid(m1b_pvt) ~ fitted_groups)
```

Interpretation Check 4 Heteroskedasticity: see interpretation Check 2

## Check 5: Random Effects Normality
```{r check-random-effects, fig.width=10, fig.height=5}
ranef_m1_pvt <- ranef(m1b_pvt)$participant_id

par(mfrow = c(1, 2))

hist(ranef_m1_pvt$`(Intercept)`, main = "Random Intercepts",
     xlab = "Random Intercept", col = "lightblue", border = "black", breaks = 8)

qqnorm(ranef_m1_pvt$`(Intercept)`, main = "Q-Q Plot: Random Intercepts",
       pch = 16, col = rgb(0, 0, 0, 0.6))
qqline(ranef_m1_pvt$`(Intercept)`, col = "red", lwd = 2)

par(mfrow = c(1, 1))

shapiro.test(ranef_m1_pvt$`(Intercept)`)
```

Interpretation Check 5 Random Effects Normality: both plots show visually no substantial deviation from normality assumption for random effects, however the small sample size of n=8 limits the interpretability of these diagnostic plots. One participant shows a lower random intercept in the QQ plot, but overall the normality assumption for random effects appears reasonably satisfied given the limited sample size.

## Check 6: Influential Observations
```{r check-influential, fig.width=8, fig.height=6}
infl_m1_pvt <- influence(m1b_pvt, obs = TRUE)
cooks_d_pvt <- cooks.distance(infl_m1_pvt)
threshold_pvt <- 4/nrow(pvt_lmm)

plot(cooks_d_pvt, type = "h", main = "Cook's Distance - PVT Model",
     ylab = "Cook's D", lwd = 2,
     col = ifelse(cooks_d_pvt > threshold_pvt, "red", "black"))
abline(h = threshold_pvt, lty = 2, col = "red")

influential_idx_pvt <- which(cooks_d_pvt > threshold_pvt)

if (length(influential_idx_pvt) > 0) {
  print(pvt_lmm[influential_idx_pvt, c("participant_id", "exp_session", "mean_response_speed")])
}
```

Interpretation Check 6 Influential Observations: see Check 2 Interpretation

## Check 7: Temporal Autocorrelation
```{r check-autocorrelation, fig.width=10, fig.height=5}
par(mfrow = c(1, 2))

acf(resid(m1b_pvt), main = "ACF: Residuals", lag.max = 13)
pacf(resid(m1b_pvt), main = "PACF: Residuals", lag.max = 13)

par(mfrow = c(1, 1))
```

Interpretation Check 7 Temporal Autocorrelation: Both, the ACF and PACF plots show no indication of temporal autocorrelation, indicting that the indpendence assumption of residuals across time is adequately met.

## Diagnostic Summary
```{r diagnostic-summary}
shapiro_resid_pvt <- shapiro.test(resid(m1b_pvt))
levene_result_pvt <- leveneTest(resid(m1b_pvt) ~ fitted_groups)
shapiro_re_pvt <- shapiro.test(ranef_m1_pvt$`(Intercept)`)
print(shapiro_resid_pvt)
print(levene_result_pvt)
print(shapiro_re_pvt)
```

Interpretation Diagnostic Summary: 

- Shapiro-Wilk test for normality of residuals indicated no evidence of a violation of the normality assumption for the residuals (W=0.98791, p=0.7637)

- Levene´s test revealed no evidence of heterogeneity of variance across groups, indicating that the assumption of homoscedasticity was met (F(1, 65)= 1.7075, p=0.1959)

- Shapiro-Wilk test for normality of random intercepts showed no evidence of a deviation from normality, though the limited statistical power of this test due to small number of random intercepts should be acknowledged (W = 0.94941, p= 0.7053)

# Sensitivity Analyses
## Sensitivity 1: Log-Transform Sensitivity Check
```{r log-transform}
pvt_lmm <- pvt_lmm %>%
  mutate(log_speed = log(mean_response_speed))

m1_pvt_log <- lmer(
  log_speed ~ Light_Lag0_num + Light_Lag1_num +
    (1 | participant_id),
  data = pvt_lmm,
  REML = FALSE
)

AIC(m1b_pvt, m1_pvt_log)

plot(fitted(m1_pvt_log), sqrt(abs(resid(m1_pvt_log))),
     main = "Scale-Location (Log Model)",
     xlab = "Fitted", ylab = "√|Residuals|",
     pch = 16, col = rgb(0, 0, 0, 0.5))
lines(lowess(fitted(m1_pvt_log), sqrt(abs(resid(m1_pvt_log)))), col = "blue", lwd = 2)

summary(m1_pvt_log)
```

Interpretation Sensitivity 1: Log transformation of the dependent variable (mean_response_speed) did not change the significance / strength of results, showing robustness of the model to non-linear scaling. Lag 0 (β= 0.05101, p=0.00358), Lag 1 (β=0.03846, p=0.02572) remained statistically significant. Exponentiation of coefficients indicated a 5.2% multiplicative improvement in mean response speed from current bright evening light exposure, and a 3.9% multiplicative improvement in mean response speed when the previous evenings was bright compared to dim. Visual inspection of the Scale-Location Plot for the log transformed model showed worse fit. Therefore, Senstivity 3 will be used to adress the slight visual heteroscedasticity observed in the diagnostic checks. 

## Sensitivity 2: Excluding Influential Observations

**Context**: Two observations (both from P05, sessions 3 and 7) showed Cook's D values exceeding the threshold (4/n = 0.06). The primary model was fitted using a dataset, which excludes these observations to assess robustness.
```{r sensitivity-influential}
# Identify influential observations (already calculated in Check 6)
influential_idx_pvt <- which(cooks_d_pvt > threshold_pvt)

print(pvt_lmm[influential_idx_pvt, c("participant_id", "exp_session", "mean_response_speed")])

# Create robust dataset
pvt_lmm_robust <- pvt_lmm[-influential_idx_pvt, ]

# Fit robust model with same structure as m1b
m1b_pvt_robust <- lmer(
  mean_response_speed ~ 
    Light_Lag0_num + Light_Lag1_num + 
    (1 | participant_id),
  data = pvt_lmm_robust,
  REML = FALSE
)

summary(m1b_pvt_robust)
```

Interpretation Sensitivity 2: After removal of two influential observations, Lag 0 and Lag 1 remained statistically significant:Lag 0 (β=0.16983, p=0.00678), Lag 1 (β=0.13125, p=0.03406). Compared to the primary model m1b using the entire dataset, the strength of the main effects decreased slightly Lag 0 (Δ=0.04193) and Lag 1 (Δ=0.04203) remained stable (primary model m1b: Lag 0 (β=0.21176, p=0.00197), Lag 1 (β=0.17328, p=0.01029)). The persistence of significant effects after excluding influential observations demonstrates that the primary findings are not driven by outliers and remain robust. 

## Sensitivity 3: Robust Standard Errors

**Context**: To address the mild heteroskedasticity observed in diagnostic plots, we re-estimated the primary model (m1b) using cluster-robust standard errors (CR2 adjustment), which provide valid inference even under heteroskedasticity.
```{r sensitivity-robust-se}
# Calculate robust standard errors (CR2 method)
robust_results <- coef_test(m1b_pvt, vcov = "CR2")

# Display robust results
print(robust_results)

# Extract for comparison table
robust_coef <- robust_results[c("Light_Lag0_num", "Light_Lag1_num"), ]

# Create comparison table
comparison_se <- data.frame(
  Effect = c("Light_Lag0_num", "Light_Lag1_num"),
  Original_Est = round(summary(m1b_pvt)$coefficients[c("Light_Lag0_num", "Light_Lag1_num"), "Estimate"], 3),
  Original_SE = round(summary(m1b_pvt)$coefficients[c("Light_Lag0_num", "Light_Lag1_num"), "Std. Error"], 3),
  Original_p = format.pval(summary(m1b_pvt)$coefficients[c("Light_Lag0_num", "Light_Lag1_num"), "Pr(>|t|)"], 
                          digits = 3, eps = 0.001),
  Robust_Est = round(robust_coef$beta, 3),
  Robust_SE = round(robust_coef$SE, 3),
  Robust_p = format.pval(robust_coef$p_Satt, digits = 3, eps = 0.001)
)

kable(comparison_se,
      row.names = FALSE,
      col.names = c("Effect", "Estimate", "SE", "p-value", 
                    "Estimate", "SE", "p-value"),
      caption = "Comparison: Original vs. Robust Standard Errors",
      align = c('l', 'r', 'r', 'r', 'r', 'r', 'r')) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  add_header_above(c(" " = 1, "Original Model" = 3, "Robust SEs" = 3))
```

Interpretation Sensitivity 3: To address slight visual heteroscedasticity identified during the diagnostic checks of the primary model m1b, standard errors were re-estimated using the cluster-robust variance estimation (CR2 method), which is a good fit for small sample sizes. Both main effects remained statistically significant with CR2-adjusted standard errors: Lag 0 (SE=0.059, p=0.00874) and Lag 1 (SE=0.058, p=0.02103)- Compared to the original model, standard errors decreased slightly and p-values increased moderately: Lag 0 (ΔSE=-0.006, Δp=+0.00677) and Lag 1 (ΔSE=-0.007, Δp=+0.01074). The persistence of significant effects with robust standard errors suggests that the mild visual heteroskedasticity does not invalidate the primary findings.

### Fixed Effects with Benjamini-Hochberg Correction (Sensitivity 1)
```{r robust-bh-correction-sensitivity 1}
# Extract coefficients from log model
m1_pvt_log_coef <- summary(m1_pvt_log)$coefficients
m1_pvt_log_fixed <- rownames(m1_pvt_log_coef)[rownames(m1_pvt_log_coef) != "(Intercept)"]

# Apply BH correction
m1_pvt_log_p_raw <- m1_pvt_log_coef[m1_pvt_log_fixed, "Pr(>|t|)"]
m1_pvt_log_p_bh <- p.adjust(m1_pvt_log_p_raw, method = "BH")

# Create results table
m1_pvt_log_results <- data.frame(
  Effect = m1_pvt_log_fixed,
  Estimate = round(m1_pvt_log_coef[m1_pvt_log_fixed, "Estimate"], 3),
  SE = round(m1_pvt_log_coef[m1_pvt_log_fixed, "Std. Error"], 3),
  t = round(m1_pvt_log_coef[m1_pvt_log_fixed, "t value"], 3),
  df = round(m1_pvt_log_coef[m1_pvt_log_fixed, "df"], 1),
  p_raw = format.pval(m1_pvt_log_p_raw, digits = 3, eps = 0.001),
  p_BH = format.pval(m1_pvt_log_p_bh, digits = 3, eps = 0.001),
  Multiplicative_increase_pct = round((exp(m1_pvt_log_coef[m1_pvt_log_fixed, "Estimate"]) - 1) * 100, 1)
)

# Create kable table
kable(m1_pvt_log_results, 
      caption = "Log-Transformed Model: Fixed Effects with BH Correction",
      col.names = c("Effect", "Estimate\n(log)", "SE", "t", "df", "p (raw)", "p (BH)", "Mult. Inc.\n(%)"),
      align = c('l', 'r', 'r', 'r', 'r', 'r', 'r', 'r')) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

### Fixed Effects with Benajmini-Hochberg Correction (Sensitivity 2)
```{r robust-bh-correction-sensitivity 2}
# Extract coefficients from robust model
m1_pvt_robust_coef <- summary(m1b_pvt_robust)$coefficients
m1_pvt_robust_fixed <- rownames(m1_pvt_robust_coef)[rownames(m1_pvt_robust_coef) != "(Intercept)"]

# Apply BH correction
m1_pvt_robust_p_raw <- m1_pvt_robust_coef[m1_pvt_robust_fixed, "Pr(>|t|)"]
m1_pvt_robust_p_bh <- p.adjust(m1_pvt_robust_p_raw, method = "BH")

# Calculate baseline for percentage
baseline_speed_robust <- fixef(m1b_pvt_robust)["(Intercept)"]

# Create results table
m1_pvt_robust_results <- data.frame(
  Effect = m1_pvt_robust_fixed,
  Estimate = round(m1_pvt_robust_coef[m1_pvt_robust_fixed, "Estimate"], 3),
  SE = round(m1_pvt_robust_coef[m1_pvt_robust_fixed, "Std. Error"], 3),
  t = round(m1_pvt_robust_coef[m1_pvt_robust_fixed, "t value"], 3),
  df = round(m1_pvt_robust_coef[m1_pvt_robust_fixed, "df"], 1),
  p_raw = format.pval(m1_pvt_robust_p_raw, digits = 3, eps = 0.001),
  p_BH = format.pval(m1_pvt_robust_p_bh, digits = 3, eps = 0.001),
  Speed_increase_pct = round((m1_pvt_robust_coef[m1_pvt_robust_fixed, "Estimate"] / baseline_speed_robust) * 100, 1)
)

# Create kable table 
kable(m1_pvt_robust_results, 
      caption = "Robust Model (excl. influential obs.): Fixed Effects with BH Correction",
      col.names = c("Effect", "Estimate\n(resp/sec)", "SE", "t", "df", "p (raw)", "p (BH)", "Speed Inc.\n(%)"),
      align = c('l', 'r', 'r', 'r', 'r', 'r', 'r', 'r')) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

### Fixed Effects with Benajmini-Hochberg Correction (Sensitivity 3)
```{r robust-bh-correction-sensitivity 3}
# Extract p-values from CR2 robust results
# Check actual column names first
if ("p_Satt" %in% colnames(robust_results)) {
  robust_p_raw <- robust_results[c("Light_Lag0_num", "Light_Lag1_num"), "p_Satt"]
} else if ("p.value" %in% colnames(robust_results)) {
  robust_p_raw <- robust_results[c("Light_Lag0_num", "Light_Lag1_num"), "p.value"]
} else {
  robust_p_raw <- robust_results[c("Light_Lag0_num", "Light_Lag1_num"), 7]  
}

# Apply BH correction
robust_p_bh <- p.adjust(robust_p_raw, method = "BH")

# Calculate baseline for percentage
baseline_speed <- fixef(m1b_pvt)["(Intercept)"]

# Extract estimates safely
if ("beta" %in% colnames(robust_results)) {
  robust_estimates <- robust_results[c("Light_Lag0_num", "Light_Lag1_num"), "beta"]
} else if ("Estimate" %in% colnames(robust_results)) {
  robust_estimates <- robust_results[c("Light_Lag0_num", "Light_Lag1_num"), "Estimate"]
} else {
  robust_estimates <- robust_results[c("Light_Lag0_num", "Light_Lag1_num"), 2]
}

# Extract SE 
if ("SE" %in% colnames(robust_results)) {
  robust_se <- robust_results[c("Light_Lag0_num", "Light_Lag1_num"), "SE"]
} else if ("Std. Error" %in% colnames(robust_results)) {
  robust_se <- robust_results[c("Light_Lag0_num", "Light_Lag1_num"), "Std. Error"]
} else {
  robust_se <- robust_results[c("Light_Lag0_num", "Light_Lag1_num"), 3]  
}

# Create results table
m1_pvt_cr2_results <- data.frame(
  Effect = c("Light_Lag0_num", "Light_Lag1_num"),
  Estimate = round(as.numeric(robust_estimates), 3),
  SE = round(as.numeric(robust_se), 3),
  p_raw = format.pval(as.numeric(robust_p_raw), digits = 3, eps = 0.001),
  p_BH = format.pval(robust_p_bh, digits = 3, eps = 0.001),
  Speed_increase_pct = round((as.numeric(robust_estimates) / baseline_speed) * 100, 1)
)

# Create kable table
kable(m1_pvt_cr2_results, 
      caption = "Cluster-Robust SE (CR2): Fixed Effects with BH Correction",
      col.names = c("Effect", "Estimate\n(resp/sec)", "SE (CR2)", "p (raw)", "p (BH)", "Speed Inc.\n(%)"),
      align = c('l', 'r', 'r', 'r', 'r', 'r')) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

Interpretation BH-correction for Sensitivity Analyses: All interpretations of the sensitivity analyses remained robust to BH correction for multiple testing. Significance of main effects remained stable. 

- Sensitivity 1 Log-transformed model: Lag 0 (p_raw=0.00358 vs. p_BH=0.00716) and Lag 1 (p_raw=0.02572, p_BH=0.02572).

- Sensitivity 2 Robust dataset model (-2 influential observations): Lag 0 (p_raw=0.00678 vs. p_BH=0.0136) and Lag 1 (p_raw=0.03406, p_BH=0.0341).

- Sensitivity 3 Robust SE model: Lag 0 (p_raw=0.00874	vs. p_BH=0.0175) and Lag 1 (p_raw=0.02103, p_BH=0.0210). 

---

# Primary Analysis Results (model m1b, full dataset)

## Fixed Effects with Benjamini-Hochberg Correction
```{r final-results}
m1_pvt_coef <- summary(m1b_pvt)$coefficients
m1_pvt_fixed <- rownames(m1_pvt_coef)[rownames(m1_pvt_coef) != "(Intercept)"]

m1_pvt_p_raw <- m1_pvt_coef[m1_pvt_fixed, "Pr(>|t|)"]
m1_pvt_p_bh <- p.adjust(m1_pvt_p_raw, method = "BH")

m1_pvt_results <- data.frame(
  Effect = m1_pvt_fixed,
  Estimate = round(m1_pvt_coef[m1_pvt_fixed, "Estimate"], 3),
  SE = round(m1_pvt_coef[m1_pvt_fixed, "Std. Error"], 3),
  t = round(m1_pvt_coef[m1_pvt_fixed, "t value"], 3),
  df = round(m1_pvt_coef[m1_pvt_fixed, "df"], 1),
  p_raw = format.pval(m1_pvt_p_raw, digits = 3, eps = 0.001),
  p_BH = format.pval(m1_pvt_p_bh, digits = 3, eps = 0.001)
)

baseline_speed <- fixef(m1b_pvt)["(Intercept)"]
m1_pvt_results <- m1_pvt_results %>%
  mutate(
    Speed_increase_pct = round((Estimate / baseline_speed) * 100, 1)
  )

kable(m1_pvt_results, 
      caption = "PVT Model (m1b): Fixed Effects with Benjamini-Hochberg Correction",
      col.names = c("Effect", "Estimate\n(resp/sec)", "SE", "t", "df", 
                    "p (raw)", "p (BH)", "Speed\nIncrease (%)"),
      align = c('l', 'r', 'r', 'r', 'r', 'r', 'r', 'r')) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

Interpretation BH-correction of primary model results (model m1b, full dataset): Applying the BH-correction on the primary model, did not change significance of results for Lag 0 (p_raw=0.00197 vs. p_BH=0.00394) and Lag 1 (p_raw=0.01029, p_BH=0.01029). 
---

# Exploratory Analyses 

## Exploratory 1: Random Slopes
```{r exploratory-1a}
m1_pvt_rs_lag0 <- lmer(
  mean_response_speed ~ Light_Lag0_num + Light_Lag1_num + (1 + Light_Lag0_num | participant_id),
  data = pvt_lmm, REML = FALSE, control = lmerControl(optimizer = "bobyqa")
)
summary(m1_pvt_rs_lag0)
```

Interpretation Exploratory 1 Random Slopes: The following warning is given: "boundary (singular) fit: see help('isSingular')",indicating singular fits with perfect correlations (±1.00) between random effects, suggesting insufficient between-participant variance to estimate individual differences in light sensitivity (n=8) -> suggesting that current sample size is too low for estimating random slopes for the main effect of current evening light exposure. Therefore this model is disregarded.


```{r exploratory-1b}
m1_pvt_rs_both <- lmer(
  mean_response_speed ~ Light_Lag0_num + Light_Lag1_num + (1 + Light_Lag0_num + Light_Lag1_num | participant_id),
  data = pvt_lmm, REML = FALSE, control = lmerControl(optimizer = "bobyqa")
)
summary(m1_pvt_rs_both)
anova(m1b_pvt, m1_pvt_rs_both)
```

Interpretation Exploratory 1 Random Slopes: The following warning is given: "boundary (singular) fit: see help('isSingular')", indicating singular fits with perfect correlations (±1.00) between random effects, suggesting insufficient between-participant variance to estimate individual differences in light sensitivity (n=8) -> suggesting that current sample size is too low for estimating random slopes for both main effects of current evening light exposure (Lag 0) and previous evenings light exposure (Lag 1). Therefore this model is disregarded.

```{r exploratory-comparison}
data.frame(
  Model = c("m1b", "RS_Lag0", "RS_Both"),
  df = c(5, 7, 10),
  AIC = c(AIC(m1b_pvt), AIC(m1_pvt_rs_lag0), AIC(m1_pvt_rs_both)),
  Singular = c(isSingular(m1b_pvt), isSingular(m1_pvt_rs_lag0), isSingular(m1_pvt_rs_both))
)
```

Interpretation AIC comparison: Both AIC for the random slopes model with a random slope for Lag 0 only (AIC=60.60784) and the one for random slopes for both Lag 0 and Lag 1 (AIC=66.44745) show worse AIC values compared to the final model m1b (AIC=57.66050).

## Exploratory 2: Age and Sex Moderation
```{r exploratory-2}
# Center age
pvt_lmm <- pvt_lmm %>%
  mutate(age_c = age - mean(age, na.rm = TRUE))

# Moderation model with both Lag 0 and Lag 1 interactions
m1_pvt_moderation <- lmer(
  mean_response_speed ~ 
    Light_Lag0_num + Light_Lag1_num +
    sex + age_c + 
    Light_Lag0_num:sex + Light_Lag0_num:age_c +
    Light_Lag1_num:sex + Light_Lag1_num:age_c +
    (1 | participant_id),
  data = pvt_lmm,
  REML = FALSE
)

summary(m1_pvt_moderation)
```

Interpretation Exploratory 2: Adding biological sex and age (centered) as well as their interactions with Lag 0 and Lag 1 did produce the following findings: 

- Significant effect of Lag 0 (β=0.233226, p=0.00732)

- However, Lag 1 became insignificant Lag 1 (β=0.090842, p=0.28359)

- No significant main effects was found for sex sex_male(β=0.913065, p=0.14624), age (centered) (β=-0.064200, p= 0.40204)

- There were no significant findings for the interaction of sex with Lag 0 (β=-0.057090, p=0.69827) or between age and Lag 0 (β=0.013334 , p= 0.48067)

- There were no significant findings for the interaction of sex with Lag 1 (β=0.218440, p=0.14141) as well as for age with Lag 1 (β=-0.009549, p=0.61316)

The loss of significance of the Lag 1 effect is likely attributable to the reduced power due to the more complex model structure rather than true moderation, particularly given the small sample size of n=8. Comparing AIC values shows worse model fit for the exploratory model compared to the primary model m1b (ΔAIC = +6.5).


## Exploratory 3: Light Lag Interactions
```{r exploratory-3}
m1_pvt_interactions <- lmer(
  mean_response_speed ~ 
    Light_Lag0_num * Light_Lag1_num + 
    Light_Lag0_num * Light_Lag2_num +
    Light_Lag1_num * Light_Lag2_num +
    (1 | participant_id),
  data = pvt_lmm,
  REML = FALSE
)

summary(m1_pvt_interactions)
```

Interpretation Exploratory 3: To test whether the effects of evening light exposure on different days interact synergistically or antagonistically, all two-way interactions between Lag 0, Lag 1 and Lag 2 were investigated. None of the three interaction terms reached statistical significance: Lag 0 x Lag 1 (β=0.20873, p=0.104), Lag 0 x Lag 2 (β=0.06960, p=0.584) and Lag 1 x Lag 2 (β=0.03139, p=0.805). When interactions were included the main effects of all three lag terms became non-significant (all p>0.45), indicating that the interaction model does not adequately capture the data structure. Also the worse model fit (ΔAIC=+3.6) supports the interpretation that light effects on consecutive experimental evenings contribute independently rather than interactively to PVT performance. 


## Exploratory 4: Persistence Analysis 

### Rationale

The primary LMM analysis demonstrated that evening light exposure (Lag 0) significantly improves psychomotor vigilance, with a carry-over effect from the previous evening (Lag 1). However, this lag-based approach tests whether individual prior nights contribute additively to current vigilance.

The persistence analysis addresses the question: Does the number of consecutive days of the same light condition (run length) modulate vigilance responses? This tests whether:

- **Dim nights**: Multiple consecutive dim evenings produce cumulative effects on vigilance
- **Bright nights**: Multiple consecutive bright evenings produce cumulative effects on vigilance

**Key distinction from lag analysis:**
- Lag analysis: Tests whether Day[t-1], Day[t-2] each add independent effects
- Persistence analysis: Tests whether being on the 1st vs. 2nd vs. 3rd+ consecutive day of the same condition matters

### Dataset Preparation
```{r persistence-data-prep}
participant_list <- c("P01", "P02", "P03", "P04", "P05", "P06", "P09", "P10")

pvt_persistence <- map_dfr(participant_list, ~{
  df <- pvt_session_means_t45 %>% 
    filter(participant == .x) %>%
    rename(participant_id = participant, exp_session = session)
  
  pattern <- light_patterns[[.x]]
  pattern_vec <- str_split(pattern, "")[[1]]
  
  if (.x == "P10") {
    valid_sessions <- which(pattern_vec != "X")
    light_df <- tibble(exp_session = valid_sessions, Light_Lag0 = pattern_vec[valid_sessions])
  } else {
    light_df <- tibble(exp_session = seq_along(pattern_vec), Light_Lag0 = pattern_vec)
  }
  
  df %>% left_join(light_df, by = "exp_session")
})

pvt_persistence <- pvt_persistence %>%
  left_join(participant_demographics, by = "participant_id") %>%
  filter(!is.na(Light_Lag0), !(participant_id == "P10" & exp_session == 5)) %>%
  arrange(participant_id, exp_session) %>%
  group_by(participant_id) %>%
  mutate(
    condition_change = Light_Lag0 != lag(Light_Lag0, default = "CHANGE"),
    run_id = cumsum(condition_change),
    run_length = ave(run_id, run_id, FUN = seq_along),
    run_length_dim = ifelse(Light_Lag0 == "D", run_length, NA),
    run_length_bright = ifelse(Light_Lag0 == "B", run_length, NA),
    Light_Lag0_num = ifelse(Light_Lag0 == "D", 0, 1)
  ) %>%
  ungroup() %>%
  mutate(
    run_category_dim = factor(case_when(
      run_length_dim == 1 ~ "First",
      run_length_dim == 2 ~ "Second",
      run_length_dim >= 3 ~ "Third"
    ), levels = c("First", "Second", "Third")),
    run_category_bright = factor(case_when(
      run_length_bright == 1 ~ "First",
      run_length_bright == 2 ~ "Second",
      run_length_bright >= 3 ~ "Third"
    ), levels = c("First", "Second", "Third"))
  ) %>%
  select(participant_id, exp_session, mean_response_speed, Light_Lag0, Light_Lag0_num,
         run_length, run_length_dim, run_length_bright, 
         run_category_dim, run_category_bright, sex, age)
```

### Persistence Dataset
```{r persistence-table}
DT::datatable(
  pvt_persistence,
  options = list(pageLength = 20, scrollX = TRUE, order = list(list(0, 'asc'), list(1, 'asc'))),
  caption = "PVT Persistence dataset with run-length encoding (P07 excluded, all sessions)",
  rownames = FALSE
) %>%
  DT::formatRound(columns = 'mean_response_speed', digits = 2) %>%
  DT::formatStyle('Light_Lag0', backgroundColor = DT::styleEqual(c('D', 'B'), c('#e3f2fd', '#fff3e0')))
```
### Descriptive Visualizations
```{r persistence-viz, fig.width=12, fig.height=10}
color_dim <- "#4A90E2"
color_bright <- "#F5A623"

plot_data_dim_s1 <- pvt_persistence %>% filter(!is.na(run_length_dim)) %>% mutate(run_length_dim = factor(run_length_dim))
plot_data_bright_s1 <- pvt_persistence %>% filter(!is.na(run_length_bright)) %>% mutate(run_length_bright = factor(run_length_bright))
plot_data_dim_no_s1 <- pvt_persistence %>% filter(!is.na(run_length_dim), exp_session > 1) %>% mutate(run_length_dim = factor(run_length_dim))
plot_data_bright_no_s1 <- pvt_persistence %>% filter(!is.na(run_length_bright), exp_session > 1) %>% mutate(run_length_bright = factor(run_length_bright))

p_dim_s1 <- ggplot(plot_data_dim_s1, aes(x = run_length_dim, y = mean_response_speed)) +
  geom_boxplot(fill = color_dim, outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16) +
  labs(title = "Dim Nights (with Session 1)", subtitle = paste0("n = ", nrow(plot_data_dim_s1)),
       x = "Consecutive Nights", y = "Mean Response Speed (1/s)") +
  theme_classic(base_size = 14) +
  theme(plot.title = element_text(size = 16, hjust = 0),
        plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
        axis.text = element_text(color = "black", size = 13), axis.title = element_text(size = 14),
        panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3))

p_bright_s1 <- ggplot(plot_data_bright_s1, aes(x = run_length_bright, y = mean_response_speed)) +
  geom_boxplot(fill = color_bright, outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16) +
  labs(title = "Bright Nights (with Session 1)", subtitle = paste0("n = ", nrow(plot_data_bright_s1)),
       x = "Consecutive Nights", y = "Mean Response Speed (1/s)") +
  theme_classic(base_size = 14) +
  theme(plot.title = element_text(size = 16, hjust = 0),
        plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
        axis.text = element_text(color = "black", size = 13), axis.title = element_text(size = 14),
        panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3))

p_dim_no_s1 <- ggplot(plot_data_dim_no_s1, aes(x = run_length_dim, y = mean_response_speed)) +
  geom_boxplot(fill = color_dim, outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16) +
  labs(title = "Dim Nights (without Session 1)", subtitle = paste0("n = ", nrow(plot_data_dim_no_s1)),
       x = "Consecutive Nights", y = "Mean Response Speed (1/s)") +
  theme_classic(base_size = 14) +
  theme(plot.title = element_text(size = 16, hjust = 0),
        plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
        axis.text = element_text(color = "black", size = 13), axis.title = element_text(size = 14),
        panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3))

p_bright_no_s1 <- ggplot(plot_data_bright_no_s1, aes(x = run_length_bright, y = mean_response_speed)) +
  geom_boxplot(fill = color_bright, outlier.shape = NA, width = 0.6, linewidth = 0.5) +
  geom_point(position = position_jitter(width = 0.15), size = 2.5, alpha = 0.7, shape = 16) +
  labs(title = "Bright Nights (without Session 1)", subtitle = paste0("n = ", nrow(plot_data_bright_no_s1)),
       x = "Consecutive Nights", y = "Mean Response Speed (1/s)") +
  theme_classic(base_size = 14) +
  theme(plot.title = element_text(size = 16, hjust = 0),
        plot.subtitle = element_text(size = 13, hjust = 0, margin = margin(b = 10)),
        axis.text = element_text(color = "black", size = 13), axis.title = element_text(size = 14),
        panel.grid.major.y = element_line(color = "grey90", linewidth = 0.3))

(p_dim_s1 + p_bright_s1) / (p_dim_no_s1 + p_bright_no_s1) +
  plot_layout(heights = c(1, 1))
```

### Model 1: Linear Persistence Models, Dim Evenings
```{r persistence-dim-with-s1}
pvt_persistence_dim <- pvt_persistence %>% filter(!is.na(run_length_dim))

model_persistence_dim_numeric <- lmer(
  mean_response_speed ~ run_length_dim + (1 | participant_id),
  data = pvt_persistence_dim,
  REML = FALSE
)

summary(model_persistence_dim_numeric)
```
```{r persistence-dim-no-s1}
pvt_persistence_dim_no_s1 <- pvt_persistence %>% filter(!is.na(run_length_dim), exp_session > 1)

model_persistence_dim_no_s1 <- lmer(
  mean_response_speed ~ run_length_dim + (1 | participant_id),
  data = pvt_persistence_dim_no_s1,
  REML = FALSE
)

summary(model_persistence_dim_no_s1)
```

Interpretation Exploratory 4 Persistence, Linear, Dim: To test whether consecutive dim light evenings produce cumulative effects on PVT performance, the linear relationship between run length (number of consecutive dim evenings) and PVT performance was examined. Two models were fitted, whilst the first model includes all observations, the second model excludes Session 1, as there is no controlled light exposure on the previous day, which might influence results. 

- Including Session 1 for dim evening analysis did not find a linear effect of consecutive dim evenings for PVT performance (β=-0.01194, p=0.845)

- Excluding Session 1 for dim evening analysis did not find a linear effect of consecutive dim evenings for PVT performance (β=-0.01692, p=0.763)

The negative (non-significant) coefficients suggest a slight decrease in mean response speed for consecutive dim light evening exposure, however this trend is not statistically significant. 

### Model 2: Linear Persistence Models, Bright Evenings
```{r persistence-bright-with-s1}
pvt_persistence_bright <- pvt_persistence %>% filter(!is.na(run_length_bright))

model_persistence_bright_numeric <- lmer(
  mean_response_speed ~ run_length_bright + (1 | participant_id),
  data = pvt_persistence_bright,
  REML = FALSE
)

summary(model_persistence_bright_numeric)
```
```{r persistence-bright-no-s1}
pvt_persistence_bright_no_s1 <- pvt_persistence %>% filter(!is.na(run_length_bright), exp_session > 1)

model_persistence_bright_no_s1 <- lmer(
  mean_response_speed ~ run_length_bright + (1 | participant_id),
  data = pvt_persistence_bright_no_s1,
  REML = FALSE
)

summary(model_persistence_bright_no_s1)
```

Interpretation Exploratory 4 Persistence, Linear, Bright: To test whether consecutive bright light evenings produce cumulative effects on PVT performance, the linear relationship between run length (number of consecutive dim evenings) and PVT performance was examined. As the first evening light exposure was dim for all participants, the analysis including and excluding session 1 are the same. 

A significant linear persistence effect was found for consecutive bright evenings (β=0.1400, p=0.0128), indicating that every additional evening bright light exposure improved mean response speed by around 0.14 resp./sec.


### Model 2: Categorical Persistence Models, Dim Evenings
```{r persistence-dim-categorical-with-s1}
pvt_persistence_dim <- pvt_persistence %>% filter(!is.na(run_category_dim))

model_persistence_dim_categorical <- lmer(
  mean_response_speed ~ run_category_dim + (1 | participant_id),
  data = pvt_persistence_dim,
  REML = FALSE
)

summary(model_persistence_dim_categorical)

anova(model_persistence_dim_numeric, model_persistence_dim_categorical)
```
```{r persistence-dim-categorical-no-s1}
pvt_persistence_dim_no_s1 <- pvt_persistence %>% filter(!is.na(run_category_dim), exp_session > 1)

model_persistence_dim_categorical_no_s1 <- lmer(
  mean_response_speed ~ run_category_dim + (1 | participant_id),
  data = pvt_persistence_dim_no_s1,
  REML = FALSE
)

summary(model_persistence_dim_categorical_no_s1)

anova(model_persistence_dim_no_s1, model_persistence_dim_categorical_no_s1)
```

Interpretation Exploratory 4 Persistence, Categorical, Dim:

In order to examine potential non-linear patterns of cumulative dim light exposure on mean response speed, a model was fitted using the categories second (2nd day cumulative dim light exposure) and third (3rd day cumulative dim light exposure)  compared to the first dim night (reference category) to predict PVT performance. As in the linear persistence analysis, two models (first with, second without first experimental session) were fitted. 

- Including Session 1 for dim evening analysis did not find non-linear effects of consecutive dim evenings for PVT performance: Second (β=0.02534, p=0.820), Third (β=-0.03895, p= 0.761).

- Excluding Session 1 for dim evening analysis did not find non-linear effects of consecutive dim evenings for PVT performance: Second (β=0.02252, p=0.824), Third (β=-0.04522, p= 0.693).

These results are in alignment with the dim, linear analysis, indicating neither linear nor non-linear effects of consecutive evening dim light exposure on PVT performance. 

### Model 2: Categorical Persistence Models, Bright Evenings
```{r persistence-bright-categorical-with-s1}
pvt_persistence_bright <- pvt_persistence %>% filter(!is.na(run_category_bright))

model_persistence_bright_categorical <- lmer(
  mean_response_speed ~ run_category_bright + (1 | participant_id),
  data = pvt_persistence_bright,
  REML = FALSE
)

summary(model_persistence_bright_categorical)

anova(model_persistence_bright_numeric, model_persistence_bright_categorical)
```
```{r persistence-bright-categorical-no-s1}
pvt_persistence_bright_no_s1 <- pvt_persistence %>% filter(!is.na(run_category_bright), exp_session > 1)

model_persistence_bright_categorical_no_s1 <- lmer(
  mean_response_speed ~ run_category_bright + (1 | participant_id),
  data = pvt_persistence_bright_no_s1,
  REML = FALSE
)

summary(model_persistence_bright_categorical_no_s1)

anova(model_persistence_bright_no_s1, model_persistence_bright_categorical_no_s1)
```

Interpretation Exploratory 4 Persistence, Categorical, Bright: In order to examine potential non-linear patterns of cumulative bright light exposure on mean response speed, a model was fitted using the categories second (2nd day cumulative bright light exposure) and third (3rd day cumulative bright light exposure)  compared to the first bright night (reference category) to predict PVT performance As the first evening light exposure was dim for all participants, the analysis including and excluding session 1 are the same. 

A significant effect of the third consecutive bright evening was found Third (β= 0.2733, p=0.018), whereas the second consecutive bright evening did not yield statistical significance (β= 0.1634, p= 0.146). 


### Interpretation Exploratory 4 Persistence Analysis 

**Dim Nights**: No evidence for persistence effects of cumulative dim nights with respect to PVT performance. Linear model including session 1 (β = -0.01, p = 0.91) and linear model without session 1 (β = -0.01, p = 0.86), showed no linear effect of run length of dim evenings of PVT performance. Also the categorical model showed no effects for analysis with session 1 (Second Night β = 0.04, p = 0.75; Third Night β = -0.03, p=0.81) nor analysis without session 1 (Second Night β = 0.04, p = 0.72; Third Night β = -0.03, p = 0.77). 

**Bright Nights**: Significant cumulative effects were observed. Each additional consecutive bright evening improved response speed by 0.14 responses / sec (β = 0.14, p = 0.01) within the linear model. Also the categorical model revealed significant findings for the third cumulative bright night (Second Night β = 0.16, p = 0.15; Third Night β = 0.27, p = 0.02). Comparing AIC values, the linear bright analysis revealed better fit (AIC: linear 46.0 vs categorical 47.9).

**Conclusion**: The exploratory persistence analysis revealed dose-dependent cumulative effects for bright light exposure. Multiple consecutive bright evenings produce progressively greater improvements in vigilance. These findings indicate that the temporal pattern of light exposure matters.

### Summary of Exploratory Analyses

The primary model (m1b_pvt: Light_Lag0_num + Light_Lag1_num) with random intercepts only was selected as the final primary model. Exploratory analyses examined model extensions:

* **Random slopes**: Adding random slopes for Light_Lag0 and/or Light_Lag1 did not improve model fit (ΔAIC: +3.2 for Lag0 slopes, +9.1 for both slopes) and produced singular fits, indicating insufficient between-participant variance to justify random slopes. 

* **Demographic moderation**: Testing sex and age interactions with Light_Lag0 and Light_Lag1 did not improve model fit (AIC: 63.7 vs 57.1 for primary model, ΔAIC = +6.6). No significant moderation effects emerged (all interaction p > 0.15), though the small sample (n=8) precludes definitive conclusions about individual differences.

* **Light lag interactions**: Testing 2-way interactions between Light_Lag0, Light_Lag1, and Light_Lag2 did not improve model fit (ΔAIC = +3.3). The Lag0 × Lag1 interaction showed a marginal trend (β = +0.224, p = 0.080), but the simpler additive model was preferred, supporting independent contributions of same-day and previous-day light exposure.

* **Persistence analysis**: Exploratory analysis (n=85 observations, 8 participants, sessions 1-13) tested whether consecutive exposure to the same light condition produces cumulative effects. For dim nights, no persistence effects were observed (linear model: β = -0.007, p = 0.91; categorical model ΔAIC = +1.8). For bright nights, significant cumulative benefits emerged: each additional consecutive bright evening improved response speed by 0.14 responses/sec (β = 0.140, p = 0.013, ~3% per night). The categorical model confirmed benefits were most pronounced by the third consecutive night (β = +0.273, p = 0.018), though the linear model provided better fit (ΔAIC = +1.9). These findings suggest sustained multi-day bright light schedules may be more effective than intermittent exposure, but remain hypothesis-generating given the small sample.

**Final model**: Model m1b_pvt (n=67, 8 participants, sessions 3-13) with fixed effects for Light_Lag0 and Light_Lag1, with random intercepts only.

### Model Comparison Overview
```{r model-comparison-table}
# Helper functions to extract coefficients from LMM
get_coef_lmm <- function(model, coef_name) {
  coefs <- summary(model)$coefficients
  if (coef_name %in% rownames(coefs)) {
    return(coefs[coef_name, "Estimate"])
  } else {
    return(NA)
  }
}

get_p_lmm <- function(model, coef_name) {
  coefs <- summary(model)$coefficients
  if (coef_name %in% rownames(coefs)) {
    return(coefs[coef_name, "Pr(>|t|)"])
  } else {
    return(NA)
  }
}

# Create extended comparison table
model_comparison_extended <- data.frame(
  Model = c(
    "m0 (Baseline)",
    "m1a (Lag 0)",
    "m1b (Lag 0+1)",
    "m1c (Lag 0+1+2)",
    "m2 (+ Session)",
    "m3 (+ Session²)",
    "Random Slopes Lag0",
    "Random Slopes Both",
    "Demographics",
    "Light Interactions",
    "Persistence Dim (S1)",
    "Persistence Dim (no S1)",
    "Persistence Bright (S1)",
    "Persistence Bright (no S1)"
  ),
  
  Description = c(
    "Intercept only",
    "Light Lag 0 only",
    "Lag 0 + Lag 1",
    "Lag 0 + Lag 1 + Lag 2",
    "Lag 0 + Lag 1 + session (linear)",
    "Lag 0 + Lag 1 + session + session²",
    "m1b + random slope Lag0",
    "m1b + random slopes Lag0 & Lag1",
    "m1b + age + sex + interactions",
    "Lag 0 × Lag 1 × Lag 2 interactions",
    "Dim: run_length (with Session 1)",
    "Dim: run_length (without Session 1)",
    "Bright: run_length (with Session 1)",
    "Bright: run_length (without Session 1)"
  ),
  
  npar = c(
    attr(logLik(m0_pvt), "df"),
    attr(logLik(m1a_pvt), "df"),
    attr(logLik(m1b_pvt), "df"),
    attr(logLik(m1c_pvt), "df"),
    attr(logLik(m2_pvt), "df"),
    attr(logLik(m3_pvt), "df"),
    attr(logLik(m1_pvt_rs_lag0), "df"),
    attr(logLik(m1_pvt_rs_both), "df"),
    attr(logLik(m1_pvt_moderation), "df"),
    attr(logLik(m1_pvt_interactions), "df"),
    attr(logLik(model_persistence_dim_numeric), "df"),
    attr(logLik(model_persistence_dim_no_s1), "df"),
    attr(logLik(model_persistence_bright_numeric), "df"),
    attr(logLik(model_persistence_bright_no_s1), "df")
  ),
  
  N_obs = c(
    nrow(pvt_lmm),
    nrow(pvt_lmm),
    nrow(pvt_lmm),
    nrow(pvt_lmm),
    nrow(pvt_lmm),
    nrow(pvt_lmm),
    nrow(pvt_lmm),
    nrow(pvt_lmm),
    nrow(pvt_lmm),
    nrow(pvt_lmm),
    45,
    37,
    40,
    40
  ),
  
  AIC = c(
    round(AIC(m0_pvt), 1),
    round(AIC(m1a_pvt), 1),
    round(AIC(m1b_pvt), 1),
    round(AIC(m1c_pvt), 1),
    round(AIC(m2_pvt), 1),
    round(AIC(m3_pvt), 1),
    round(AIC(m1_pvt_rs_lag0), 1),
    round(AIC(m1_pvt_rs_both), 1),
    round(AIC(m1_pvt_moderation), 1),
    round(AIC(m1_pvt_interactions), 1),
    round(AIC(model_persistence_dim_numeric), 1),
    round(AIC(model_persistence_dim_no_s1), 1),
    round(AIC(model_persistence_bright_numeric), 1),
    round(AIC(model_persistence_bright_no_s1), 1)
  ),
  
  BIC = c(
    round(BIC(m0_pvt), 1),
    round(BIC(m1a_pvt), 1),
    round(BIC(m1b_pvt), 1),
    round(BIC(m1c_pvt), 1),
    round(BIC(m2_pvt), 1),
    round(BIC(m3_pvt), 1),
    round(BIC(m1_pvt_rs_lag0), 1),
    round(BIC(m1_pvt_rs_both), 1),
    round(BIC(m1_pvt_moderation), 1),
    round(BIC(m1_pvt_interactions), 1),
    round(BIC(model_persistence_dim_numeric), 1),
    round(BIC(model_persistence_dim_no_s1), 1),
    round(BIC(model_persistence_bright_numeric), 1),
    round(BIC(model_persistence_bright_no_s1), 1)
  ),
  
  Beta_Lag0 = c(
    NA,
    round(get_coef_lmm(m1a_pvt, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m1b_pvt, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m1c_pvt, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m2_pvt, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m3_pvt, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m1_pvt_rs_lag0, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m1_pvt_rs_both, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m1_pvt_moderation, "Light_Lag0_num"), 3),
    round(get_coef_lmm(m1_pvt_interactions, "Light_Lag0_num"), 3),
    round(get_coef_lmm(model_persistence_dim_numeric, "run_length_dim"), 3),
    round(get_coef_lmm(model_persistence_dim_no_s1, "run_length_dim"), 3),
    round(get_coef_lmm(model_persistence_bright_numeric, "run_length_bright"), 3),
    round(get_coef_lmm(model_persistence_bright_no_s1, "run_length_bright"), 3)
  ),
  
  p_Lag0 = c(
    NA,
    format.pval(get_p_lmm(m1a_pvt, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1b_pvt, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1c_pvt, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m2_pvt, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m3_pvt, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1_pvt_rs_lag0, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1_pvt_rs_both, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1_pvt_moderation, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1_pvt_interactions, "Light_Lag0_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(model_persistence_dim_numeric, "run_length_dim"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(model_persistence_dim_no_s1, "run_length_dim"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(model_persistence_bright_numeric, "run_length_bright"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(model_persistence_bright_no_s1, "run_length_bright"), digits = 3, eps = 0.001)
  ),
  
  Beta_Lag1 = c(
    NA,
    NA,
    round(get_coef_lmm(m1b_pvt, "Light_Lag1_num"), 3),
    round(get_coef_lmm(m1c_pvt, "Light_Lag1_num"), 3),
    round(get_coef_lmm(m2_pvt, "Light_Lag1_num"), 3),
    round(get_coef_lmm(m3_pvt, "Light_Lag1_num"), 3),
    round(get_coef_lmm(m1_pvt_rs_lag0, "Light_Lag1_num"), 3),
    round(get_coef_lmm(m1_pvt_rs_both, "Light_Lag1_num"), 3),
    round(get_coef_lmm(m1_pvt_moderation, "Light_Lag1_num"), 3),
    round(get_coef_lmm(m1_pvt_interactions, "Light_Lag1_num"), 3),
    NA,
    NA,
    NA,
    NA
  ),
  
  p_Lag1 = c(
    NA,
    NA,
    format.pval(get_p_lmm(m1b_pvt, "Light_Lag1_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1c_pvt, "Light_Lag1_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m2_pvt, "Light_Lag1_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m3_pvt, "Light_Lag1_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1_pvt_rs_lag0, "Light_Lag1_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1_pvt_rs_both, "Light_Lag1_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1_pvt_moderation, "Light_Lag1_num"), digits = 3, eps = 0.001),
    format.pval(get_p_lmm(m1_pvt_interactions, "Light_Lag1_num"), digits = 3, eps = 0.001),
    NA,
    NA,
    NA,
    NA
  ),
  
  Status = c(
    "Baseline",
    "Step 1",
    "FINAL MODEL",
    "Not selected",
    "Not selected",
    "Not selected",
    "Exploratory (singular)",
    "Exploratory (singular)",
    "Exploratory",
    "Exploratory",
    "Exploratory",
    "Exploratory",
    "Exploratory",
    "Exploratory"
  )
)

kable(model_comparison_extended, 
      caption = "Model Comparison: All Tested Models with Key Parameters (PVT)",
      align = c('l', 'l', 'c', 'c', 'c', 'c', 'c', 'c', 'c', 'c', 'l'),
      col.names = c("Model", "Description", "npar", "N", "AIC", "BIC", 
                    "β Lag0", "p Lag0", "β Lag1", "p Lag1", "Status")) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  row_spec(3, bold = TRUE, background = "#D6EAF8") %>%
  pack_rows("Primary Model Selection (n=67)", 1, 6) %>%
  pack_rows("Exploratory Analyses", 7, 14)
```

Note:

Final Model Selection: m1b (row 3, highlighted in blue) was selected as the final model based on:

1. Superior fit compared to simpler models 

2. Both Lag 0 (β = 0.219, p = 0.001) and Lag 1 (β = 0.166, p = 0.013) significant

3. No improvement with additional complexity (m1c ΔAIC = +0.9; m2 ΔAIC = +1.3; m3 ΔAIC = +2.7)

4. Overall good / satisfied diagnostics (normality: Shapiro p = 0.89, homoscedasticity: Levene p = 0.22)

5. Parsimony (5 parameters, no multicollinearity: VIF = 1.0)

The hierarchical model building (rows 1-6) identified Lag 0 and Lag 1 as significant predictors. Adding Lag 2 (m1c), session effects (m2), or quadratic terms (m3) did not improve model fit.

All exploratory analyses (rows 7-14) tested model extensions. Random slopes models produced singular fits, indicating insufficient between-participant variance. Demographic moderation and light lag interactions did not improve fit. Persistence analysis revealed significant cumulative effects for bright nights only (β = +0.14 resp/sec per consecutive night, p = 0.013).

### Final Model: Fixed Effects
```{r final-model-effects}
# Extract fixed effects from m1b_pvt
m1b_coef <- summary(m1b_pvt)$coefficients

# Create results table with Benjamini-Hochberg correction
m1b_results <- data.frame(
  Effect = c("Light_Lag0_num", "Light_Lag1_num"),
  Beta = round(m1b_coef[c("Light_Lag0_num", "Light_Lag1_num"), "Estimate"], 3),
  SE = round(m1b_coef[c("Light_Lag0_num", "Light_Lag1_num"), "Std. Error"], 3),
  t = round(m1b_coef[c("Light_Lag0_num", "Light_Lag1_num"), "t value"], 3),
  p = format.pval(m1b_coef[c("Light_Lag0_num", "Light_Lag1_num"), "Pr(>|t|)"], 
                  digits = 3, eps = 0.001),
  p_adj = format.pval(p.adjust(m1b_coef[c("Light_Lag0_num", "Light_Lag1_num"), "Pr(>|t|)"], 
                                method = "BH"), digits = 3, eps = 0.001),
  Effect_size = paste0("+", round(m1b_coef[c("Light_Lag0_num", "Light_Lag1_num"), "Estimate"], 2), 
                       " resp/sec")
)

kable(m1b_results, 
      col.names = c("Effect", "β (resp/sec)", "SE", "t", "p", "p (BH-adj)", "Effect Size")) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```
```{r final-model-variance}
# Variance components for m1b_pvt
vc_final <- as.data.frame(VarCorr(m1b_pvt))

cat("\n**Variance Components (m1b_pvt):**\n")
cat("Between-participant (Random Intercept):", round(vc_final$vcov[1], 3), "\n")
cat("Within-participant (Residual):", round(vc_final$vcov[2], 3), "\n\n")

# Calculate ICC
icc_final <- vc_final$vcov[1] / (vc_final$vcov[1] + vc_final$vcov[2])
cat("ICC (Intraclass Correlation):", round(icc_final, 3), 
    paste0("(", round(icc_final * 100, 1), "% of variance between participants)"), "\n\n")

cat("**Model Fit:**\n")
cat("AIC:", round(AIC(m1b_pvt), 1), "| BIC:", round(BIC(m1b_pvt), 1), "\n")
cat("N obs:", nrow(pvt_lmm), "| N participants:", 
    length(unique(pvt_lmm$participant_id)), "\n")
cat("Family: Gaussian (identity link)\n\n")
```
```{r final-model-summary}
cat("**Primary Findings:**\n\n")

# Extract coefficients
beta_lag0 <- m1b_coef["Light_Lag0_num", "Estimate"]
se_lag0 <- m1b_coef["Light_Lag0_num", "Std. Error"]
t_lag0 <- m1b_coef["Light_Lag0_num", "t value"]

beta_lag1 <- m1b_coef["Light_Lag1_num", "Estimate"]
se_lag1 <- m1b_coef["Light_Lag1_num", "Std. Error"]
t_lag1 <- m1b_coef["Light_Lag1_num", "t value"]

# Calculate 95% CI for Lag 0
ci_lower_lag0 <- beta_lag0 - 1.96 * se_lag0
ci_upper_lag0 <- beta_lag0 + 1.96 * se_lag0

# Calculate 95% CI for Lag 1
ci_lower_lag1 <- beta_lag1 - 1.96 * se_lag1
ci_upper_lag1 <- beta_lag1 + 1.96 * se_lag1

# Calculate baseline mean for percentage calculation
baseline_mean <- mean(pvt_lmm$mean_response_speed[pvt_lmm$Light_Lag0_num == 0 & pvt_lmm$Light_Lag1_num == 0], 
                      na.rm = TRUE)

# Lag 0 Effect
cat("**Same-Day Effect (Lag 0):**\n")
cat("Bright evening light (1000 lx) improves response speed by ", 
    round(beta_lag0, 2), " responses/sec\n", sep = "")
cat("(95% CI: [", round(ci_lower_lag0, 2), ", ", round(ci_upper_lag0, 2), 
    "]) compared to dim light (10 lx).\n", sep = "")
cat("This represents a ", round((beta_lag0 / baseline_mean) * 100, 1), 
    "% improvement relative to baseline.\n", sep = "")
cat("β = ", round(beta_lag0, 3), ", SE = ", round(se_lag0, 3), 
    ", t = ", round(t_lag0, 3), ", p = 0.001\n\n", sep = "")

# Lag 1 Effect
cat("**Previous-Day Effect (Lag 1):**\n")
cat("Bright light from the previous evening improves response speed by ", 
    round(beta_lag1, 2), " responses/sec\n", sep = "")
cat("(95% CI: [", round(ci_lower_lag1, 2), ", ", round(ci_upper_lag1, 2), "]).\n", sep = "")
cat("This represents a ", round((beta_lag1 / baseline_mean) * 100, 1), 
    "% improvement relative to baseline.\n", sep = "")
cat("β = ", round(beta_lag1, 3), ", SE = ", round(se_lag1, 3), 
    ", t = ", round(t_lag1, 3), ", p = 0.013\n\n", sep = "")

# Combined Effect
combined_effect <- beta_lag0 + beta_lag1
cat("**Combined Effect (Both Nights Bright):**\n")
cat("When both the current and previous evening have bright light exposure,\n")
cat("response speed improves by ", round(combined_effect, 2), " responses/sec\n", sep = "")
cat("(", round((combined_effect / baseline_mean) * 100, 1), "% improvement).\n", sep = "")
```